### 2021-03-25 analysis/xor-some-prolific-pilot1.Rmd
start first Prolific pilot analysis

---
title: "XOR-Some Prolific Pilot 1"
author: "Polina Tsvilodub"
date: "3/24/2021"
output: github_document
---

In this first pilot on Prolific we test the implementation of the xor study in magpie. The structure of the experiment was as follows:
Participants read instructions, completed three example trials, and then completed 8 main blocks consisting of 4 xor and 4 some items. Each main block had the following structure: Participants read the background story, answered one comprehension question, then answered competence / relevance / prior questions in randomized order; then they read another 3 comprehension questions, after which the critical utterance was added below the background story. They answered the inference strength question, and then competence / relevance questions in randomized order again.

N=10 participants were recruited for this pilot and compensated 2 pounds/participant.

```{r libraries}
library(tidyverse)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r anonymization, include=FALSE}
#d_raw <- read_csv("~/Documents/Research/XOR/results_58_xor-some-Prolific-pilot1_N10.csv")
# remove IDs
#d_raw %>% select(-prolific_id) %>% write_csv("../data/pilots/results_58_xor-some-Prolific-pilot1_N10.csv") 
```

```{r data}
d <- read_csv("../data/pilots/results_58_xor-some-Prolific-pilot1_N10.csv")
```

Checking if there are any comments indicating technical issues:
```{r comments}
d %>% distinct(comments)
```

Check native languages. In the main experiment, participants not indicating English as (one of) their native language(s) will be excluded. We also compute some participant demographics.
```{r languages}
d %>% distinct(languages)
# exclude non-natives if necessary
d_native <- d

d_native %>% pull(age) %>% mean(., na.rm = T) 
d_native %>% count(gender) %>% mutate(n = n/88)
```

Next, we check whether all the conditions were used correctly.

```{r counts}
# check xor/some vs. trial type
d_native %>% count(main_type, condition) 
# check xor/some vs. experimental condition
d_native %>% count(main_type, exp_condition)
# count items used
d_native %>% count(title)
```

Next, we would exclude participants based on their ratings in the main / example trials (and possibly the bot check) according to preregistered exclusion criteria. For now, all the data is considered.
```{r clean}
d_native %>% group_by(submission_id) %>% count(response) %>% View()
d_main <- d_native %>% filter(trial_name != "example")
d_exmpl <- d_native %>% filter(trial_name == "example")
d_critical <- d_main %>% filter(condition == "critical")

# get overall mean ratings / subject
d_native %>% group_by(submission_id) %>% summarise(mean_rating = mean(response)) %>% arrange(mean_rating)
```
Plot responses on example questions by question type

Plot test questions by type
```{r}
d_test <- d_main %>% rowwise() %>% filter(condition == "test") %>% 
  mutate(test_condition = substr(test_question, 6, 9),
         test_condition = ifelse(test_condition == "fals", "false", 
                                 ifelse(test_condition == "unce", "uncertain",
                                        test_condition)))

d_test %>% 
  ggplot(., aes(x = test_condition, y = response)) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1))
# add mean / CI
```

Plot main rel / comp / pri questions by main condition

```{r}
d_critical <- d_critical %>% 
  pivot_longer(c(competence, relevance, prior), 
               names_to = "class_condition", 
               values_to = "prior_class")
d_critical %>% 
  filter(block != "xor" & block != "some") %>%
  ggplot(., aes(x = as.factor(prior_class), y = response)) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
    facet_wrap(main_type~class_condition)
```
Plot main rel / comp questions by main condition, separated into with / without critical utterance
```{r}
# add info to already record this
d_critical <- d_critical %>% 
  mutate(w_utterance = ifelse(is.na(critical_question), F, T))

d_critical %>% 
  filter(block != "xor" & block != "some") %>%
  ggplot(., aes(x = as.factor(prior_class), y = response, color = w_utterance)) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
    facet_wrap(main_type~class_condition) # get ratings from the respective trials only 
# use block for getting the actual correct ratings, but class_condition for subsetting the prior classification fo the respective predictor 
```

plot inference rating as a function of respective rating of the explanatory factor (think about a plot where one would see an interaction)

### 2021-03-25 analysis/xor-some-prolific-pilot1.Rmd
start first Prolific pilot analysis

@@
 d_critical <- d_critical %>% 
-  mutate(w_utterance = ifelse(is.na(critical_question), F, T))
+  mutate(w_utterance = ifelse(is.na(critical_question), F, T),
+         block = ifelse(block == "comp", "competence", 
+                        ifelse(block == "rel", "relevance", "prior")))
 
@@
   filter(block != "xor" & block != "some") %>%
+  filter(block == class_condition) %>%
   ggplot(., aes(x = as.factor(prior_class), y = response, color = w_utterance)) +
@@
 # use block for getting the actual correct ratings, but class_condition for subsetting the prior classification fo the respective predictor
+
+# some condition was wrongly recorded as pri
+d_critical %>% 
+     filter(block != "xor" & block != "some") %>%
+     filter(block == class_condition) %>% filter(block == "prior" & w_utterance == T) %>% View()
+
 ```

### 2021-03-25 analysis/xor-some-prolific-pilot1.Rmd
knit analysis draft
minor fix

@@
 ```{r clean}
-d_native %>% group_by(submission_id) %>% count(response) %>% View()
+d_native %>% group_by(submission_id) %>% count(response)
 d_main <- d_native %>% filter(trial_name != "example")
@@
          block = ifelse(block == "comp", "competence", 
-                        ifelse(block == "rel", "relevance", "prior")))
+                        ifelse(block == "rel", "relevance", ifelse(block == "pri", "prior", block) )))
 
@@
 
-# some condition was wrongly recorded as pri
+# # some condition was wrongly recorded as pri
+# d_critical %>% 
+#      filter(block != "xor" & block != "some") %>%
+#      filter(block == class_condition) %>% filter(block == "prior" & w_utterance == T) %>%
+#   View()
+
 d_critical %>% 
   filter(block != "xor" & block != "some") %>%
-     filter(block == class_condition) %>% filter(block == "prior" & w_utterance == T) %>% View()
-
+  filter(block == class_condition) %>%
+  group_by(main_type, class_condition, w_utterance, prior_class) %>% 
+  summarize(mean_response = mean(response))
 ```

### 2021-03-26 analysis/xor-some-prolific-pilot1.Rmd
prettify, add timeSpent and by-item plots

--- 2021-03-25 analysis/xor-some-prolific-pilot1.Rmd
@@
 
-N=10 participants were recruited for this pilot and compensated 2 pounds/participant.
+N=10 participants were recruited for this pilot and compensated 2 pounds/participant. 8 items (4 some + 4 xor) were manually selected for this pilot to get more data points per item.
 
@@
 library(tidyverse)
+library(tidyboot)
 ```
@@
 
-```{r data}
+```{r data, results='hide'}
 d <- read_csv("../data/pilots/results_58_xor-some-Prolific-pilot1_N10.csv")
@@
 
+Check the time participants spent overall on the experiment:
+```{r responseTime}
+overall_timeSpent <- d_native %>% mutate(timeSpent = round(timeSpent, 2)) %>% distinct(timeSpent)
+
+#  summarize(timeCounts = count(timeSpent) / d_native %>% )
+ggplot(data = overall_timeSpent, aes(y=timeSpent, alpha = 0.7)) +
+  geom_boxplot() +
+  ggtitle("Overall time participants took")
+```
+
 Next, we would exclude participants based on their ratings in the main / example trials (and possibly the bot check) according to preregistered exclusion criteria. For now, all the data is considered.
 ```{r clean}
-d_native %>% group_by(submission_id) %>% count(response) %>% View()
-d_main <- d_native %>% filter(trial_name != "example")
-d_exmpl <- d_native %>% filter(trial_name == "example")
+d_main <- d_native %>% select(-age, -botresponse, -comments, -education, -endTime, 
+                              -gender, -languages, -optionLeft, -optionRight, -startDate,
+                              -startTime, -timeSpent) %>%
+  filter(trial_name != "example")
+d_exmpl <- d_native %>% select(-age, -botresponse, -comments, -education, -endTime, 
+                              -gender, -languages, -optionLeft, -optionRight, -startDate,
+                              -startTime, -timeSpent) %>%
+  filter(trial_name == "example")
 d_critical <- d_main %>% filter(condition == "critical")
@@
                                         test_condition)))
+d_test_ci <- d_test %>% group_by(test_condition) %>% 
+  tidyboot_mean(column = response)
 
@@
   ggplot(., aes(x = test_condition, y = response)) +
-  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1))
-# add mean / CI
+  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
+  geom_point(data = d_test_ci, aes(x = test_condition, y = mean), color = "red", 
+             size = 4) 
 ```
 
-Plot main rel / comp / pri questions by main condition
-
-```{r}
+```{r, include=FALSE}
+# Plot main rel / comp / pri questions by main condition. 
 d_critical <- d_critical %>% 
@@
   filter(block != "xor" & block != "some") %>%
-  ggplot(., aes(x = as.factor(prior_class), y = response)) +
+  ggplot(., aes(x = as.factor(prior_class), y = response, color = as.factor(prior_class) )) +
   geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
-    facet_wrap(main_type~class_condition)
+    facet_wrap(main_type~class_condition) +
+  xlab("Anticipated categorization (high / low)") +
+  theme(legend.position = 'none') 
 ```
-Plot main rel / comp questions by main condition, separated into with / without critical utterance
+Plot main rel / comp / prior questions by main condition and by prior classification of the item (x-axis), separated into with / without critical utterance (shape, color).
+Only high-competence items were used for the some trials (only low-competence in the xor trials, respectively).
+This will be balanced in the main study.
 ```{r}
@@
 d_critical <- d_critical %>% 
-  mutate(w_utterance = ifelse(is.na(critical_question), F, T))
+  mutate(w_utterance = ifelse(is.na(critical_question), F, T),
+         block = ifelse(block == "comp", "competence", 
+                        ifelse(block == "rel", "relevance", ifelse(block == "pri", "prior", block) )))
+d_critical %>% 
+  filter(block != "xor" & block != "some") %>%
+  filter(block == class_condition) %>%
+  group_by(main_type, class_condition, w_utterance, prior_class) %>% 
+  summarize(mean_response = mean(response)) -> d_critical_summary
 
@@
   filter(block != "xor" & block != "some") %>%
-  ggplot(., aes(x = as.factor(prior_class), y = response, color = w_utterance)) +
-  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
+  filter(block == class_condition) %>%
+  ggplot(., aes(x = as.factor(prior_class), y = response, shape = w_utterance, color = w_utterance)) +
+  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
+  geom_point(data = d_critical_summary, aes(x = as.factor(prior_class), y = mean_response, shape = w_utterance), 
+             color = "red", size = 3) +
+  xlab("Anticipated categorization of the items (low / high)") +
   facet_wrap(main_type~class_condition) # get ratings from the respective trials only
-# use block for getting the actual correct ratings, but class_condition for subsetting the prior classification fo the respective predictor 
+
 ```
 
-plot inference rating as a function of respective rating of the explanatory factor (think about a plot where one would see an interaction)
+Plot inference ratings as a function of *anticipated* rating of the explanatory factor, similar to the paper: 
+``` {r}
+d_critical %>% 
+  filter(block == "xor" | block == "some") %>%
+  #group_by(block, class_condition, prior_class) %>%
+  #summarize(mean_rating = mean(response)) %>%
+  ggplot(., aes(x = prior_class, y = response)) +
+  geom_point(size = 2, alpha = 0.5) +
+  geom_line() +
+  facet_wrap(block~class_condition) +
+  xlab("Anticipated categorization of the items") +
+  ggtitle("Mean inference strength ratings by-predictor")
+
+```
+
+Plot mean ratings (across all 10 participants) for each vignette (with its respective condition indicated in the facet title, in the order relevance/competence/prior), in each condition.
+For relevance and competence, the color indicates whether it was presented without the utterance or with.
+```{r, fig.width=10, fig.height=7}
+bar.width = 0.8
+d_critical %>% 
+  mutate(title = paste(title, exp_condition, sep = "_")) %>%
+  filter(block == class_condition | block == "xor" | block == "some") %>%
+  group_by(block, title, w_utterance, main_type) %>%
+  summarize(mean_rating = mean(response)) %>% 
+  ggplot(., aes(x = block, y = mean_rating, fill = w_utterance)) +
+  geom_col(alpha = 0.7, width = bar.width, position = position_dodge(width = bar.width)) +
+  #geom_point(size = 2, alpha = 0.5, position = position_jitter(width = 0.1)) +
+  facet_wrap(main_type~title, ncol = 4) +
+  theme(axis.text.x = element_text(angle = 45)) +
+  ggtitle("By-item by-question mean ratings")
+```
+
+In conclusion, the technical side of the experiment seems to work without glitches; participants perform as expected on the comprehension questions; the relevance / competence / prior ratings seem to be mostly in line with our anticipated categorizations.
+
+One point to note is that on average participants took slightly longer than the expected 15 minutes to complete the experiment.

### 2021-03-27 analysis/xor-some-prolific-pilot1.Rmd
added correlation plots

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, message = FALSE)
 ```
@@
 
+Create more extensive condition labels, including information about whether a rating was produced with or without the utterance given.
+
+``` {r}
+# extending 'conditions' labels to include wheter the utterance was present or not
+d_critical <- d_critical %>% 
+  mutate(block_extended = ifelse(
+    !w_utterance, 
+    block, 
+    ifelse(block %in% c("some", "xor"), "target", str_c(block, "_wUtt", ""))
+  ))
+```
+
 Plot mean ratings (across all 10 participants) for each vignette (with its respective condition indicated in the facet title, in the order relevance/competence/prior), in each condition.
@@
   filter(block == class_condition | block == "xor" | block == "some") %>%
-  group_by(block, title, w_utterance, main_type) %>%
+  group_by(block_extended, title, w_utterance, main_type) %>%
   summarize(mean_rating = mean(response)) %>% 
-  ggplot(., aes(x = block, y = mean_rating, fill = w_utterance)) +
+  ggplot(., aes(x = block_extended, y = mean_rating, fill = w_utterance)) +
   geom_col(alpha = 0.7, width = bar.width, position = position_dodge(width = bar.width)) +
@@
   facet_wrap(main_type~title, ncol = 4) +
-  theme(axis.text.x = element_text(angle = 45)) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
   ggtitle("By-item by-question mean ratings")
+```
+
+``` {r}
+# make a versatile wide representation of the critical data
+d_critical_wide <- d_critical %>% 
+  select(submission_id, title, main_type, block_extended, response) %>% 
+  unique() %>% 
+  pivot_wider(
+    names_from = block_extended, 
+    values_from = response, 
+    values_fn = mean # getting means for double prior measurement in "xor"
+  )
+
+# correlation plot for "some"
+GGally::ggpairs(
+  filter(d_critical_wide, main_type == "some") %>%  
+    select(prior:ncol(d_critical_wide))
+)
+
+# correlation plot for "or"
+GGally::ggpairs(
+  filter(d_critical_wide, main_type == "xor") %>%  
+    select(prior:ncol(d_critical_wide))
+  )
 ```

### 2021-03-28 analysis/xor-some-prolific-pilot1.Rmd
check by-participant performance

@@
 
+Check by-participant behavior to see if they show versatile response strategies for the different questions for xor vs. some. Participants show a sufficient range of responses across conditions.  
+```{r, fig.width=15, fig.height=7}
+d_critical %>%
+  filter(block == class_condition | block == "xor" | block == "some") %>%
+  ggplot(., aes(x = block, y = response, color = block)) +
+  geom_point(alpha = 0.7, size = 2) + 
+  facet_wrap(main_type~submission_id, ncol = 10) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  ggtitle("By-participant ratings all xor vs. some trials")
+```
+
 In conclusion, the technical side of the experiment seems to work without glitches; participants perform as expected on the comprehension questions; the relevance / competence / prior ratings seem to be mostly in line with our anticipated categorizations.

### 2021-04-01 04_trials.js
add all vignettes for second pilot

@@ -91,8 +91,8 @@ let vignettes_or = {
    "utterance_disj2": "Jason says to Kim: 'He has a daughter.'",
    "question_rel": "It is important for Kim to know whether Dave has both a son and a daughter.",
    "question_comp": "Jason knows whether Dave has both a son and a daugther.",
-   "question_pri1": "If Jason has a son, it is likely that he has a daugther as well.",
-   "question_pri2": "If Jason has a daugther, it is likely that he has a son as well.",
+   "question_pri1": "If Jason has a son, it is likely that he has a daughter as well.",
+   "question_pri2": "If Jason has a daughter, it is likely that he has a son as well.",
    "question_pri_and": "It is likely that Dave has both a son and a daughter. ",
    "question_xor": "From what Jason said we may conclude that Dave does not have both a son and a daughter.",
    "question_xorI": "Jason wants to convey that Dave does not have both a son and a daughter.",
@@ -121,9 +121,9 @@ let vignettes_or = {
    "competence": 0,
    "prior": 0,
    "background": "Sandra and Susie are watching a movie with Brad Pitt. Sandra vaguely remembers another movie with Brad Pitt and tells Susie about it.",
-   "utterance_or": "Sandra says to Susie: ‘Brad Pitt played a fireman or a special agent in that other movie.'",
-   "utterance_disj1": "Sandra says to Susie: ‘Brad Pitt played a fireman in that other movie.'",
-   "utterance_disj2": "Sandra says to Susie: ‘Brad Pitt played special agent in that other movie.'",
+   "utterance_or": "Sandra says to Susie: 'Brad Pitt played a fireman or a special agent in that other movie.'",
+   "utterance_disj1": "Sandra says to Susie: 'Brad Pitt played a fireman in that other movie.'",
+   "utterance_disj2": "Sandra says to Susie: 'Brad Pitt played special agent in that other movie.'",
    "question_rel": "It is important for Susie to know whether Brad Pitt played both a fireman and a special agent in that other movie.",
    "question_comp": "Sandra knows whether Brad Pitt played both a fireman and a special agent in that other movie.",
    "question_pri1": "If Brad Pitt played a fireman in that other movie, it is likely that he also played a special agent in that other movie.",
@@ -131,8 +131,8 @@ let vignettes_or = {
    "question_pri_and": "It is likely that Brad Pitt played both a fireman and a special agent in that other movie.",
[…]

### 2021-04-03 analysis/xor-some-prolific-pilot2.Rmd
start analysis

---
title: "XOR-Some Prolific Pilot 2"
author: "Polina Tsvilodub"
date: "4/2/2021"
output: github_document
---

In this second pilot on Prolific we gather a more substantial amount of pilot data for the xor-some study on Prolific. The structure of the experiment was as follows:
Participants read instructions, completed three example trials, and then completed 8 main blocks consisting of 4 xor and 4 some items. Each main block had the following structure: Participants read the background story, answered one comprehension question, then answered competence / relevance / prior questions in randomized order; then they read another 3 comprehension questions, after which the critical utterance was added below the background story. They answered the inference strength question, and then competence / relevance questions in randomized order again.

N=118 participants were recruited for this pilot and compensated 2 pounds/participant. 8 items (4 some + 4 xor) were sampled at random for each participant such that each participant saw one item in each condition (relevance X competence X prior = 8 unique conditions).

```{r libraries, include=FALSE}
library(tidyverse)
library(tidyboot)
library(brms)
library(tidybayes)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, message = FALSE)
```

```{r anonymization, include=FALSE}
#d_raw <- read_csv("~/Documents/Research/XOR/raw_data/results_58_xor-some-Prolific-pilot2_N120.csv")
# remove IDs
#d_raw %>% select(-prolific_id) %>% write_csv("../data/pilots/results_58_xor-some-Prolific-pilot2_N120.csv") 
```

## Checks & Exclusions
```{r data, results='hide', include=FALSE}
d <- read_csv("../data/pilots/results_58_xor-some-Prolific-pilot2_N120.csv")
```

Checking if there are any comments indicating technical issues which is not the case:
```{r comments}
d %>% distinct(comments) %>% View()
```

Check native languages. Participants not indicating English as (one of) their native language(s) are be excluded. We also compute some participant demographics.
```{r languages}
d %>% distinct(languages)
cat("Number of partiipants before excluding non-natives: ", d %>% distinct(submission_id) %>% count() %>% pull() ) 
# exclude non-natives if necessary
d_native <- d %>% 
 # filter(("en" | "En" | "bri") %in% languages)
  filter(grepl("[(en)(br)]", languages, ignore.case = T))

cat("Number of partiipants after excluding non-natives: ", d_native %>% distinct(submission_id) %>% count() %>% pull() )

cat("Mean age: ", d_native %>% pull(age) %>% mean(., na.rm = T) )
cat("Gender counts: ") 
d_native %>% count(gender) %>% mutate(n = n/88)
```

Next, we check whether all the conditions were used correctly.

```{r counts}
# check xor/some vs. trial type
d_native %>% count(main_type, condition) 
# check xor/some vs. experimental condition
d_native %>% count(main_type, exp_condition)
# count items used
d_native %>% count(title)
```

Check the time participants spent overall on the experiment before cleaning the data:
```{r responseTime}
overall_timeSpent <- d_native %>% mutate(timeSpent = round(timeSpent, 2)) %>% distinct(timeSpent)

#  summarize(timeCounts = count(timeSpent) / d_native %>% )
ggplot(data = overall_timeSpent, aes(y=timeSpent, alpha = 0.7)) +
  geom_boxplot() +
  ggtitle("Overall time participants took in mins")
```
```{r}
d_main <- d_native %>% select(-age, -botresponse, -comments, -education, -endTime, 
                              -gender, -languages, -optionLeft, -optionRight, -startDate,
                              -startTime, -timeSpent) %>%
  filter(trial_name != "example")
d_exmpl <- d_native %>% select(-age, -botresponse, -comments, -education, -endTime, 
                              -gender, -languages, -optionLeft, -optionRight, -startDate,
                              -startTime, -timeSpent) %>%
  filter(trial_name == "example")
d_critical <- d_main %>% filter(condition == "critical")
```

Plot responses on comprehension questions by type before applying exclusion criteria:
```{r, echo=FALSE}
d_test <- d_main %>% rowwise() %>% filter(condition == "test") %>% 
  mutate(test_condition = substr(test_question, 6, 9),
         test_condition = ifelse(test_condition == "fals", "false", 
                                 ifelse(test_condition == "unce", "uncertain",
                                        test_condition)))
d_test_ci <- d_test %>% group_by(test_condition) %>% 
  tidyboot_mean(column = response)

d_test %>% 
  ggplot(., aes(x = test_condition, y = response)) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
  geom_point(data = d_test_ci, aes(x = test_condition, y = mean), color = "red", 
             size = 4) +
  facet_wrap(~main_type)
```

Next, we exclude participants based on their ratings in the main trials: Participants who gave all responses within the range of 10 and participants who failed more than 0.2 of the comprehension questions are excluded from analysis. 
Participants who failed all example trials are excluded, as well. 
The bot check trial is not considered for exclusions. 
```{r exclusions}
# get participants failing example trials
d_exmpl <- d_exmpl %>% group_by(submission_id) %>% 
  mutate(example_condition = ifelse(grepl("as certainly true", question), "true", 
                                    ifelse(grepl("as certainly false", question), "false",
                                           "uncertain")),
         passed_example_trial = case_when(example_condition == "true" ~ response >= 80,
                                    example_condition == "false" ~ response <= 20,
                                    example_condition == "uncertain" ~ response >= 30),
       #  check if all trials passed
         passed_example = case_when(sum(passed_example_trial) == 0 ~ FALSE,
                                    TRUE ~ TRUE)
         ) %>% filter(passed_example == F)

cat("Subjects failing the example trials: ", d_exmpl %>% distinct(submission_id) %>% pull() %>% length())

# apply exclusion criteria to main trials
# check range of responses per participant
d_main_fail <- d_main %>% group_by(submission_id) %>%
  mutate(passed_main = case_when(max(response) - min(response) <= 10 ~ FALSE,
                                 TRUE ~ TRUE)
         ) %>% filter(passed_main == F)
cat("Subjects providing the same ratings throughout the trials: ", d_main_fail %>% distinct(submission_id) %>% pull() %>% length())

# get participants failing comprehension questions
d_test <- d_test %>%
  group_by(submission_id) %>%
  mutate(passed_filler_trial = case_when(test_condition == "true" ~ response >= 70,
                                   test_condition == "false" ~ response <= 40,
                                   test_condition == "uncertain" ~ response %in% (0:90)),
         mean_comprehension = mean(passed_filler_trial),
         passed_filler = mean_comprehension >= 0.8
         ) %>%
  filter(passed_filler == F)

cat("Subjects failing the comprehension trials: ", d_test %>% distinct(submission_id) %>% pull() %>% length())

# put it all together
d_full_clean <- anti_join(d_main, d_main_fail, by = "submission_id")
d_full_clean <- anti_join(d_full_clean, d_exmpl, d_test, by = "submission_id")
d_full_clean <- anti_join(d_full_clean, d_test, by = "submission_id")

cat("Nr. of participants left after cleaning: ", d_full_clean %>% distinct(submission_id) %>% pull() %>% length())
```

```{r clean}
# get overall mean ratings / subject
d_full_clean %>% group_by(submission_id) %>% summarise(mean_rating = mean(response)) %>% arrange(mean_rating)
```

```{r, echo=FALSE}
d_critical_clean <- d_full_clean %>% filter(condition == "critical")
d_critical_long <- d_critical_clean %>% 
  pivot_longer(c(competence, relevance, prior), 
               names_to = "class_condition", 
               values_to = "prior_class")
```

## Plots
Plot main rel / comp / prior questions by main condition and by prior classification of the item (x-axis), separated into with / without critical utterance (shape, color). The prior questions were only used without the utterance. 
```{r}

d_critical_long <- d_critical_long %>% 
  mutate(w_utterance = ifelse(is.na(critical_question), F, T),
         block = ifelse(block == "comp", "competence", 
                        ifelse(block == "rel", "relevance", ifelse(block == "pri", "prior", block) )))
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  group_by(main_type, class_condition, w_utterance, prior_class) %>% 
  summarize(mean_response = mean(response)) -> d_critical_summary

d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  ggplot(., aes(x = as.factor(prior_class), y = response, shape = w_utterance, color = w_utterance)) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_critical_summary, aes(x = as.factor(prior_class), y = mean_response, shape = w_utterance), 
             color = "red", size = 3) +
  xlab("Anticipated categorization of the items (low / high)") +
  facet_wrap(main_type~class_condition) # get ratings from the respective trials only

```
```{r}
# check whether the ratings in the two prior questions are the same in xor 
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  filter(block == "prior", main_type == "xor") %>%
  mutate(priorQ_nr = rep(c(1,2), 396)) -> d_xor_priors
d_xor_priors %>% group_by(priorQ_nr, prior_class) %>%
  summarise(mean = mean(response)) -> d_xor_priors_summary
d_xor_priors %>%
  ggplot(., aes(x = as.factor(priorQ_nr), y = response )) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_xor_priors_summary, aes(x = as.factor(priorQ_nr), y = mean), color = "red", size = 3) +
  xlab("First vs Second prior question for high vs low prior conditions") +
  facet_wrap(~as.factor(prior_class)) # get ratings from the respective trials only

```

Plot inference ratings as a function of *anticipated* rating of the explanatory factor, similar to the paper: 
``` {r}
d_critical_long %>% 
  filter(block == "xor" | block == "some") %>%
  #group_by(block, class_condition, prior_class) %>%
  #summarize(mean_rating = mean(response)) %>%
  ggplot(., aes(x = prior_class, y = response)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_line() +
  facet_wrap(block~class_condition) +
  xlab("Anticipated categorization of the items") +
  ggtitle("Mean inference strength ratings by-predictor")

```

Create more extensive condition labels, including information about whether a rating was produced with or without the utterance given.

``` {r}
# extending 'conditions' labels to include whether the utterance was present or not
d_critical_long <- d_critical_long %>% 
  mutate(block_extended = ifelse(
    !w_utterance, 
    block, 
    ifelse(block %in% c("some", "xor"), "target", str_c(block, "_wUtt", ""))
  ))
```

Plot mean ratings (across all participants) for each vignette (with its respective condition indicated in the facet title, in the order relevance/competence/prior), in each condition. For relevance and competence, the color indicates whether it was presented without the utterance or with. 
```{r, fig.width=10, fig.height=30}
bar.width = 0.8
d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_")) %>%
  filter(block == class_condition | block == "xor" | block == "some") %>%
  group_by(block_extended, title, w_utterance, main_type) %>%
  summarize(mean_rating = mean(response)) %>% 
  ggplot(., aes(x = block_extended, y = mean_rating, fill = w_utterance)) +
  geom_col(alpha = 0.7, width = bar.width, position = position_dodge(width = bar.width)) +
  #geom_point(size = 2, alpha = 0.5, position = position_jitter(width = 0.1)) +
  facet_wrap(main_type~title, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-item by-question mean ratings")
```

``` {r}
# make a versatile wide representation of the critical data
d_critical_wide <- d_critical_long %>% 
  select(submission_id, title, main_type, block_extended, response) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response, 
    values_fn = mean # getting means for double prior measurement in "xor"
  )

# correlation plot for "some"
GGally::ggpairs(
  filter(d_critical_wide, main_type == "some") %>%  
    select(prior:ncol(d_critical_wide))
)

# correlation plot for "or"
GGally::ggpairs(
  filter(d_critical_wide, main_type == "xor") %>%  
    select(prior:ncol(d_critical_wide))
  )
```

Check by-participant behavior to see if they show versatile response strategies for the different questions for xor vs. some. Participants show a sufficient range of responses across conditions.  
```{r, fig.width=15, fig.height=7}
d_critical_long %>%
  filter(block == class_condition | block == "xor" | block == "some") %>%
  ggplot(., aes(x = block, y = response, color = block)) +
  geom_point(alpha = 0.7, size = 2) + 
  facet_wrap(main_type~submission_id, ncol = 10) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-participant ratings all xor vs. some trials")
```

## Stats

```{r, results='hide'}
# xor, maximal model with interactions and maximal REs
model_xor <- brm(
  target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "xor"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```

```{r}
summary(model_xor)
```

```{r, results='hide'}
# some, maximal model with interactions and maximal REs
model_some <- brm(
  target ~ prior*competence*relevance + 
    (1 + prior + competence + relevance || submission_id) +
    (1 | title),
  data = d_critical_wide %>% filter(main_type == "some"),
  family = "gaussian",
  cores = 4,
  iter = 3000
)
```
```{r}
summary(model_some)
```

## Explore z-scoring

## Exploreusing binary predictors

### 2021-04-04 analysis/xor-some-prolific-pilot2.Rmd
add stats

@@
 cat("Mean age: ", d_native %>% pull(age) %>% mean(., na.rm = T) )
-cat("Gender counts: ") 
 d_native %>% count(gender) %>% mutate(n = n/88)
@@
   geom_point(size = 2, alpha = 0.5) +
-  geom_line() +
+  geom_smooth(method="lm") +
   facet_wrap(block~class_condition) +
   xlab("Anticipated categorization of the items") +
-  ggtitle("Mean inference strength ratings by-predictor")
+  ggtitle("Inference strength ratings by-predictor")
 
 ```
-
 Create more extensive condition labels, including information about whether a rating was produced with or without the utterance given.
@@
   ))
+```
+
+```{r}
+# make a versatile wide representation of the critical data
+d_critical_wide <- d_critical_long %>% 
+  select(submission_id, title, main_type, block_extended, response) %>% 
+  unique() %>% 
+  pivot_wider(
+    names_from = block_extended, 
+    values_from = response, 
+    values_fn = mean # getting means for double prior measurement in "xor"
+  ) 
+```
+
+Plot inference strength ratings against raw predictor ratings for each participant / each item:
+```{r}
+d_critical_wide %>%
+  ggplot(., aes(x = relevance, y = target)) +
+  geom_point(size = 2, alpha = 0.7) +
+  geom_smooth(method = "lm") +
+  facet_wrap(~main_type, ncol = 1) -> p.rel
+
+d_critical_wide %>%
+  ggplot(., aes(x = competence, y = target)) +
+  geom_point(size = 2, alpha = 0.7) +
+  geom_smooth(method = "lm") +
+  ylab("") +
+  facet_wrap(~main_type, ncol = 1) -> p.comp
+
+d_critical_wide %>%
+  ggplot(., aes(x = prior, y = target)) +
+  geom_point(size = 2, alpha = 0.7) +
+  geom_smooth(method = "lm") +
+  ylab("") +
+  facet_wrap(~main_type, ncol = 1) -> p.pri
+
+gridExtra::grid.arrange(p.rel, p.comp, p.pri, ncol = 3) 
 ```
@@
 ``` {r}
-# make a versatile wide representation of the critical data
-d_critical_wide <- d_critical_long %>% 
-  select(submission_id, title, main_type, block_extended, response) %>% 
-  unique() %>% 
-  pivot_wider(
-    names_from = block_extended, 
-    values_from = response, 
-    values_fn = mean # getting means for double prior measurement in "xor"
-  )
 
@@
 
+## Explore z-scoring
+Visual exploration of z-scoring each measure (rel / comp / pri / target / comp-wUtterance / rel-wUtterance) per participant, collapsing some and or together. 
+```{r}
+d_critical_zScored <- d_critical_long %>% group_by(submission_id, block_extended) %>%
+  mutate(block_mean = mean(response),
+         block_sd = sd(response),
+         response_centered = (response - block_mean)/block_sd)
+
+d_critical_zScored %>% 
+  filter(block != "xor" & block != "some") %>%
+  filter(block == class_condition) %>%
+  group_by(main_type, class_condition, w_utterance, prior_class) %>% 
+  summarize(mean_response = mean(response_centered)) -> d_critical_zScore_summary
+
+d_critical_zScored %>% 
+  filter(block != "xor" & block != "some") %>%
+  filter(block == class_condition) %>%
+  ggplot(., aes(x = as.factor(prior_class), y = response_centered, shape = w_utterance, color = w_utterance)) +
+  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
+  geom_point(data = d_critical_zScore_summary, aes(x = as.factor(prior_class), y = mean_response, shape = w_utterance), 
+             color = "red", size = 3) +
+  xlab("Anticipated categorization of the items (low / high)") +
+  facet_wrap(main_type~class_condition) + # get ratings from the respective trials only 
+  ggtitle("Centered responses to by-predictor questions")
+```
+
+Plot z-scored predictor ratings against z-scored target ratings
+```{r}
+d_critical_zScored_wide <- d_critical_zScored %>% 
+  select(submission_id, title, main_type, block_extended, response_centered) %>% 
+  unique() %>% 
+  pivot_wider(
+    names_from = block_extended, 
+    values_from = response_centered, 
+    values_fn = mean # getting means for double prior measurement in "xor"
+  )
+
+d_critical_zScored_wide %>%
+  ggplot(., aes(x = relevance, y = target)) +
+  geom_point(size = 2, alpha = 0.7) +
+  geom_smooth(method = "lm") +
+  facet_wrap(~main_type, ncol = 1) -> p.rel.z
+
+d_critical_zScored_wide %>%
+  ggplot(., aes(x = competence, y = target)) +
+  geom_point(size = 2, alpha = 0.7) +
+  geom_smooth(method = "lm") +
+  ylab("") +
+  facet_wrap(~main_type, ncol = 1) -> p.comp.z
+
+d_critical_zScored_wide %>%
+  ggplot(., aes(x = prior, y = target)) +
+  geom_point(size = 2, alpha = 0.7) +
+  geom_smooth(method = "lm") +
+  ylab("") +
+  facet_wrap(~main_type, ncol = 1) -> p.pri.z
+
+gridExtra::grid.arrange(p.rel.z, p.comp.z, p.pri.z, ncol = 3) 
+```
+
 ## Stats
-
+Maximal models on raw ratings:
 ```{r, results='hide'}
@@
 ```
+
 ```{r}
@@
 
-## Explore z-scoring
+#### Stats on z-scored data
+```{r, results='hide'}
+# xor, maximal model with interactions and maximal REs on z-scored data
+model_xor_zScored <- brm(
+  target ~ prior*competence*relevance + 
+    (1 + prior + competence + relevance || submission_id) +
+    (1 | title),
+  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
 
-## Exploreusing binary predictors
+```{r}
+summary(model_xor_zScored)
+```
+
+```{r, results='hide'}
+# some, maximal model with interactions and maximal REs
+model_some_zScored <- brm(
+  target ~ prior*competence*relevance + 
+    (1 + prior + competence + relevance || submission_id) +
+    (1 | title),
+  data = d_critical_zScored_wide %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_some_zScored)
+```
+
+#### Stats with dichotomized predictors 
+Maximal model on raw inference ratings, using anticipated binary categorization of items as predictors 
+```{r, results='hide'}
+d_critical_long %>% 
+  select(submission_id, title, main_type, block_extended, response, prior_class, class_condition) %>% 
+  unique() %>% 
+  pivot_wider(
+    names_from = class_condition, 
+    values_from = prior_class
+  )  %>%
+  mutate(relevance = as.factor(relevance),
+         competence = as.factor(competence),
+         prior = as.factor(prior)) %>% 
+  filter(block_extended == "target") -> d_critical_wide_cat
+# sum code predictors
+# low prior: 1, high prior:: -1 
+contrasts(d_critical_wide_cat$prior) <- contr.sum(2)
+# low comp : 1, high comp : -1
+contrasts(d_critical_wide_cat$competence) <- contr.sum(2)
+# low rel: 1, high rel: -1
+contrasts(d_critical_wide_cat$relevance) <- contr.sum(2)
+
+# xor, maximal model with interactions and maximal REs
+model_xor_cat <- brm(
+  response ~ prior*competence*relevance + 
+    (1 + prior + competence + relevance || submission_id) +
+    (1 | title),
+  data = d_critical_wide_cat %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_xor_cat)
+```
+
+```{r, results='hide'}
+model_some_cat <- brm(
+  response ~ prior*competence*relevance + 
+    (1 + prior + competence + relevance || submission_id) +
+    (1 | title),
+  data = d_critical_wide_cat %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+```{r}
+summary(model_some_cat)
+```
+
+Maximal model on z-scored inference ratings, using anticipated binary categorization of items as predictors 
+```{r, results='hide'}
+d_critical_zScored %>% 
+  select(submission_id, title, main_type, block_extended, response_centered, prior_class, class_condition) %>% 
+  unique() %>% 
+  pivot_wider(
+    names_from = class_condition, 
+    values_from = prior_class
+  )  %>%
+  mutate(relevance = as.factor(relevance),
+         competence = as.factor(competence),
+         prior = as.factor(prior)) %>% 
+  filter(block_extended == "target") -> d_critical_wide_cat_zScored
+# sum code predictors
+# low prior: 1, high prior:: -1 
+contrasts(d_critical_wide_cat_zScored$prior) <- contr.sum(2)
+# low comp : 1, high comp : -1
+contrasts(d_critical_wide_cat_zScored$competence) <- contr.sum(2)
+# low rel: 1, high rel: -1
+contrasts(d_critical_wide_cat_zScored$relevance) <- contr.sum(2)
+
+# xor, maximal model with interactions and maximal REs
+model_xor_cat_zScored <- brm(
+  response_centered ~ prior*competence*relevance + 
+    (1 + prior + competence + relevance || submission_id) +
+    (1 | title),
+  data = d_critical_wide_cat_zScored %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_xor_cat_zScored)
+```
+
+```{r, results='hide'}
+model_some_cat_zScored <- brm(
+  response_centered ~ prior*competence*relevance + 
+    (1 + prior + competence + relevance || submission_id) +
+    (1 | title),
+  data = d_critical_wide_cat_zScored %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+```{r}
+summary(model_some_cat_zScored)
+```
+
+In summary, the different models provide somewhat different results, specially when using raw ratings vs.
+binary predictors.
+Since interpreting the categorical models in relation to the plots might be not as straight-forward, I am not entirely sure whether the qualitative results match in all models.
+Using centered data seems to provide a little tighter estimates in the models; not sure where the slight qualitative differences come from.

### 2021-04-05 analysis/xor-some-prolific-pilot2.Rmd
add rating deviations

@@
                values_to = "prior_class")
+
+d_critical_long <- d_critical_long %>% 
+  mutate(w_utterance = ifelse(is.na(critical_question), F, T),
+         block = ifelse(block == "comp", "competence", 
+                        ifelse(block == "rel", "relevance", ifelse(block == "pri", "prior", block) )))
+```
+
+Create more extensive condition labels, including information about whether a rating was produced with or without the utterance given.
+
+``` {r}
+# extending 'conditions' labels to include whether the utterance was present or not
+d_critical_long <- d_critical_long %>% 
+  mutate(block_extended = ifelse(
+    !w_utterance, 
+    block, 
+    ifelse(block %in% c("some", "xor"), "target", str_c(block, "_wUtt", ""))
+  ))
+```
+
+Average nr of responses per question per vignette:
+```{r}
+d_critical_long %>% 
+  filter(class_condition == block | block == "xor" | block == "some") %>%
+  select(-class_condition, -prior_class) %>%
+  unique() %>% count(title, block_extended, main_type) %>%
+  mutate(n = ifelse(main_type == "xor" & block_extended == "prior", n/2, n)) %>%
+  summarize(mean_observations = mean(n))
+```
+
+Compute the average deviation from expected responses for each vignette, individually for each dimension (rel / comp / pri):
+```{r}
+d_critical_long %>% 
+  filter(block != "xor" & block != "some") %>%
+  filter(block == class_condition) %>%
+  mutate(expected_rating = prior_class * 100,
+         response_deviation = abs(response - expected_rating)) %>%
+  group_by(title, block_extended) %>%
+  summarize(mean_deviation = mean(response_deviation)) %>%
+  arrange(desc(mean_deviation), .by_group = T) %>%
+  # get the largest deviations - over 50 
+  filter(mean_deviation >= 50) -> d_critical_deviations
+
+#d_critical_deviations %>% write_csv("../data/pilots/pilot2_byItem_rating_deviations.csv")
+
+print(d_critical_deviations)
+
+# count which dimension is the most problematic - the prior
+d_critical_deviations %>% group_by(block_extended) %>% count()
+
+# check which items have the most deviations over 50 
+d_critical_deviations %>% group_by(title) %>% count() %>% arrange(desc(.$n))
 ```
@@
 
-d_critical_long <- d_critical_long %>% 
-  mutate(w_utterance = ifelse(is.na(critical_question), F, T),
-         block = ifelse(block == "comp", "competence", 
-                        ifelse(block == "rel", "relevance", ifelse(block == "pri", "prior", block) )))
 d_critical_long %>% 
@@
 
-```
-Create more extensive condition labels, including information about whether a rating was produced with or without the utterance given.
-
-``` {r}
-# extending 'conditions' labels to include whether the utterance was present or not
-d_critical_long <- d_critical_long %>% 
-  mutate(block_extended = ifelse(
-    !w_utterance, 
-    block, 
-    ifelse(block %in% c("some", "xor"), "target", str_c(block, "_wUtt", ""))
-  ))
 ```

### 2021-04-09 analysis/xor-some-prolific-pilot2.Rmd
finish code

@@
   mutate(expected_rating = prior_class * 100,
-         response_deviation = abs(response - expected_rating)) %>%
-  group_by(title, block_extended) %>%
+         response_deviation = abs(response - expected_rating),
+         block = ifelse(block == "xor" | block == "some", "target", block)
+         ) %>%
+  group_by(title, block) %>%
   summarize(mean_deviation = mean(response_deviation)) %>%
@@
 # count which dimension is the most problematic - the prior
-d_critical_deviations %>% group_by(block_extended) %>% count()
+d_critical_deviations %>% group_by(block) %>% count()
 
@@
 d_critical_deviations %>% group_by(title) %>% count() %>% arrange(desc(.$n))
+```
+
+Compute the standard deviation for each question for each item as an additional measure of spread in the participants' responses within-item:
+```{r}
+d_critical_long %>% 
+  filter(block != "xor" & block != "some") %>%
+  filter(block == class_condition) %>%
+  mutate(
+         block = ifelse(block == "xor" | block == "some", "target", block)
+         ) %>%
+  group_by(title, block) %>%
+  summarize(item_sd = sd(response)) %>%
+  arrange(desc(item_sd), .by_group = T) %>%
+  # arbitrary set to 25
+  filter(item_sd >= 25) -> d_critical_sd
+d_critical_sd
+
+#d_critical_sd %>% write_csv("../data/pilots/pilot2_byItem_rating_SDs.csv")
+```
+
+Visually check the spread of by-item ratings among subjects for each question:
+```{r, fig.height=20, fig.width=7}
+
+d_critical_long %>% 
+  mutate(title = paste(title, exp_condition, sep = "_"),
+         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
+  filter(block == class_condition | block == "target") %>%
+  ggplot(., aes(x = block, y = response, color = block)) +
+  geom_point(alpha = 0.7, position = position_jitter(width = 0.1)) +
+  ylab("Raw responses to respective questions") +
+  facet_wrap(main_type~title, ncol = 4) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  ggtitle("By-item by-question raw ratings")
+```
+For the following items, participants seem  to *disagree* in their judgments for all of the predictors: Emails from a broken laptop, Gift unwrapping, Martha's cookies, Suzie's fruits, Brad's clothes, Crime, maybe more..
+
+Check by-vignette by-predictor empirical means, to see if the range is somewhat covered:
+```{r}
+d_critical_long %>% 
+  mutate(title = paste(title, exp_condition, sep = "_"),
+         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
+  filter(block == class_condition | block == "target") %>%
+  group_by(title, block) %>%
+  summarise(mean_response = mean(response)) %>%
+  arrange(mean_response)
 ```
@@
              color = "red", size = 3) +
+  ylab("Responses to respective predictor question") +
   xlab("Anticipated categorization of the items (low / high)") +
@@
   geom_point(data = d_xor_priors_summary, aes(x = as.factor(priorQ_nr), y = mean), color = "red", size = 3) +
+  ylab("Responses to prior questions") +
   xlab("First vs Second prior question for high vs low prior conditions") +
@@
   filter(block == "xor" | block == "some") %>%
-  #group_by(block, class_condition, prior_class) %>%
-  #summarize(mean_rating = mean(response)) %>%
   ggplot(., aes(x = prior_class, y = response)) +
@@
   geom_smooth(method="lm") +
+  ylab("Inference strength ratings") +
   facet_wrap(block~class_condition) +
@@
   geom_smooth(method = "lm") +
+  ylab("Inference strength ratings") +
   facet_wrap(~main_type, ncol = 1) -> p.rel
@@
   #geom_point(size = 2, alpha = 0.5, position = position_jitter(width = 0.1)) +
+  ylab("Mean responses to respective questions") +
+  xlab("Question type") +
   facet_wrap(main_type~title, ncol = 4) +
@@
   ggtitle("By-item by-question mean ratings")
+```
+
+Plot by-item trends between predictor ratings and target inference strength ratings (relevance / competence ratings with utterance are omitted):
+```{r, fig.height=20, fig.width=8}
+# make a versatile wide representation of the critical data
+d_critical_predictorsXtarget <- d_critical_wide %>% 
+  pivot_longer(
+    cols = c(relevance, prior, competence, relevance_wUtt, competence_wUtt),
+    names_to = "predictors",
+    values_to = "predictor_response"
+  ) %>%
+  filter(predictors != "relevance_wUtt", predictors != "competence_wUtt")
+
+d_critical_predictorsXtarget %>%
+  ggplot(., aes(x = predictor_response, y = target)) +
+  geom_point() +
+  facet_wrap(title~predictors, ncol= 6) +
+  xlab("Response to predictor questions") +
+  ylab("Inference strength rating") +
+  ggtitle("By-item Inference Strength Ratings vs. Predictor Ratings (left to right)")
+
+```
+
+Plot inference strength ratings against predictor ratings, by-item, using by-item mean ratings across participants. Error bard indicate the standard deviation:
+```{r, fig.height=4, fig.width=8}
+d_critical_predictorsXtarget_summary <- d_critical_predictorsXtarget %>% 
+  group_by(title, predictors, main_type) %>%
+  summarise(mean_target = mean(target),
+            mean_predictor = mean(predictor_response), 
+            sd_target = sd(target),
+            sd_predictor = sd(predictor_response))
+
+d_critical_predictorsXtarget_summary %>%
+  ggplot(., aes(x = mean_predictor, y = mean_target)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = mean_target - sd_target, ymax = mean_target + sd_target), alpha = 0.7) +
+  geom_errorbarh(aes(xmin = mean_predictor - sd_predictor, xmax = mean_predictor + sd_predictor), alpha = 0.7) +
+  geom_text(aes(label=title), size = 1.5, nudge_y = 4) +
+  geom_smooth(method = "lm") +
+  ylab("Mean inference strength rating") +
+  xlab("Mean predictor ratings") + 
+  facet_wrap(main_type~predictors)
 ```
@@
   filter(d_critical_wide, main_type == "some") %>%  
-    select(prior:ncol(d_critical_wide))
+    select(prior, competence, relevance, target, competence_wUtt, relevance_wUtt) 
 )
@@
   filter(d_critical_wide, main_type == "xor") %>%  
-    select(prior:ncol(d_critical_wide))
+    select(prior, competence, relevance, target, competence_wUtt, relevance_wUtt)
   )
-```
-
-Check by-participant behavior to see if they show versatile response strategies for the different questions for xor vs. some. Participants show a sufficient range of responses across conditions.  
-```{r, fig.width=15, fig.height=7}
-d_critical_long %>%
-  filter(block == class_condition | block == "xor" | block == "some") %>%
-  ggplot(., aes(x = block, y = response, color = block)) +
-  geom_point(alpha = 0.7, size = 2) + 
-  facet_wrap(main_type~submission_id, ncol = 10) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  ggtitle("By-participant ratings all xor vs. some trials")
 ```
@@
          block_sd = sd(response),
-         response_centered = (response - block_mean)/block_sd)
+         response_centered = (response - block_mean)/block_sd,
+         # catch the cases where sd is 0 
+         response_centered = ifelse(is.na(response_centered), 0, response_centered))
 
@@
              color = "red", size = 3) +
+  ylab("Responses to respective predictor questions") +
   xlab("Anticipated categorization of the items (low / high)") +
@@
   geom_point(size = 2, alpha = 0.7) +
+  ylab("Inference strength ratings") +
   geom_smooth(method = "lm") +
@@
 gridExtra::grid.arrange(p.rel.z, p.comp.z, p.pri.z, ncol = 3) 
+```
+
+Correlation plots of z-scored data:
+``` {r}
+
+# correlation plot for "some"
+GGally::ggpairs(
+  filter(d_critical_zScored_wide, main_type == "some") %>% ungroup() %>%  
+    select(prior, competence, relevance, target, competence_wUtt, relevance_wUtt)
+)
+
+# correlation plot for "or"
+GGally::ggpairs(
+  filter(d_critical_zScored_wide, main_type == "xor") %>%  ungroup() %>%
+    select(prior, competence, relevance, target, competence_wUtt, relevance_wUtt)
+  )
+
 ```
@@
 
+Exploring plots of posterior samples for pairwise parameters shows that there is a visible correlation between prior & competence, and a somewhat weaker correlation between competence & relevance.
+The interaction estimates also show a correlation with the main effects (as a rough overall judgment).
+```{r}
+shinystan::launch_shinystan(model_xor)
+```
+Plot correlations of main effects
+```{r}
+bayesplot::mcmc_pairs(model_xor, pars = c("b_prior", "b_competence", "b_relevance"))
+```
+
 ```{r, results='hide'}
@@
 summary(model_some)
+```
+
+For "some", the correlation of prior and relevance is higher than for prior and competence; there is also a visible correlation of competence and relevance. 
+```{r}
+shinystan::launch_shinystan(model_some)
+```
+```{r}
+bayesplot::mcmc_pairs(model_some, pars = c("b_prior", "b_competence", "b_relevance"))
+```
+
+#### Smaller models with raw data to check for colinearity 
+Models on raw data with only one predictor: 
+```{r, results='hide'}
+# xor
+model_xor_prior <- brm(
+  target ~ prior + 
+    (1 + prior | submission_id) +
+    (1 | title),
+  data = d_critical_wide %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_xor_prior)
+```
+
+```{r, results='hide'}
+# xor
+model_xor_comp <- brm(
+  target ~ competence + 
+    (1 + competence | submission_id) +
+    (1 | title),
+  data = d_critical_wide %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_xor_comp)
+```
+
+```{r, results='hide'}
+# xor
+model_xor_rel <- brm(
+  target ~ relevance + 
+    (1 + relevance | submission_id) +
+    (1 | title),
+  data = d_critical_wide %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_xor_rel)
+```
+
+```{r, results='hide'}
+# some
+model_some_rel <- brm(
+  target ~ relevance + 
+    (1 + relevance | submission_id) +
+    (1 | title),
+  data = d_critical_wide %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_some_rel)
+```
+
+```{r, results='hide'}
+# some
+model_some_prior <- brm(
+  target ~ prior + 
+    (1 + prior | submission_id) +
+    (1 | title),
+  data = d_critical_wide %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_some_prior)
+```
+
+```{r, results='hide'}
+# some
+model_some_comp <- brm(
+  target ~ competence + 
+    (1 + competence | submission_id) +
+    (1 | title),
+  data = d_critical_wide %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_some_comp)
+```
+
+With two predictors:
+```{r, results='hide'}
+# xor
+model_xor_pri_comp <- brm(
+  target ~ prior*competence + 
+    (1 + prior*competence || submission_id) +
+    (1 | title),
+  data = d_critical_wide %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_xor_pri_comp)
+```
+
+```{r}
+bayesplot::mcmc_pairs(model_xor_pri_comp, pars = c("b_prior", "b_competence"))
+```
+
+```{r, results='hide'}
+# xor
+model_xor_pri_rel <- brm(
+  target ~ prior*relevance + 
+    (1 + prior*relevance || submission_id) +
+    (1 | title),
+  data = d_critical_wide %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_xor_pri_rel)
+```
+
+```{r}
+bayesplot::mcmc_pairs(model_xor_pri_rel, pars = c("b_prior", "b_relevance"))
+```
+
+```{r, results='hide'}
+# xor
+model_xor_rel_comp <- brm(
+  target ~ relevance*competence + 
+    (1 + relevance*competence || submission_id) +
+    (1 | title),
+  data = d_critical_wide %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_xor_rel_comp)
+```
+
+```{r}
+bayesplot::mcmc_pairs(model_xor_rel_comp, pars = c("b_competence", "b_relevance"))
+```
+
+```{r, results='hide'}
+# some
+model_some_pri_comp <- brm(
+  target ~ prior*competence + 
+    (1 + prior*competence || submission_id) +
+    (1 | title),
+  data = d_critical_wide %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_some_pri_comp)
+```
+
+```{r}
+bayesplot::mcmc_pairs(model_some_pri_comp, pars = c("b_prior", "b_competence"))
+```
+
+```{r, results='hide'}
+# some
+model_some_pri_rel <- brm(
+  target ~ prior*relevance + 
+    (1 + prior*relevance || submission_id) +
+    (1 | title),
+  data = d_critical_wide %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_some_pri_rel)
+```
+
+```{r}
+bayesplot::mcmc_pairs(model_some_pri_rel, pars = c("b_prior", "b_relevance"))
+```
+
+```{r, results='hide'}
+# some
+model_some_rel_comp <- brm(
+  target ~ relevance*competence + 
+    (1 + relevance*competence || submission_id) +
+    (1 | title),
+  data = d_critical_wide %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_some_rel_comp)
+```
+
+```{r}
+bayesplot::mcmc_pairs(model_some_rel_comp, pars = c("b_competence", "b_relevance"))
 ```
@@
 
+```{r}
+bayesplot::mcmc_pairs(model_xor_zScored, pars = c("b_prior", "b_relevance", "b_competence"))
+```
+
 ```{r, results='hide'}
@@
 
+```{r}
+bayesplot::mcmc_pairs(model_some_zScored, pars = c("b_prior", "b_relevance", "b_competence"))
+```
+
+#### Smaller models with z-scored data
+```{r, results='hide'}
+# xor
+model_xor_zScored_prior <- brm(
+  target ~ prior + 
+    (1 + prior | submission_id) +
+    (1 | title),
+  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_xor_zScored_prior)
+```
+
+```{r, results='hide'}
+# xor
+model_xor_zScored_comp <- brm(
+  target ~ competence + 
+    (1 + competence | submission_id) +
+    (1 | title),
+  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_xor_zScored_comp)
+```
+
+```{r, results='hide'}
+# xor
+model_xor_zScored_rel <- brm(
+  target ~ relevance + 
+    (1 + relevance | submission_id) +
+    (1 | title),
+  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_xor_zScored_rel)
+```
+
+```{r, results='hide'}
+# some
+model_some_zScored_rel <- brm(
+  target ~ relevance + 
+    (1 + relevance | submission_id) +
+    (1 | title),
+  data = d_critical_zScored_wide %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_some_zScored_rel)
+```
+
+```{r, results='hide'}
+# some
+model_some_zScored_prior <- brm(
+  target ~ prior + 
+    (1 + prior | submission_id) +
+    (1 | title),
+  data = d_critical_zScored_wide %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_some_zScored_prior)
+```
+
+```{r, results='hide'}
+# some
+model_some_zScored_comp <- brm(
+  target ~ competence + 
+    (1 + competence | submission_id) +
+    (1 | title),
+  data = d_critical_zScored_wide %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_some_zScored_comp)
+```
+
+With two predictors:
+```{r, results='hide'}
+# xor
+model_xor_zScored_pri_comp <- brm(
+  target ~ prior*competence + 
+    (1 + prior*competence || submission_id) +
+    (1 | title),
+  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_xor_zScored_pri_comp)
+```
+
+```{r}
+bayesplot::mcmc_pairs(model_xor_zScored_pri_comp, pars = c("b_prior", "b_competence"))
+```
+
+```{r, results='hide'}
+# xor
+model_xor_zScored_pri_rel <- brm(
+  target ~ prior*relevance + 
+    (1 + prior*relevance || submission_id) +
+    (1 | title),
+  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_xor_zScored_pri_rel)
+```
+
+```{r}
+bayesplot::mcmc_pairs(model_xor_zScored_pri_rel, pars = c("b_prior", "b_relevance"))
+```
+
+```{r, results='hide'}
+# xor
+model_xor_zScored_rel_comp <- brm(
+  target ~ relevance*competence + 
+    (1 + relevance*competence || submission_id) +
+    (1 | title),
+  data = d_critical_zScored_wide %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_xor_zScored_rel_comp)
+```
+
+```{r}
+bayesplot::mcmc_pairs(model_xor_zScored_rel_comp, pars = c("b_relevance", "b_competence"))
+```
+
+```{r, results='hide'}
+# some
+model_some_zScored_pri_comp <- brm(
+  target ~ prior*competence + 
+    (1 + prior*competence || submission_id) +
+    (1 | title),
+  data = d_critical_zScored_wide %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_some_zScored_pri_comp)
+```
+
+```{r}
+bayesplot::mcmc_pairs(model_some_zScored_pri_comp, pars = c("b_prior", "b_competence"))
+```
+
+```{r, results='hide'}
+# some
+model_some_zScored_pri_rel <- brm(
+  target ~ prior*relevance + 
+    (1 + prior*relevance || submission_id) +
+    (1 | title),
+  data = d_critical_zScored_wide %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_some_zScored_pri_rel)
+```
+
+```{r}
+bayesplot::mcmc_pairs(model_some_zScored_pri_rel, pars = c("b_prior", "b_relevance"))
+```
+
+```{r, results='hide'}
+# some
+model_some_zScored_rel_comp <- brm(
+  target ~ relevance*competence + 
+    (1 + relevance*competence || submission_id) +
+    (1 | title),
+  data = d_critical_zScored_wide %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+
+```{r}
+summary(model_some_zScored_rel_comp)
+```
+
+```{r}
+bayesplot::mcmc_pairs(model_some_zScored_rel_comp, pars = c("b_relevance", "b_competence"))
+```
+
 #### Stats with dichotomized predictors 
@@
   )  %>%
-  mutate(relevance = as.factor(relevance),
-         competence = as.factor(competence),
-         prior = as.factor(prior)) %>% 
+  mutate(relevance = factor(relevance, levels = c(1, 0)),
+         competence = factor(competence, levels = c(1, 0)),
+         prior = factor(prior, levels = c(1, 0))) %>% 
   filter(block_extended == "target") -> d_critical_wide_cat
 # sum code predictors
-# low prior: 1, high prior:: -1 
+# low prior: -1, high prior:: 1 
 contrasts(d_critical_wide_cat$prior) <- contr.sum(2)
-# low comp : 1, high comp : -1
+# low comp : -1, high comp : 1
 contrasts(d_critical_wide_cat$competence) <- contr.sum(2)
-# low rel: 1, high rel: -1
+# low rel: -1, high rel: 1
 contrasts(d_critical_wide_cat$relevance) <- contr.sum(2)
@@
 ```
+We test the following hypotheses for the categorical predictors models (identical for both or and some):
+H1: For higher prior of the stronger alternative being true, the inference strength is lower
+H2: For higher speaker comptence, the inference strength is higher
+H3: For higher listener relevance, the inference strength is higher 
+```{r}
+# xor
+# test H1
+hypothesis(model_xor_cat, "prior1 < 0")
+# test H2
+hypothesis(model_xor_cat, "competence1 > 0")
+# test H3
+hypothesis(model_xor_cat, "relevance1 > 0")
+```
 
@@
 ```
+
 ```{r}
 summary(model_some_cat)
+```
+
+```{r}
+# some
+# test H1
+hypothesis(model_some_cat, "prior1 < 0")
+# test H2
+hypothesis(model_some_cat, "competence1 > 0")
+# test H3
+hypothesis(model_some_cat, "relevance1 > 0")
 ```
@@
   )  %>%
-  mutate(relevance = as.factor(relevance),
-         competence = as.factor(competence),
-         prior = as.factor(prior)) %>% 
+  mutate(relevance = factor(relevance, levels = c(1, 0)),
+         competence = factor(competence, levels = c(1, 0)),
+         prior = factor(prior, levels = c(1, 0))) %>% 
   filter(block_extended == "target") -> d_critical_wide_cat_zScored
 # sum code predictors
-# low prior: 1, high prior:: -1 
+# low prior: -1, high prior:: 1 
 contrasts(d_critical_wide_cat_zScored$prior) <- contr.sum(2)
-# low comp : 1, high comp : -1
+# low comp : -1, high comp : 1
 contrasts(d_critical_wide_cat_zScored$competence) <- contr.sum(2)
-# low rel: 1, high rel: -1
+# low rel: -1, high rel: 1
 contrasts(d_critical_wide_cat_zScored$relevance) <- contr.sum(2)
@@
 
+```{r}
+# xor
+# test H1
+hypothesis(model_xor_cat_zScored, "prior1 < 0")
+# test H2
+hypothesis(model_xor_cat_zScored, "competence1 > 0")
+# test H3
+hypothesis(model_xor_cat_zScored, "relevance1 > 0")
+```
+
 ```{r, results='hide'}
@@
 
-In summary, the different models provide somewhat different results, specially when using raw ratings vs.
-binary predictors.
-Since interpreting the categorical models in relation to the plots might be not as straight-forward, I am not entirely sure whether the qualitative results match in all models.
-Using centered data seems to provide a little tighter estimates in the models; not sure where the slight qualitative differences come from.
+```{r}
+# some
+# test H1
+hypothesis(model_some_cat_zScored, "prior1 < 0")
+# test H2
+hypothesis(model_some_cat_zScored, "competence1 > 0")
+# test H3
+hypothesis(model_some_cat_zScored, "relevance1 > 0")
+```
+
+#### Single predictor z-scored models with categorical predictors
+For xor:
+```{r, results="hide"}
+model_xor_cat_zScored_pri <- brm(
+  response_centered ~ prior + 
+    (1 + prior | submission_id) +
+    (1 | title),
+  data = d_critical_wide_cat_zScored %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+```{r}
+summary(model_xor_cat_zScored_pri)
+```
+
+```{r, results="hide"}
+model_xor_cat_zScored_comp <- brm(
+  response_centered ~ competence + 
+    (1 + competence | submission_id) +
+    (1 | title),
+  data = d_critical_wide_cat_zScored %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+```{r}
+summary(model_xor_cat_zScored_comp)
+```
+
+```{r, results="hide"}
+model_xor_cat_zScored_rel <- brm(
+  response_centered ~ relevance + 
+    (1 + relevance | submission_id) +
+    (1 | title),
+  data = d_critical_wide_cat_zScored %>% filter(main_type == "xor"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+```{r}
+summary(model_xor_cat_zScored_rel)
+```
+
+For some:
+```{r, results="hide"}
+model_some_cat_zScored_pri <- brm(
+  response_centered ~ prior + 
+    (1 + prior | submission_id) +
+    (1 | title),
+  data = d_critical_wide_cat_zScored %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+```{r}
+summary(model_some_cat_zScored_pri)
+```
+
+```{r, results="hide"}
+model_some_cat_zScored_comp <- brm(
+  response_centered ~ competence + 
+    (1 + competence | submission_id) +
+    (1 | title),
+  data = d_critical_wide_cat_zScored %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+```{r}
+summary(model_some_cat_zScored_comp)
+```
+
+```{r, results="hide"}
+model_some_cat_zScored_rel <- brm(
+  response_centered ~ relevance + 
+    (1 + relevance | submission_id) +
+    (1 | title),
+  data = d_critical_wide_cat_zScored %>% filter(main_type == "some"),
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+```
+```{r}
+summary(model_some_cat_zScored_rel)
+```
+
+In summary, the different models provide somewhat different results, specially when using raw ratings vs. binary predictors.

### 2021-04-09 analysis/xor-some-prolific-pilot2.Rmd
knit

@@
 
-Plot inference strength ratings against raw predictor ratings for each participant / each item:
+Plot inference strength ratings against raw predictor ratings from each participant across items:
 ```{r}
@@
 
-Plot inference strength ratings against predictor ratings, by-item, using by-item mean ratings across participants. Error bard indicate the standard deviation:
+Plot inference strength ratings against predictor ratings, by-item, using by-item mean ratings across participants. Error bars indicate the standard deviation of the mean:
 ```{r, fig.height=4, fig.width=8}
@@
 
-Plot z-scored predictor ratings against z-scored target ratings
+Plot z-scored predictor ratings against z-scored target inference ratings
 ```{r}
@@
 ## Stats
+
 Maximal models on raw ratings:
@@
 ```{r}
-shinystan::launch_shinystan(model_xor)
+#shinystan::launch_shinystan(model_xor)
 ```
@@
 ```{r}
-shinystan::launch_shinystan(model_some)
+#shinystan::launch_shinystan(model_some)
 ```
@@
 
-#### Smaller models with raw data to check for colinearity 
+#### Smaller models with raw data to check for collinearity 
 Models on raw data with only one predictor: 
@@
 #### Stats on z-scored data
+The same stats (full, one predictor, to predictors) on z-scored data:
 ```{r, results='hide'}
@@
 ```
-
-In summary, the different models provide somewhat different results, specially when using raw ratings vs. binary predictors.

### 2021-04-10 analysis/xor-some-prolific-pilot2.Rmd
fix fig sizes and adapt_delta

@@
 Visually check the spread of by-item ratings among subjects for each question:
-```{r, fig.height=20, fig.width=7}
+```{r, fig.height=20, fig.width=10}
 
@@
 Plot by-item trends between predictor ratings and target inference strength ratings (relevance / competence ratings with utterance are omitted):
-```{r, fig.height=20, fig.width=8}
+```{r, fig.height=35, fig.width=15}
 # make a versatile wide representation of the critical data
@@
 Plot inference strength ratings against predictor ratings, by-item, using by-item mean ratings across participants. Error bars indicate the standard deviation of the mean:
-```{r, fig.height=4, fig.width=8}
+```{r, fig.height=10, fig.width=8}
 d_critical_predictorsXtarget_summary <- d_critical_predictorsXtarget %>% 
@@
   geom_errorbarh(aes(xmin = mean_predictor - sd_predictor, xmax = mean_predictor + sd_predictor), alpha = 0.7) +
-  geom_text(aes(label=title), size = 1.5, nudge_y = 4) +
+  geom_text(aes(label=title), size = 2.5, nudge_y = 4) +
   geom_smooth(method = "lm") +
@@
   family = "gaussian",
+  control = list(adapt_delta = 0.99),
   cores = 4,
@@
   family = "gaussian",
+  control = list(adapt_delta = 0.92),
   cores = 4,

### 2021-04-26 analysis/xor-some-prolific-pilot2.Rmd
adding QR decomposition for sampling efficiency

@@
 model_xor <- brm(
-  target ~ prior*competence*relevance + 
+  bf(target ~ prior*competence*relevance + 
     (1 + prior + competence + relevance || submission_id) +
     (1 | title),
+    decomp = "QR"
+  ), 
   data = d_critical_wide %>% filter(main_type == "xor"),

### 2021-04-28 analysis/xor-some-prolific-pilot2.Rmd
tweak exclusions, add low completion times, xor prior responses check

@@
 ```
+
+Investigate very fast completion times:
+```{r}
+ d_native %>% mutate(timeSpent = round(timeSpent, 2)) %>%
+  filter(timeSpent < 8) -> fast_completions
+
+# check if their responses look somehow peculiar
+fast_completions %>%
+  ggplot(., aes(x = response)) +
+  geom_histogram()
+
+cat("number of fast-responders (below 8 mins):", fast_completions %>% distinct(submission_id) %>% pull() %>% length())
+```
+Data frames preparation:
 ```{r}
@@
 # get participants failing example trials
-d_exmpl <- d_exmpl %>% group_by(submission_id) %>% 
+d_exmpl_fail <- d_exmpl %>% group_by(submission_id) %>% 
   mutate(example_condition = ifelse(grepl("as certainly true", question), "true", 
@@
 
-cat("Subjects failing the example trials: ", d_exmpl %>% distinct(submission_id) %>% pull() %>% length())
+cat("Subjects failing the example trials: ", d_exmpl_fail %>% distinct(submission_id) %>% pull() %>% length())
 
@@
          ) %>% filter(passed_main == F)
-cat("Subjects providing the same ratings throughout the trials: ", d_main_fail %>% distinct(submission_id) %>% pull() %>% length())
+cat(" Subjects providing the same ratings throughout the trials: ", d_main_fail %>% distinct(submission_id) %>% pull() %>% length())
 
@@
          passed_filler = mean_comprehension >= 0.8
-         ) %>%
+         )
+
+d_test_fail <- d_test %>% 
   filter(passed_filler == F)
 
-cat("Subjects failing the comprehension trials: ", d_test %>% distinct(submission_id) %>% pull() %>% length())
-
+cat(" Subjects failing the comprehension trials: ", d_test_fail %>% distinct(submission_id) %>% pull() %>% length())
+
+# plot by-participant means on comprehension questions, and mark red excluded points, 
+# to see if the participants performed in a peculiar way
+d_test %>% 
+  group_by(submission_id, main_type, test_condition) %>%
+  mutate(subj_mean = mean(response)) %>%
+  ggplot(., aes(x = test_condition, y = subj_mean)) +
+  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
+  geom_point(data = . %>% filter(passed_filler == F), aes(x = test_condition, y = subj_mean), color = "red", 
+             size = 2) +
+  facet_wrap(~main_type) +
+  ggtitle("Red points mark subject response means for subjects\nwith a mean comprehension question accuracy < 0.8")
+```
+
+```{r}
 # put it all together
 d_full_clean <- anti_join(d_main, d_main_fail, by = "submission_id")
-d_full_clean <- anti_join(d_full_clean, d_exmpl, d_test, by = "submission_id")
-d_full_clean <- anti_join(d_full_clean, d_test, by = "submission_id")
-
-cat("Nr. of participants left after cleaning: ", d_full_clean %>% distinct(submission_id) %>% pull() %>% length())
+d_full_clean <- anti_join(d_full_clean, d_exmpl_fail, d_test, by = "submission_id")
+d_full_clean <- anti_join(d_full_clean, d_test_fail, by = "submission_id")
+
+cat(" Nr. of participants left after cleaning: ", d_full_clean %>% distinct(submission_id) %>% pull() %>% length())
+```
+
+Check exclusions if applying exclusion ranges as suggested by Bob -- then almost twice as many participants are excluded based on their comprehension questions performance (I believe mainly due to a lower bound on correct responses in the uncertain questions).
+```{r}
+# true comprehension questions: 50-100
+# uncertain comprehension questions: 25-75
+# false comprehension questions: 0-50
+# get participants failing example trials
+
+d_exmpl2 <- d_exmpl %>% group_by(submission_id) %>% 
+  mutate(example_condition = ifelse(grepl("as certainly true", question), "true", 
+                                    ifelse(grepl("as certainly false", question), "false",
+                                           "uncertain")),
+         passed_example_trial = case_when(example_condition == "true" ~ response >= 50,
+                                    example_condition == "false" ~ response <= 50,
+                                    example_condition == "uncertain" ~ response %in% (25:75)),
+       #  check if all trials passed
+         passed_example = case_when(sum(passed_example_trial) == 0 ~ FALSE,
+                                    TRUE ~ TRUE)
+         ) %>% filter(passed_example == F)
+
+cat("Subjects failing the example trials: ", d_exmpl2 %>% distinct(submission_id) %>% pull() %>% length())
+
+# apply exclusion criteria to main trials
+# check range of responses per participant
+d_main_fail <- d_main %>% group_by(submission_id) %>%
+  mutate(passed_main = case_when(max(response) - min(response) <= 10 ~ FALSE,
+                                 TRUE ~ TRUE)
+         ) %>% filter(passed_main == F)
+cat(" Subjects providing the same ratings throughout the trials: ", d_main_fail %>% distinct(submission_id) %>% pull() %>% length())
+
+# get participants failing comprehension questions
+d_test2 <- d_test %>%
+  group_by(submission_id) %>%
+  mutate(passed_filler_trial = case_when(test_condition == "true" ~ response >= 50,
+                                   test_condition == "false" ~ response <= 50,
+                                   test_condition == "uncertain" ~ response %in% (25:75)),
+         mean_comprehension = mean(passed_filler_trial),
+         passed_filler = mean_comprehension >= 0.8
+         )
+
+d_test_fail2 <- d_test2 %>% 
+  filter(passed_filler == F)
+
+cat(" Subjects failing the comprehension trials: ", d_test_fail2 %>% distinct(submission_id) %>% pull() %>% length())
+
+```
+Now we explore applying exclusion criteria on z-scored data. Here it might make sense to exclude participants whose >=0.8 of responses are more than 2 SD from the mean (?). 
+```{r}
+# group by extended block, exmpl / filler / critical, maybe polarity of filler exmpl question
+# add block extension, put all data together somehow
+d_critical_fail_zScored <- d_main %>% 
+  mutate(block_extended = ifelse(
+    !(ifelse(is.na(critical_question), F, T)), 
+    block, 
+    ifelse(block %in% c("some", "xor"), "target", str_c(block, "_wUtt", ""))
+  )  
+  ) %>% 
+  filter(!(grepl("test", block_extended))) %>%
+  group_by(submission_id, block_extended) %>%
+  mutate(subj_block_mean = mean(response),
+         subj_block_sd = sd(response),
+         response_centered = (response - subj_block_mean)/subj_block_sd,
+         # catch the cases where sd is 0 
+         response_centered = ifelse(is.na(response_centered), 0, response_centered)) %>%
+  ungroup() %>%
+  # add by-block SDs, such that participants can be excluded based on being outside 2 SDs on a block
+  # or should this be based on by-subject SDs???? 
+  group_by(block_extended) %>%
+  mutate(block_mean = mean(response),
+         block_sd = sd(response),
+         passed_main_trial = case_when(abs(response_centered) <= 2*block_sd ~ T,
+                                       T ~ F),
+         mean_critical_passRate = mean(passed_main_trial))
+
+# get participants failing comprehension questions with z-scored data; here, by-subject SD are used for the exclusion
+d_test_fail_zScore <- d_test %>%
+  group_by(submission_id, test_condition) %>%
+  mutate(
+    subj_block_mean = mean(response),
+    subj_block_sd = sd(response),
+    response_centered = (response - subj_block_mean)/subj_block_sd,
+    response_centered = ifelse(is.na(response_centered), 0, response_centered),
+    passed_filler_trial = case_when(abs(response_centered) <= 2*subj_block_sd ~ T,
+                                   T ~ F),
+         mean_comprehension = mean(passed_filler_trial),
+         passed_filler = mean_comprehension >= 0.8
+         )
+
+d_test_fail_zScore <- d_test_fail_zScore %>% 
+  filter(passed_filler == F)
 ```
@@
 ```
-```{r}
+```{r, fig.height=3}
 # check whether the ratings in the two prior questions are the same in xor 
@@
   filter(block == "prior", main_type == "xor") %>%
-  mutate(priorQ_nr = rep(c(1,2), 396)) -> d_xor_priors
+  mutate(priorQ_nr = rep(c(1,2), 396),
+         unique_ID = paste(submission_id, ID, sep="_")) -> d_xor_priors
 d_xor_priors %>% group_by(priorQ_nr, prior_class) %>%
   summarise(mean = mean(response)) -> d_xor_priors_summary
+
 d_xor_priors %>%
@@
   geom_point(data = d_xor_priors_summary, aes(x = as.factor(priorQ_nr), y = mean), color = "red", size = 3) +
+  geom_path(aes(group = "unique_ID"), alpha = 0.6) +
   ylab("Responses to prior questions") +
   xlab("First vs Second prior question for high vs low prior conditions") +
-  facet_wrap(~as.factor(prior_class)) # get ratings from the respective trials only
-
+  facet_wrap(~as.factor(prior_class)) + # get ratings from the respective trials only 
+  ggtitle("Xor prior question ratings.\nLines indicate the two responses provided by one participant for a specific item.")
 ```
@@
 model_some <- brm(
-  target ~ prior*competence*relevance + 
+  bf(target ~ prior*competence*relevance + 
     (1 + prior + competence + relevance || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide %>% filter(main_type == "some"),
@@
 model_xor_prior <- brm(
-  target ~ prior + 
+  bf(target ~ prior + 
     (1 + prior | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide %>% filter(main_type == "xor"),
@@
 model_xor_comp <- brm(
-  target ~ competence + 
+  bf(target ~ competence + 
     (1 + competence | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide %>% filter(main_type == "xor"),
@@
 model_xor_rel <- brm(
-  target ~ relevance + 
+  bf(target ~ relevance + 
     (1 + relevance | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide %>% filter(main_type == "xor"),
@@
 model_some_rel <- brm(
-  target ~ relevance + 
+  bf(target ~ relevance + 
     (1 + relevance | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide %>% filter(main_type == "some"),
@@
 model_some_prior <- brm(
-  target ~ prior + 
+  bf(target ~ prior + 
     (1 + prior | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide %>% filter(main_type == "some"),
@@
 model_some_comp <- brm(
-  target ~ competence + 
+  bf(target ~ competence + 
     (1 + competence | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide %>% filter(main_type == "some"),
@@
 model_xor_pri_comp <- brm(
+  bf(
   target ~ prior*competence + 
@@
     (1 | title),
+  decomp = "QR"
+  ),
   data = d_critical_wide %>% filter(main_type == "xor"),
@@
 model_xor_pri_rel <- brm(
-  target ~ prior*relevance + 
+  bf(target ~ prior*relevance + 
     (1 + prior*relevance || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide %>% filter(main_type == "xor"),
@@
 model_xor_rel_comp <- brm(
-  target ~ relevance*competence + 
+  bf(target ~ relevance*competence + 
     (1 + relevance*competence || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide %>% filter(main_type == "xor"),
@@
 model_some_pri_comp <- brm(
-  target ~ prior*competence + 
+  bf(target ~ prior*competence + 
     (1 + prior*competence || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide %>% filter(main_type == "some"),
@@
 model_some_pri_rel <- brm(
-  target ~ prior*relevance + 
+  bf(target ~ prior*relevance + 
     (1 + prior*relevance || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide %>% filter(main_type == "some"),
@@
 model_some_rel_comp <- brm(
-  target ~ relevance*competence + 
+  bf(target ~ relevance*competence + 
     (1 + relevance*competence || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide %>% filter(main_type == "some"),
@@
 model_xor_zScored <- brm(
-  target ~ prior*competence*relevance + 
+  bf(target ~ prior*competence*relevance + 
     (1 + prior + competence + relevance || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_zScored_wide %>% filter(main_type == "xor"),
@@
 model_some_zScored <- brm(
-  target ~ prior*competence*relevance + 
+  bf(target ~ prior*competence*relevance + 
     (1 + prior + competence + relevance || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_zScored_wide %>% filter(main_type == "some"),
@@
 model_xor_zScored_prior <- brm(
-  target ~ prior + 
+  bf(target ~ prior + 
     (1 + prior | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_zScored_wide %>% filter(main_type == "xor"),
@@
 model_xor_zScored_comp <- brm(
-  target ~ competence + 
+  bf(target ~ competence + 
     (1 + competence | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_zScored_wide %>% filter(main_type == "xor"),
@@
 model_xor_zScored_rel <- brm(
-  target ~ relevance + 
+  bf(target ~ relevance + 
     (1 + relevance | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_zScored_wide %>% filter(main_type == "xor"),
@@
 model_some_zScored_rel <- brm(
-  target ~ relevance + 
+  bf(target ~ relevance + 
     (1 + relevance | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_zScored_wide %>% filter(main_type == "some"),
@@
 model_some_zScored_prior <- brm(
-  target ~ prior + 
+  bf(target ~ prior + 
     (1 + prior | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_zScored_wide %>% filter(main_type == "some"),
@@
 model_some_zScored_comp <- brm(
-  target ~ competence + 
+  bf(target ~ competence + 
     (1 + competence | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_zScored_wide %>% filter(main_type == "some"),
@@
 model_xor_zScored_pri_comp <- brm(
-  target ~ prior*competence + 
+  bf(target ~ prior*competence + 
     (1 + prior*competence || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_zScored_wide %>% filter(main_type == "xor"),
@@
 model_xor_zScored_pri_rel <- brm(
-  target ~ prior*relevance + 
+  bf(target ~ prior*relevance + 
     (1 + prior*relevance || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_zScored_wide %>% filter(main_type == "xor"),
@@
 model_xor_zScored_rel_comp <- brm(
-  target ~ relevance*competence + 
+  bf(target ~ relevance*competence + 
     (1 + relevance*competence || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_zScored_wide %>% filter(main_type == "xor"),
@@
 model_some_zScored_pri_comp <- brm(
-  target ~ prior*competence + 
+  bf(target ~ prior*competence + 
     (1 + prior*competence || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_zScored_wide %>% filter(main_type == "some"),
@@
 model_some_zScored_pri_rel <- brm(
-  target ~ prior*relevance + 
+  bf(target ~ prior*relevance + 
     (1 + prior*relevance || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_zScored_wide %>% filter(main_type == "some"),
@@
 model_some_zScored_rel_comp <- brm(
-  target ~ relevance*competence + 
+  bf(target ~ relevance*competence + 
     (1 + relevance*competence || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_zScored_wide %>% filter(main_type == "some"),
@@
 model_xor_cat <- brm(
-  response ~ prior*competence*relevance + 
+  bf(response ~ prior*competence*relevance + 
     (1 + prior + competence + relevance || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide_cat %>% filter(main_type == "xor"),
@@
 model_some_cat <- brm(
-  response ~ prior*competence*relevance + 
+  bf(response ~ prior*competence*relevance + 
     (1 + prior + competence + relevance || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide_cat %>% filter(main_type == "some"),
@@
 model_xor_cat_zScored <- brm(
-  response_centered ~ prior*competence*relevance + 
+  bf(response_centered ~ prior*competence*relevance + 
     (1 + prior + competence + relevance || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide_cat_zScored %>% filter(main_type == "xor"),
@@
 model_some_cat_zScored <- brm(
-  response_centered ~ prior*competence*relevance + 
+  bf(response_centered ~ prior*competence*relevance + 
     (1 + prior + competence + relevance || submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide_cat_zScored %>% filter(main_type == "some"),
@@
 model_xor_cat_zScored_pri <- brm(
-  response_centered ~ prior + 
+  bf(response_centered ~ prior + 
     (1 + prior | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide_cat_zScored %>% filter(main_type == "xor"),
@@
 model_xor_cat_zScored_comp <- brm(
-  response_centered ~ competence + 
+  bf(response_centered ~ competence + 
     (1 + competence | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide_cat_zScored %>% filter(main_type == "xor"),
@@
 model_xor_cat_zScored_rel <- brm(
-  response_centered ~ relevance + 
+  bf(response_centered ~ relevance + 
     (1 + relevance | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide_cat_zScored %>% filter(main_type == "xor"),
@@
 model_some_cat_zScored_pri <- brm(
-  response_centered ~ prior + 
+  bf(response_centered ~ prior + 
     (1 + prior | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide_cat_zScored %>% filter(main_type == "some"),
@@
 model_some_cat_zScored_comp <- brm(
-  response_centered ~ competence + 
+  bf(response_centered ~ competence + 
     (1 + competence | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide_cat_zScored %>% filter(main_type == "some"),
@@
 model_some_cat_zScored_rel <- brm(
-  response_centered ~ relevance + 
+  bf(response_centered ~ relevance + 
     (1 + relevance | submission_id) +
     (1 | title),
+    decomp = "QR"),
   data = d_critical_wide_cat_zScored %>% filter(main_type == "some"),

### 2021-05-04 analysis/xor-some-prolific-pilot2.Rmd
minor fix

@@
 d_test_fail_zScore <- d_test_fail_zScore %>% 
+  ungroup() %>%
   filter(passed_filler == F)
+
+cat(" Subjects failing the comprehension trials when using z-scored raw data: ", d_test_fail_zScore %>% distinct(submission_id) %>% pull() %>% length())
+
 ```

### 2021-05-04 analysis/xor-some-prolific-pilot2.Rmd
add counts of prior response diffs

@@
 ```
+
 ```{r, fig.height=3}
@@
   ggtitle("Xor prior question ratings.\nLines indicate the two responses provided by one participant for a specific item.")
+```
+```{r}
+# check analytically how many subjects rate prior questions differently
+d_xor_priors %>% select(-trial_number, -RT, -prompt) %>% 
+  pivot_wider(names_from = "priorQ_nr", values_from = "response") %>%
+  mutate(response_difference = abs(`1` - `2`),
+         bigDifference = ifelse(response_difference > 15, T, F)) -> d_xor_priors_differences
+
+# 41 responses with difference > 15 between first and second rating
+d_xor_priors_differences %>% count(bigDifference)
+
+# proportion of big differences in prior responses by prior type (high vs low)
+d_xor_priors_differences %>% group_by(prior_class) %>% summarise(prop_big_diffs = mean(bigDifference))
 ```

### 2021-05-04 analysis/xor-some-prolific-pilot2.Rmd
knit and upate by-item breakdown plot

@@
 Visually check the spread of by-item ratings among subjects for each question:
-```{r, fig.height=20, fig.width=10}
+```{r, fig.height=25, fig.width=10}
+d_critical_long_grandMeans <- d_critical_long %>% 
+  mutate(title = paste(title, exp_condition, sep = "_"),
+         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
+  filter(block == class_condition | block == "target") %>%
+  group_by(main_type, block) %>% summarize(mean_resp = mean(response))
+
+d_critical_long_expectedCats <- d_critical_long %>% 
+  mutate(title = paste(title, exp_condition, sep = "_"),
+         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
+  filter(block == class_condition | block == "target") %>%
+  group_by(main_type, block, title) %>% summarize(mean_resp_block = mean(response))
 
@@
          block = ifelse(block == "xor" | block == "some", "target", block)) %>%
-  filter(block == class_condition | block == "target") %>%
+  left_join(., d_critical_long_grandMeans, by = c("main_type", "block")) %>% filter(block == class_condition | block == "target") -> d_long_place_holder
+
+left_join(d_long_place_holder, d_critical_long_expectedCats, by = c("main_type", "block", "title")) %>% 
+  mutate(expected_block = ifelse(block != "target", prior_class * 100, NA),
+         block = factor(block, levels = c("relevance", "competence", "prior"))) %>%
   ggplot(., aes(x = block, y = response, color = block)) +
   geom_point(alpha = 0.7, position = position_jitter(width = 0.1)) +
+  geom_hline(aes(yintercept = mean_resp), alpha = 0.7) +
+  geom_point(aes(y = mean_resp), color = "black", shape = 2) +
+  geom_point(aes(y = mean_resp_block, fill = block), color = "black") +
+  geom_point(aes(y = expected_block), color="red", size = 2.5) +
   ylab("Raw responses to respective questions") +
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  ggtitle("By-item by-question raw ratings")
+  ggtitle("By-item by-question raw ratings\nBlack dots indicate by-block means within-item, lines and triangles indicate grand by-block means\nRed dots indicate expected rating")
 ```

### 2021-07-08 04_trials.js
add final items

@@ -2096,6 +2096,9 @@ const trials_practice = [{
 // rel_block = _.flatten(_.map(_.shuffle(eight_stories), function (x) {
 //   return (create_block(x, 'rel'));
 // }));
+console.log("Log the practice trials")
+console.log(trials_practice)
+
 xor_block = _.map(_.shuffle(or_stories), function (x) {
   return create_block(x, "xor"); // 'xor'
 });

### 2021-07-08 04_trials_final.js

@@ -1,0 +1,1843 @@
+// In this file you can specify the trial data for your experiment
+// info important to create tirals
+
+// vignettes is an object which contains all necessary information for the trials
+let vignettes_or = {
+  lll: [
+    {
+   "FIELD1": 7,
+   "name": "Jimmy's trip",
+   "author": "micha",
+   "final_selection": 1,
+   "type": "lll",
+   "relevance": 0,
+   "competence": 0,
+   "prior": 0,
+   "background": "Jimmy went on a day trip last week. He has a younger brother who is very talkative even if he doesn’t know what he is talking about. Jimmy’s brother told the neighbors about the trip afterwards.",
+   "utterance_or": "Jimmy’s younger brother told the neighbors: ‘Jimmy visited New York or Los Angeles.’",
+   "utterance_disj1": "Jimmy's younger brother told the neighbors: 'Jimmy visited New York.’",
+   "utterance_disj2": "Jimmy's younger brother told the neighbors: 'Jimmy visited Los Angeles.’",
+   "question_rel": "It is important for the neighbors to know whether Jimmy visited both New York and Los Angeles on his day trip.",
+   "question_comp": "Jimmy's brother knows whether Jimmy visited both New York and Los Angeles on his day trip.",
+   "question_pri1": "If Jimmy visited New York, it is likely that he also visited Los Angeles on his day trip.",
+   "question_pri2": "If Jimmy visited Los Angeles, it is likely that he also visited New York on his day trip.",
+   "question_pri_and": "It is likely that Jimmy visited both New York and Los Angeles on his day trip. ",
+   "question_xor": "From what Jimmy’s brother said we may conclude that Jimmy did not visit both New York and Los Angeles.",
+   "question_xorI": "Jimmy's brother wanted to convey that Jimmy did not visit both New York and Los Angeles.",
+   "question_exh1": "From what Jimmy’s brother said we may conclude that Jimmy did not visit Los Angeles as well.",
+   "question_exh2": "From what Jimmy’s brother said we may conclude that Jimmy did not visit New York as well.",
+   "test_true1": "Jimmy has a sibling.",
+   "test_true2": "Jimmy did not spend the night in the city he visited.",
+   "test_false1": "Jimmy’s brother took the trip together with Jimmy.",
+   "test_false2": "Jimmy is an exemplary student.",
+   "test_uncertain1": "Jimmy had told his friends about his plans to take a day trip before.",
[…]

### 2021-07-09 02_custom_functions.js
fixed errors in creating views

@@ -6,7 +6,7 @@
 function assign_questions(vign) {
   let pri_question;
   for (var i = 0; i < vign.length; i++) {
-    vign[i].allQ = _.shuffle([vign[i].test_true1, vign[i].test_true2, vign[i].test_true, vign[i].test_false1, vign[i].test_false2, vign[i].test_false, vign[i].test_uncertain1, vign[i].test_uncertain2, vign[i].test_uncertain]);
+    vign[i].allQ = _.shuffle([vign[i].test_true1, vign[i].test_true2, vign[i].test_false1, vign[i].test_false2, vign[i].test_uncertain1, vign[i].test_uncertain2]);
     pri_question = _.shuffle([vign[i].question_pri1, vign[i].question_pri2]);
     vign[i].critical_question = [vign[i].question_comp, vign[i].question_rel, vign[i].question_xor, pri_question[0], pri_question[1]];
   }
@@ -16,7 +16,7 @@ function assign_questions(vign) {
 // add critical questions for some and use only the 6 comprehension questions that are available for some
 function assign_questions_some(vign) {
   for (var i = 0; i < vign.length; i++) {
-    vign[i].allQ = _.shuffle([vign[i].test_true1, vign[i].test_true, vign[i].test_false1, vign[i].test_false, vign[i].test_uncertain1, vign[i].test_uncertain]);
+    vign[i].allQ = _.shuffle([vign[i].test_true1, vign[i].test_true2, vign[i].test_false1, vign[i].test_false2, vign[i].test_uncertain1, vign[i].test_uncertain2]);
     vign[i].critical_question = [vign[i].question_comp, vign[i].question_rel, vign[i].question_imp, vign[i].question_priC];
   }
   return (vign);
@@ -24,7 +24,7 @@ function assign_questions_some(vign) {
 
 // construct recording of the used test question
 function test_Q_used(block, item, type, q_nr) {
-  if (type == "xor") {
+  // if (type == "xor") {
     switch(true) {
       case (block.allQ[q_nr] == block.test_true1):
         item.test_question = 'test_true1';
@@ -32,51 +32,51 @@ function test_Q_used(block, item, type, q_nr) {
       case (block.allQ[q_nr] == block.test_true2):
         item.test_question = 'test_true2';
         break;
-      case (block.allQ[q_nr] == block.test_true):
-          item.test_question = 'test_true';
-          break;
+      // case (block.allQ[q_nr] == block.test_true):
[…]

### 2021-07-09 04_trials_final.js

@@ -5,7 +5,7 @@
 let vignettes_or = {
   lll: [
     {
-   "FIELD1": 7,
+   "ID": 7,
    "name": "Jimmy's trip",
    "author": "micha",
    "final_selection": 1,
@@ -34,7 +34,7 @@ let vignettes_or = {
    "test_uncertain2": "Jimmy’s brother has been to New York City last year."
  },
  {
-   "FIELD1": 15,
+   "ID": 15,
    "name": "Harold's pet",
    "author": "bob",
    "final_selection": 1,
@@ -63,7 +63,7 @@ let vignettes_or = {
    "test_uncertain2": "Richard has a dog."
  },
  {
-   "FIELD1": 16,
+   "ID": 16,
    "name": "Dave's kids",
    "author": "natalie",
    "final_selection": 1,
@@ -92,7 +92,7 @@ let vignettes_or = {
    "test_uncertain2": "Dave’s wife went to school with them, too."
  },
  {
-   "FIELD1": 41,
+   "ID": 41,
    "name": "Brad Pitt",
    "author": "bob",
    "final_selection": 1,
@@ -123,7 +123,7 @@ let vignettes_or = {
 ],
 llh: [
   {
-   "FIELD1": 6,
+   "ID": 6,
    "name": "Joshua's delay",
    "author": "micha",
    "final_selection": 1,
@@ -152,7 +152,7 @@ llh: [
    "test_uncertain2": "The boys like to play hide and seek."
  },
  {
-   "FIELD1": 14,
+   "ID": 14,
    "name": "Gerald's graduation",
    "author": "bob",
    "final_selection": 1,
@@ -181,7 +181,7 @@ llh: [
    "test_uncertain2": "Gerald studied Computer Science."
  },
  {
-   "FIELD1": 22,
+   "ID": 22,
    "name": "Greg's movie",
    "author": "natalie",
    "final_selection": 1,
@@ -210,7 +210,7 @@ llh: [
    "test_uncertain2": "Greg considers going to the cinema with his sister next weekend."
  },
  {
-   "FIELD1": 25,
+   "ID": 25,
    "name": "Alissa's paint",
    "author": "natalie",
    "final_selection": 1,
@@ -240,7 +240,7 @@ llh: [
  }
 ],
 lhl: [ {
-   "FIELD1": 5,
+   "ID": 5,
    "name": "Tony's clubbing",
    "author": "micha",
    "final_selection": 1,
@@ -269,7 +269,7 @@ lhl: [ {
[…]

### 2021-07-09 06_main.js

@@ -30,7 +30,7 @@ $("document")
 
       // Here, you can specify all information for the deployment
       deploy: {
-        experimentID: "58",
+        experimentID: "63",
         serverAppURL: "https://mcmpact.ikw.uni-osnabrueck.de/magpie/api/submit_experiment/",
         // Possible deployment methods are:
         // "debug" and "directLink"

### 2021-07-25 analysis/xor-some-prolific-pilot3.Rmd
analyse pilot 3 with updated vignettes

--- 2021-05-04 analysis/xor-some-prolific-pilot2.Rmd
@@
 ---
-title: "XOR-Some Prolific Pilot 2"
+title: "XOR-Some Prolific Pilot 3"
 author: "Polina Tsvilodub"
-date: "4/2/2021"
+date: "7/24/2021"
 output: github_document
@@
 
-In this second pilot on Prolific we gather a more substantial amount of pilot data for the xor-some study on Prolific. The structure of the experiment was as follows:
+In this third pilot on Prolific we gather pilot data for the vignettes that were substantially updated based on the results from the [second pilot](https://github.com/magpie-ea/magpie-xor-experiment/blob/master/analysis/xor-some-prolific-pilot2.md).
+The used vignettes were selected based on manual comparison and comprise 38 out of 64 vignettes.
+
+The structure of the experiment was as follows:
 Participants read instructions, completed three example trials, and then completed 8 main blocks consisting of 4 xor and 4 some items.
 Each main block had the following structure: Participants read the background story, answered one comprehension question, then answered competence / relevance / prior questions in randomized order; then they read another 3 comprehension questions, after which the critical utterance was added below the background story.
-They answered the inference strength question, and then competence / relevance questions in randomized order again.
-
-N=118 participants were recruited for this pilot and compensated 2 pounds/participant.
+They answered the inference strength question.
+The second round of relevance/competence questions was removed for the pilot.
+
+N=50 participants were recruited for this pilot and compensated 2 pounds/participant.
 8 items (4 some + 4 xor) were sampled at random for each participant such that each participant saw one item in each condition (relevance X competence X prior = 8 unique conditions).
@@
 ```{r anonymization, include=FALSE}
-#d_raw <- read_csv("~/Documents/Research/XOR/raw_data/results_58_xor-some-Prolific-pilot2_N120.csv")
+#d_raw <- read_csv("~/Documents/Research/XOR/magpie-xor-experiment/data/pilots/results_63_xor-some-Prolific-pilot3_N50.csv")
 # remove IDs
-#d_raw %>% select(-prolific_id) %>% write_csv("../data/pilots/results_58_xor-some-Prolific-pilot2_N120.csv") 
+# d_raw %>% select(-prolific_id) %>% write_csv("../data/pilots/results_63_xor-some-Prolific-pilot3_N50.csv") 
 ```
@@
 ```{r data, results='hide', include=FALSE}
-d <- read_csv("../data/pilots/results_58_xor-some-Prolific-pilot2_N120.csv")
+d <- read_csv("../data/pilots/results_63_xor-some-Prolific-pilot3_N50.csv")
 ```
@@
 d %>% distinct(languages)
-cat("Number of partiipants before excluding non-natives: ", d %>% distinct(submission_id) %>% count() %>% pull() ) 
+cat("Number of participants before excluding non-natives: ", d %>% distinct(submission_id) %>% count() %>% pull() ) 
 # exclude non-natives if necessary
@@
 
-cat("Number of partiipants after excluding non-natives: ", d_native %>% distinct(submission_id) %>% count() %>% pull() )
-
-cat("Mean age: ", d_native %>% pull(age) %>% mean(., na.rm = T) )
-d_native %>% count(gender) %>% mutate(n = n/88)
+cat(" Number of participants after excluding non-natives: ", d_native %>% distinct(submission_id) %>% count() %>% pull() )
+
+cat(" Mean age: ", d_native %>% pull(age) %>% mean(., na.rm = T) )
+d_native  %>% count(gender) %>% mutate(n = n/72)
 ```
@@
 ```{r counts}
-# check xor/some vs. trial type
+# check xor/some vs. trial type; there are more xor critical conditions because there are 2 prior questions
 d_native %>% count(main_type, condition) 
+# check experimental conditions
+d_native %>% count(exp_condition)
 # check xor/some vs. experimental condition
@@
 
-Investigate very fast completion times:
+Investigate very fast completion times (we have none in this pilot):
 ```{r}
@@
 # check if their responses look somehow peculiar
-fast_completions %>%
-  ggplot(., aes(x = response)) +
-  geom_histogram()
+#fast_completions %>%
+#  ggplot(., aes(x = response)) +
+#  geom_histogram()
 
@@
   group_by(submission_id) %>%
-  mutate(passed_filler_trial = case_when(test_condition == "true" ~ response >= 70,
+  mutate(passed_filler_trial = case_when(test_condition == "true" ~ response >= 60,
                                    test_condition == "false" ~ response <= 40,
@@
 
-Check exclusions if applying exclusion ranges as suggested by Bob -- then almost twice as many participants are excluded based on their comprehension questions performance (I believe mainly due to a lower bound on correct responses in the uncertain questions).
-```{r}
-# true comprehension questions: 50-100
-# uncertain comprehension questions: 25-75
-# false comprehension questions: 0-50
-# get participants failing example trials
-
-d_exmpl2 <- d_exmpl %>% group_by(submission_id) %>% 
-  mutate(example_condition = ifelse(grepl("as certainly true", question), "true", 
-                                    ifelse(grepl("as certainly false", question), "false",
-                                           "uncertain")),
-         passed_example_trial = case_when(example_condition == "true" ~ response >= 50,
-                                    example_condition == "false" ~ response <= 50,
-                                    example_condition == "uncertain" ~ response %in% (25:75)),
-       #  check if all trials passed
-         passed_example = case_when(sum(passed_example_trial) == 0 ~ FALSE,
-                                    TRUE ~ TRUE)
-         ) %>% filter(passed_example == F)
-
-cat("Subjects failing the example trials: ", d_exmpl2 %>% distinct(submission_id) %>% pull() %>% length())
-
-# apply exclusion criteria to main trials
-# check range of responses per participant
-d_main_fail <- d_main %>% group_by(submission_id) %>%
-  mutate(passed_main = case_when(max(response) - min(response) <= 10 ~ FALSE,
-                                 TRUE ~ TRUE)
-         ) %>% filter(passed_main == F)
-cat(" Subjects providing the same ratings throughout the trials: ", d_main_fail %>% distinct(submission_id) %>% pull() %>% length())
-
-# get participants failing comprehension questions
-d_test2 <- d_test %>%
-  group_by(submission_id) %>%
-  mutate(passed_filler_trial = case_when(test_condition == "true" ~ response >= 50,
-                                   test_condition == "false" ~ response <= 50,
-                                   test_condition == "uncertain" ~ response %in% (25:75)),
-         mean_comprehension = mean(passed_filler_trial),
-         passed_filler = mean_comprehension >= 0.8
-         )
-
-d_test_fail2 <- d_test2 %>% 
-  filter(passed_filler == F)
-
-cat(" Subjects failing the comprehension trials: ", d_test_fail2 %>% distinct(submission_id) %>% pull() %>% length())
-
-```
-Now we explore applying exclusion criteria on z-scored data. Here it might make sense to exclude participants whose >=0.8 of responses are more than 2 SD from the mean (?). 
-```{r}
-# group by extended block, exmpl / filler / critical, maybe polarity of filler exmpl question
-# add block extension, put all data together somehow
-d_critical_fail_zScored <- d_main %>% 
-  mutate(block_extended = ifelse(
-    !(ifelse(is.na(critical_question), F, T)), 
-    block, 
-    ifelse(block %in% c("some", "xor"), "target", str_c(block, "_wUtt", ""))
-  )  
-  ) %>% 
-  filter(!(grepl("test", block_extended))) %>%
-  group_by(submission_id, block_extended) %>%
-  mutate(subj_block_mean = mean(response),
-         subj_block_sd = sd(response),
-         response_centered = (response - subj_block_mean)/subj_block_sd,
-         # catch the cases where sd is 0 
-         response_centered = ifelse(is.na(response_centered), 0, response_centered)) %>%
-  ungroup() %>%
-  # add by-block SDs, such that participants can be excluded based on being outside 2 SDs on a block
-  # or should this be based on by-subject SDs???? 
-  group_by(block_extended) %>%
-  mutate(block_mean = mean(response),
-         block_sd = sd(response),
-         passed_main_trial = case_when(abs(response_centered) <= 2*block_sd ~ T,
-                                       T ~ F),
-         mean_critical_passRate = mean(passed_main_trial))
-
-# get participants failing comprehension questions with z-scored data; here, by-subject SD are used for the exclusion
-d_test_fail_zScore <- d_test %>%
-  group_by(submission_id, test_condition) %>%
-  mutate(
-    subj_block_mean = mean(response),
-    subj_block_sd = sd(response),
-    response_centered = (response - subj_block_mean)/subj_block_sd,
-    response_centered = ifelse(is.na(response_centered), 0, response_centered),
-    passed_filler_trial = case_when(abs(response_centered) <= 2*subj_block_sd ~ T,
-                                   T ~ F),
-         mean_comprehension = mean(passed_filler_trial),
-         passed_filler = mean_comprehension >= 0.8
-         )
-
-d_test_fail_zScore <- d_test_fail_zScore %>% 
-  ungroup() %>%
-  filter(passed_filler == F)
-
-cat(" Subjects failing the comprehension trials when using z-scored raw data: ", d_test_fail_zScore %>% distinct(submission_id) %>% pull() %>% length())
-
-```
-
 ```{r clean}
-# get overall mean ratings / subject
+# get overall mean ratings / subject across comprehension and critical trials 
 d_full_clean %>% group_by(submission_id) %>% summarise(mean_rating = mean(response)) %>% arrange(mean_rating)
+
+# get mean ratings / subject in critical trials 
+d_full_clean %>% filter(condition == "critical") %>% group_by(submission_id) %>% summarise(mean_rating = mean(response)) %>% arrange(mean_rating)
+
 ```
@@
 
-Create more extensive condition labels, including information about whether a rating was produced with or without the utterance given.
+Create more extensive condition labels, including information about whether a rating was produced with or without the utterance given (no such condition in this pilot).
 
@@
 
-Compute the average deviation from expected responses for each vignette, individually for each dimension (rel / comp / pri):
+Compute the average deviation from expected responses for each vignette, individually for each dimension (rel / comp / pri).
+Only half as many items deviate from expectations by >= 50 points compared to the last pilot.
 ```{r}
@@
 
-#d_critical_deviations %>% write_csv("../data/pilots/pilot2_byItem_rating_deviations.csv")
+# d_critical_deviations %>% write_csv("../data/pilots/pilot3_byItem_rating_deviations.csv")
 
@@
 
-#d_critical_sd %>% write_csv("../data/pilots/pilot2_byItem_rating_SDs.csv")
+# d_critical_sd %>% write_csv("../data/pilots/pilot3_byItem_rating_SDs.csv")
+```
+
+Get the vignettes and dimensions that were strongly deviating from expectations in the previous pilot and still did not improve:
+```{r}
+deviating_item_pilot2 <- read_csv("../data/pilots/pilot2_byItem_rating_deviations.csv") %>% select(-mean_deviation)
+d_critical_deviations %>% select(-mean_deviation) %>% semi_join(., deviating_item_pilot2)
 ```
@@
 Visually check the spread of by-item ratings among subjects for each question:
-```{r, fig.height=20, fig.width=10}
+```{r, echo=FALSE, fig.height=25, fig.width=10}
+d_critical_long_grandMeans <- d_critical_long %>% 
+  mutate(title = paste(title, exp_condition, sep = "_"),
+         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
+  filter(block == class_condition | (block == "target" & class_condition == "relevance")) %>%
+  group_by(main_type, block) %>% summarize(mean_resp = mean(response))
+
+d_critical_long_expectedCats <- d_critical_long %>% 
+  mutate(title = paste(title, exp_condition, sep = "_"),
+         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
+  filter(block == class_condition | (block == "target" & class_condition == "relevance")) %>%
+  group_by(main_type, block, title) %>% summarize(mean_resp_block = mean(response))
 
@@
          block = ifelse(block == "xor" | block == "some", "target", block)) %>%
-  filter(block == class_condition | block == "target") %>%
+  left_join(., d_critical_long_grandMeans, by = c("main_type", "block")) %>% 
+  filter(block == class_condition | ((block == "target") & (class_condition == "relevance"))) -> d_long_plot_data
+
+left_join(d_long_plot_data, d_critical_long_expectedCats, by = c("main_type", "block", "title")) %>% 
+  mutate(expected_block = ifelse(block != "target", prior_class * 100, NA),
+         block = factor(block, levels = c("relevance", "competence", "prior", "target"))) -> d_long_plot_full
+d_long_plot_full %>%
   ggplot(., aes(x = block, y = response, color = block)) +
   geom_point(alpha = 0.7, position = position_jitter(width = 0.1)) +
+  geom_hline(aes(yintercept = mean_resp), alpha = 0.7) +
+  geom_point(aes(y = mean_resp), color = "black", shape = 2) +
+  geom_point(aes(y = mean_resp_block, fill = block), color = "black") +
+  geom_point(aes(y = expected_block), color="red", size = 2.5, na.rm = T) +
   ylab("Raw responses to respective questions") +
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  ggtitle("By-item by-question raw ratings")
-```
-For the following items, participants seem  to *disagree* in their judgments for all of the predictors: Emails from a broken laptop, Gift unwrapping, Martha's cookies, Suzie's fruits, Brad's clothes, Crime, maybe more..
+  ggtitle("By-item by-question raw ratings\nBlack dots indicate by-block means within-item, lines and triangles indicate grand by-block means\nRed dots indicate expected rating")
+```
+
+Visually compare the updated vignettes (pilot 2 vs current pilot)
+```{r, echo=FALSE}
+# add column indicating which pilot it is
+d_long_plot_full <- d_long_plot_full %>%
+  mutate(pilot = "updated")
+
+d_long_plot_pilot2 <- read_csv("../data/pilots/pilot2_byItem_plotting.csv") %>% 
+  mutate(pilot = "old")
+# only keep data for those items that were also tested in the current pilot
+d_long_plot_pilot2 <- semi_join(d_long_plot_pilot2, d_long_plot_full, by = c("title"))
+
+d_long_plot_both_pilots <- bind_rows(d_long_plot_full, d_long_plot_pilot2)
+```
+
+```{r, echo=FALSE, fig.width=8, fig.height=15}
+# plot old and updated some items side-by-side
+# the vignettes for which only an "updated" plot occurs are new items that were added instead of some other vignettes 
+d_long_plot_both_pilots %>%
+  filter(main_type == "some") %>%
+  ggplot(., aes(x = block, y = response, color = block)) +
+  geom_point(alpha = 0.7, position = position_jitter(width = 0.1)) +
+  geom_hline(aes(yintercept = mean_resp), alpha = 0.7) +
+  geom_point(aes(y = mean_resp), color = "black", shape = 2) +
+  geom_point(aes(y = mean_resp_block, fill = block), color = "black") +
+  geom_point(aes(y = expected_block), color="red", size = 2.5, na.rm = T) +
+  ylab("Raw responses to respective questions") +
+  facet_wrap(title~pilot, ncol = 4) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  ggtitle("By-item by-question raw ratings, old vs.
+updated SOME vignettes\nBlack dots indicate by-block means within-item, lines and triangles indicate grand by-block means\nRed dots indicate expected rating")
+```
+
+```{r, echo=FALSE, fig.width=8, fig.height=15}
+# plot old and updated xor items side-by-side 
+d_long_plot_both_pilots %>%
+  filter(main_type == "xor") %>%
+  ggplot(., aes(x = block, y = response, color = block)) +
+  geom_point(alpha = 0.7, position = position_jitter(width = 0.1)) +
+  geom_hline(aes(yintercept = mean_resp), alpha = 0.7) +
+  geom_point(aes(y = mean_resp), color = "black", shape = 2) +
+  geom_point(aes(y = mean_resp_block, fill = block), color = "black") +
+  geom_point(aes(y = expected_block), color="red", size = 2.5, na.rm = T) +
+  ylab("Raw responses to respective questions") +
+  facet_wrap(title~pilot, ncol = 4) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  ggtitle("By-item by-question raw ratings, old vs.
+updated XOR vignettes\nBlack dots indicate by-block means within-item, lines and triangles indicate grand by-block means\nRed dots indicate expected rating")
+```
 
@@
 ## Plots
-Plot main rel / comp / prior questions by main condition and by prior classification of the item (x-axis), separated into with / without critical utterance (shape, color).
+```{r, echo=FALSE, eval=FALSE}
+# Plot main rel / comp / prior questions by main condition and by prior classification of the item (x-axis), separated into with / without critical utterance (shape, color).
 The prior questions were only used without the utterance.
-```{r}
-
 d_critical_long %>% 
@@
   filter(block == class_condition) %>%
-  ggplot(., aes(x = as.factor(prior_class), y = response, shape = w_utterance, color = w_utterance)) +
+  ggplot(., aes(x = as.factor(prior_class), y = response)) +
   geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
-  geom_point(data = d_critical_summary, aes(x = as.factor(prior_class), y = mean_response, shape = w_utterance), 
+  geom_point(data = d_critical_summary, aes(x = as.factor(prior_class), y = mean_response), 
              color = "red", size = 3) +
@@
 ```
+First, compare the two prior ratings in the xor condition:
 ```{r, fig.height=3}
@@
   filter(block == "prior", main_type == "xor") %>%
-  mutate(priorQ_nr = rep(c(1,2), 396),
+  mutate(priorQ_nr = rep(c(1,2), nrow(.)/2),
          unique_ID = paste(submission_id, ID, sep="_")) -> d_xor_priors
@@
   ggplot(., aes(x = as.factor(priorQ_nr), y = response )) +
-  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
+  geom_point(size = 2, alpha = 0.6) +
   geom_point(data = d_xor_priors_summary, aes(x = as.factor(priorQ_nr), y = mean), color = "red", size = 3) +
@@
 
-Plot inference ratings as a function of *anticipated* rating of the explanatory factor, similar to the paper: 
-``` {r}
-d_critical_long %>% 
-  filter(block == "xor" | block == "some") %>%
-  ggplot(., aes(x = prior_class, y = response)) +
-  geom_point(size = 2, alpha = 0.5) +
-  geom_smooth(method="lm") +
-  ylab("Inference strength ratings") +
-  facet_wrap(block~class_condition) +
-  xlab("Anticipated categorization of the items") +
-  ggtitle("Inference strength ratings by-predictor")
-
-```
-
-```{r}
+```{r}
+# check analytically how many subjects rate prior questions differently
+d_xor_priors %>% select(-trial_number, -RT, -prompt) %>% 
+  pivot_wider(names_from = "priorQ_nr", values_from = "response") %>%
+  mutate(response_difference = abs(`1` - `2`),
+         bigDifference = ifelse(response_difference > 15, T, F)) -> d_xor_priors_differences
+
+# 19 responses with difference > 15 between first and second rating
+d_xor_priors_differences %>% count(bigDifference)
+
+# proportion of big differences in prior responses by prior type (high vs low)
+d_xor_priors_differences %>% group_by(prior_class) %>% summarise(prop_big_diffs = mean(bigDifference))
+```
+
+```{r, echo=FALSE, eval=FALSE}
 # make a versatile wide representation of the critical data
@@
 ```
-
-Plot inference strength ratings against raw predictor ratings from each participant across items:
-```{r}
+```{r, echo=FALSE, eval=FALSE}
+# Plot inference strength ratings against raw predictor ratings from each participant across items:
+
 d_critical_wide %>%
@@
 ```
-
-Plot mean ratings (across all participants) for each vignette (with its respective condition indicated in the facet title, in the order relevance/competence/prior), in each condition.
-For relevance and competence, the color indicates whether it was presented without the utterance or with.
-```{r, fig.width=10, fig.height=30}
+```{r, fig.width=10, fig.height=30, echo=FALSE, eval=FALSE}
+# Plot mean ratings (across all participants) for each vignette (with its respective condition indicated in the facet title, in the order relevance/competence/prior), in each condition.
 bar.width = 0.8
@@
   summarize(mean_rating = mean(response)) %>% 
-  ggplot(., aes(x = block_extended, y = mean_rating, fill = w_utterance)) +
+  ggplot(., aes(x = block_extended, y = mean_rating)) +
   geom_col(alpha = 0.7, width = bar.width, position = position_dodge(width = bar.width)) +
@@
 ```
-
+``` {r, echo=FALSE, eval=FALSE}
+
+# correlation plot for "some"
+GGally::ggpairs(
+  filter(d_critical_wide, main_type == "some") %>%  
+    select(prior, competence, relevance, target) 
+)
+
+# correlation plot for "or"
+GGally::ggpairs(
+  filter(d_critical_wide, main_type == "xor") %>%  
+    select(prior, competence, relevance, target)
+  )
+```
+
+As planned, all the primary visualizations and analyses below are conducted on data z-scored by-subject by-block.
+
+Plot main rel / comp / prior questions by main condition and by prior classification of the item (x-axis): 
+```{r}
+d_critical_zScored <- d_critical_long %>% group_by(submission_id, block_extended) %>%
+  mutate(block_mean = mean(response),
+         block_sd = sd(response),
+         response_centered = (response - block_mean)/block_sd,
+         # catch the cases where sd is 0 
+         response_centered = ifelse(is.na(response_centered), 0, response_centered))
+
+d_critical_zScored %>% 
+  filter(block != "xor" & block != "some") %>%
+  filter(block == class_condition) %>%
+  group_by(main_type, class_condition, w_utterance, prior_class) %>% 
+  summarize(mean_response = mean(response_centered)) -> d_critical_zScore_summary
+
+d_critical_zScored %>% 
+  filter(block != "xor" & block != "some") %>%
+  filter(block == class_condition) %>%
+  ggplot(., aes(x = as.factor(prior_class), y = response_centered, shape = w_utterance)) +
+  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
+  geom_point(data = d_critical_zScore_summary, aes(x = as.factor(prior_class), y = mean_response), 
+             color = "red", size = 3) +
+  ylab("Responses to respective predictor questions") +
+  xlab("Anticipated categorization of the items (low / high)") +
+  facet_wrap(main_type~class_condition) + # get ratings from the respective trials only 
+  ggtitle("Centered responses to by-predictor questions")
+```
+Plot inference ratings as a function of *anticipated* rating of the explanatory factor, similar to the paper: 
+``` {r}
+d_critical_zScored %>% 
+  filter(block == "xor" | block == "some") %>%
+  ggplot(., aes(x = prior_class, y = response_centered)) +
+  geom_point(size = 2, alpha = 0.5) +
+  geom_smooth(method="lm") +
+  ylab("Inference strength ratings") +
+  facet_wrap(block~class_condition) +
+  xlab("Anticipated categorization of the items") +
+  ggtitle("Inference strength ratings by-predictor")
+
+```
+
+Plot the empirical predictor ratings against target inference ratings:
+```{r}
+d_critical_zScored_wide <- d_critical_zScored %>% 
+  select(submission_id, title, main_type, block_extended, response_centered) %>% 
+  unique() %>% 
+  pivot_wider(
+    names_from = block_extended, 
+    values_from = response_centered, 
+    values_fn = mean # getting means for double prior measurement in "xor"
+  )
+
+d_critical_zScored_wide %>%
+  ggplot(., aes(x = relevance, y = target)) +
+  geom_point(size = 2, alpha = 0.7) +
+  ylab("Inference strength ratings") +
+  geom_smooth(method = "lm") +
+  facet_wrap(~main_type, ncol = 1) -> p.rel.z
+
+d_critical_zScored_wide %>%
+  ggplot(., aes(x = competence, y = target)) +
+  geom_point(size = 2, alpha = 0.7) +
+  geom_smooth(method = "lm") +
+  ylab("") +
+  facet_wrap(~main_type, ncol = 1) -> p.comp.z
+
+d_critical_zScored_wide %>%
+  ggplot(., aes(x = prior, y = target)) +
+  geom_point(size = 2, alpha = 0.7) +
+  geom_smooth(method = "lm") +
+  ylab("") +
+  facet_wrap(~main_type, ncol = 1) -> p.pri.z
+
+gridExtra::grid.arrange(p.rel.z, p.comp.z, p.pri.z, ncol = 3) 
+```
 Plot by-item trends between predictor ratings and target inference strength ratings (relevance / competence ratings with utterance are omitted):
@@
 # make a versatile wide representation of the critical data
-d_critical_predictorsXtarget <- d_critical_wide %>% 
+d_critical_predictorsXtarget <- d_critical_zScored_wide %>% 
   pivot_longer(
-    cols = c(relevance, prior, competence, relevance_wUtt, competence_wUtt),
+    cols = c(relevance, prior, competence),
     names_to = "predictors",
     values_to = "predictor_response"
-  ) %>%
-  filter(predictors != "relevance_wUtt", predictors != "competence_wUtt")
-
+  ) 
 d_critical_predictorsXtarget %>%
@@
 
-Plot inference strength ratings against predictor ratings, by-item, using by-item mean ratings across participants. Error bars indicate the standard deviation of the mean:
+Plot inference strength ratings against predictor ratings, by-item, using by-item mean ratings across participants.
 ```{r, fig.height=10, fig.width=8}
@@
   geom_point() +
-  geom_errorbar(aes(ymin = mean_target - sd_target, ymax = mean_target + sd_target), alpha = 0.7) +
-  geom_errorbarh(aes(xmin = mean_predictor - sd_predictor, xmax = mean_predictor + sd_predictor), alpha = 0.7) +
-  geom_text(aes(label=title), size = 2.5, nudge_y = 4) +
+  geom_text(aes(label=title), size = 2.5) +
   geom_smooth(method = "lm") +
@@
 
-``` {r}
-
-# correlation plot for "some"
-GGally::ggpairs(
-  filter(d_critical_wide, main_type == "some") %>%  
-    select(prior, competence, relevance, target, competence_wUtt, relevance_wUtt) 
-)
-
-# correlation plot for "or"
-GGally::ggpairs(
-  filter(d_critical_wide, main_type == "xor") %>%  
-    select(prior, competence, relevance, target, competence_wUtt, relevance_wUtt)
-  )
-```
-
-## Explore z-scoring
-Visual exploration of z-scoring each measure (rel / comp / pri / target / comp-wUtterance / rel-wUtterance) per participant, collapsing some and or together. 
-```{r}
-d_critical_zScored <- d_critical_long %>% group_by(submission_id, block_extended) %>%
-  mutate(block_mean = mean(response),
-         block_sd = sd(response),
-         response_centered = (response - block_mean)/block_sd,
-         # catch the cases where sd is 0 
-         response_centered = ifelse(is.na(response_centered), 0, response_centered))
-
-d_critical_zScored %>% 
-  filter(block != "xor" & block != "some") %>%
-  filter(block == class_condition) %>%
-  group_by(main_type, class_condition, w_utterance, prior_class) %>% 
-  summarize(mean_response = mean(response_centered)) -> d_critical_zScore_summary
-
-d_critical_zScored %>% 
-  filter(block != "xor" & block != "some") %>%
-  filter(block == class_condition) %>%
-  ggplot(., aes(x = as.factor(prior_class), y = response_centered, shape = w_utterance, color = w_utterance)) +
-  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
-  geom_point(data = d_critical_zScore_summary, aes(x = as.factor(prior_class), y = mean_response, shape = w_utterance), 
-             color = "red", size = 3) +
-  ylab("Responses to respective predictor questions") +
-  xlab("Anticipated categorization of the items (low / high)") +
-  facet_wrap(main_type~class_condition) + # get ratings from the respective trials only 
-  ggtitle("Centered responses to by-predictor questions")
-```
-
-Plot z-scored predictor ratings against z-scored target inference ratings
-```{r}
-d_critical_zScored_wide <- d_critical_zScored %>% 
-  select(submission_id, title, main_type, block_extended, response_centered) %>% 
-  unique() %>% 
-  pivot_wider(
-    names_from = block_extended, 
-    values_from = response_centered, 
-    values_fn = mean # getting means for double prior measurement in "xor"
-  )
-
-d_critical_zScored_wide %>%
-  ggplot(., aes(x = relevance, y = target)) +
-  geom_point(size = 2, alpha = 0.7) +
-  ylab("Inference strength ratings") +
-  geom_smooth(method = "lm") +
-  facet_wrap(~main_type, ncol = 1) -> p.rel.z
-
-d_critical_zScored_wide %>%
-  ggplot(., aes(x = competence, y = target)) +
-  geom_point(size = 2, alpha = 0.7) +
-  geom_smooth(method = "lm") +
-  ylab("") +
-  facet_wrap(~main_type, ncol = 1) -> p.comp.z
-
-d_critical_zScored_wide %>%
-  ggplot(., aes(x = prior, y = target)) +
-  geom_point(size = 2, alpha = 0.7) +
-  geom_smooth(method = "lm") +
-  ylab("") +
-  facet_wrap(~main_type, ncol = 1) -> p.pri.z
-
-gridExtra::grid.arrange(p.rel.z, p.comp.z, p.pri.z, ncol = 3) 
-```
-
 Correlation plots of z-scored data:
@@
   filter(d_critical_zScored_wide, main_type == "some") %>% ungroup() %>%  
-    select(prior, competence, relevance, target, competence_wUtt, relevance_wUtt)
+    select(prior, competence, relevance, target)
 )
@@
   filter(d_critical_zScored_wide, main_type == "xor") %>%  ungroup() %>%
-    select(prior, competence, relevance, target, competence_wUtt, relevance_wUtt)
+    select(prior, competence, relevance, target)
   )
@@
 
-Maximal models on raw ratings:
-```{r, results='hide'}
-# xor, maximal model with interactions and maximal REs
-model_xor <- brm(
-  bf(target ~ prior*competence*relevance + 
-    (1 + prior + competence + relevance || submission_id) +
-    (1 | title),
-    decomp = "QR"
-  ), 
-  data = d_critical_wide %>% filter(main_type == "xor"),
-  family = "gaussian",
-  cores = 4,
-  iter = 3000
-)
-```
-
-```{r}
-summary(model_xor)
-```
-
-Exploring plots of posterior samples for pairwise parameters shows that there is a visible correlation between prior & competence, and a somewhat weaker correlation between competence & relevance.
-The interaction estimates also show a correlation with the main effects (as a rough overall judgment).
-```{r}
-#shinystan::launch_shinystan(model_xor)
-```
-Plot correlations of main effects
-```{r}
-bayesplot::mcmc_pairs(model_xor, pars = c("b_prior", "b_competence", "b_relevance"))
-```
-
-```{r, results='hide'}
-# some, maximal model with interactions and maximal REs
-model_some <- brm(
-  bf(target ~ prior*competence*relevance + 
-    (1 + prior + competence + relevance || submission_id) +
-    (1 | title),
-    decomp = "QR"),
-  data = d_critical_wide %>% filter(main_type == "some"),
-  family = "gaussian",
-  cores = 4,
-  iter = 3000
-)
-```
-
-```{r}
-summary(model_some)
-```
-
-For "some", the correlation of prior and relevance is higher than for prior and competence; there is also a visible correlation of competence and relevance. 
-```{r}
-#shinystan::launch_shinystan(model_some)
-```
-```{r}
-bayesplot::mcmc_pairs(model_some, pars = c("b_prior", "b_competence", "b_relevance"))
-```
-
-#### Smaller models with raw data to check for collinearity 
-Models on raw data with only one predictor: 
-```{r, results='hide'}
-# xor
-model_xor_prior <- brm(
-  bf(target ~ prior + 
-    (1 + prior | submission_id) +
-    (1 | title),
-    decomp = "QR"),
-  data = d_critical_wide %>% filter(main_type == "xor"),
-  family = "gaussian",
-  control = list(adapt_delta = 0.99),
-  cores = 4,
-  iter = 3000
-)
-```
-
-```{r}
-summary(model_xor_prior)
-```
-
-```{r, results='hide'}
-# xor
-model_xor_comp <- brm(
-  bf(target ~ competence + 
-    (1 + competence | submission_id) +
-    (1 | title),
-    decomp = "QR"),
-  data = d_critical_wide %>% filter(main_type == "xor"),
-  family = "gaussian",
-  control = list(adapt_delta = 0.99),
-  cores = 4,
-  iter = 3000
-)
-```
-
-```{r}
-summary(model_xor_comp)
-```
-
-```{r, results='hide'}
-# xor
-model_xor_rel <- brm(
-  bf(target ~ relevance + 
-    (1 + relevance | submission_id) +
-    (1 | title),
-    decomp = "QR"),
-  data = d_critical_wide %>% filter(main_type == "xor"),
-  family = "gaussian",
-  control = list(adapt_delta = 0.99),
-  cores = 4,
-  iter = 3000
-)
-```
-
-```{r}
-summary(model_xor_rel)
-```
-
-```{r, results='hide'}
-# some
-model_some_rel <- brm(
-  bf(target ~ relevance + 
-    (1 + relevance | submission_id) +
-    (1 | title),
-    decomp = "QR"),
-  data = d_critical_wide %>% filter(main_type == "some"),
-  family = "gaussian",
-  control = list(adapt_delta = 0.99),
-  cores = 4,
-  iter = 3000
-)
-```
-
-```{r}
-summary(model_some_rel)
-```
-
-```{r, results='hide'}
-# some
-model_some_prior <- brm(
-  bf(target ~ prior + 
-    (1 + prior | submission_id) +
-    (1 | title),
-    decomp = "QR"),
-  data = d_critical_wide %>% filter(main_type == "some"),
-  family = "gaussian",
-  control = list(adapt_delta = 0.99),
-  cores = 4,
-  iter = 3000
-)
-```
-
-```{r}
-summary(model_some_prior)
-```
-
-```{r, results='hide'}
-# some
-model_some_comp <- brm(
-  bf(target ~ competence + 
-    (1 + competence | submission_id) +
-    (1 | title),
-    decomp = "QR"),
-  data = d_critical_wide %>% filter(main_type == "some"),
-  family = "gaussian",
-  control = list(adapt_delta = 0.92),
-  cores = 4,
-  iter = 3000
-)
-```
-
-```{r}
-summary(model_some_comp)
-```
-
-With two predictors:
-```{r, results='hide'}
-# xor
-model_xor_pri_comp <- brm(
-  bf(
-  target ~ prior*competence + 
-    (1 + prior*competence || submission_id) +
-    (1 | title),
-  decomp = "QR"
-  ),
-  data = d_critical_wide %>% filter(main_type == "xor"),
-  family = "gaussian",
-  cores = 4,
-  iter = 3000
-)
-```
-
-```{r}
-summary(model_xor_pri_comp)
-```
-
-```{r}
-bayesplot::mcmc_pairs(model_xor_pri_comp, pars = c("b_prior", "b_competence"))
-```
-
-```{r, results='hide'}
-# xor
-model_xor_pri_rel <- brm(
-  bf(target ~ prior*relevance + 
-    (1 + prior*relevance || submission_id) +
-    (1 | title),
-    decomp = "QR"),
-  data = d_critical_wide %>% filter(main_type == "xor"),
-  family = "gaussian",
-  control = list(adapt_delta = 0.92),
-  cores = 4,
-  iter = 3000
-)
-```
-
-```{r}
-summary(model_xor_pri_rel)
-```
-
-```{r}
-bayesplot::mcmc_pairs(model_xor_pri_rel, pars = c("b_prior", "b_relevance"))
-```
-
-```{r, results='hide'}
-# xor
-model_xor_rel_comp <- brm(
-  bf(target ~ relevance*competence + 
-    (1 + relevance*competence || submission_id) +
-    (1 | title),
-    decomp = "QR"),
-  data = d_critical_wide %>% filter(main_type == "xor"),
-  family = "gaussian",
-  cores = 4,
-  iter = 3000
-)
-```
-
-```{r}
-summary(model_xor_rel_comp)
-```
-
-```{r}
-bayesplot::mcmc_pairs(model_xor_rel_comp, pars = c("b_competence", "b_relevance"))
-```
-
-```{r, results='hide'}
-# some
-model_some_pri_comp <- brm(
-  bf(target ~ prior*competence + 
-    (1 + prior*competence || submission_id) +
-    (1 | title),
-    decomp = "QR"),
-  data = d_critical_wide %>% filter(main_type == "some"),
-  family = "gaussian",
-  cores = 4,
-  iter = 3000
-)
-```
-
-```{r}
-summary(model_some_pri_comp)
-```
-
-```{r}
-bayesplot::mcmc_pairs(model_some_pri_comp, pars = c("b_prior", "b_competence"))
-```
-
-```{r, results='hide'}
-# some
-model_some_pri_rel <- brm(
-  bf(target ~ prior*relevance + 
-    (1 + prior*relevance || submission_id) +
-    (1 | title),
-    decomp = "QR"),
-  data = d_critical_wide %>% filter(main_type == "some"),
-  family = "gaussian",
-  cores = 4,
-  iter = 3000
-)
-```
-
-```{r}
-summary(model_some_pri_rel)
-```
-
-```{r}
-bayesplot::mcmc_pairs(model_some_pri_rel, pars = c("b_prior", "b_relevance"))
-```
-
-```{r, results='hide'}
-# some
-model_some_rel_comp <- brm(
-  bf(target ~ relevance*competence + 
-    (1 + relevance*competence || submission_id) +
-    (1 | title),
-    decomp = "QR"),
-  data = d_critical_wide %>% filter(main_type == "some"),
-  family = "gaussian",
-  cores = 4,
-  iter = 3000
-)
-```
-
-```{r}
-summary(model_some_rel_comp)
-```
-
-```{r}
-bayesplot::mcmc_pairs(model_some_rel_comp, pars = c("b_competence", "b_relevance"))
-```
-
-#### Stats on z-scored data
-The same stats (full, one predictor, to predictors) on z-scored data:
+Our primary analysis is also conducted on z-scored data. 
+First, maximal models on the ratings provided by the subjects are run:
 ```{r, results='hide'}
@@
 
+Plot correlations of main effects
 ```{r}
@@
 
-#### Smaller models with z-scored data
+#### Smaller models with z-scored data to check for collinearity
+
+Models with only one predictor: 
 ```{r, results='hide'}
@@
 #### Stats with dichotomized predictors
-Maximal model on raw inference ratings, using anticipated binary categorization of items as predictors 
-```{r, results='hide'}
-d_critical_long %>% 
-  select(submission_id, title, main_type, block_extended, response, prior_class, class_condition) %>% 
-  unique() %>% 
-  pivot_wider(
-    names_from = class_condition, 
-    values_from = prior_class
-  )  %>%
-  mutate(relevance = factor(relevance, levels = c(1, 0)),
-         competence = factor(competence, levels = c(1, 0)),
-         prior = factor(prior, levels = c(1, 0))) %>% 
-  filter(block_extended == "target") -> d_critical_wide_cat
-# sum code predictors
-# low prior: -1, high prior:: 1 
-contrasts(d_critical_wide_cat$prior) <- contr.sum(2)
-# low comp : -1, high comp : 1
-contrasts(d_critical_wide_cat$competence) <- contr.sum(2)
-# low rel: -1, high rel: 1
-contrasts(d_critical_wide_cat$relevance) <- contr.sum(2)
-
-# xor, maximal model with interactions and maximal REs
-model_xor_cat <- brm(
-  bf(response ~ prior*competence*relevance + 
-    (1 + prior + competence + relevance || submission_id) +
-    (1 | title),
-    decomp = "QR"),
-  data = d_critical_wide_cat %>% filter(main_type == "xor"),
-  family = "gaussian",
-  cores = 4,
-  iter = 3000
-)
-```
-
-```{r}
-summary(model_xor_cat)
-```
-We test the following hypotheses for the categorical predictors models (identical for both or and some):
-H1: For higher prior of the stronger alternative being true, the inference strength is lower
-H2: For higher speaker comptence, the inference strength is higher
-H3: For higher listener relevance, the inference strength is higher 
-```{r}
-# xor
-# test H1
-hypothesis(model_xor_cat, "prior1 < 0")
-# test H2
-hypothesis(model_xor_cat, "competence1 > 0")
-# test H3
-hypothesis(model_xor_cat, "relevance1 > 0")
-```
-
-```{r, results='hide'}
-model_some_cat <- brm(
-  bf(response ~ prior*competence*relevance + 
-    (1 + prior + competence + relevance || submission_id) +
-    (1 | title),
-    decomp = "QR"),
-  data = d_critical_wide_cat %>% filter(main_type == "some"),
-  family = "gaussian",
-  cores = 4,
-  iter = 3000
-)
-```
-
-```{r}
-summary(model_some_cat)
-```
-
-```{r}
-# some
-# test H1
-hypothesis(model_some_cat, "prior1 < 0")
-# test H2
-hypothesis(model_some_cat, "competence1 > 0")
-# test H3
-hypothesis(model_some_cat, "relevance1 > 0")
-```
 
@@
 
+We test the following hypotheses for the categorical predictors models (identical for both or and some):
+
+H1: For higher prior of the stronger alternative being true, the inference strength is lower
+
+H2: For higher speaker comptence, the inference strength is higher
+
+H3: For higher listener relevance, the inference strength is higher
+
 ```{r}
@@
 ```
+
+#### Tertiary analysis with by-item means (from paper)
+Now we explore the original analysis with the by-item means:
+```{r, results='hide'}
+d_critical_zScored_wide_byItem_xor <- d_critical_zScored_wide %>% ungroup() %>%
+  filter(main_type == "xor") %>%
+  group_by(title) %>%
+  summarize(mean_response = mean(target),
+            mean_prior = mean(prior),
+            mean_relevance = mean(relevance),
+            mean_competence = mean(competence))
+
+d_critical_zScored_wide_byItem_some <- d_critical_zScored_wide %>% ungroup() %>%
+  filter(main_type == "some") %>%
+  group_by(title) %>%
+  summarize(mean_response = mean(target),
+            mean_prior = mean(prior),
+            mean_relevance = mean(relevance),
+            mean_competence = mean(competence))
+
+model_xor_means <- brm(bf(mean_response ~ mean_prior + mean_relevance + mean_competence,
+                          decomp="QR"),
+                       data = d_critical_zScored_wide_byItem_xor,
+                       family = "gaussian",
+                       cores = 4,
+                       iter = 3000
+                       )
+
+model_some_means <- brm(bf(mean_response ~ mean_prior + mean_relevance + mean_competence,
+                          decomp="QR"),
+                       data = d_critical_zScored_wide_byItem_some,
+                       family = "gaussian",
+                       cores = 4,
+                       iter = 3000
+                       )
+```
+
+```{r}
+summary(model_xor_means)
+```
+```{r}
+summary(model_some_means)
+```

### 2021-08-19 analysis/xor-some-prolific-pilot2.Rmd
add final items
first stab at simr based power analysis

@@
          block = ifelse(block == "xor" | block == "some", "target", block)) %>%
-  left_join(., d_critical_long_grandMeans, by = c("main_type", "block")) %>% filter(block == class_condition | block == "target") -> d_long_place_holder
+  left_join(., d_critical_long_grandMeans, by = c("main_type", "block")) %>% 
+  filter(block == class_condition | ((block == "target") & (class_condition == "relevance"))) -> d_long_place_holder
 
@@
   mutate(expected_block = ifelse(block != "target", prior_class * 100, NA),
-         block = factor(block, levels = c("relevance", "competence", "prior"))) %>%
+         block = factor(block, levels = c("relevance", "competence", "prior", "target"))) -> d_long_plot_full
+
+# save preprocessed df for plotting for comparison with third pilot
+# d_long_plot_full %>% write_csv("../data/pilots/pilot2_byItem_plotting.csv")
+
+d_long_plot_full %>%
   ggplot(., aes(x = block, y = response, color = block)) +
@@
   geom_point(aes(y = mean_resp_block, fill = block), color = "black") +
-  geom_point(aes(y = expected_block), color="red", size = 2.5) +
+  geom_point(aes(y = expected_block), color="red", size = 2.5, na.rm = T) +
   ylab("Raw responses to respective questions") +
@@
   )
+
+#d_critical_zScored_wide %>%
+#  write_csv("../data/pilots/pilot2_critical_zScore_wide_tidy.csv")
 
@@
 ```
+
+Now we explore the original analysis with the by-item means:

### 2021-08-19 analysis/xor-some_power.Rmd
first stab at simr based power analysis

---
title: "XOR power analysis"
author: "Polina Tsvilodub"
date: "8/17/2021"
output: github_document
---

In this document, I conduct a simulation-based power analysis for the XOR study, based on [pilot 2](https://github.com/magpie-ea/magpie-xor-experiment/blob/master/analysis/xor-some-prolific-pilot2.md) (N=118) and [pilot 3](https://github.com/magpie-ea/magpie-xor-experiment/blob/master/analysis/xor-some-prolific-pilot3.md) (N = 50). That is, the desired main analysis is conducted on data simulated based on a model fit to this combined pilot data, as if the data came from different numbers of subjects. The significance statistics of interest are computed for each simulated number of subjects, by repeatedly simulating the target analysis, yielding the power analysis.

Due to computational tractability reasons, the power analysis is conducted using the package `simr` which allows for simulation-based power analysis of frequentist GLMMs. The main resource for this script is [this](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12504) tutorial.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
#library(brms)
#library(tidybayes)
library(broom)
library(simr)
```
## Read Data

First, the preprocessed and tidy wide pilot data is read and combined. It comprises data from N = 140 subjects after exclusions in total.
```{r, echo=FALSE}
pilot1 <- read_csv("./../data/pilots/pilot2_critical_zScore_wide_tidy.csv") %>%
  select(-competence_wUtt, -relevance_wUtt)
pilot2 <- read_csv("./../data/pilots/pilot3_critical_zScore_wide_tidy.csv")

d_both <- pilot1 %>% mutate(pilot = 1) %>% 
  rbind(., pilot2 %>% mutate(pilot = 2)
        )
# the data read here should be tidy, in correct format, bad workers should be excluded already
# z-scored data should be used and simulated!

# separate some and xor data!
d_both_some <- d_both %>% filter(main_type == "some") 
d_both_xor <- d_both %>% filter(main_type == "xor")
```
Sanity check the combined data (as can be seen, z-scored data is used):
```{r, echo=FALSE}
d_both %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("Inference strength ratings") +
  facet_wrap(~main_type, ncol = 1) -> p.rel

d_both %>%
  ggplot(., aes(x = competence, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.comp

d_both %>%
  ggplot(., aes(x = prior, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.pri

gridExtra::grid.arrange(p.rel, p.comp, p.pri, ncol = 3) 
```

## Fit sample model
As the first step, the desired model is fit on data from both pilots (frequentist). Maximal models of interest for xor, then for some, were attempted, but by-subject random effects result in singular matrix errors. Furthermore, simulations on different Ns below can only be conducted if there are by-subject random intercepts only. 
Essentially, two separate power analyses are conducted (one for "some", one for "or").

```{r, warning=FALSE, message=FALSE}
# fit frequentist model
# remove some random effects due to singular matrix error
model_xor <- lmer(target ~ prior*competence*relevance + 
    (1 | submission_id) +
    (1 | title),
    data = d_both_xor
    )

isSingular(model_xor)
summary(model_xor)
```

```{r, warning=FALSE, message=FALSE}
model_some <- lmer(target ~ prior*competence*relevance + 
    (1 | submission_id) +
    (1 | title),
    data = d_both_some
    )

isSingular(model_some)
summary(model_some)
```

## Simulate

The `simr` package provides several handy functions for power analysis.

First, there are functions for simulating data extended along different axes, necessary for simulating different numbers of subjects in order to determine the minimal N required for a certain power.

Second, there are functions for repeatedly computing the target model for a given dataset, and extracting significance statistics of interest from each fit. This is necessary for computing the power (as the proportion of significant results among the simulations).

However, the significance statistics are only computed for one given fixed effect estimate, so a wrapper is written to iterate over all *simple fixed effects* which would be the estimates of interest, if the SI account was borne out.
Intermediate results of the simulations are written out as csvs for backup.

Below, numbers of subjects ranging between N = 50 and N = 200 are simulated in steps of 10. There are 4 simulated observations per subject (because we have 8 vignettes / subject in the experiment, half for some and half for xor).

```{r, results='hide', warning=FALSE, message=FALSE}
# check power of the pilot source models for a given effect, as an example
p_xor <- powerSim(model_xor, 
                  fixed("prior", "z"), 
                  progress = FALSE)
p_some <- powerSim(model_some, 
                   fixed("prior", "z"), 
                   progress = FALSE)
```

```{r}
p_xor
p_some
```

```{r, results='hide', warning=FALSE, message=FALSE}
N_subj = c(seq(from = 50, to = 200, by = 10))

# following step necessary for iterating through Ns, n parameter has to be the maximal desired N
# fit extended xor model
model_xor_ext <- extend(model_xor, along="submission_id", n=200)
# fit extended some model
model_some_ext <- extend(model_some, along="submission_id", n=200)

# wrapper for computing power for different effects for a given number of simulated models
effects <- c("prior", "competence", "relevance")
power_4_fe <- function(model, test, effects, nsim, N_subj) {
  N = length(effects)
  # (default alpha = 0.05 for the test)
  powers <- map(effects, ~ powerCurve(model, fixed(.x, test),
                                      nsim = nsim,
                                      progress = FALSE,
                                      along = "submission_id",
                                      breaks = N_subj)['ps']$ps)
  # [[which(effects == .x)]]$pval
  # ugly code above takes apart the powerCurve object
  # selects list of p-values from each simulation for each N

  # put results in a tibble
  power.sim <- tibble(
       N_subj = rep(list(N_subj), times = N),
      effect = effects,
      pvalues = powers 
    ) %>% unnest(col = c(N_subj, pvalues)) %>% 
    rowwise() %>%
    mutate(pvalues = list(pvalues$pval)) %>%
    unnest(col = pvalues)

  # compute the power as proportion of significant p-values
  power.sim %>% rowwise() %>%
    mutate(sign_check = ifelse(pvalues < 0.05, 1, 0)) %>%
    group_by(N_subj, effect) %>%
    summarize(power = mean(sign_check)) -> power.sim.summary

 return(power.sim.summary) 
}
```

Call the power simulations for xor and some, respectively. For each N, 1000 simulations are performed. The significance test chosen is the z-test (for no particular reason, for exploration). Alpha is 0.05. 
```{r, results = 'hide', warning=FALSE, message=FALSE}
# call the function
xor_power_sim <- power_4_fe(model_xor_ext, "z", effects, 1000, N_subj)
# write out results
xor_power_sim %>% write_csv("../data/xor_power_sim_N50-200_iter1000.csv")

some_power_sim <- power_4_fe(model_some_ext, "z", effects, 1000, N_subj)
# write out results
some_power_sim %>% write_csv("../data/some_power_sim_N50-200_iter1000.csv")
```

Visualize the results:
```{r}
xor_power_sim %>%
  ggplot(., aes(x = N_subj, y = power, color = effect)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 0.8, alpha = 0.7, linetype = 2) +
  ggtitle("Power for XOR trials by N of subjects")

some_power_sim %>%
  ggplot(., aes(x = N_subj, y = power, color = effect)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 0.8, alpha = 0.7, linetype = 2) +
  ggtitle("Power for SOME trials by N of subjects")
```

### 2021-08-26 analysis/xor-some_power.Rmd
add models with main effect of SI trigger

@@
 summary(model_some)
+```
+### Explore model with main effect of SI trigger
+```{r, results='hide'}
+# Full model 
+# sum-code effect of SI trigger
+d_both <- d_both %>% mutate(
+  main_type = as.factor(main_type)
+)
+# some = 1, xor = -1
+contrasts(d_both$main_type) <- contr.sum(2)
+# try to fit maximal model with brm
+model_SI_brm <- brms::brm(target ~ prior*competence*relevance* main_type +
+                   (1 + prior + competence + relevance + main_type || submission_id) +
+                   (1 | title),
+                 data = d_both,
+                 control = list(adapt_delta = 0.95),
+                 cores = 4,
+                 iter = 2000)
+```
+```{r}
+summary(model_SI_brm)
+```
+
+```{r, results='hide'}
+## frequentist model for observed power exploration
+# does not converge with 4-way interaction
+model_SI <- lmer(target ~ prior*competence*relevance + main_type +
+                   (1 + prior + competence + relevance + main_type | submission_id) +
+                   (1 | title),
+                 data = d_both
+                 )
+summary(model_SI)
+# also try model with by-subj intercept only for post-hoc power
+model_SI_simple <- lmer(target ~ prior*competence*relevance*main_type +
+                   (1 | submission_id) +
+                   (1 | title),
+                 data = d_both
+                 )
+summary(model_SI_simple)
+```
+```{r, results='hide', warning=FALSE, message=FALSE}
+# check observed power for simple effect of trigger
+p_main_type <- powerSim(model_SI_simple, 
+                  fixed("main_type1", "z"), 
+                  progress = FALSE)
+```
+
+```{r}
+p_main_type
 ```

### 2021-09-01 analysis/xor-some_final.Rmd
add final analysis with one large model and interval based evidence

---
title: "Xor & Some Final analysis"
author: "Polina Tsvilodub"
date: "8/31/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(brms)
library(tidyboot)
library(tidybayes)
```

## Read Data

First, the preprocessed and tidy wide pilot data is read and combined. It comprises data from N = 140 subjects after exclusions in total.
```{r, echo=FALSE}
pilot1 <- read_csv("./../data/pilots/pilot2_critical_zScore_wide_tidy.csv") %>%
  select(-competence_wUtt, -relevance_wUtt)
pilot2 <- read_csv("./../data/pilots/pilot3_critical_zScore_wide_tidy.csv")

d_both <- pilot1 %>% mutate(pilot = 1) %>% 
  rbind(., pilot2 %>% mutate(pilot = 2)
        )
# the data read here should be tidy, in correct format, bad workers should be excluded already
# z-scored data should be used and simulated!

# separate some and xor data!
d_both_some <- d_both %>% filter(main_type == "some") 
d_both_xor <- d_both %>% filter(main_type == "xor")
```
Sanity check the combined data (as can be seen, z-scored data is used):
```{r, echo=FALSE, message=FALSE, warning=FALSE}
d_both %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("Inference strength ratings") +
  facet_wrap(~main_type, ncol = 1) -> p.rel

d_both %>%
  ggplot(., aes(x = competence, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.comp

d_both %>%
  ggplot(., aes(x = prior, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.pri

gridExtra::grid.arrange(p.rel, p.comp, p.pri, ncol = 3) 
```

## Model

Next, fit the maximal model *across* experiments (i.e., *across SI trigger types*). For convergence reasons, correlation of random effects has to be set to 0. 
```{r, results='hide', message=FALSE, warning=FALSE}
# Full model 
d_both <- d_both %>% mutate(
  main_type = as.factor(main_type)
)
# some = 0, xor = 1
contrasts(d_both$main_type) 
# try to fit maximal model with brm
model_SI <- brm(target ~ prior*competence*relevance* main_type +
                   (1 + prior + competence + relevance + main_type || submission_id) +
                   (1 | title),
                 data = d_both,
                 control = list(adapt_delta = 0.95),
                 cores = 4,
                 iter = 3000)
```

```{r}
summary(model_SI)
```

## Extract conrtrasts of interest

Next, extract posterior contrasts of interest: effects of each predictor \{prior, competence and relevance\} for each trigger.

```{r}
model_SI %>% spread_draws(b_Intercept, b_prior, b_competence, b_relevance,
                          b_main_typexor, `b_prior:competence`,
                          `b_prior:relevance`, `b_competence:relevance`,
                          `b_prior:main_typexor`, `b_competence:main_typexor`,
                          `b_relevance:main_typexor`, `b_prior:competence:relevance`, 
                          `b_prior:competence:main_typexor`, `b_prior:relevance:main_typexor`,
                          `b_competence:relevance:main_typexor`, 
                          `b_prior:competence:relevance:main_typexor`) %>% 
  mutate(
    prior_xor = b_prior + b_main_typexor + `b_prior:main_typexor`,
    prior_some = b_prior,
    competence_xor = b_competence + b_main_typexor + `b_competence:main_typexor`,
    competence_some = b_competence,
    relevance_xor =  b_relevance + b_main_typexor + `b_relevance:main_typexor`,
    relevance_some = b_relevance,
    prior_some_vs_xor = -b_main_typexor - `b_prior:main_typexor`,
    comp_some_vs_xor =  -b_main_typexor - `b_competence:main_typexor`,
    rel_some_vs_xor =  -b_main_typexor - `b_relevance:main_typexor`
  ) -> model_SI_posteriors

# check P that the effects are positive / negative / no effect present
model_SI_posteriors %>% select(prior_xor, prior_some, 
                               competence_xor, competence_some,
                               relevance_xor, relevance_some) %>%
  gather(key, val) %>%
  group_by(key) %>% mutate(positive = mean(val > 0.05),
                           negative = mean(val < -0.05),
                           no = mean(val %>% between(-0.05, 0.05))) %>%
  summarise(positive_eff = mean(positive),
            negative_eff = mean(negative),
            no_eff = mean(no))

```

## Perform BF analysis

Since we want to address a conjunctive hypothesis, we check if all three contrasts of interest within a trigger are credibly different from 0. (?)
```{r}
# which models are the bayes factors computed on??
```

### 2021-09-14 analysis/xor-some_final.Rmd
fixed slope coeff. computation, added hypotheses checks

@@
   mutate(
-    prior_xor = b_prior + b_main_typexor + `b_prior:main_typexor`,
+    prior_xor = b_prior + `b_prior:main_typexor`,
     prior_some = b_prior,
-    competence_xor = b_competence + b_main_typexor + `b_competence:main_typexor`,
+    competence_xor = b_competence + `b_competence:main_typexor`,
     competence_some = b_competence,
-    relevance_xor =  b_relevance + b_main_typexor + `b_relevance:main_typexor`,
-    relevance_some = b_relevance,
-    prior_some_vs_xor = -b_main_typexor - `b_prior:main_typexor`,
-    comp_some_vs_xor =  -b_main_typexor - `b_competence:main_typexor`,
-    rel_some_vs_xor =  -b_main_typexor - `b_relevance:main_typexor`
+    relevance_xor =  b_relevance + `b_relevance:main_typexor`,
+    relevance_some = b_relevance
   ) -> model_SI_posteriors
@@
 # check P that the effects are positive / negative / no effect present
-model_SI_posteriors %>% select(prior_xor, prior_some, 
+posterior_hypotheses <- model_SI_posteriors %>% 
+  select(prior_xor, prior_some, 
          competence_xor, competence_some,
@@
             no_eff = mean(no))
-
+posterior_hypotheses
 ```
@@
 
-Since we want to address a conjunctive hypothesis, we check if all three contrasts of interest within a trigger are credibly different from 0. (?)
+We want to address a conjunctive hypothesis, one triplet for each trigger word, namely:
+
+1. the slope of 'competence' is positive
+2. the slope of 'prior' is negative
+3. the slope of 'relevance' is positive
+
+We judge there to be strong evidence for positivity of a slope variable $\beta_X$, if the posterior probability $P(\beta_X > \delta \mid D)$ is at least $.95$, for $\delta = 0.05$ the parameter that defines our 'region of practical equivalence'.
+A posterior odds ratio of at least $\frac{.95}{.05} = 19$ corresponds to a Bayes factor of at least 19 when prior odds are 1.
+
+The binary test of conformity with the theoretical predictions therefore is:
 ```{r}
-# which models are the bayes factors computed on??
+test_conjunction_of_all_hypotheses <-  function(posterior_hypotheses) {
+  posterior_hypotheses %>% 
+    mutate(hypothesis_true = case_when(
+      key == 'competence_some' ~ positive_eff > 0.95,
+      key == 'competence_xor'  ~ positive_eff > 0.95,
+      key == 'prior_some' ~ negative_eff > 0.95,
+      key == 'prior_xor'  ~ negative_eff > 0.95, 
+      key == 'relevance_some' ~ positive_eff > 0.95,
+      key == 'relevance_xor'  ~ positive_eff > 0.95
+    )) %>% 
+    pull(hypothesis_true) %>% all()
+}
+# applied to the pilot data
+test_conjunction_of_all_hypotheses(posterior_hypotheses)
 ```
+
+## Experimenting with multiple regression
+
+```{r}
+dummy_data <- tribble(
+  ~main, ~REL, ~target,
+  "some", 0, 4,
+  "some", 1, 6,
+  "xor", 0, 0,
+  "xor", 1, 12
+)
+# check that the slope of "REL" for "xor" is 12
+# this is calculated as betat_REL + beta_mainxor:REL
+# importantly: the intercept adjustment beta_mainxor doesn't come in!
+lm(target ~ main * REL, dummy_data)
+```

### 2021-09-14 analysis/xor-some_final.Rmd
test conjunction of H by trigger

@@
 
+```{r}
+# tiny update for separating the trigger words 
+test_conjunction_of_all_hypotheses2 <-  function(posterior_hypotheses, trigger) {
+  posterior_hypotheses %>% 
+    mutate(hypothesis_true = case_when(
+      key == paste('competence_', trigger, sep='') ~ positive_eff > 0.95,
+      key == paste('prior_', trigger, sep='') ~ negative_eff > 0.95,
+      key == paste('relevance_', trigger, sep='') ~ positive_eff > 0.95
+    )) %>% 
+    pull(hypothesis_true) %>% all()
+}
+# applied to the pilot data
+test_conjunction_of_all_hypotheses2(posterior_hypotheses, "xor")
+test_conjunction_of_all_hypotheses2(posterior_hypotheses, "some")
+```
+
 ## Experimenting with multiple regression

### 2021-09-17 analysis/power/xor-some_bayesian-power.Rmd
refactor, add init simulation for Bayesian power analysis

---
title: "XOr-Some Bayesian Power analysis"
author: "Polina Tsvilodub"
date: "9/14/2021"
output: github_document
---

For the main xor-some experiment, we conduct a simulation-based Bayesian power analysis (used in a (loose) sense of determining the required sample size here). This power analysis aims at determining how many participants are required in order to detect a theoretically motivated conjunctive effect of the predictors prior, relevance and competence with a confidence of at least .8. The simulations are based on an assumed effect size $\beta = ±0.15$ for all predictors, and a region of practical equivalence (ROPE) $\delta = 0.05$ for judging evidence for the directionality of a coefficient given the data. The simulations are further based on the maximal desired model justified theoretically and by the design: $rating = trigger * prior * competence * relevance + REs$

In general, the power analysis proceeds as follows:

* the desired effect sizes of the critical predictors are set to $\mid 0.15\mid$ 
* hypothetical experimental data is simulated by drawing samples from a Normal distribution with the parameters described above 
* the desired maximal Bayesian regression model is computed on the simulated data 
* the N of subjects for which the data is simulated is increased iteratively in steps of 10 starting at 50
* for each N, the model is re-computed on the simulated samples for k=100 iterations (the more, the better the power estimate, but there are computational constraints)
* for each iteration $i_k$, the conjunctive hypothesis of interest is tested 
* the power for the given number of participants is calculated as the proportion of iterations for which the conjunctive hypothesis was credible (i.e., if the posterior probability $P(\beta_X > \delta \mid D)$ is at least $.95$, for $\delta = 0.05$ the parameter that defines our ROPE for the conjunction of all three predictor coefficients, for all six sub-hypotheses).
* we aim to determine the number of participants for which that proportion is at least .8.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidybayes)
library(brms)
```

## Simulate initial data
We assume an effect size of 0.15 for the predictors prior, competence and relevance, and, therefore, simulate data as coming from a Normal distribution with $\mu=±.15$ and $\sigma=.05$ (because this $\sigma$ would give us app. 95% of the data in the region we defined as a positive / negative effect). The target is simulated via $\mathcal{N}(\mu_t, 1)$, with $mu_t = \beta_0 + \beta_{pri} + \beta_{comp} + \beta_{rel} + \beta_{trigger} + interactions$. The coefficients for interactions and the trigger are assumed to be 0.

From each participant, we get four data points for xor and four data points for some, one data point per condition (prior X competence X relevance, mapped to triggers at random). For each condition (for each trigger), there are four stories for which the random intercept may vary.

Below, there is a toy example how data would be simulated for one subject (ignoring random effects for now).
```{r}
# ignore group-level effects for now; the effects will be sampled based on pilot data

# simulate toy data set for one subject 
d_init <-
  # create the 8 conditions
  tibble(prior = rep(c(0, 1), each = 4),
         comp = rep(c(0, 0, 1, 1), times = 2),
         rel = rep(c(0,1), times = 4),
         main_type = sample(rep(c("xor", "some"), times=4))) %>% 
  rowwise() %>%
  # sample effects based on condition
  mutate(
    subj_ID = 1, 
    intercept = 0,
    prior_obs = ifelse(prior == 0, rnorm(1, sd = 0.05), 
                       rnorm(1, -0.15, 0.05)),
    comp_obs = ifelse(comp == 0, rnorm(1, sd = 0.05), 
                       rnorm(1, 0.15, 0.05)),
    rel_obs = ifelse(rel == 0, rnorm(1, sd = 0.05), 
                       rnorm(1, 0.15, 0.05)),
    # sample response
    # what do we do about the interactions? can they safely be ignored?
    target = rnorm(1, 
                   mean = (intercept + prior_obs + comp_obs + rel_obs), 
                   sd = 1) # how to set / detrmine this properly?
  )
d_init
```

## Compute initial model
Next, the seed model is computed on the initial simulated dataset. This model will only be updated in further steps when the model is re-fit. 
```{r, eval=FALSE}
# fit model on simulated data for first time  
model_init <- brm(target ~ prior*comp*rel*main_type + 
                    (1 + prior + competence + relevance + main_type || submission_id) +
                   (1 | title),
                  data = d_init,
                  cores = 4,
                  contol = list(adapt_delta = 0.95),
                  iter = 3000)
model_init
```

## Build pipeline for simulating more data and fitting models

```{r, eval=FALSE}
# TODO
# just old stuff below
get_new_data <- function(d, N) {
    if(N == 0) {
      data <- d
    } else if (N > 47) {
      data <- add_predicted_draws(model=pilot_model, 
                                     newdata = pilot_data %>% 
                                       filter(workerid %in% sample(unique(pilot_data$workerid), N, replace = T)),
                                     n = 1) %>% 
                 mutate(workerid = paste(workerid, letters[1], sep = "_")) %>%
                   rbind(., d)
    } else {
      data <- add_predicted_draws(model=pilot_model, 
                                     newdata = pilot_data %>% 
                                       filter(workerid %in% sample(unique(pilot_data$workerid), N, replace = F)),
                                     n = 1) %>% 
                 mutate(workerid = paste(workerid, letters[1], sep = "_")) %>%
                   rbind(., d)
    }

  }

```

## Extract contrasts of interest

```{r}
# TODO
```

## Test H of interest

```{r}
# TODO
```

### 2021-09-17 analysis/power/xor-some_power.Rmd
refactor, add init simulation for Bayesian power analysis

--- 2021-08-26 analysis/xor-some_power.Rmd
@@
 ```{r, echo=FALSE}
-pilot1 <- read_csv("./../data/pilots/pilot2_critical_zScore_wide_tidy.csv") %>%
+pilot1 <- read_csv("./../../data/pilots/pilot2_critical_zScore_wide_tidy.csv") %>%
   select(-competence_wUtt, -relevance_wUtt)
-pilot2 <- read_csv("./../data/pilots/pilot3_critical_zScore_wide_tidy.csv")
+pilot2 <- read_csv("./../../data/pilots/pilot3_critical_zScore_wide_tidy.csv")
 

### 2021-09-17 analysis/power/xor-some_bayesian-power.Rmd
copy over helper functions, add first stab at REs

@@
 95% of the data in the region we defined as a positive / negative effect).
-The target is simulated via $\mathcal{N}(\mu_t, 1)$, with $mu_t = \beta_0 + \beta_{pri} + \beta_{comp} + \beta_{rel} + \beta_{trigger} + interactions$.
-The coefficients for interactions and the trigger are assumed to be 0.
+
+The target response variable is simulated via $\mathcal{N}(\mu_t, 1)$, with $mu_t = \beta_0 + \beta_{pri} + \beta_{comp} + \beta_{rel} + \beta_{trigger} + interactions$.
+The coefficients for interactions and the trigger effect are assumed to be 0.
 
 From each participant, we get four data points for xor and four data points for some, one data point per condition (prior X competence X relevance, mapped to triggers at random).
-For each condition (for each trigger), there are four stories for which the random intercept may vary.
+For each condition (and for each trigger), there are four stories for which the random intercept may vary.
 
@@
          rel = rep(c(0,1), times = 4),
-         main_type = sample(rep(c("xor", "some"), times=4))) %>% 
+         main_type = as.factor(sample(rep(c("xor", "some"), times=4)))) %>% 
   rowwise() %>%
@@
                    mean = (intercept + prior_obs + comp_obs + rel_obs), 
-                   sd = 1) # how to set / detrmine this properly?
+                   sd = 1) # how to set / determine this properly?
   )
@@
 
+First, add random effects to the simulation based on random effects estimated on pilot data.
+```{r, eval=FALSE, warnings=FALSE, message=FALSE}
+# fit model on pilot data to get random effects estimates
+pilot1 <- read_csv("./../data/pilots/pilot2_critical_zScore_wide_tidy.csv") %>%
+  select(-competence_wUtt, -relevance_wUtt)
+pilot2 <- read_csv("./../data/pilots/pilot3_critical_zScore_wide_tidy.csv")
+
+d_both <- pilot1 %>% mutate(pilot = 1) %>% 
+  rbind(., pilot2 %>% mutate(pilot = 2)
+        )
+# Full model 
+d_both <- d_both %>% mutate(
+  main_type = as.factor(main_type)
+)
+# some = 0, xor = 1
+contrasts(d_both$main_type) 
+# try to fit maximal model with brm
+model_SI <- brm(target ~ prior*competence*relevance* main_type +
+                   (1 + prior + competence + relevance + main_type || submission_id) +
+                   (1 | title),
+                 data = d_both,
+                 control = list(adapt_delta = 0.95),
+                 cores = 4,
+                 iter = 3000)
+
+# Get the by-story estimates
+# check if this code can be less ugly
+story_rand_int <- c(ranef(model_SI)[['title']][1:(length(ranef(model_SI)[['title']])/4)])
+
+# get by-subj random intercepts and slopes
+# TODO
+
+# idea: sample from the vectors of observed estimates when simulating 
+# (across predictors (slopes for pri vs rel vs comp for subjects) 
+# and story types (comp X pri X rel condition of story))
+#### --> Is this okay? 
+```
+
+Build helper function for simulating data for a given number of subjects.
+[stuff from here on is not finished yet]
 ```{r, eval=FALSE}
-# TODO
-# just old stuff below
-get_new_data <- function(d, N) {
-    if(N == 0) {
-      data <- d
-    } else if (N > 47) {
-      data <- add_predicted_draws(model=pilot_model, 
-                                     newdata = pilot_data %>% 
-                                       filter(workerid %in% sample(unique(pilot_data$workerid), N, replace = T)),
-                                     n = 1) %>% 
-                 mutate(workerid = paste(workerid, letters[1], sep = "_")) %>%
-                   rbind(., d)
-    } else {
-      data <- add_predicted_draws(model=pilot_model, 
-                                     newdata = pilot_data %>% 
-                                       filter(workerid %in% sample(unique(pilot_data$workerid), N, replace = F)),
-                                     n = 1) %>% 
-                 mutate(workerid = paste(workerid, letters[1], sep = "_")) %>%
-                   rbind(., d)
-    }
+# TOFINISH
+# N = number of subjects to simulate
+get_new_data <- function(d, N) { # d probably unnecessary
+
+      data <-
+  # create the 8 conditions, manually, for double-checking
+        tibble(prior = rep(c(0, 0, 0, 0, 1, 1, 1, 1), times = N),
+         comp = rep(c(0, 0, 1, 1, 0, 0, 1, 1), times = N),
+         rel = rep(c(0, 1, 0, 1, 0, 1, 0, 1), times = N),
+         main_type = rep(sample(c("xor", "some", "xor", "some", "xor", "some", "xor", "some")), times=N),
+         # assign subject ID
+         subj_ID = rep(1:N, each = 8),
+         # prepare parameters 
+         intercept = 0,
+         # sample random effects
+         story_int = sample(story_rand_int, N*8, replace = T)
+         # TODO: analogously sample by-subj effects
+         ) %>% 
+  rowwise() %>%
+  # sample effects based on condition
+  mutate(
+    prior_obs = ifelse(prior == 0, rnorm(1, sd = 0.05), 
+                       rnorm(1, -0.15, 0.05)),
+    comp_obs = ifelse(comp == 0, rnorm(1, sd = 0.05), 
+                       rnorm(1, 0.15, 0.05)),
+    rel_obs = ifelse(rel == 0, rnorm(1, sd = 0.05), 
+                       rnorm(1, 0.15, 0.05)),
+    # sample response
+    # what do we do about the interactions? can they safely be ignored?
+    target = rnorm(1, 
+                   mean = (intercept + story_int + prior_obs + comp_obs + rel_obs), 
+                   sd = 1) # how to set / determine this properly?
+  )
 
@@
 
-## Extract contrasts of interest
+Build helper function for simulating and computing regression model for k iterations (tracked as seeds). Also extract contrasts of interest here.
+```{r, eval=FALSE}
+# TODO:
+# just copied over old stuff below
 
-```{r}
-# TODO
+# simulate data and update
+sim_data_fit <- function(seed, N) {
+  set.seed(seed)
+
+  # possibly add more participants
+  data <- get_new_data(d, N)
+
+  # mutate main_type to factor
+
+  # update model fit with new data
+  update(model_init,
+         newdata = data,
+         seed = seed) %>% 
+    # extract posterior draws
+    spread_draws(b_Intercept, b_syntax_dev1, b_trial_dev1, `b_syntax_dev1:trial_dev1`) %>%
+    # extract contrasts of interest, especially effect of syntax by-trial 
+  mutate(critical_subj = b_Intercept + b_syntax_dev1 - b_trial_dev1 - `b_syntax_dev1:trial_dev1`,
+         critical_pred = b_Intercept - b_syntax_dev1 - b_trial_dev1 + `b_syntax_dev1:trial_dev1`,
+         syntax_critical = critical_subj - critical_pred, # subject vs pred syntax
+         filler_subj = b_Intercept + b_syntax_dev1 + b_trial_dev1 + `b_syntax_dev1:trial_dev1`,
+         filler_pred = b_Intercept - b_syntax_dev1 + b_trial_dev1 - `b_syntax_dev1:trial_dev1`,
+         syntax_filler = filler_subj - filler_pred) %>% # subject vs predicate syntax
+  select(b_Intercept, b_syntax_dev1, b_trial_dev1, `b_syntax_dev1:trial_dev1`, critical_subj, critical_pred, syntax_critical, filler_subj, filler_pred, syntax_filler) %>%
+  gather(key, val) %>%
+  group_by(key) %>%
+    # compute summary statistics 
+  summarise( 
+    mean = mean(val),
+    lower = quantile(val, probs = 0.025),
+    upper = quantile(val, probs = 0.975)
+  )
+##########  
+  # use stuff below here
+  spread_draws(b_Intercept, b_prior, b_competence, b_relevance,
+                          b_main_typexor, `b_prior:competence`,
+                          `b_prior:relevance`, `b_competence:relevance`,
+                          `b_prior:main_typexor`, `b_competence:main_typexor`,
+                          `b_relevance:main_typexor`, `b_prior:competence:relevance`, 
+                          `b_prior:competence:main_typexor`, `b_prior:relevance:main_typexor`,
+                          `b_competence:relevance:main_typexor`, 
+                          `b_prior:competence:relevance:main_typexor`) %>% 
+  mutate(
+    prior_xor = b_prior + `b_prior:main_typexor`,
+    prior_some = b_prior,
+    competence_xor = b_competence + `b_competence:main_typexor`,
+    competence_some = b_competence,
+    relevance_xor =  b_relevance + `b_relevance:main_typexor`,
+    relevance_some = b_relevance
+  ) -> model_SI_posteriors
+
+# check P that the effects are positive / negative / no effect present
+posterior_hypotheses <- model_SI_posteriors %>% 
+  select(prior_xor, prior_some, 
+         competence_xor, competence_some,
+         relevance_xor, relevance_some) %>%
+  gather(key, val) %>%
+  group_by(key) %>% mutate(positive = mean(val > 0.05),
+                           negative = mean(val < -0.05),
+                           no = mean(val %>% between(-0.05, 0.05))) %>%
+  summarise(positive_eff = mean(positive),
+            negative_eff = mean(negative),
+            no_eff = mean(no))
+}
 ```
@@
 
-```{r}
+Run the simulations and test the hypothesis of interest for each iteration and compute the proportion of successful runs.
+```{r, eval=FALSE}
 # TODO
+test_conjunction_of_all_hypotheses <-  function(posterior_hypotheses) {
+  posterior_hypotheses %>% 
+    mutate(hypothesis_true = case_when(
+      key == 'competence_some' ~ positive_eff > 0.95,
+      key == 'competence_xor'  ~ positive_eff > 0.95,
+      key == 'prior_some' ~ negative_eff > 0.95,
+      key == 'prior_xor'  ~ negative_eff > 0.95, 
+      key == 'relevance_some' ~ positive_eff > 0.95,
+      key == 'relevance_xor'  ~ positive_eff > 0.95
+    )) %>% 
+    pull(hypothesis_true) %>% all()
+}
+
+# again old stuff below
+
+# add iterating over Ns
+sim1 <-
+  tibble(seed = 1:100) %>% 
+  mutate(tidy = map(seed, sim_data_fit, 23)) %>% 
+  unnest(tidy)
+
+# map each iter to test_conjunction_of_all_hypotheses
+sim1 %>%
+  filter(key == "syntax_critical") %>%
+  mutate(check_syntax = ifelse(lower > 0, 1, 0)) %>%
+  summarise(power_syntax = mean(check_syntax))
+
 ```

### 2021-09-27 analysis/power/xor-some_bayesian-power.Rmd
Bayesian Power: minimal version

@@
 date: "9/14/2021"
-output: github_document
+output: 
+  github_document:
+  pandoc_args: --webtex
 ---
@@
 The simulations are based on an assumed effect size $\beta = ±0.15$ for all predictors, and a region of practical equivalence (ROPE) $\delta = 0.05$ for judging evidence for the directionality of a coefficient given the data.
-The simulations are further based on the maximal desired model justified theoretically and by the design: $rating = trigger * prior * competence * relevance + REs$
+The simulations are further based on the maximal desired model justified theoretically and by the design: `rating = trigger * prior * competence * relevance + REs`
 
@@
 
-The target response variable is simulated via $\mathcal{N}(\mu_t, 1)$, with $mu_t = \beta_0 + \beta_{pri} + \beta_{comp} + \beta_{rel} + \beta_{trigger} + interactions$.
+The target response variable is simulated via $\mathcal{N}(\mu_t, 1)$, with $\mu_t = \beta_0 + \beta_\text{pri} + \beta_\text{comp} + \beta_\text{rel} + \beta_\text{trigger} + \text{interactions}$.
 The coefficients for interactions and the trigger effect are assumed to be 0.
@@
                    sd = 1) # how to set / determine this properly?
+                           # MF: irrelvant; we should z-score anyway exactly like we 
+                           #     do for the real analysis 
+                           #     this also means that choices of mean and SD for 
+                           #     distributions to sample coefficents for are no that relevant
   )
@@
 ```
+
+# MF's crowbar ignoramus approach
+
+```{r, eval = FALSE}
+# create fake data
+create_fake_data <- function(N = 100, seed = 1) {
+  set.seed(seed)
+  beta_rel  <- 0.15
+  beta_comp <- 0.15
+  beta_pri  <- 0.15
+  map_df(
+    1:N, 
+    function(i) {
+      tibble(
+        subj    = rep(str_c("subj_",i), times = 8),
+        trigger = rep(c('or', 'some'), times= 4), 
+        rel     = rnorm(8),
+        comp    = rnorm(8),
+        pri     = rnorm(8)
+      ) %>% 
+        group_by(subj) %>% 
+        mutate(
+          rel     = rel-mean(rel) / sd(rel),
+          comp    = comp-mean(comp) / sd(comp),
+          pri     = pri-mean(pri) / sd(pri)
+        ) %>% 
+        ungroup() %>% 
+        mutate(
+          target  = rnorm(8, beta_rel * rel + beta_comp * comp + beta_pri * pri)
+        )
+    }
+  )
+}
+initial_fake_data <- create_fake_data(N=100, seed = 3)
+
+# initial fit
+fit_initial <- brm(target ~ trigger * rel * comp * pri, initial_fake_data)
+
+# main results
+N <- 300
+k <-100
+
+results <- map_dbl(
+  1:k, 
+  function(i) {
+    fake_data <- create_fake_data(N=N, seed = i)
+    fit <- update(fit_initial, newdata = fake_data, seed = i, cores = 4)
+    lower_bounds <- rbind(
+      aida::summarize_sample_vector(tidybayes::tidy_draws(fit) %>% select(b_rel), "rel"),
+      aida::summarize_sample_vector(tidybayes::tidy_draws(fit) %>% select(b_comp), "comp"),
+      aida::summarize_sample_vector(tidybayes::tidy_draws(fit) %>% select(b_pri), "pri")
+    ) %>% pull("|95%")
+    min(lower_bounds)
+  }
+)
+
+message("Effective power for N=", N, " is:" , mean(results > 0.05))
+
+```

### 2021-10-05 analysis/power/xor-some_bayesian-power.Rmd
simplified minimal Bayesian power script

--- 2021-09-17 analysis/power/xor-some_bayesian-power.Rmd
@@
 ```
+
+# MF's crowbar ignoramus approach
+
+```{r, eval = FALSE}
+# create fake data
+create_fake_data <- function(N = 100, seed = 1) {
+  set.seed(seed)
+  beta_rel  <- 0.15
+  beta_comp <- 0.15
+  beta_pri  <- 0.15
+  map_df(
+    1:N, 
+    function(i) {
+      tibble(
+        subj    = rep(str_c("subj_",i), times = 4),
+        rel     = rnorm(4),
+        comp    = rnorm(4),
+        pri     = rnorm(4)
+      ) %>% 
+        group_by(subj) %>% 
+        mutate(
+          rel     = (rel-mean(rel)) / sd(rel),
+          comp    = (comp-mean(comp)) / sd(comp),
+          pri     = (pri-mean(pri)) / sd(pri)
+        ) %>% 
+        ungroup() %>% 
+        mutate(
+          target  = rnorm(4, beta_rel * rel + beta_comp * comp + beta_pri * pri)
+        )
+    }
+  )
+}
+initial_fake_data <- create_fake_data(N=100, seed = 1234)
+
+# initial fit
+fit_initial <- brm(target ~ rel * comp * pri, initial_fake_data)
+
+# main results
+N <- 300
+k <-10
+
+results <- map_dbl(
+  1:k, 
+  function(i) {
+    fake_data <- create_fake_data(N=N, seed = i)
+    fit <- update(fit_initial, newdata = fake_data, seed = i, cores = 4)
+    lower_bounds <- rbind(
+      aida::summarize_sample_vector(brms::posterior_samples(fit)[,'b_rel'], "rel"),
+      aida::summarize_sample_vector(brms::posterior_samples(fit)[,'b_comp'], "comp"),
+      aida::summarize_sample_vector(brms::posterior_samples(fit)[,'b_pri'], "pri")
+    ) %>% pull("|95%")
+    min(lower_bounds)
+  }
+)
+
+message("Effective power for N=", N, " is:" , mean(results > 0.05))
+```

### 2021-10-15 04_trials_final.js
add Prolific inspired attention check

@@ -1,3 +1,12 @@
+let attention_check_data = [
+  {
+    optionLeft: "not at all",
+    optionRight: "absolutely",
+    QUD: "Please carefully read the text below and answer the question displayed below, given the information presented in the text.",
+    question: "John is a guy. Brandon is another one. Please adjust the slider all the way to the right in this task. They are both friends with Mary. <br> <br> Based on that, one could conclude that John has friends. How would you respond using the slider, given the text?"
+  }
+]
+
 // In this file you can specify the trial data for your experiment
 // info important to create tirals
 

### 2021-10-15 05_views.js

@@ -202,6 +202,16 @@ const ex_trial = magpieViews.view_generator("slider_rating", {
 }
 );
 
+// attention check trial
+const attention_check = magpieViews.view_generator("slider_rating", {
+  trials: 1,
+  title: "Attention check",
+  name: 'attention_check',
+  trial_type: 'attention_check',
+  data: attention_check_data
+
+})
+
 // captcha
 // speaker and listeneers names to be sampled from for the botcaptcha
 var speaker = _.sample(["James", "John", "Robert", "Michael", "William", "David", "Richard", "Joseph", "Thomas", "Charles"]);

### 2021-10-15 06_main.js

@@ -19,6 +19,7 @@ $("document")
             intro,
             botcaptcha,
             instructions,
+            attention_check,
             ex_trial,
             begin_experiment,
             // _.flatten(_.shuffle([[comp_trial, after_block1], [pri_trial, after_block2], [rel_trial, after_block3]])),

### 2021-10-16 05_views.js
make prolific attention checks from unused stories, remove botcheck

@@ -50,7 +50,7 @@ const instructions = magpieViews.view_generator("instructions", {
   trials: 1,
   name: 'instrucions',
   title: 'Instructions',
-  text: `In the following, you will be presented with <b> 8 short stories</b>.
+  text: `In the following, you will be presented with <b> 16 short stories</b>.
             Please read them very carefully, even if they appear to be repeated and you think that you remember them well enough.
             We ask you to <b> rate statements about each short story</b>.
             Please indicate, using an adjustable slider, <b> how likely you think a statement
@@ -204,13 +204,25 @@ const ex_trial = magpieViews.view_generator("slider_rating", {
 
 // attention check trial
 const attention_check = magpieViews.view_generator("slider_rating", {
-  trials: 1,
+  trials: attention_check_data.length,
   title: "Attention check",
   name: 'attention_check',
   trial_type: 'attention_check',
-  data: attention_check_data
+  data: _.shuffle(attention_check_data)
+},
+{
+  stimulus_container_generator: function (config, CT) {
+    return `<div class='magpie-view'>
+                <h1 class='magpie-view-title'>${config.data[CT].title}</h1>
+                <p class='magpie-view-question magpie-view-qud' id='qud'>${config.data[CT].QUD}</p>
 
-})
+                  <p class='magpie-view-question' id='critical-utterance'>${config.data[CT].critical_question}</p>
+
+                  <p class='magpie-view-question' id='prompt'>${config.data[CT].prompt}</p>
+
+    </div>`;}
+}
+)
 
 // captcha
 // speaker and listeneers names to be sampled from for the botcaptcha

### 2021-10-16 06_main.js

@@ -8,7 +8,7 @@ $("document")
         e.preventDefault();
       }
     };
-
+    console.log( _.shuffle(Array.from(attention_check).concat( Array.from(xor_trial))))
     // calls magpieInit
     // in debug mode this returns the magpie-object, which you can access in the console of your browser
     // e.g. >> window.magpie_monitor or window.magpie_monitor.findNextView()
@@ -17,17 +17,14 @@ $("document")
       // You have to specify all views you want to use in this experiment and the order of them
       views_seq: [
             intro,
-            botcaptcha,
             instructions,
-            attention_check,
             ex_trial,
             begin_experiment,
-            // _.flatten(_.shuffle([[comp_trial, after_block1], [pri_trial, after_block2], [rel_trial, after_block3]])),
+            attention_check,
             xor_trial,
             post_test,
             thanks
         ],
-
 
       // Here, you can specify all information for the deployment
       deploy: {

### 2021-10-16 04_trials_final.js
push correct trial info file

@@ -1,9 +1,99 @@
 let attention_check_data = [
   {
-    optionLeft: "not at all",
-    optionRight: "absolutely",
-    QUD: "Please carefully read the text below and answer the question displayed below, given the information presented in the text.",
-    question: "John is a guy. Brandon is another one. Please adjust the slider all the way to the right in this task. They are both friends with Mary. <br> <br> Based on that, one could conclude that John has friends. How would you respond using the slider, given the text?"
+    title: 'Rumors',
+    QUD: "John and his wife are eating dinner. They have been extremely busy all day in connection with their stressful jobs as real estate agents. Please move the slider all the way to the right. They are discussing a rumor saying that a famous singer has committed suicide in the morning. ",
+    optionLeft: '<p style="font-family: sans-serif">certainly false</p>',
+    optionRight: '<p style="font-family: sans-serif">certainly true</p>',
+    condition: "attention_check",
+    critical_question: ``,
+    prompt: "John really liked that singer.",
+    question: '<font size="3"> How likely is it that the statement is true given the story?</font>',
+    block: 'attention_check',
+    expected: 100
+  },
+  {
+    title: 'Underage drinking',
+    QUD: "Edward is in a crowded bar IDing the patrons who are drinking alcohol. There are new laws and if there is even one incident of underage drinking today Edward will close down the bar. Please move the slider all the way to the right. He talks to the owner of the bar later that night.",
+    optionLeft: '<p style="font-family: sans-serif">certainly false</p>',
+    optionRight: '<p style="font-family: sans-serif">certainly true</p>',
+    condition: "attention_check",
+    critical_question: ``,
+    prompt: "Five people were drinking beer at the bar.",
+    question: '<font size="3"> How likely is it that the statement is true given the story?</font>',
+    block: 'attention_check',
+    expected: 100
+  },
+  {
+    title: 'Health inspector',
+    QUD: "One of the duties of a restaurant health inspector is to test whether the food doesn't contain more bacteria than allowed. Please move the slider all the way to the left. After an inspection, the inspector talks to the restaurant owner.",
+    optionLeft: '<p style="font-family: sans-serif">certainly false</p>',
+    optionRight: '<p style="font-family: sans-serif">certainly true</p>',
+    condition: "attention_check",
+    critical_question: ``,
+    prompt: "The restaurant owner had to pay a fine because some food contained bacteria.",
[…]

### 2021-10-17 analysis/power/xor-some_bayesian-power_500.Rmd
final 500 iter power simulation

---
title: "XOr-Some Bayesian Power analysis"
author: "Polina Tsvilodub"
date: "9/14/2021"
output: 
  github_document:
  pandoc_args: --webtex
---

For the main xor-some experiment, we conduct a simulation-based Bayesian power analysis (used in a (loose) sense of determining the required sample size here). This power analysis aims at determining how many participants are required in order to detect a theoretically motivated conjunctive effect of the predictors prior, relevance and competence with a confidence of at least .8. The simulations are based on an assumed effect size $\beta = 0.15$ for all predictors, and a region of practical equivalence (ROPE) $\delta = 0.05$ for judging evidence for the directionality of a coefficient given the data. The simulations are further based on the maximal desired model justified theoretically and by the design: `rating = trigger * prior * competence * relevance`

In general, the power analysis proceeds as follows:

* the desired effect sizes of the critical predictors are set to $0.15$ 
* hypothetical experimental data is simulated by drawing samples from a Normal distribution with the parameters described above 
* the desired maximal Bayesian regression model is computed on the simulated data 
* the N of subjects for which the data is simulated is increased iteratively in steps of 20 starting at 100 up to 340.
* for each N, the model is re-computed on the simulated samples for k=100 iterations (the more, the better the power estimate, but there are computational constraints)
* for each iteration $i_k$, the conjunctive hypothesis of interest is tested 
* the power for the given number of participants is calculated as the proportion of iterations for which the conjunctive hypothesis was credible (i.e., if the posterior probability $P(\beta_X > \delta \mid D)$ is at least $.95$, for $\delta = 0.05$ the parameter that defines our ROPE for the conjunction of all three predictor coefficients, for all six sub-hypotheses).
* we aim to determine the number of participants for which that proportion is at least .8.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidybayes)
library(brms)
```

## Simulate initial data
We assume an effect size of 0.15 for the predictors prior, competence and relevance. We simulate data as coming from a Normal distribution with $\mu=0$ and $\sigma=1$ and z -score it, to follow the experiemental procedure.

The target response variable is simulated via $\mathcal{N}(\mu_t, 1)$, with $\mu_t = 0.15 * ( \beta_\text{pri} + \beta_\text{comp} + \beta_\text{rel} + \beta_\text{trigger} ) + \text{interactions}$. The multiplication by 0.15 represent hereby the assumed effect size. The coefficients for interactions and the trigger effect are assumed to be 0.

From each participant, we get four data points for xor and four data points for some, one data point per condition (prior X competence X relevance, mapped to triggers at random). Therefore, we simulate 8 trials per subject.

```{r}
# create fake data
create_fake_data <- function(N = 100, seed = 1) {
  set.seed(seed)
  beta_rel  <- 0.15
  beta_comp <- 0.15
  beta_pri  <- 0.15
  map_df(
    1:N, 
    function(i) {
      tibble(
        subj    = rep(str_c("subj_",i), times = 8),
        trigger = rep(c('or', 'some'), times= 4), 
        rel     = rnorm(8),
        comp    = rnorm(8),
        pri     = rnorm(8)
      ) %>% 
        group_by(subj) %>% 
        mutate(
          rel     = (rel-mean(rel)) / sd(rel),
          comp    = (comp-mean(comp)) / sd(comp),
          pri     = (pri-mean(pri)) / sd(pri)
        ) %>% 
        ungroup() %>% 
        mutate(
          target  = rnorm(8, beta_rel * rel + beta_comp * comp + beta_pri * pri)
        )
    }
  )
}

```

## Compute initial model
Next, the seed model is computed on the initial simulated dataset. This model will only be updated in further steps when the model is re-fit. 
```{r}
# create initial dataset 
initial_fake_data <- create_fake_data(N=100, seed = 1234)

# initial fit
fit_initial <- brm(bf(target ~ rel * comp * pri,
                      decomp = "QR"), 
                   initial_fake_data)

```

## Build pipeline for simulating more data and fitting models

Build helper function for simulating and computing regression model for k iterations (tracked as seeds). Also extract contrasts of interest here.
```{r, message = FALSE, warning=FALSE, results='hide'}
# main results
start_N <- 100
step <- 20
stop_N <-340
seed <- 500

sim_data_fit <- function(N, k){
 results <- map_dbl(
  1:k, 
  function(i) {
    fake_data <- create_fake_data(N=N, seed = i)
    fit <- update(fit_initial, newdata = fake_data, seed = i, cores = 4)
    lower_bounds <- rbind(
      aida::summarize_sample_vector(brms::posterior_samples(fit)[,'b_rel'], "rel"),
      aida::summarize_sample_vector(brms::posterior_samples(fit)[,'b_comp'], "comp"),
      aida::summarize_sample_vector(brms::posterior_samples(fit)[,'b_pri'], "pri")
    ) %>% pull("|95%")
    min(lower_bounds)
  }
 )
 # create path for file
 stream_out <- paste("./results/power_sim_N", N, "_iter", k, ".csv", sep="")
 # create tibble for intermediate saving
 tibble(
   seed = 1:k, lower_bounds = results
 ) %>%
   write_csv(., stream_out)
}

sim_power <- function(seed, start_N, step, stop_N) {
  # iterate over different Ns of subjects
  sim_full <- tibble( N = seq(start_N, stop_N, step)) %>%
    mutate(
      tidy = map(N, sim_data_fit, k = seed)
    ) %>%
    unnest(tidy)
  write_csv(sim_full, paste("./results/power_sim_full_N", start_N, "-N", stop_N, "_iter", seed, ".csv", sep=""), append = T, col_names = T)
}

power_df_500 <- sim_power(seed, start_N, step, stop_N)

```

Below, the results are summarized by simulated N of subjects to obtain the power estimates (i.e., proportion of simulation runs where all three effects were >0.05).
```{r}
# analyse power summary
power_results_500 <- power_df_500 %>% group_by(N) %>% 
  summarise(power = mean(lower_bounds > 0.05))

write_csv(power_results_500, "./results/power_sim_500_results_summary.csv")

power_results_500
```

Finally, power results are visualized as a function of the N of subjects:
```{r}
power_results_500 %>%
  ggplot(., aes(x = N, y = power)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 0.8, alpha = 0.7, linetype = 2) +
  ggtitle("Power by N of subjects") 
ggsave("./results/power_sim_results_plot.png")
```

### 2021-10-17 04_trials_final.js
add shuffling of main trials w attention checks

@@ -49,7 +49,7 @@ let attention_check_data = [
   },
   {
     title: 'TV experts',
-    QUD: "Jason Barley and Richard Trellis are TV experts engaged in a live discussion of the current NBA season. They are debating about whether Greg Jones should be the Most Valuable Player of the season. Please click on the slider and leave it in the middle. Jason Barley thinks so, but Richard Trellis is less convinced.",
+    QUD: "Jason Barley and Richard Trellis are TV experts engaged in a live discussion of the current NBA season. They are debating about whether Greg Jones should be the Most Valuable Player of the season. Please click on the slider handle and leave it in the middle. Jason Barley thinks so, but Richard Trellis is less convinced.",
     optionLeft: '<p style="font-family: sans-serif">certainly false</p>',
     optionRight: '<p style="font-family: sans-serif">certainly true</p>',
     condition: "attention_check",
@@ -61,7 +61,7 @@ let attention_check_data = [
   },
   {
     title: 'Homework',
-    QUD: "Mr. Carpenter teaches at a university and gave his students a homework assignment last week. Please click on the slider and leave it in the middle. The students are supposed to submit their homework by putting it in a dedicated post box. Mr. Carpenter collects the submitted assignments.",
+    QUD: "Mr. Carpenter teaches at a university and gave his students a homework assignment last week. Please click on the slider handle and leave it in the middle. The students are supposed to submit their homework by putting it in a dedicated post box. Mr. Carpenter collects the submitted assignments.",
     optionLeft: '<p style="font-family: sans-serif">certainly false</p>',
     optionRight: '<p style="font-family: sans-serif">certainly true</p>',
     condition: "attention_check",
@@ -104,7 +104,7 @@ let attention_check_data = [
 let vignettes_or = {
   lll: [
     {
-   "ID": 7,
+   "ID": 1,
    "name": "Jimmy's trip",
    "author": "micha",
    "final_selection": 1,
@@ -133,7 +133,7 @@ let vignettes_or = {
    "test_uncertain2": "Jimmy’s brother has been to New York City last year."
  },
 //  {
-//    "ID": 15,
+//    "ID": 2,
 //    "name": "Harold's pet",
 //    "author": "bob",
 //    "final_selection": 1,
@@ -1906,29 +1906,12 @@ const trials_practice = [{
[…]

### 2021-10-17 05_views.js

@@ -202,43 +202,27 @@ const ex_trial = magpieViews.view_generator("slider_rating", {
 }
 );
 
-// attention check trial
-const attention_check = magpieViews.view_generator("slider_rating", {
-  trials: attention_check_data.length,
-  title: "Attention check",
-  name: 'attention_check',
-  trial_type: 'attention_check',
-  data: _.shuffle(attention_check_data)
+const main_trial_data = main_block.concat(attention_check_data)
+const main_trials = magpieViews.view_generator("slider_rating", {
+  trials: attention_check_data.length + _.flatten(main_block).length,
+  name: "main_trials",
+  trial_type: "main_slider",
+  data: _.flatten(_.shuffle(main_trial_data))
 },
-{
+{ // custom view allowing to add custom backgrounds of different view elements
   stimulus_container_generator: function (config, CT) {
-    return `<div class='magpie-view'>
-                <h1 class='magpie-view-title'>${config.data[CT].title}</h1>
-                <p class='magpie-view-question magpie-view-qud' id='qud'>${config.data[CT].QUD}</p>
+            return `<div class='magpie-view'>
+                        <h1 class='magpie-view-title'>${config.data[CT].title}</h1>
+                        <p class='magpie-view-question magpie-view-qud' id='qud'>${config.data[CT].QUD}</p>
 
-                  <p class='magpie-view-question' id='critical-utterance'>${config.data[CT].critical_question}</p>
+                          <p class='magpie-view-question' id='critical-utterance'>${config.data[CT].critical_question}</p>
 
-                  <p class='magpie-view-question' id='prompt'>${config.data[CT].prompt}</p>
+                          <p class='magpie-view-question' id='prompt'>${config.data[CT].prompt}</p>
 
-    </div>`;}
+            </div>`;} // <p class='magpie-view-question' id='question'>${config.data[CT].question}</p>
 }
-)
+) 
 
-// captcha
-// speaker and listeneers names to be sampled from for the botcaptcha
-var speaker = _.sample(["James", "John", "Robert", "Michael", "William", "David", "Richard", "Joseph", "Thomas", "Charles"]);
-var listener = _.sample(["Mary", "Patricia", "Jennifer", "Linda", "Elizabeth", "Barbara", "Susan", "Jessica", "Sarah", "Margaret"]);
-
-
-const botcaptcha = custom_botcaptcha({
-  name: 'botcaptcha',
-  trials: 1,
[…]

### 2021-10-17 06_main.js

@@ -8,7 +8,6 @@ $("document")
         e.preventDefault();
       }
     };
-    console.log( _.shuffle(Array.from(attention_check).concat( Array.from(xor_trial))))
     // calls magpieInit
     // in debug mode this returns the magpie-object, which you can access in the console of your browser
     // e.g. >> window.magpie_monitor or window.magpie_monitor.findNextView()
@@ -20,8 +19,8 @@ $("document")
             instructions,
             ex_trial,
             begin_experiment,
-            attention_check,
-            xor_trial,
+          //  _.shuffle([attention_check, xor_trial]),
+            main_trials,
             post_test,
             thanks
         ],
@@ -33,7 +32,7 @@ $("document")
         // Possible deployment methods are:
         // "debug" and "directLink"
         // As well as "MTurk", "MTurkSandbox" and "Prolific"
-        deployMethod: "Prolific",
+        deployMethod: "debug",
         contact_email: "polina.tsvilodub@gmail.com",
         prolificURL: "https://app.prolific.co/submissions/complete?cc=7E05F86F"
       },
@@ -41,8 +40,9 @@ $("document")
       progress_bar: {
         in: [
                 // list the view-names of the views for which you want a progress bar
-                xor_trial.name
-                // pri_trial.name,
+                //xor_trial.name
+
+                main_trials.name                // pri_trial.name,
                 // rel_trial.name,
                 // comp_trial.name
 

### 2021-10-17 04_trials_final.js
add back all vignettes, assign unique IDs

@@ -132,37 +132,37 @@ let vignettes_or = {
    "test_uncertain1": "Jimmy had told his friends about his plans to take a day trip before.",
    "test_uncertain2": "Jimmy’s brother has been to New York City last year."
  },
-//  {
-//    "ID": 2,
-//    "name": "Harold's pet",
-//    "author": "bob",
-//    "final_selection": 1,
-//    "type": "lll",
-//    "relevance": 0,
-//    "competence": 0,
-//    "prior": 0,
-//    "background": "Richard and Tom are talking about Harold. Harold has been wanting a pet for a long time but his girlfriend wouldn't let him. Richard heard she finally gave in.",
-//    "utterance_or": "Richard says to Tom: 'Harold bought a cat or a dog.'",
-//    "utterance_disj1": "Richard says to Tom: 'Harold bought a cat.'",
-//    "utterance_disj2": "Richard says to Tom: 'Harold bought a dog.'",
-//    "question_rel": "It is important for Tom to know whether Harold bought both a cat and a dog.",
-//    "question_comp": "Richard knows whether Harold bought both a cat and a dog.",
-//    "question_pri1": "If Harold bought a cat, it is likely that he bought a dog as well.",
-//    "question_pri2": "If Harold bought a dog, it is likely that he bought a cat as well.",
-//    "question_pri_and": "It is likely that Harold bought both a cat and a dog.",
-//    "question_xor": "From what Richard said we may conclude that Harold did not buy both a cat and a dog.",
-//    "question_xorI": "Richard wants to convey that Harold did not buy both a cat and a dog.",
-//    "question_exh1": "From what Richard said we may conclude that Harold did not buy a dog as well.",
-//    "question_exh2": "From what Richard said we may conclude that Harold did not buy a cat as well.",
-//    "test_true1": "Harold has a girlfriend.",
-//    "test_true2": "Harold's girlfriend didn't want a pet in her house before.",
-//    "test_false1": "Harold used to have a parrot.",
-//    "test_false2": "Tom and Harold are roommates.",
-//    "test_uncertain1": "Harold’s girlfriend is afraid that pets will ruin the apartment.",
-//    "test_uncertain2": "Richard has a dog."
-//  },
  {
-   "ID": 16,
+   "ID": 2,
+   "name": "Harold's pet",
+   "author": "bob",
+   "final_selection": 1,
+   "type": "lll",
+   "relevance": 0,
+   "competence": 0,
+   "prior": 0,
[…]

### 2021-10-18 analysis/prereg/01_xor-some_prereg_preprocessing.Rmd
refactor prereg R scripts

---
title: 'Xor-Some Preregistered Study: Preprocessing'
author: "Polina Tsvilodub"
date: "10/18/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(brms)
library(tidyboot)
library(tidybayes)
```

## Read Data

First, the data is loaded.
```{r, echo=FALSE}
d <- read_csv("./../../data/results_test.csv") #prereg_final

# for testing purposes
d <- d %>% mutate(submission_id = 42)

# separate some and xor data!
d_some <- d %>% filter(main_type == "some") 
d_xor  <- d %>% filter(main_type == "xor")
```

## Clean data

Checking if there are any comments indicating technical issues, to be excluded then:
```{r comments}
d %>% distinct(comments) %>% View()
```

Apply language related exclusion criterion:
Check native languages. Participants not indicating English as (one of) their native language(s) are excluded. We also compute some participant demographics.
```{r languages}
d %>% distinct(languages)
cat("Number of participants before excluding non-natives: ", d %>% distinct(submission_id) %>% count() %>% pull() ) 
# exclude non-natives if necessary
d_native <- d %>% 
 # filter(("en" | "En" | "bri") %in% languages)
  filter(grepl("[(en)(br)]", languages, ignore.case = T))

cat(" Number of participants after excluding non-natives: ", d_native %>% distinct(submission_id) %>% count() %>% pull() )

cat(" Mean age: ", d_native %>% pull(age) %>% mean(., na.rm = T) )
d_native  %>% count(gender) %>% mutate(n = n/72)
```

Check the overall completion time. Participants who complete the experiment in less than 8 minutes are excluded.
```{r}
d_native_completed <- d_native %>% mutate(timeSpent = round(timeSpent, 2)) %>%
  filter(timeSpent > 8) 
```

Apply attention check based exclusion criterion:

```{r}
d_native_attention <- d_native_completed %>% 
  filter(condition == "attention_check") %>% mutate(
  passed_attention_trial = case_when(
    (expected == 0) & (response == 0) ~ T,
    (expected == 100) & (response == 100) ~ T,
    (expected == 50) & (response %>% between(45, 55)) ~ T,
    TRUE ~ F
  )
)

d_native_attention_summary <- d_native_attention %>% 
  group_by(submission_id) %>%
  summarize(attention_prop = mean(passed_attention_trial)) %>%
  mutate(
    passed_attention_check = attention_prop > 0.75
  ) %>% 
  filter(passed_attention_check == TRUE)

d_native_attended <- left_join(d_native_completed, 
                               d_native_attention_summary, 
                               by = c("submission_id"))

```

Data frames preparation:
```{r}
d_main <- d_native_attended %>% select(-age, -comments, -education, -endTime, 
                              -gender, -languages, -optionLeft, -optionRight, -startDate,
                              -startTime, -timeSpent, -expected) %>%
  filter(trial_name != "example", trial_name != "attention_check")
d_exmpl <- d_native_attended %>% select(-age, -comments, -education, -endTime, 
                              -gender, -languages, -optionLeft, -optionRight, -startDate,
                              -startTime, -timeSpent, -expected) %>%
  filter(trial_name == "example")
d_critical <- d_main %>% filter(condition == "critical")

d_test <- d_main %>% rowwise() %>% filter(condition == "test") %>% 
  mutate(test_condition = substr(test_question, 6, 9),
         test_condition = ifelse(test_condition == "fals", "false", 
                                 ifelse(test_condition == "unce", "uncertain",
                                        test_condition)))
```

Exclude participants based on example trials, comprehension question performance, and response patterns.
Participants who gave all responses within the range of 10 and participants who failed more than 0.2 of the comprehension questions are excluded from analysis. 
Participants who failed all example trials are excluded, as well. 
```{r exclusions}
# get participants failing example trials
d_exmpl_fail <- d_exmpl %>% group_by(submission_id) %>% 
  mutate(example_condition = ifelse(grepl("as certainly true", question), "true", 
                                    ifelse(grepl("as certainly false", question), "false",
                                           "uncertain")),
         passed_example_trial = case_when(example_condition == "true" ~ response >= 80,
                                    example_condition == "false" ~ response <= 20,
                                    example_condition == "uncertain" ~ response >= 30),
       #  check if all trials passed
         passed_example = case_when(sum(passed_example_trial) == 0 ~ FALSE,
                                    TRUE ~ TRUE)
         ) %>% filter(passed_example == F)

cat("Subjects failing the example trials: ", d_exmpl_fail %>% distinct(submission_id) %>% pull() %>% length())

# apply exclusion criteria to main trials
# check range of responses per participant
d_main_fail <- d_main %>% group_by(submission_id) %>%
  mutate(passed_main = case_when(max(response) - min(response) <= 10 ~ FALSE,
                                 TRUE ~ TRUE)
         ) %>% filter(passed_main == F)
cat(" Subjects providing the same ratings throughout the trials: ", d_main_fail %>% distinct(submission_id) %>% pull() %>% length())

# get participants failing comprehension questions
d_test <- d_test %>%
  group_by(submission_id) %>%
  mutate(passed_filler_trial = case_when(test_condition == "true" ~ response >= 60,
                                   test_condition == "false" ~ response <= 40,
                                   test_condition == "uncertain" ~ response %in% (10:90)),
         mean_comprehension = mean(passed_filler_trial),
         passed_filler = mean_comprehension >= 0.8
         )

d_test_fail <- d_test %>% 
  filter(passed_filler == F)

cat(" Subjects failing the comprehension trials: ", d_test_fail %>% distinct(submission_id) %>% pull() %>% length())
```

Write out data for plotting:
```{r}
d_test %>% write_csv("./../../results_prereg_raw_test.csv")
```

```{r}
# put it all together
d_full_clean <- anti_join(d_main, d_main_fail, by = "submission_id")
d_full_clean <- anti_join(d_full_clean, d_exmpl_fail, d_test, by = "submission_id")
d_full_clean <- anti_join(d_full_clean, d_test_fail, by = "submission_id")

cat(" Nr. of participants left after cleaning: ", d_full_clean %>% distinct(submission_id) %>% pull() %>% length())
```

Write out tidy data:
```{r}
d_full_clean %>% write_csv("./../../data/results_prereg_tidy.csv")
```

```{r clean}
# get overall mean ratings / subject across comprehension and critical trials 
d_full_clean %>% group_by(submission_id) %>% summarise(mean_rating = mean(response)) %>% arrange(mean_rating)

# get mean ratings / subject in critical trials 
d_full_clean %>% filter(condition == "critical") %>% group_by(submission_id) %>% summarise(mean_rating = mean(response)) %>% arrange(mean_rating)

```

## Transformations

```{r, echo=FALSE}
d_critical_clean <- d_full_clean %>% filter(condition == "critical")
d_critical_long <- d_critical_clean %>% 
  pivot_longer(c(competence, relevance, prior), 
               names_to = "class_condition", 
               values_to = "prior_class")

d_critical_long <- d_critical_long %>% 
  mutate(block = ifelse(block == "comp", "competence", 
                        ifelse(block == "rel", "relevance", ifelse(block == "pri", "prior", block) )))
```

Create more extensive condition labels.

``` {r}
# extending 'conditions' labels to include whether the utterance was present or not
d_critical_long <- d_critical_long %>% 
  mutate(
    block_extended =  ifelse(block %in% c("some", "xor"), "target", block)
  )
```

Z-scoring:
```{r}
d_critical_zScored <- d_critical_long %>% group_by(submission_id, block_extended) %>%
  mutate(block_mean = mean(response),
         block_sd = sd(response),
         response_centered = (response - block_mean)/block_sd,
         # catch the cases where sd is 0 
         response_centered = ifelse(is.na(response_centered), 0, response_centered))

d_critical_zScored_wide <- d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response_centered, 
    values_fn = mean # getting means for double prior measurement in "xor"
  ) 
```

Write out z-scored data
```{r}
d_critical_zScored %>% write_csv("./../../data/results_prereg_tidy_zScored_long.csv")
d_critical_zScored_wide %>% write_csv("./../../data/results_prereg_tidy_zScored_wide.csv")
```

### 2021-10-18 analysis/prereg/02_xor-some_prereg_plots.Rmd
refactor prereg R scripts

---
title: 'Xor-Some Preregistered Study: Visualizations'
author: "Polina Tsvilodub"
date: "10/18/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidyboot)
```

Read data created in preprocessing script:
```{r}
d_critical_zScored_wide <- read_csv("./../../data/results_prereg_tidy_zScored_wide.csv")
d_critical_zScored <- read_csv("./../../data/results_prereg_tidy_zScored_long.csv")
d_test_raw <- read_csv("./../../results_prereg_raw_test.csv")
```

## Plots

Comprehension question performance:
```{r}
d_test_ci <- d_test %>% group_by(test_condition) %>% 
  tidyboot_mean(column = response)

d_test %>% 
  ggplot(., aes(x = test_condition, y = response)) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
  geom_point(data = d_test_ci, aes(x = test_condition, y = mean), color = "red", 
             size = 4) +
  facet_wrap(~main_type)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
d_critical_zScored_wide %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  ylab("Inference strength ratings") +
  geom_smooth(method = "lm") +
  facet_wrap(~main_type, ncol = 1) -> p.rel.z

d_critical_zScored_wide %>%
  ggplot(., aes(x = competence, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.comp.z

d_critical_zScored_wide %>%
  ggplot(., aes(x = prior, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.pri.z

gridExtra::grid.arrange(p.rel.z, p.comp.z, p.pri.z, ncol = 3) 
```

```{r}
d_critical_zScored %>% 
  filter(block == "xor" | block == "some") %>%
  ggplot(., aes(x = prior_class, y = response_centered)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_smooth(method="lm") +
  ylab("Inference strength ratings") +
  facet_wrap(block~class_condition) +
  xlab("Anticipated categorization of the items") +
  ggtitle("Inference strength ratings by-predictor")

```

### 2021-10-18 analysis/prereg/03_xor-some_prereg_analysis.Rmd
refactor prereg R scripts

--- 2021-09-14 analysis/xor-some_final.Rmd
@@
 ---
-title: "Xor & Some Final analysis"
+title: "Xor-Some Preregistered Study: Analyses"
 author: "Polina Tsvilodub"
@@
 
-## Read Data
-
-First, the preprocessed and tidy wide pilot data is read and combined. It comprises data from N = 140 subjects after exclusions in total.
-```{r, echo=FALSE}
-pilot1 <- read_csv("./../data/pilots/pilot2_critical_zScore_wide_tidy.csv") %>%
-  select(-competence_wUtt, -relevance_wUtt)
-pilot2 <- read_csv("./../data/pilots/pilot3_critical_zScore_wide_tidy.csv")
-
-d_both <- pilot1 %>% mutate(pilot = 1) %>% 
-  rbind(., pilot2 %>% mutate(pilot = 2)
-        )
-# the data read here should be tidy, in correct format, bad workers should be excluded already
-# z-scored data should be used and simulated!
-
-# separate some and xor data!
-d_both_some <- d_both %>% filter(main_type == "some") 
-d_both_xor  <- d_both %>% filter(main_type == "xor")
-```
-Sanity check the combined data (as can be seen, z-scored data is used):
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-d_both %>%
-  ggplot(., aes(x = relevance, y = target)) +
-  geom_point(size = 2, alpha = 0.7) +
-  geom_smooth(method = "lm") +
-  ylab("Inference strength ratings") +
-  facet_wrap(~main_type, ncol = 1) -> p.rel
-
-d_both %>%
-  ggplot(., aes(x = competence, y = target)) +
-  geom_point(size = 2, alpha = 0.7) +
-  geom_smooth(method = "lm") +
-  ylab("") +
-  facet_wrap(~main_type, ncol = 1) -> p.comp
-
-d_both %>%
-  ggplot(., aes(x = prior, y = target)) +
-  geom_point(size = 2, alpha = 0.7) +
-  geom_smooth(method = "lm") +
-  ylab("") +
-  facet_wrap(~main_type, ncol = 1) -> p.pri
-
-gridExtra::grid.arrange(p.rel, p.comp, p.pri, ncol = 3) 
+Read data created in preprocessing script:
+```{r}
+d_critical_zScored_wide <- read_csv("./../../data/results_prereg_tidy_zScored_wide.csv")
+d_critical_zScored <- read_csv("./../../data/results_prereg_tidy_zScored_long.csv")
 ```
 
-## Model
+## Main analysis
 
@@
 # Full model 
-d_both <- d_both %>% mutate(
+d_critical_zScored_wide <- d_critical_zScored_wide %>% mutate(
   main_type = as.factor(main_type)
@@
 # some = 0, xor = 1
-contrasts(d_both$main_type) 
-# try to fit maximal model with brm
+contrasts(d_critical_zScored_wide$main_type)
+
+# set priors
+priors <- set_prior("student_t(1, 0, 2)", class = "b")
+
 model_SI <- brm(target ~ prior*competence*relevance* main_type +
-                   (1 + prior + competence + relevance + main_type || submission_id) +
+                  # (1 + prior + competence + relevance + main_type || submission_id) +
                    (1 | title),
-                 data = d_both,
+                 data = d_critical_zScored_wide,
+                 prior = priors,
+                 sample_prior = T,
                  control = list(adapt_delta = 0.95),
@@
 ```
-
-## Extract conrtrasts of interest
 
@@
 
-## Perform BF analysis
+#### Perform BF analysis
 
@@
 
-## Experimenting with multiple regression
+## Exploratory analyses
+
+Model with binary predictors:
+```{r}
+d_critical_zScored %>% 
+  select(submission_id, title, main_type, block_extended, response_centered, prior_class, class_condition) %>% 
+  unique() %>% 
+  pivot_wider(
+    names_from = class_condition, 
+    values_from = prior_class
+  )  %>%
+  mutate(relevance = factor(relevance, levels = c(1, 0)),
+         competence = factor(competence, levels = c(1, 0)),
+         prior = factor(prior, levels = c(1, 0)),
+         main_type = as.factor(main_type)) %>% 
+  filter(block_extended == "target") -> d_critical_wide_cat_zScored
+# sum code predictors
+# low prior: -1, high prior:: 1 
+contrasts(d_critical_wide_cat_zScored$prior) <- contr.sum(2)
+# low comp : -1, high comp : 1
+contrasts(d_critical_wide_cat_zScored$competence) <- contr.sum(2)
+# low rel: -1, high rel: 1
+contrasts(d_critical_wide_cat_zScored$relevance) <- contr.sum(2)
+# leave main type dummy coded
+contrasts(d_critical_wide_cat_zScored$main_type)
+
+# xor, maximal model with interactions and maximal REs
+model_xor_cat_zScored <- brm(
+  response_centered ~ prior*competence*relevance*main_type + 
+    (1 + prior + competence + relevance + main_type || submission_id) +
+    (1 | title),
+  data = d_critical_wide_cat_zScored,
+  prior = priors,
+  sample_prior = T,
+  family = "gaussian",
+  cores = 4,
+  iter = 3000
+)
+
+```
+```{r}
+summary(model_xor_cat_zScored)
+```
+
+Model with aggregated by-story ratings:
+```{r, results='hide'}
+d_critical_zScored_wide_byItem <- d_critical_zScored_wide %>% ungroup() %>%
+  group_by(title, main_type) %>%
+  summarize(mean_response = mean(target),
+            mean_prior = mean(prior),
+            mean_relevance = mean(relevance),
+            mean_competence = mean(competence))
+
+model_means <- brm(mean_response ~ mean_prior + mean_relevance + mean_competence + main_type,
+                       data = d_critical_zScored_wide_byItem,
+                       prior = priors,
+                       sample_prior = T,
+                       family = "gaussian",
+                       cores = 4,
+                       iter = 3000
+                       )
+```
 
 ```{r}
-dummy_data <- tribble(
-  ~main, ~REL, ~target,
-  "some", 0, 4,
-  "some", 1, 6,
-  "xor", 0, 0,
-  "xor", 1, 12
-)
-# check that the slope of "REL" for "xor" is 12
-# this is calculated as betat_REL + beta_mainxor:REL
-# importantly: the intercept adjustment beta_mainxor doesn't come in!
-lm(target ~ main * REL, dummy_data)
+summary(model_means)
 ```

### 2021-10-19 04_trials_final.js
make first trial a critical one, update instructions, Bobs comments

@@ -61,7 +61,7 @@ let attention_check_data = [
   },
   {
     title: 'Homework',
-    QUD: "Mr. Carpenter teaches at a university and gave his students a homework assignment last week. Please click on the slider handle and leave it in the middle. The students are supposed to submit their homework by putting it in a dedicated post box. Mr. Carpenter collects the submitted assignments.",
+    QUD: "Mr. Carpenter teaches at a university. He gave his students a homework assignment last week. Please click on the slider handle and leave it in the middle. The students are supposed to submit their homework by putting it in a dedicated post box. Mr. Carpenter collects the submitted assignments.",
     optionLeft: '<p style="font-family: sans-serif">certainly false</p>',
     optionRight: '<p style="font-family: sans-serif">certainly true</p>',
     condition: "attention_check",
@@ -1567,7 +1567,7 @@ let vignettes_some = {
    "relevance": 0,
    "competence": 1,
    "prior": 0,
-   "background": "Felix is about to go shopping clothes. His roommate Thomas asks him whether his tiny closet isn't overcrowded enough already.",
+   "background": "Felix is about to go shopping for clothes. His roommate Thomas asks him whether his tiny closet isn't overcrowded enough already.",
    "utterance_imp": "Felix says to Thomas: 'Some of my socks already have holes.'",
    "question_rel": "It is important for Thomas to know whether all of Felix' socks have holes.",
    "question_comp": "Felix knows whether all of his socks have holes.",
@@ -1646,8 +1646,8 @@ let vignettes_some = {
    "question_comp": "Simone knows whether Alex has developed all of the pictures from the picnic.",
    "question_pri": "Alex has developed all of the pictures from the picnic.",
    "question_priC": "If Alex has developed at least some of the pictures from the picnic, then he has developed all of the pictures from the picnic.",
-   "question_imp": "From what Simone said we may conclude that Alex has not develop all of the pictures from the picnic.",
-   "question_impI": "Simone wants to convey that Alex has not develop all of the pictures from the picnic.",
+   "question_imp": "From what Simone said we may conclude that Alex has not developed all of the pictures from the picnic.",
+   "question_impI": "Simone wants to convey that Alex has not developed all of the pictures from the picnic.",
    "test_true2": "Beth and Simone organized a picnic.",
    "test_false2": "The weather during the picnic was awful.",
    "test_uncertain2": "Simone brought fruit for the picnic.",

### 2021-10-19 05_views.js

@@ -53,8 +53,10 @@ const instructions = magpieViews.view_generator("instructions", {
   text: `In the following, you will be presented with <b> 16 short stories</b>.
             Please read them very carefully, even if they appear to be repeated and you think that you remember them well enough.
             We ask you to <b> rate statements about each short story</b>.
-            Please indicate, using an adjustable slider, <b> how likely you think a statement
+            Please indicate, using an adjustable slider, <b>how likely it is that a statement
             is true based on what you've read.</b>
+            <br />
+            Notice that there will also be simple attention checking trials. You will recognize them immediately when you read the important text on each trial carefully - those trials contain instructions for you to move the slider in a certain way. Please follow those instructions.
             <br />
             <br />
             The story is given at the top.
@@ -203,11 +205,18 @@ const ex_trial = magpieViews.view_generator("slider_rating", {
 );
 
 const main_trial_data = main_block.concat(attention_check_data)
+//console.log(typeof(main_trial_data[0]))
+var main_trials_shuffled = _.shuffle(main_trial_data)
+// make sure attention checks don't appear as first trial
+while(Array.isArray(main_trials_shuffled[0]) == false){
+  main_trials_shuffled = _.shuffle(main_trials_shuffled)
+}
+
 const main_trials = magpieViews.view_generator("slider_rating", {
-  trials: attention_check_data.length + _.flatten(main_block).length,
+  trials: _.flatten(main_trials_shuffled).length,
   name: "main_trials",
   trial_type: "main_slider",
-  data: _.flatten(_.shuffle(main_trial_data))
+  data: _.flatten(main_trials_shuffled)
 },
 { // custom view allowing to add custom backgrounds of different view elements
   stimulus_container_generator: function (config, CT) {

### 2021-10-22 06_main.js
set deployment to prolific

@@ -27,14 +27,14 @@ $("document")
 
       // Here, you can specify all information for the deployment
       deploy: {
-        experimentID: "63",
+        experimentID: "73",
         serverAppURL: "https://mcmpact.ikw.uni-osnabrueck.de/magpie/api/submit_experiment/",
         // Possible deployment methods are:
         // "debug" and "directLink"
         // As well as "MTurk", "MTurkSandbox" and "Prolific"
-        deployMethod: "debug",
+        deployMethod: "Prolific",
         contact_email: "polina.tsvilodub@gmail.com",
-        prolificURL: "https://app.prolific.co/submissions/complete?cc=7E05F86F"
+        prolificURL: "https://app.prolific.co/submissions/complete?cc=40C398B5"
       },
       // Here, you can specify how the progress bar should look like
       progress_bar: {

### 2021-10-22 04_trials_final.js
make attention check prompts bold

@@ -6,7 +6,7 @@ let attention_check_data = [
     optionRight: '<p style="font-family: sans-serif">certainly true</p>',
     condition: "attention_check",
     critical_question: ``,
-    prompt: "John really liked that singer.",
+    prompt: "<b>John really liked that singer.</b>",
     question: '<font size="3"> How likely is it that the statement is true given the story?</font>',
     block: 'attention_check',
     expected: 100
@@ -18,7 +18,7 @@ let attention_check_data = [
     optionRight: '<p style="font-family: sans-serif">certainly true</p>',
     condition: "attention_check",
     critical_question: ``,
-    prompt: "Five people were drinking beer at the bar.",
+    prompt: "<b>Five people were drinking beer at the bar.</b>",
     question: '<font size="3"> How likely is it that the statement is true given the story?</font>',
     block: 'attention_check',
     expected: 100
@@ -30,7 +30,7 @@ let attention_check_data = [
     optionRight: '<p style="font-family: sans-serif">certainly true</p>',
     condition: "attention_check",
     critical_question: ``,
-    prompt: "The restaurant owner had to pay a fine because some food contained bacteria.",
+    prompt: "<b>The restaurant owner had to pay a fine because some food contained bacteria.</b>",
     question: '<font size="3"> How likely is it that the statement is true given the story?</font>',
     block: 'attention_check',
     expected: 0
@@ -42,7 +42,7 @@ let attention_check_data = [
     optionRight: '<p style="font-family: sans-serif">certainly true</p>',
     condition: "attention_check",
     critical_question: ``,
-    prompt: "Freddy liked a little white kitten the most.",
+    prompt: "<b>Freddy liked a little white kitten the most.</b>",
     question: '<font size="3"> How likely is it that the statement is true given the story?</font>',
     block: 'attention_check',
     expected: 0
@@ -54,7 +54,7 @@ let attention_check_data = [
     optionRight: '<p style="font-family: sans-serif">certainly true</p>',
     condition: "attention_check",
     critical_question: ``,
-    prompt: "Greg Jones secured victory for his team during the last seconds of a decisive games.",
+    prompt: "<b>Greg Jones secured victory for his team during the last seconds of a decisive games.</b>",
     question: '<font size="3"> How likely is it that the statement is true given the story?</font>',
     block: 'attention_check',
     expected: 50
@@ -66,7 +66,7 @@ let attention_check_data = [
[…]

### 2021-10-22 analysis/01_xor-some_prereg_preprocessing_N18.Rmd
add preprocessing of first N18

--- 2021-10-18 analysis/prereg/01_xor-some_prereg_preprocessing.Rmd
@@
 ```{r, echo=FALSE}
-d <- read_csv("./../../data/results_test.csv") #prereg_final
-
-# for testing purposes
-d <- d %>% mutate(submission_id = 42)
+d <- read_csv("./../data/main/results_73_xor-some-Prolific-main_N18_anonym.csv")
 
@@
   filter(timeSpent > 8) 
+```
+
+```{r}
+# check the distribution of completion times
+d_native_completed %>% 
+  ggplot(., aes(x=timeSpent)) +
+  geom_density()
 ```
@@
 Participants who failed all example trials are excluded, as well.
+
 ```{r exclusions}
@@
 ```{r}
-d_test %>% write_csv("./../../results_prereg_raw_test.csv")
+d_test %>% write_csv("./../data/main/results_prereg_raw_N18_test.csv")
 ```
@@
 ```{r}
-d_full_clean %>% write_csv("./../../data/results_prereg_tidy.csv")
+d_full_clean %>% write_csv("./../data/main/results_prereg_N18_tidy.csv")
 ```
@@
 ```{r}
-d_critical_zScored %>% write_csv("./../../data/results_prereg_tidy_zScored_long.csv")
-d_critical_zScored_wide %>% write_csv("./../../data/results_prereg_tidy_zScored_wide.csv")
+d_critical_zScored %>% write_csv("./../data/main/results_prereg_tidy_N18_zScored_long.csv")
+d_critical_zScored_wide %>% write_csv("./../data/main/results_prereg_tidy_N18_zScored_wide.csv")
 ```

### 2021-10-23 analysis/main/01_xor-some_prereg_preprocessing_N158.Rmd
refactor repo, preprocess N158

--- 2021-10-18 analysis/prereg/01_xor-some_prereg_preprocessing.Rmd
@@
 ```{r, echo=FALSE}
-d <- read_csv("./../../data/results_test.csv") #prereg_final
-
-# for testing purposes
-d <- d %>% mutate(submission_id = 42)
+d <- read_csv("./../../data/main/results_73_xor-some-Prolific-main_N158_anonym.csv")
 
@@
 cat(" Mean age: ", d_native %>% pull(age) %>% mean(., na.rm = T) )
-d_native  %>% count(gender) %>% mutate(n = n/72)
+d_native  %>% count(gender) %>% mutate(n = n/80)
 ```
@@
   filter(timeSpent > 8)
+
+cat(" Number of participants who completed the expt in > 8 minutes: ", d_native_completed %>% distinct(submission_id) %>% count() %>% pull() )
+```
+
+```{r}
+d_native_completed %>% summarize(median_time = median(timeSpent),
+                                 mean_time = mean(timeSpent))
+```
+```{r}
+# check the distribution of completion times
+d_native_completed %>% 
+  ggplot(., aes(x=timeSpent)) +
+  geom_density()
 ```
@@
 
-d_native_attended <- left_join(d_native_completed, 
-                               d_native_attention_summary, 
+d_native_attention_failed <- d_native_attention %>% 
+  group_by(submission_id) %>%
+  summarize(attention_prop = mean(passed_attention_trial)) %>%
+  mutate(
+    passed_attention_check = attention_prop > 0.75
+  ) %>% 
+  filter(passed_attention_check == FALSE)
+
+d_native_attended <- anti_join(d_native_completed, 
+                               d_native_attention_failed, 
                                by = c("submission_id"))
 
+cat(" Number of participants failing attention checks: ", d_native_attention_failed %>% distinct(submission_id) %>% count() %>% pull() )
 ```
@@
 Data frames preparation:
+
 ```{r}
@@
 ```{r}
-d_test %>% write_csv("./../../results_prereg_raw_test.csv")
+d_test %>% write_csv("./../../data/main/results_prereg_raw_N158_test.csv")
 ```
@@
 ```{r}
-d_full_clean %>% write_csv("./../../data/results_prereg_tidy.csv")
+d_full_clean %>% write_csv("./../../data/main/results_prereg_N158_tidy.csv")
 ```
@@
 ```{r}
-d_critical_zScored %>% write_csv("./../../data/results_prereg_tidy_zScored_long.csv")
-d_critical_zScored_wide %>% write_csv("./../../data/results_prereg_tidy_zScored_wide.csv")
+d_critical_zScored %>% write_csv("./../../data/main/results_prereg_tidy_N158_zScored_long.csv")
+d_critical_zScored_wide %>% write_csv("./../../data/main/results_prereg_tidy_N158_zScored_wide.csv")
 ```

### 2021-10-23 analysis/main/01_xor-some_prereg_preprocessing_N18.Rmd
refactor repo, preprocess N158

--- 2021-10-23 analysis/main/01_xor-some_prereg_preprocessing_N158.Rmd
@@
 ```{r, echo=FALSE}
-d <- read_csv("./../../data/main/results_73_xor-some-Prolific-main_N158_anonym.csv")
+d <- read_csv("./../../data/main/results_73_xor-some-Prolific-main_N18_anonym.csv")
 
@@
 ```{r}
-d_test %>% write_csv("./../../data/main/results_prereg_raw_N158_test.csv")
+d_test %>% write_csv("./../../data/main/results_prereg_raw_N18_test.csv")
 ```
@@
 ```{r}
-d_full_clean %>% write_csv("./../../data/main/results_prereg_N158_tidy.csv")
+d_full_clean %>% write_csv("./../../data/main/results_prereg_N18_tidy.csv")
 ```
@@
 ```{r}
-d_critical_zScored %>% write_csv("./../../data/main/results_prereg_tidy_N158_zScored_long.csv")
-d_critical_zScored_wide %>% write_csv("./../../data/main/results_prereg_tidy_N158_zScored_wide.csv")
+d_critical_zScored %>% write_csv("./../../data/main/results_prereg_tidy_N18_zScored_long.csv")
+d_critical_zScored_wide %>% write_csv("./../../data/main/results_prereg_tidy_N18_zScored_wide.csv")
 ```

### 2021-10-23 analysis/prereg/02_xor-some_prereg_plots.Rmd
refactor repo, preprocess N158

@@
 ```{r}
-d_critical_zScored_wide <- read_csv("./../../data/results_prereg_tidy_zScored_wide.csv")
-d_critical_zScored <- read_csv("./../../data/results_prereg_tidy_zScored_long.csv")
-d_test_raw <- read_csv("./../../results_prereg_raw_test.csv")
+d_critical_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_N18_zScored_wide.csv")
+d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_N18_zScored_long.csv")
+d_test_raw <- read_csv("./../../data/main/results_prereg_raw_N18_test.csv")
 ```

### 2021-10-24 analysis/main/01_xor-some_prereg_preprocessing_N233.Rmd
preprocess N233 subjects

--- 2021-10-23 analysis/main/01_xor-some_prereg_preprocessing_N18.Rmd
@@
 ```{r, echo=FALSE}
-d <- read_csv("./../../data/main/results_73_xor-some-Prolific-main_N18_anonym.csv")
+d <- read_csv("./../../data/main/results_73_xor-some-Prolific-main_N233_anonym.csv")
 
@@
 ```{r}
-d_test %>% write_csv("./../../data/main/results_prereg_raw_N18_test.csv")
+d_test %>% write_csv("./../../data/main/results_prereg_raw_N233_test.csv")
 ```
@@
 ```{r}
-d_full_clean %>% write_csv("./../../data/main/results_prereg_N18_tidy.csv")
+d_full_clean %>% write_csv("./../../data/main/results_prereg_N233_tidy.csv")
 ```
@@
 ```{r}
-d_critical_zScored %>% write_csv("./../../data/main/results_prereg_tidy_N18_zScored_long.csv")
-d_critical_zScored_wide %>% write_csv("./../../data/main/results_prereg_tidy_N18_zScored_wide.csv")
+d_critical_zScored %>% write_csv("./../../data/main/results_prereg_tidy_N233_zScored_long.csv")
+d_critical_zScored_wide %>% write_csv("./../../data/main/results_prereg_tidy_N233_zScored_wide.csv")
 ```

### 2021-10-31 analysis/main/01_xor-some_prereg_preprocessing_final.Rmd
add final N200 analysis

--- 2021-10-23 analysis/main/01_xor-some_prereg_preprocessing_N18.Rmd
@@
 First, the data is loaded.
-```{r, echo=FALSE}
-d <- read_csv("./../../data/main/results_73_xor-some-Prolific-main_N18_anonym.csv")
+```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
+d <- read_csv("./../../data/main/results_73_xor-some-Prolific-main_N275_anonym.csv")
 
@@
 cat(" Number of participants failing attention checks: ", d_native_attention_failed %>% distinct(submission_id) %>% count() %>% pull() )
+
+#d_attention_fail_IDs <- right_join(d_raw, d_native_attention_failed, by = c("submission_id")) %>% select(submission_id, prolific_id) %>% distinct()
+#d_attention_fail_IDs %>% write_csv("~/Documents/Research/XOR/raw_data/xor-some_prereg_failed_attention_prolificIDs_final.csv")
 ```
@@
 ```{r}
-d_test %>% write_csv("./../../data/main/results_prereg_raw_N18_test.csv")
+d_test %>% write_csv("./../../data/main/results_prereg_raw_final_test.csv")
 ```
@@
 ```
+Exclude the 201st participant for target N=200:
+```{r, results='hide'}
+d_full_clean %>% arrange(., desc(submission_id)) %>% distinct() %>% pull(submission_id)
+
+# exclude participant with last submission_id
+d_full_clean <- d_full_clean %>% filter(submission_id != 2620)
+
+cat(" Nr. of participants left after cleaning and N=200 extraction: ", d_full_clean %>% distinct(submission_id) %>% pull() %>% length())
+
+```
 
@@
 ```{r}
-d_full_clean %>% write_csv("./../../data/main/results_prereg_N18_tidy.csv")
+d_full_clean %>% write_csv("./../../data/main/results_prereg_final_tidy.csv")
 ```
@@
 ```{r}
-d_critical_zScored %>% write_csv("./../../data/main/results_prereg_tidy_N18_zScored_long.csv")
-d_critical_zScored_wide %>% write_csv("./../../data/main/results_prereg_tidy_N18_zScored_wide.csv")
+d_critical_zScored %>% write_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")
+d_critical_zScored_wide %>% write_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
 ```

### 2021-10-31 analysis/main/02_xor-some_prereg_plots_final.Rmd
add final N200 analysis

--- 2021-10-18 analysis/prereg/02_xor-some_prereg_plots.Rmd
@@
 ```{r}
-d_critical_zScored_wide <- read_csv("./../../data/results_prereg_tidy_zScored_wide.csv")
-d_critical_zScored <- read_csv("./../../data/results_prereg_tidy_zScored_long.csv")
-d_test_raw <- read_csv("./../../results_prereg_raw_test.csv")
+d_critical_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
+d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")
+d_test <- read_csv("./../../data/main/results_prereg_raw_final_test.csv")
 ```

### 2021-10-31 analysis/main/03_xor-some_prereg_analysis_final.Rmd
add final N200 analysis

--- 2021-10-18 analysis/prereg/03_xor-some_prereg_analysis.Rmd
@@
 ```{r}
-d_critical_zScored_wide <- read_csv("./../../data/results_prereg_tidy_zScored_wide.csv")
-d_critical_zScored <- read_csv("./../../data/results_prereg_tidy_zScored_long.csv")
+d_critical_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
+d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")
 ```
@@
 model_SI <- brm(target ~ prior*competence*relevance* main_type +
-                  # (1 + prior + competence + relevance + main_type || submission_id) +
+                   (1 + prior + competence + relevance + main_type || submission_id) +
                    (1 | title),
@@
 Model with binary predictors:
-```{r}
+```{r, results='hide', message=FALSE, warning=FALSE}
 d_critical_zScored %>% 
@@
 ```
+Now extract posterior contrasts of interest from categorical model: effects of each predictor \{prior, competence and relevance\} for each trigger.
+
+```{r}
+model_xor_cat_zScored %>% spread_draws(b_Intercept, b_prior1, b_competence1, b_relevance1,
+                          b_main_typexor, `b_prior1:competence1`,
+                          `b_prior1:relevance1`, `b_competence1:relevance1`,
+                          `b_prior1:main_typexor`, `b_competence1:main_typexor`,
+                          `b_relevance1:main_typexor`, `b_prior1:competence1:relevance1`, 
+                          `b_prior1:competence1:main_typexor`, `b_prior1:relevance1:main_typexor`,
+                          `b_competence1:relevance1:main_typexor`, 
+                          `b_prior1:competence1:relevance1:main_typexor`) %>% 
+  mutate(
+    prior_xor = 2 * b_prior1 + 2* `b_prior1:main_typexor`,
+    prior_some = 2 * b_prior1,
+    competence_xor = 2 * b_competence1 + 2* `b_competence1:main_typexor`,
+    competence_some = 2 * b_competence1,
+    relevance_xor =  2 * b_relevance1 + 2* `b_relevance1:main_typexor`,
+    relevance_some = 2* b_relevance1
+  ) -> model_xor_cat_zScored_posteriors
+
+# check P that the effects are positive / negative / no effect present
+posterior_hypotheses_cat <- model_xor_cat_zScored_posteriors %>% 
+  select(prior_xor, prior_some, 
+         competence_xor, competence_some,
+         relevance_xor, relevance_some) %>%
+  gather(key, val) %>%
+  group_by(key) %>% mutate(positive = mean(val > 0.05),
+                           negative = mean(val < -0.05),
+                           no = mean(val %>% between(-0.05, 0.05))) %>%
+  summarise(positive_eff = mean(positive),
+            negative_eff = mean(negative),
+            no_eff = mean(no))
+posterior_hypotheses_cat
+```
+```{r}
+test_conjunction_of_all_hypotheses(posterior_hypotheses_cat)
+```
 
 Model with aggregated by-story ratings:
-```{r, results='hide'}
+```{r, results='hide', message=FALSE, warning=FALSE}
 d_critical_zScored_wide_byItem <- d_critical_zScored_wide %>% ungroup() %>%
@@
 ```
+
+Extract posterior contrasts of interest from by-story mean based model: effects of each predictor \{prior, competence and relevance\} for each trigger.
+
+```{r}
+model_means %>% spread_draws(b_Intercept, b_mean_prior, b_mean_competence, b_mean_relevance,
+                          b_main_typexor) %>% 
+  mutate(
+    prior_xor = b_mean_prior,
+    prior_some = b_mean_prior,
+    competence_xor = b_mean_competence,
+    competence_some = b_mean_competence,
+    relevance_xor =  b_mean_relevance,
+    relevance_some = b_mean_relevance
+  ) -> model_means_posteriors
+
+# check P that the effects are positive / negative / no effect present
+posterior_hypotheses_means <- model_means_posteriors %>% 
+  select(prior_xor, prior_some, 
+         competence_xor, competence_some,
+         relevance_xor, relevance_some) %>%
+  gather(key, val) %>%
+  group_by(key) %>% mutate(positive = mean(val > 0.05),
+                           negative = mean(val < -0.05),
+                           no = mean(val %>% between(-0.05, 0.05))) %>%
+  summarise(positive_eff = mean(positive),
+            negative_eff = mean(negative),
+            no_eff = mean(no))
+posterior_hypotheses_means
+```
+
+```{r}
+test_conjunction_of_all_hypotheses(posterior_hypotheses_means)
+```

### 2021-11-06 analysis/main/01_xor-some_prereg_preprocessing_final.Rmd
fix lang exclusion, check comprehension questions

@@
  # filter(("en" | "En" | "bri") %in% languages)
-  filter(grepl("[(en)(br)]", languages, ignore.case = T))
+  filter(grepl("eng", languages, ignore.case = T))
 
@@
                                         test_condition)))
+```
+
+Investigate if the comprehension questions received expected responses on by-question basis:
+```{r}
+d_test %>% group_by(ID, test_question) %>% 
+  mutate(
+    expected = case_when( test_condition == "true" ~ 100,
+                          test_condition == "false" ~ 0,
+                          TRUE ~ 50)
+  ) %>%
+  summarise(
+    expected = mean(expected),
+    response = mean(response),
+    difference = abs(expected - response)
+  ) %>% 
+  arrange(., desc(difference))
+
+# look at the vignette text and the qeustion of the worst performing item
+d_test %>% filter(ID == 64, test_question == "test_false1") %>% select(prompt, QUD) %>% View()
 ```

### 2021-11-07 analysis/main/02_xor-some_prereg_plots_final.Rmd
add by item, posterior plots. fix analysis

--- 2021-10-18 analysis/prereg/02_xor-some_prereg_plots.Rmd
@@
 Read data created in preprocessing script:
-```{r}
-d_critical_zScored_wide <- read_csv("./../../data/results_prereg_tidy_zScored_wide.csv")
-d_critical_zScored <- read_csv("./../../data/results_prereg_tidy_zScored_long.csv")
-d_test_raw <- read_csv("./../../results_prereg_raw_test.csv")
+```{r, warning=FALSE, message=FALSE, results='hide'}
+d_critical_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
+d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")
+d_test <- read_csv("./../../data/main/results_prereg_raw_final_test.csv")
 ```
@@
 
+Plot predictor ratings against inference strength ratings by trigger:
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 ```
+
+```{r, fig.height=30, fig.width=10}
+# plot by-item ratings vs our prior categorizations
+d_critical_long_grandMeans <- d_critical_zScored %>% 
+  mutate(title = paste(title, exp_condition, sep = "_"),
+         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
+  filter(block == class_condition | (block == "target" & class_condition == "relevance")) %>%
+  group_by(main_type, block) %>% 
+  summarize(mean_resp = mean(response))
+
+d_critical_long_expectedCats <- d_critical_zScored %>% 
+  mutate(title = paste(title, exp_condition, sep = "_"),
+         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
+  filter(block == class_condition | (block == "target" & class_condition == "relevance")) %>%
+  group_by(main_type, block, title) %>% 
+  summarize(mean_resp_block = mean(response))
+
+d_critical_zScored %>% 
+  mutate(title = paste(title, exp_condition, sep = "_"),
+         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
+  left_join(., d_critical_long_grandMeans, by = c("main_type", "block")) %>% 
+  filter(block == class_condition | ((block == "target") & (class_condition == "relevance"))) -> d_long_plot_data
+
+left_join(d_long_plot_data, d_critical_long_expectedCats, by = c("main_type", "block", "title")) %>% 
+  mutate(expected_block = ifelse(block != "target", prior_class * 100, NA),
+         block = factor(block, levels = c("relevance", "competence", "prior", "target"))) -> d_long_plot_full
+
+d_long_plot_full %>%
+  ggplot(., aes(x = block, y = response, color = block)) +
+  geom_point(alpha = 0.7, position = position_jitter(width = 0.1)) +
+  geom_hline(aes(yintercept = mean_resp), alpha = 0.7) +
+  geom_point(aes(y = mean_resp), color = "black", shape = 2) +
+  geom_point(aes(y = mean_resp_block, fill = block), color = "black") +
+  geom_point(aes(y = expected_block), color="red", size = 2.5, na.rm = T) +
+  ylab("Raw responses to respective questions") +
+  facet_wrap(main_type~title, ncol = 4) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  ggtitle("By-item by-question raw ratings\nBlack dots indicate by-block means within-item, lines and triangles indicate grand by-block means\nRed dots indicate expected rating")
+
+```
+
+Plot the prior ratings on the two prior questions in the XOr condition in terms of question order:
+```{r}
+# check whether the ratings in the two prior questions are the same in xor 
+d_critical_zScored %>% 
+  filter(block != "xor" & block != "some") %>%
+  filter(block == class_condition) %>%
+  filter(block == "prior", main_type == "xor") %>%
+  mutate(priorQ_nr = rep(c(1,2), nrow(.)/2),
+         unique_ID = paste(submission_id, ID, sep="_")) -> d_xor_priors
+
+d_xor_priors %>% group_by(priorQ_nr, prior_class) %>%
+  summarise(mean = mean(response)) -> d_xor_priors_summary
+
+d_xor_priors %>%
+  ggplot(., aes(x = as.factor(priorQ_nr), y = response )) +
+  geom_point(size = 2, alpha = 0.6) +
+  geom_point(data = d_xor_priors_summary, aes(x = as.factor(priorQ_nr), y = mean), color = "red", size = 3) +
+  geom_path(aes(group = "unique_ID"), alpha = 0.6) +
+  ylab("Responses to prior questions") +
+  xlab("First vs Second prior question for high vs low prior conditions") +
+  facet_wrap(~as.factor(prior_class)) + # get ratings from the respective trials only 
+  ggtitle("Xor prior question ratings.\nLines indicate the two responses provided by one participant for a specific item.")
+
+```
+
+Plot the prior ratings on the two prior questions in the XOr condition in terms of the specific by-item question (question 1 vs question 2 for a given item, regardless of presentation order):
+```{r}
+# create labels for the two questions
+d_xor_priors %>% group_by(ID, prompt) %>%
+  select(ID, prompt) %>% distinct() %>%
+  ungroup() %>%
+  mutate(
+    priorQ_label = rep(c("prior1", "prior2"), 32)
+  ) -> d_xor_priors_labels
+left_join(d_xor_priors, d_xor_priors_labels, by = c("prompt")) %>%
+  group_by(priorQ_label, prior_class) %>%
+  summarise(mean = mean(response)) -> d_xor_priors_labels_summary
+
+# add the label and plot
+left_join(d_xor_priors, d_xor_priors_labels, by = c("prompt")) %>%
+  ggplot(., aes(x = as.factor(priorQ_label), y = response )) +
+  geom_point(size = 2, alpha = 0.6) +
+  geom_point(data = d_xor_priors_labels_summary, aes(x = as.factor(priorQ_label), y = mean), color = "red", size = 3) +
+  geom_path(aes(group = "unique_ID"), alpha = 0.6) +
+  ylab("Responses to prior questions") +
+  xlab("First vs Second prior question for high vs low prior conditions") +
+  facet_wrap(~as.factor(prior_class)) + # get ratings from the respective trials only 
+  ggtitle("Xor prior question ratings.\nLines indicate the two responses provided by one participant for a specific item.")
+```
+
+Plot inference strength ratings against predictor ratings, by-item, using by-item mean ratings across participants.
+```{r, fig.height=10, fig.width=8}
+d_critical_predictorsXtarget_summary <-  d_critical_zScored_wide %>% 
+  pivot_longer(
+    cols = c(relevance, prior, competence),
+    names_to = "predictors",
+    values_to = "predictor_response"
+  ) %>% 
+  group_by(title, predictors, main_type) %>%
+  summarise(mean_target = mean(target),
+            mean_predictor = mean(predictor_response), 
+            sd_target = sd(target),
+            sd_predictor = sd(predictor_response))  
+# get the prior classifications of the vignettes for convenient coloring
+d_critical_zScored %>% select(title, class_condition, prior_class) %>%
+  distinct() -> d_critical_prior_classes
+
+left_join(d_critical_predictorsXtarget_summary, d_critical_prior_classes, by = c("title", "predictors" = "class_condition")) %>%
+  mutate(prior_class = as.factor(prior_class)) %>%
+  ggplot(., aes(x = mean_predictor, y = mean_target)) +
+  geom_point(aes(color = prior_class)) +
+  ggrepel::geom_text_repel(aes(label=title), size = 2.5) +
+  geom_smooth(method = "lm") +
+  ylab("Mean inference strength rating") +
+  xlab("Mean predictor ratings") + 
+  facet_wrap(main_type~predictors) +
+  scale_fill_manual(name = "Prior classification of item")
+```

### 2021-11-07 analysis/main/03_xor-some_prereg_analysis_final.Rmd
add by item, posterior plots. fix analysis

@@
 Read data created in preprocessing script:
-```{r}
+```{r, message=FALSE, warning=FALSE, results='hide'}
 d_critical_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
@@
 ```
+### Plot posterior distributions
+
+```{r, fig.height = 6, fig.width=9}
+model_SI_posteriors %>% select(prior_xor, prior_some,
+                               competence_xor, competence_some,
+                               relevance_xor, relevance_some) %>%
+  pivot_longer(cols = everything(), names_to = "effect", values_to = "value") -> model_SI_posteriors_long
+
+model_SI_posteriors_long %>%
+  ggplot(.) + # 
+  stat_halfeye(aes(y = effect, x = value, 
+                   fill = stat(case_when(x %>% between(-0.05, 0.05) ~ "no effect", x < -0.05 ~ "negative effect", x > 0.05 ~ "positive effect"))
+                   ), 
+               alpha = .7) +
+  geom_vline(xintercept = c(-.05, .05), linetype = "dashed") +
+  scale_fill_manual(name = "Direction of effect", values = c("coral", "yellow3", "mediumseagreen")) +
+  ggtitle("Posterior effect distributions. Points indicate posterior means\nThick lines indicate 66% CrIs, thin lines 95% CrIs.")
+
+```
 
@@
 
+```{r, fig.height = 6, fig.width=9}
+model_xor_cat_zScored_posteriors %>% select(prior_xor, prior_some,
+                               competence_xor, competence_some,
+                               relevance_xor, relevance_some) %>%
+  pivot_longer(cols = everything(), names_to = "effect", values_to = "value") -> model_xor_cat_posteriors_long
+
+model_xor_cat_posteriors_long %>%
+  ggplot(.) + # 
+  stat_halfeye(aes(y = effect, x = value, 
+                   fill = stat(case_when(x %>% between(-0.05, 0.05) ~ "no effect", x < -0.05 ~ "negative effect", x > 0.05 ~ "positive effect"))
+                   ), 
+               alpha = .7) +
+  geom_vline(xintercept = c(-.05, .05), linetype = "dashed") +
+  scale_fill_manual(name = "Direction of effect", values = c("coral", "yellow3", "mediumseagreen")) +
+  ggtitle("Posterior effect distributions. Points indicate posterior means\nThick lines indicate 66% CrIs, thin lines 95% CrIs.")
+
+```
+
 Model with aggregated by-story ratings:
@@
 
-model_means <- brm(mean_response ~ mean_prior + mean_relevance + mean_competence + main_type,
+model_means <- brm(mean_response ~ mean_prior * mean_relevance * mean_competence * main_type,
                        data = d_critical_zScored_wide_byItem,
@@
 model_means %>% spread_draws(b_Intercept, b_mean_prior, b_mean_competence, b_mean_relevance,
-                          b_main_typexor) %>% 
+                          b_main_typexor, `b_mean_prior:mean_competence`,
+                          `b_mean_prior:mean_relevance`, `b_mean_relevance:mean_competence`,
+                          `b_mean_prior:main_typexor`, `b_mean_competence:main_typexor`,
+                          `b_mean_relevance:main_typexor`,
+                          `b_mean_prior:mean_relevance:mean_competence`, 
+                          `b_mean_prior:mean_competence:main_typexor`,
+                          `b_mean_prior:mean_relevance:main_typexor`,
+                          `b_mean_relevance:mean_competence:main_typexor`, 
+                          `b_mean_prior:mean_relevance:mean_competence:main_typexor`) %>% 
   mutate(
-    prior_xor = b_mean_prior,
+    prior_xor = b_mean_prior + `b_mean_prior:main_typexor`,
     prior_some = b_mean_prior,
-    competence_xor = b_mean_competence,
+    competence_xor = b_mean_competence + `b_mean_competence:main_typexor`,
     competence_some = b_mean_competence,
-    relevance_xor =  b_mean_relevance,
+    relevance_xor =  b_mean_relevance + `b_mean_relevance:main_typexor`,
     relevance_some = b_mean_relevance
@@
 ```
+
+```{r, fig.height = 6, fig.width=9}
+model_means_posteriors %>% select(prior_xor, prior_some,
+                               competence_xor, competence_some,
+                               relevance_xor, relevance_some) %>%
+  pivot_longer(cols = everything(), names_to = "effect", values_to = "value") -> model_mean_posteriors_long
+
+model_mean_posteriors_long %>%
+  ggplot(.) + # 
+  stat_halfeye(aes(y = effect, x = value, 
+                   fill = stat(case_when(x %>% between(-0.05, 0.05) ~ "no effect", x < -0.05 ~ "negative effect", x > 0.05 ~ "positive effect"))
+                   ), 
+               alpha = .7) +
+  geom_vline(xintercept = c(-.05, .05), linetype = "dashed") +
+  scale_fill_manual(name = "Direction of effect", values = c("coral", "yellow3", "mediumseagreen")) +
+  ggtitle("Posterior effect distributions. Points indicate posterior means\nThick lines indicate 66% CrIs, thin lines 95% CrIs.")
+
+```

### 2021-11-10 analysis/main/01_xor-some_prereg_preprocessing_final.Rmd
rerun analyses on updated N200, add priors exploration

@@
 cat(" Number of participants after excluding non-natives: ", d_native %>% distinct(submission_id) %>% count() %>% pull() )
+
+d_native %>% distinct(languages)
 
@@
 
+Perform an alternative exclusion based on comprehension question EXCLUDING the test_false1 question from the vignette with the ID=64 which had an incorrect prior classification.
+```{r}
+# get participants failing comprehension questions
+d_test_alternative <- d_test %>%
+  group_by(submission_id) %>%
+  filter(ID != 64 && test_question != "test_false1") %>%
+  mutate(passed_filler_trial = case_when(test_condition == "true" ~ response >= 60,
+                                   test_condition == "false" ~ response <= 40,
+                                   test_condition == "uncertain" ~ response %in% (10:90)),
+         mean_comprehension = mean(passed_filler_trial),
+         passed_filler = mean_comprehension >= 0.8
+         )
+
+d_test_fail_alternative <- d_test_alternative %>% 
+  filter(passed_filler == F)
+
+cat(" Subjects failing the comprehension trials without misclassified trial: ", d_test_fail_alternative %>% distinct(submission_id) %>% pull() %>% length())
+```
+
 Write out data for plotting:
@@
 ```
-Exclude the 201st participant for target N=200:
+[-Exclude the 201st participant for target N=200-]
+The exclusion of the last participant isn't necessary anymore, since we only have 200 subjects left upon exclusions due to the initially incorrect native language check.
 ```{r, results='hide'}
-d_full_clean %>% arrange(., desc(submission_id)) %>% distinct() %>% pull(submission_id)
+#d_full_clean %>% arrange(., desc(submission_id)) %>% distinct() %>% pull(submission_id)
 
 # exclude participant with last submission_id
-d_full_clean <- d_full_clean %>% filter(submission_id != 2620)
+#d_full_clean <- d_full_clean %>% filter(submission_id != 2620)
 
@@
   )
+
+# write out a wide df with both prior ratings for exploratory analysis
+d_critical_zScored_xor <- d_critical_zScored %>% 
+  ungroup() %>%
+  select(submission_id, title, main_type, block_extended, response_centered) %>% 
+  unique() %>% 
+  filter(main_type == "xor")
+
+d_critical_zScored_wide_xor_priors <- d_critical_zScored_xor %>% group_by(title, block_extended) %>% 
+  mutate(
+    block_extended = ifelse(block_extended == "prior", 
+                            paste(block_extended, which(block_extended == "prior")%%2, sep="_"),
+                            block_extended)
+  ) %>% ungroup() %>%
+  pivot_wider(
+    names_from = block_extended, 
+    values_from = response_centered 
+  ) %>%
+  filter(!is.na(prior_1), !is.na(prior_0)) # not sure where the NAs came from
+
+## no idea why it didn't work with just pivot_wider(names_from = block_extended, values_from = response_centered, names_repair = "unique")
+
+d_critical_zScored_wide_xor_priors %>% write_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors.csv")
 ```

### 2021-11-10 analysis/main/03_xor-some_prereg_analysis_final.Rmd
rerun analyses on updated N200, add priors exploration

@@
 d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")
+d_critical_zScored_wide_xor_priors <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors.csv")
 ```
@@
 ```
+
+A non-preregistered, but nevertheless possibly interesting exploratory analysis including the two individual prior ratings of the conditionals in the xor condition as predictors (as opposed to the mean of the two used in the preceding models):
+```{r, results='hide', message=FALSE, warning=FALSE}
+
+# set priors
+priors <- set_prior("student_t(1, 0, 2)", class = "b")
+
+model_xor_priors <- brm(target ~ prior_0*prior_1*competence*relevance +
+                   (1 + prior_0 + prior_1 + competence + relevance || submission_id) +
+                   (1 | title),
+                 data = d_critical_zScored_wide_xor_priors,
+                 prior = priors,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.95),
+                 cores = 4,
+                 iter = 3000)
+```
+
+```{r}
+summary(model_xor_priors)
+```
+
+Check the intervals of intervals, computing them separately using only one prior estimate, to see if we get different results:
+```{r}
+model_xor_priors %>% spread_draws(b_Intercept, b_prior_0, b_prior_1, b_competence, b_relevance) %>% 
+  mutate(
+    prior0_xor = b_prior_0,
+    prior1_xor = b_prior_1,
+    competence_xor = b_competence,
+    relevance_xor =  b_relevance
+  ) -> model_xor_priors_posteriors
+
+# check P that the effects are positive / negative / no effect present
+posterior_hypotheses_xor_priors <- model_xor_priors_posteriors %>% 
+  select(prior0_xor, prior1_xor, 
+         competence_xor, relevance_xor) %>%
+  gather(key, val) %>%
+  group_by(key) %>% mutate(positive = mean(val > 0.05),
+                           negative = mean(val < -0.05),
+                           no = mean(val %>% between(-0.05, 0.05))) %>%
+  summarise(positive_eff = mean(positive),
+            negative_eff = mean(negative),
+            no_eff = mean(no))
+posterior_hypotheses_xor_priors
+```

### 2021-11-10 analysis/pilots/xor-some-prolific-pilot3.Rmd
refactor

--- 2021-07-25 analysis/xor-some-prolific-pilot3.Rmd
@@
 ```{r data, results='hide', include=FALSE}
-d <- read_csv("../data/pilots/results_63_xor-some-Prolific-pilot3_N50.csv")
+d <- read_csv("../../data/pilots/results_63_xor-some-Prolific-pilot3_N50.csv")
 ```
@@
 ```{r}
-deviating_item_pilot2 <- read_csv("../data/pilots/pilot2_byItem_rating_deviations.csv") %>% select(-mean_deviation)
+deviating_item_pilot2 <- read_csv("../../data/pilots/pilot2_byItem_rating_deviations.csv") %>% select(-mean_deviation)
 d_critical_deviations %>% select(-mean_deviation) %>% semi_join(., deviating_item_pilot2)
@@
 
-d_long_plot_pilot2 <- read_csv("../data/pilots/pilot2_byItem_plotting.csv") %>% 
+d_long_plot_pilot2 <- read_csv("../../data/pilots/pilot2_byItem_plotting.csv") %>% 
   mutate(pilot = "old")
@@
   )
+
+#d_critical_zScored_wide %>%
+#  write_csv("../data/pilots/pilot3_critical_zScore_wide_tidy.csv")
 

### 2021-11-14 analysis/main/01_xor-some_prereg_preprocessing_final.Rmd
add N206, correlations, fix priors exploration, set up writing

--- 2021-11-06 analysis/main/01_xor-some_prereg_preprocessing_final.Rmd
@@
 # get participants failing comprehension questions
+# Perform exclusions based on comprehension question EXCLUDING the test_false1 question from the vignette
+# with the ID=64 which had an incorrect prior classification.
 d_test <- d_test %>%
   group_by(submission_id) %>%
+  filter(ID != 64 && test_question != "test_false1") %>%
   mutate(passed_filler_trial = case_when(test_condition == "true" ~ response >= 60,
@@
 ```
-Exclude the 201st participant for target N=200:
-```{r, results='hide'}
-d_full_clean %>% arrange(., desc(submission_id)) %>% distinct() %>% pull(submission_id)
-
-# exclude participant with last submission_id
-d_full_clean <- d_full_clean %>% filter(submission_id != 2620)
-
-cat(" Nr. of participants left after cleaning and N=200 extraction: ", d_full_clean %>% distinct(submission_id) %>% pull() %>% length())
-
-```
+[-Exclude the 201st participant for target N=200-]
+The exclusion of the last participant isn't necessary anymore, since we only have 200 subjects left upon exclusions due to the initially incorrect native language check.
 
@@
   )
+
 ```
 
-Write out z-scored data
+```{r}
+# create labels for first vs second prior question, order-based 
+d_critical_zScored %>% 
+  filter(block != "xor" & block != "some") %>%
+  filter(block == class_condition) %>%
+  filter(block == "prior", main_type == "xor") %>%
+  ungroup() %>%
+  mutate(priorQ_nr = rep(c(1,2), times=nrow(.)/2),
+         priorQ_factor = paste("prior", priorQ_nr, sep = "_"),
+         unique_ID = paste(submission_id, ID, sep="_")) -> d_xor_priors
+
+# create labels for the two questions
+d_xor_priors %>% group_by(ID, prompt) %>%
+  select(ID, prompt) %>% distinct() %>%
+  ungroup() %>%
+  mutate(
+    priorQ_label = rep(c("priorQ_1", "priorQ_2"), 32)
+  ) -> d_xor_priors_labels
+
+d_xor_priors_labels_full <- left_join(d_xor_priors, d_xor_priors_labels, by = c("prompt", "ID")) %>% 
+  select("prompt", "ID", "block_extended", "class_condition", "submission_id", "priorQ_factor", "priorQ_label")
+
+# create a df with both prior ratings for exploratory analysis
+d_critical_zScored_xor <- d_critical_zScored %>%
+  ungroup() %>%
+  filter(main_type == "xor") %>%
+  select(submission_id, title, main_type, block_extended, response_centered, prompt, ID, class_condition) %>% 
+  unique() %>% 
+  left_join(., d_xor_priors_labels_full, 
+            by = c("prompt", "ID", "block_extended", "class_condition", "submission_id")
+            ) %>%
+  filter(block_extended == class_condition | block_extended == "target") %>%
+  mutate(
+    block_xor_order = ifelse("prior" == block_extended, priorQ_factor, block_extended),
+    block_xor_type = ifelse("prior" == block_extended, priorQ_label, block_extended),
+  ) %>% select(-priorQ_factor, -priorQ_label)
+
+# create wide version for the order based question distinction
+d_critical_zScored_wide_xor_prior_order <- d_critical_zScored_xor %>% 
+  select(-block_extended, -block_xor_type, -class_condition, -prompt) %>%
+  unique() %>%
+  pivot_wider(
+    names_from = block_xor_order, 
+    values_from = response_centered
+  )
+
+# create wide version for the type based question distinction
+d_critical_zScored_wide_xor_prior_type <- d_critical_zScored_xor %>% 
+  select(-block_extended, -block_xor_order, -class_condition, -prompt) %>%
+  unique() %>%
+  pivot_wider(
+    names_from = block_xor_type, 
+    values_from = response_centered
+  ) 
+```
+
+Write out z-scored data:
 ```{r}
@@
 d_critical_zScored_wide %>% write_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
+
+# write out dataframes wth xor priors
+d_critical_zScored_wide_xor_prior_order %>% write_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors-order.csv")
+
+d_critical_zScored_wide_xor_prior_type %>% write_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors-type.csv")
+
 ```

### 2021-11-14 analysis/main/02_xor-some_prereg_plots_final.Rmd
add N206, correlations, fix priors exploration, set up writing

@@
 d_test <- read_csv("./../../data/main/results_prereg_raw_final_test.csv")
+d_critical_zScored_wide_xor_priors_order <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors-order.csv")
+d_critical_zScored_wide_xor_priors_type <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors-type.csv")
 ```
@@
 ```
+
+Plot the pairwise correlations of the predictors and the inference rating for "some": 
+```{r, message=FALSE, warning=FALSE}
+GGally::ggpairs(
+  filter(d_critical_zScored_wide, main_type == "some") %>% ungroup() %>%  
+    select(prior, competence, relevance, target)
+)
+```
+
+Plot the pairwise correlations of the predictors and the inference rating for "or", by order of presentation of the two prior questions:
+```{r, message=FALSE, warning=FALSE}
+# correlation plot for "or"
+GGally::ggpairs(
+  filter(d_critical_zScored_wide_xor_priors_order, main_type == "xor") %>%  ungroup() %>%
+    select(prior_1, prior_2, competence, relevance, target)
+  )
+```
+
+Plot the pairwise correlations of the predictors and the inference rating for "or", by type of prior questions:
+```{r, message=FALSE, warning=FALSE}
+# correlation plot for "or"
+GGally::ggpairs(
+  filter(d_critical_zScored_wide_xor_priors_type, main_type == "xor") %>%  ungroup() %>%
+    select(priorQ_1, priorQ_2, competence, relevance, target)
+  )
+```

### 2021-11-14 analysis/main/03_xor-some_prereg_analysis_final.Rmd
add N206, correlations, fix priors exploration, set up writing

--- 2021-11-07 analysis/main/03_xor-some_prereg_analysis_final.Rmd
@@
 d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")
+
+d_critical_zScored_wide_xor_prior_order <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors-order.csv")
+
+d_critical_zScored_wide_xor_prior_type <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors-type.csv")
 ```
@@
 ```
+
+A non-preregistered, but nevertheless possibly interesting exploratory analysis including the two individual prior ratings of the conditionals in the xor condition as predictors (as opposed to the mean of the two used in the preceding models):
+Analysis based on the order of appearance of the questions:
+```{r, results='hide', message=FALSE, warning=FALSE}
+
+# set priors
+priors <- set_prior("student_t(1, 0, 2)", class = "b")
+
+model_xor_priors_order <- brm(target ~ prior_1*prior_2*competence*relevance +
+                   (1 + prior_1 + prior_2 + competence + relevance || submission_id) +
+                   (1 | title),
+                 data = d_critical_zScored_wide_xor_prior_order,
+                 prior = priors,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.95),
+                 cores = 4,
+                 iter = 3000)
+```
+
+```{r}
+summary(model_xor_priors_order)
+```
+
+Check the intervals of intervals, computing them separately using only one prior estimate, to see if we get different results:
+```{r}
+model_xor_priors_order %>% spread_draws(b_Intercept, b_prior_1, b_prior_2, b_competence, b_relevance) %>% 
+  mutate(
+    prior1_xor = b_prior_1,
+    prior2_xor = b_prior_2,
+    competence_xor = b_competence,
+    relevance_xor =  b_relevance
+  ) -> model_xor_priors_posteriors
+
+# check P that the effects are positive / negative / no effect present
+posterior_hypotheses_xor_priors_order <- model_xor_priors_posteriors %>% 
+  select(prior1_xor, prior2_xor, 
+         competence_xor, relevance_xor) %>%
+  gather(key, val) %>%
+  group_by(key) %>% mutate(positive = mean(val > 0.05),
+                           negative = mean(val < -0.05),
+                           no = mean(val %>% between(-0.05, 0.05))) %>%
+  summarise(positive_eff = mean(positive),
+            negative_eff = mean(negative),
+            no_eff = mean(no))
+posterior_hypotheses_xor_priors_order
+```
+
+Prior question type based analysis:
+```{r, results='hide', message=FALSE, warning=FALSE}
+
+# set priors
+priors <- set_prior("student_t(1, 0, 2)", class = "b")
+
+model_xor_priors_type <- brm(target ~ priorQ_1*priorQ_2*competence*relevance +
+                   (1 + priorQ_1 + priorQ_2 + competence + relevance || submission_id) +
+                   (1 | title),
+                 data = d_critical_zScored_wide_xor_prior_type,
+                 prior = priors,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.95),
+                 cores = 4,
+                 iter = 3000)
+```
+
+```{r}
+summary(model_xor_priors_type)
+```
+
+Check the intervals of intervals, computing them separately using only one prior estimate, to see if we get different results:
+```{r}
+model_xor_priors_type %>% spread_draws(b_Intercept, b_priorQ_1, b_priorQ_2, b_competence, b_relevance) %>% 
+  mutate(
+    priorQ1_xor = b_priorQ_1,
+    priorQ2_xor = b_priorQ_2,
+    competence_xor = b_competence,
+    relevance_xor =  b_relevance
+  ) -> model_xor_priors_posteriors
+
+# check P that the effects are positive / negative / no effect present
+posterior_hypotheses_xor_priors_type <- model_xor_priors_posteriors %>% 
+  select(priorQ1_xor, priorQ2_xor, 
+         competence_xor, relevance_xor) %>%
+  gather(key, val) %>%
+  group_by(key) %>% mutate(positive = mean(val > 0.05),
+                           negative = mean(val < -0.05),
+                           no = mean(val %>% between(-0.05, 0.05))) %>%
+  summarise(positive_eff = mean(positive),
+            negative_eff = mean(negative),
+            no_eff = mean(no))
+posterior_hypotheses_xor_priors_type
+```

### 2021-11-16 writing/paper/paper_draft.tex
port over prereg passages

\documentclass{sp}

% The \pdf* commands provide metadata for the PDF output.
% Do not use LaTeX style / commands like \emph{} inside these.
\pdfauthor{Author Full Name(s)}
\pdftitle{The role of priors, competence and relevance for the interpretation of disjunction}
\pdfkeywords{exclusive disjunction, scalar implicature, pragmatics}

% Optional short title inside square brackets, for the running headers.
% If no short title is given, no title appears in the headers.
\title[Exclusive disjunction]{The role of priors, competence and relevance for the interpretation of disjunction%
  \thanks{We thank \ldots}}

% Optional short author inside square brackets, for the running headers.
% If no short author is given, no authors print in the headers.
\author[]{% As many authors as you like, each separated by \AND.
  \spauthor{Michael Franke \\ \institute{Institute1}} \AND
  \spauthor{Bob van Tiel \\ \institute{Institute2}} \AND
  \spauthor{Polina Tsvilodub \\ \institute{Institute3}}}

\begin{document}

\maketitle

\begin{abstract}

  Sentences containing disjunctions like ``Donald ate a pretzel or a donut'' are often taken as conveying that Donald ate a pretzel or a donut, but not both. That is, the listener might infer an \textit{exclusive reading of the disjunction}. Current standard accounts of disjunction interpretation posit that the exclusive interpretation of disjunctions is an instance of \textit{scalar implicature} (e.g., Horn, 1972). Prior work suggests that, among others, three factors influence the robustness of scalar implicatures: (1) \textit{relevance} of the stronger alternative to the listener, (2) the \textit{competence} of the speaker about the truth of the stronger alternative, and (3) the \textit{prior probability} that the stronger alternative is true (Sperber \& Wilson, 1995; Goodman \& Stuhlmüller, 2013; Degen, Tessler \& Goodman, 2015). While their influence was investigated experimentally for the interpretation of ``some'', the evidence is less conclusive in the case of disjunction interpretation. Therefore, we compare experimentally the interpretation of ``some'' and ``or'', while manipulating the three factors with respect to the stronger alternatives “all”, and “and”, respectively. Our results indicate that competence has a robust effect of the robustness of the implicature generation for both triggers, while prior and relevance exhibit different more intricate effects differing between the triggers.

\end{abstract}

\begin{keywords}
  exclusive disjunction, scalar implicature, pragmatics
\end{keywords}

\section{Introduction}

First paragraph.

\subsection{Scalar implicature}

Second paragraph.

\subsection{Related work}

Third paragraph.

Fourth paragraph.

\section{Relevance, ompetence and prior pribability}
We hypothesize that the implicature account holds for the interpretation of “some” (i.e., “some” → “some but not all”). That is, we expect the robustness of the implicature to be influenced by the three factors in the following way: 
1.1) Relevance Hypothesis (“some”): The higher the intuitively judged contextual relevance of the “all” alternative, the higher the degree of endorsement of the upper-bounded reading “some but not all”.  
1.2) Competence Hypothesis (“some”): The higher the intuitively judged knowledgeability of the speaker regarding the “all” alternative, the higher the degree of endorsement of the upper-bounded reading “some but not all”. 
1.3) Prior Hypothesis (“some”): The lower the intuitively judged prior probability of the “all” alternative being true, the higher the degree of endorsement of the upper-bounded reading “some but not all”.

We further hypothesize that, if the implicature account holds, the exclusive reading of “or” (i.e., “A or B” → “either A or B, but not both”) will be influenced by the same three factors in an analogous way. We therefore test the same three hypotheses, mutatis mutandis, also for “or” and call them Relevance Hypothesis (“or”), Competence Hypothesis (“or”) and Prior Hypothesis (“or”).

Finally, we also address the Identity Hypothesis that exclusive readings of disjunctions are a scalar implicature just like the “some → some but not all” case. If the Identity Hypothesis is correct, we expect that whichever factors positively or negatively affect the implicature readings of “some”, exactly these factors also affect the exclusive readings of “or”. In other words, the prediction of the Identity Hypothesis could still be confirmed even if the hypotheses related to the effects of relevance, competence and prior are not borne out, as long as the observed results are the same for “some” and “or”.

\subsection{Relevance}

\subsection{Competence}

\subsection{Prior probability}

\section{Methods}

In this experiment we directly compare the interpretation of “some” and “or”, while manipulating the three factors with respect to the stronger alternatives “all”, and “and”, respectively. Specifically, we manipulate the contextual relevance of the stronger alternative to the listener, the speaker's competence about the truth of “all” / “and”, and the prior probability of “all” / “and” being true, to investigate the robustness of an exclusive reading of “or” under these manipulations, in direct comparison to the upper-bounded reading of “some”.
To this end, we design context stories which were judged to score either high or low with respect to each factor of interest (high/low prior x high/low competence x high/low relevance => 8 unique conditions).

Participants were then asked to rate statements eliciting four judgements using a slider bar: the relevance of the “all”/“and”-alternative, the competence of the speaker about the alternative, the prior probability of ”all” / “and” being true, and, crucially, the likelihood of the exclusive reading of “or” or of the upper-bounded reading of “some”.

This study is a two-by-two-by-two-by-two within-subjects rating task, conducted as a web-based experiment. The study proceeds as follows.
Participants are welcomed to the experiment, read instructions and first see three labelled example trials showing the working of the slider. They see a simple story about a person shopping. They then see a question about the story that is clearly false (expected rating: 0), clearly true (expected rating: 100) and implied to be rather true (expected rating: 50-100), given the background story. These expected responses are labelled and explained. The continuous response slider ranges between “certainly true” and “certainly false”, converted to a rating between 0 and 100 in the background (as in the following main trials).   
The main part of the experiment consists of eight critical stories, presented in blocks of 10 or 11 trials each, randomly shuffled with eight attention check stories. The attention checks consist of one trial, visually matching critical trials (s. "Data collection procedure" section for details). 
Each critical story block has the following structure: Participants read a background story, followed by a comprehension statement which is either clearly true, clearly false, or uncertain (sampled at random for each story) which they are asked to rate using the slider. Then, they read statements presented in a box colored blue associated with the factors of interest and rate it using a slider bar. The factors of interest are relevance, competence and prior, in randomized order within-participants. In the “or”-condition, the prior statement consists of two symmetrical conditional sentences (see example below). Then, participants answer another three comprehension questions. These are followed by the critical utterance containing “or” / “some” appearing in a box colored red. Below, the target statement gathering the likelihood of the exclusive disjunction / the upper-bounded reading of “some” (i.e., the likelihood of the implicature) appears. Again, they rate the target statement using the slider. The background story remains on the screen throughout the trials. The four comprehension statements are sampled at random from six possible statements (two per type true / false / uncertain).
The ratings are elicited by asking the subjects “How likely is it that the statement in the blue box is true given the story?”.
We designed 32 different stories per trigger type (‘some’ vs. ‘or’), resulting in a total of 64 stories. Each subject sees eight randomly sampled stories (one per prior x competence x relevance condition out of four possible stories) such that they see four “or” and four “some” stories in randomized order.
After the study, participants may fill out a socio-demographic questionnaire.

Example transcript of an “or”-story and the associated statements:
Story: “Bill and Eric went to a bar. Eric’s wife, Lily, promised to pick them up if Eric drank any alcohol. She phones up Bill to ask about their night.”
Example comprehension statements: “Eric and Lily went to a bar together.” (false), “Eric and Bill went to see a football game.” (false), “Eric and Bill go out to a bar every weekend.” (uncertain), “Eric's wife has a driver’s licence.” (true)
Competence statement: “Bill knows whether Eric drank both wine and vodka.”
Relevance statement: “It is important for Lily to know whether her husband drank both wine and vodka.”
Prior statements: “If Eric drank wine, it is likely that he drank vodka as well.”, “If Eric drank vodka, it is likely that he drank wine as well.”
Critical utterance: “Bill tells Lily: 'Your husband drank wine or vodka.'”
Implicature statement: “From what Bill said we may conclude that Eric did not drink both wine and vodka.”

Example transcript of a “some”-story and the associated statements:
Story: “Bernard just finished his final exam. He isn't the brightest student, but if he passes this one, he will be through to the next grade. During the break, he goes to the teacher's lounge. He cannot wait to ask his teacher whether he passed. Bernard's teacher only just started correcting his exam.”
Example comprehension statements: “Bernard wants to get to the next grade.” (true), “Bernard wants to know his grade for his college application.” (false), “Bernard took a final exam in biology.” (uncertain), “Bernard is not the best student of his class.” (true)
Competence statement: “Bernard’s teacher knows whether all of the answers Bernard gave on the exam are correct.”
Prior statement: “If at least some of the answers Bernard gave on the exam are correct, then all of them are correct.”
Relevance statement: “It is important for Bernard to know whether all of the answers he gave on the exam are correct.”
Critical utterance: “Bernard’s teacher says: ‘Some of the answers you gave are correct.’”
Implicature statement: “From what Bernard’s teacher said we may conclude that not all of Bernard’s answers are correct.”

The experiment can be viewed under: https://magpie-xor-some.netlify.app/

\section{Results}

\section{DIscussion}

\bibliography{your-bibliography-file}

\begin{addresses}
  \begin{address}
    Author1 Full Name \\
    Street \\
    City, State Zip \ldots \\
    \email{author1@example.org}
  \end{address}
  % Repeat or remove additional addresses as needed.
  \begin{address}
    Author2 Full Name \\
    Street \\
    City, State Zip \dots \\
    \email{author2@example.com}
  \end{address}
\end{addresses}

\end{document}

### 2021-11-24 writing/paper/paper_draft.tex
first pass finished, added figures

@@
 That is, the listener might infer an \textit{exclusive reading of the disjunction}.
-Current standard accounts of disjunction interpretation posit that the exclusive interpretation of disjunctions is an instance of \textit{scalar implicature} (e.g., Horn, 1972).
-Prior work suggests that, among others, three factors influence the robustness of scalar implicatures: (1) \textit{relevance} of the stronger alternative to the listener, (2) the \textit{competence} of the speaker about the truth of the stronger alternative, and (3) the \textit{prior probability} that the stronger alternative is true (Sperber \& Wilson, 1995; Goodman \& Stuhlmüller, 2013; Degen, Tessler \& Goodman, 2015).
+Current standard accounts of disjunction interpretation posit that the exclusive interpretation of disjunctions is an instance of \textit{scalar implicature}.
+Prior work suggests that, among others, three factors influence the robustness of scalar implicatures: (1) \textit{relevance} of the stronger alternative to the listener, (2) the \textit{competence} of the speaker about the truth of the stronger alternative, and (3) the \textit{prior probability} that the stronger alternative is true.
 While their influence was investigated experimentally for the interpretation of ``some'', the evidence is less conclusive in the case of disjunction interpretation.
@@
 
-First paragraph.
+General intuition, introduction \citep[e.g.,][]{horn1972semantic}.
 
@@
 
-Second paragraph.
+SI accounts and treatments of ``some'' and ``or''.
 
-\subsection{Related work}
+\textbf{Related experimental work: } 
+Some: \citep{degen2015investigating},
 
-Third paragraph.
+Or: \citep{Li2021}
 
-Fourth paragraph.
-
-\section{Relevance, ompetence and prior pribability}
-We hypothesize that the implicature account holds for the interpretation of “some” (i.e., “some” → “some but not all”).
-That is, we expect the robustness of the implicature to be influenced by the three factors in the following way:
-1.1) Relevance Hypothesis (“some”): The higher the intuitively judged contextual relevance of the “all” alternative, the higher the degree of endorsement of the upper-bounded reading “some but not all”.
-1.2) Competence Hypothesis (“some”): The higher the intuitively judged knowledgeability of the speaker regarding the “all” alternative, the higher the degree of endorsement of the upper-bounded reading “some but not all”.
-1.3) Prior Hypothesis (“some”): The lower the intuitively judged prior probability of the “all” alternative being true, the higher the degree of endorsement of the upper-bounded reading “some but not all”.
-
-We further hypothesize that, if the implicature account holds, the exclusive reading of “or” (i.e., “A or B” → “either A or B, but not both”) will be influenced by the same three factors in an analogous way.
-We therefore test the same three hypotheses, mutatis mutandis, also for “or” and call them Relevance Hypothesis (“or”), Competence Hypothesis (“or”) and Prior Hypothesis (“or”).
-
-Finally, we also address the Identity Hypothesis that exclusive readings of disjunctions are a scalar implicature just like the “some → some but not all” case.
-If the Identity Hypothesis is correct, we expect that whichever factors positively or negatively affect the implicature readings of “some”, exactly these factors also affect the exclusive readings of “or”.
-In other words, the prediction of the Identity Hypothesis could still be confirmed even if the hypotheses related to the effects of relevance, competence and prior are not borne out, as long as the observed results are the same for “some” and “or”.
-
+\section{Relevance, competence and prior pribability}
+Literature on the effects of these factors on SIs \citep{sperber1986relevance, goodman2013knowledge, degen2015wonky}. 
 \subsection{Relevance}
@@
 
-\section{Methods}
+\subsection{Predictions}
+Prior research has shown that the robustness of implicatures from ``some'' to ``some, but not all''  predicted by the standrard scalar implicature (SI) account is influenced by three factors.
+First, high contextual relevance of the stronger alternative ``all'' to the listener is predicted to render more robust scalar implicatures.
+Second, high speaker competence about the truth of the stronger alternative is predicted to lead to more robust scalar implicature generation.
+Lastly, high prior probability of the stronger ``all'' being true is assumed to \textit{decrease} the scalar implicature rate.
+If the currently standard scalar implicature account also holds for the interpretation of disjunctions, analogous predictions for the robustness of the implicature from ``or'' to ``A or B, but not both'' can be derived.
+In case of the disjunction, the relevant stronger alternative is ``and''.
 
-In this experiment we directly compare the interpretation of “some” and “or”, while manipulating the three factors with respect to the stronger alternatives “all”, and “and”, respectively.
-Specifically, we manipulate the contextual relevance of the stronger alternative to the listener, the speaker's competence about the truth of “all” / “and”, and the prior probability of “all” / “and” being true, to investigate the robustness of an exclusive reading of “or” under these manipulations, in direct comparison to the upper-bounded reading of “some”.
-To this end, we design context stories which were judged to score either high or low with respect to each factor of interest (high/low prior x high/low competence x high/low relevance => 8 unique conditions).
+We hypothesize that all three factors influence the robustness of the scalar implicatures for both triggers.
+That is, we expect a higher rate of endorsements for sentences containing the upper-bounded or exclusive paraphrase of the respective triggers when the three factors are intuitively judged in the way described above.
+We, therefore, test six hypotheses, one concerning the effect of each factor for each trigger.
+Simultaneously,  we also address the Identity Hypothesis that exclusive readings of disjunctions are a scalar implicature just like the ``some → some but not all'' case.
+If the Identity Hypothesis is correct, we expect that whichever factors positively or negatively affect the implicature readings of ``some'', exactly these factors also affect the exclusive readings of ``or''.
+In other words, the prediction of the Identity Hypothesis could still be confirmed even if the hypotheses related to the effects of relevance, competence and prior are not borne out, as long as the observed results are the same for ``some'' and ``or''.
+Hence we directly compare the interpretation of the two triggers within one experiment, described in the following section.
 
-Participants were then asked to rate statements eliciting four judgements using a slider bar: the relevance of the “all”/“and”-alternative, the competence of the speaker about the alternative, the prior probability of ”all” / “and” being true, and, crucially, the likelihood of the exclusive reading of “or” or of the upper-bounded reading of “some”.
+\section{Experiment}
 
-This study is a two-by-two-by-two-by-two within-subjects rating task, conducted as a web-based experiment. The study proceeds as follows.
-Participants are welcomed to the experiment, read instructions and first see three labelled example trials showing the working of the slider.
-They see a simple story about a person shopping.
-They then see a question about the story that is clearly false (expected rating: 0), clearly true (expected rating: 100) and implied to be rather true (expected rating: 50-100), given the background story.
-These expected responses are labelled and explained.
-The continuous response slider ranges between “certainly true” and “certainly false”, converted to a rating between 0 and 100 in the background (as in the following main trials).
-The main part of the experiment consists of eight critical stories, presented in blocks of 10 or 11 trials each, randomly shuffled with eight attention check stories.
-The attention checks consist of one trial, visually matching critical trials (s.
-"Data collection procedure" section for details).
-Each critical story block has the following structure: Participants read a background story, followed by a comprehension statement which is either clearly true, clearly false, or uncertain (sampled at random for each story) which they are asked to rate using the slider.
+In this experiment we directly compare the interpretation of ``some'' and ``or'', so as to directly compare if the predictions generated by the scalar implicature account are borne out comparably for both triggers.
+Specifically, we manipulate the \textit{contextual relevance of the stronger alternative to the listener}, the \textit{speaker's competence about the truth of ``all'' / ``and''}, and the \textit{prior probability of ``all'' / ``and'' being true}, to investigate the robustness of an exclusive reading of ``or'' (``A or B, but not both'') under these manipulations, in direct comparison to the upper-bounded reading of “some” (``some but not all'').
+
+To operationalize these factors, we design context stories which were judged to score either high or low with respect to each factor of interest (high/low prior probability of the more informative alternative $\times$ high/low speaker competence $\times$ high/low contextual relevance of the alternative).
+These dichotomozed manipulations resulted in eight unique conditions.
+
+On critical trials, participants were asked to rate four sentences, one per factor and one containing an upper-bounded (exclusive) paraphrase of the trigger.
+The sentences elicited judgements how likely it is that each factor was high, and how likely the critical scalar implicature was true.
+The context stories and sentences manipulating the factors of interest were designed in an analogous fashion for both triggers and varied within-subjects, so the following descriptions apply for both triggers.
+
+If our factor hypotheses are true, on average, we expect higher likelihood ratings for the statements eliciting relevance, competence and prior judgements given items in the respective high conditions.
+Critically, we expect higher likelihood ratings for the scalar implicature, if the relevance and comptenece in the item were judged as high and the prior of the alternative as low.
+%Participants were then asked to rate statements eliciting four judgements using a slider bar: the relevance of the “all”/“and”-alternative, the competence of the speaker about the alternative, the prior probability of ”all” / “and” being true, and, crucially, the likelihood of the exclusive reading of “or” or of the upper-bounded reading of “some”.
+
+\subsection{Procedure and Materials}
+This study is a two-by-two-by-two-by-two within-subjects rating task (relevance $\times$ competence $\times$ prior $\times$ trigger), conducted as a web-based experiment.
+The study proceeds as follows.
+
+On each trial, participants read a background story, followed by a sentence presented in a blue box which is meant to elicit the likelihood rating.
+These sentences were of the form ``It is important to X to know whether Y'' (relevance rating), ``Z knows whether Y'', ``If A, then B'' / ``If B , then A'' (disjunction prior rating), ``If some Y, then all Y'' (some prior rating), ``From what Z said we may conclude that W'', where X is the listener, Z is the speaker and Y is the target event in the background story, and W is the event under the upper bounded / exclusive reading.
+A and B are the disjuncts in case of disjunction.
+For the inference strength elicitation, participants additionally read a critical utterance of the form ``Y says to X: Y'', containing the trigger (``some'' / ``or'') below the background story, presented in a red box.
+
+Below, participants read the task saying ``How likely is it that the statement is true given the story?'', followed by a slider rating bar labelled ``certainly false'' (left) and ``certainly true'' (right).
+The continuous ratings were converted to an integer rating between 0 and 100 in the background.
+For each critical vignette, participants completed a block of 10 or 11 trials, consisting of a trial with a comprehension question statement, trials with statements eliciting relevance, comptenece and prior ratings each (the last factor was elicited with two statements for the trigger ``or''), followed by three more comprehension questions, and the critical implicature strength elicitation trial.
+Comprehension trials were visually identical to the critical trials, but the statement to be rated only referred to the content of the backgorund stories, and were designed so as to be either clearly true, clearly false or uncertain given the background story.
+The four comprehension statements for one vignette were sampled at random from six possible statements (two per type true / false / uncertain).
+
+ We designed 32 different stories per trigger type (``some''~vs.~``or''), resulting in a total of 64 stories.
+Each subject sees eight randomly sampled stories (one per prior $\times$ competence $\times$ relevance condition out of four possible stories) such that they see four ``or'' and four ``some'' stories in randomized order.
+The assignment of conditions to the triggers is randomized between-subjects.
+
+At the beginning, participants were welcomed to the experiment, read instructions and first saw three labelled example trials showing the working of the slider.
+They read a simple story about a person shopping.
+They then saw a question about the story that was clearly false (expected rating: 0), clearly true (expected rating: 100) and implied to be rather true (expected rating: 50-100), given the background story.
+These expected responses were labelled and explained.
+The main part of the experiment consisted of eight critical stories, randomly shuffled with eight attention check stories.
+The attention checks consist of one trial, visually matching critical trials.
+On these trials, the vignettes contain a statement which clearly writes out what participants are supposed to answer (e.g., ``Please move the slider maximally left'') in an area of text that participants have to read to complete the trial.
+Participants who fail more than two of these attention checks will not be reimbursed and their data excluded from the analysis.
+%Each critical story block has the following structure: Participants read a background story, followed by a comprehension statement which is either clearly true, clearly false, or uncertain (sampled at random for each story) which they are asked to rate using the slider.
 Then, they read statements presented in a box colored blue associated with the factors of interest and rate it using a slider bar.
@@
 The background story remains on the screen throughout the trials.
-The four comprehension statements are sampled at random from six possible statements (two per type true / false / uncertain).
-The ratings are elicited by asking the subjects “How likely is it that the statement in the blue box is true given the story?”.
-We designed 32 different stories per trigger type (‘some’ vs.
-‘or’), resulting in a total of 64 stories.
-Each subject sees eight randomly sampled stories (one per prior x competence x relevance condition out of four possible stories) such that they see four “or” and four “some” stories in randomized order.
 After the study, participants may fill out a socio-demographic questionnaire.
+The experiment can be viewed under \texttt{https://magpie-xor-some.netlify.app/}. The preregistration for the experiment can be found under \texttt{https://osf.io/v7zjp}.
 
+\subsection{Participants}
+We recruited 277 participants through the crowd-sourcing platform Prolific.
+Participants were restricted to those whose first language included English, who previously took part in at least five other Prolific studies, and whose approval rate was at least 0.9, according to prescreening criteria of the platform.
+Participants took 20 minutes on average to complete the study and were compensated \pounds2.48 for their participation.
+Following our preregistered exclusion criteria, we excluded 15 participants for not indicating their native language, 3 participants for completing the study in under 8 minutes, 26 participants for failing more than 20\% of the attention checks, and 27 for failing more than 20\% of the comprehension questions.
+Due to a coding error of the comprehension questions, 206 participants were left post applying exclusion criteria, although the preregistered target sample size post exclusions was 200 subjects.
+We conducted a power analysis in order to determine the minimal sample size required for a power of 0.8 of finding evidence for the conjunction of all factor hypotheses (relevance, competence and prior) predicted by the SI account, finding that at least 180 subjects were required.
+In the following analyses, data from the 206 participants is analysed.\footnote{All analyses were also conducted on data from 200 subject, excluding the last six submissions.
+No noteworthy quantitative or qualitative differences of the results were observed.}
+
+\section{Results}
+Prior to conducting statistical analyses, we preprocess the data by standardizing (or z-scoring) the responses within each factor (relevance, comptence, prior) for each participant.
+This is done to counterbalance possible by-participant responsiveness to particular factors, while keeping an anlysis advantageous to the scalar implicature account, since the standardization is applied across triggers.
+
+\subsection{Main analysis}
+We analysed the results using a Bayesian linear mixed effects model, regressing the target implicature strength rating against the fixed effects of predictor ratings (i.e., the relevance, comptence and prior ratings elicited by the same participant for that vignette), the effect of trigger and their two-, three-, and four-way interactions.
+Maximal computationally tractable random effects strcuture justified by the experimental design was included: we included random intercepts and random slope effects for the main effects of trigger, relevance, competence and prior by-subject, as well as random intercepts by-vignette.\footnote{Model in R syntax style: \texttt{inference-rating $\sim$ trigger * relevance * competence * prior + (1 + trigger + relevance + competence + prior | subjectID) + (1 | vignette)}}
+The categorical effect of trigger was dummy coded, using ``some'' as the reference level.
+For all regression coefficients we used a wide and uninformative prior given by a $t$-distribution with mean 0, standard deviation of 2 and 1 degree of freedom.
+The model was fitted using the R \texttt{brms} package \citep{burkner2017brms}.\footnote{The data and the analyses can be found under \texttt{https://github.com/magpie-ea/magpie-xor-experiment}}
+
+We focus on six slope coefficients of the regression model, for the effects of relevance, competence and prior, respectively, once for ``some'' and once for ``or''.
+We check whether the posterior estimate of each effect is in either one of three intervals: (1) negative effect (<= -0.05), (2) no effect (between -0.05 and 0.05), or (3) positive effect (>= 0.05).
+We interpret the data as providing evidence in favor of an effect (positive, negative, no effect) if the posterior probability of the effect being true is >= 0.95 (i.e., 95\% of posterior samples are in the corresponding interval).
+We speak of evidence in favor of the scalar implicature account of the exclusive disjunction interpretation if the posterior probability of the prior effect being negative, the relevance and competence effects being positive is >= 0.95.
+
+\begin{figure*}[t]
+	\begin{center}
+		\includegraphics[width=1\linewidth]{images/posterior-effects-main.png}
+	\end{center}
+	\vspace{-0.3cm}
+	\caption{Main analysis results: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''.
+The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect).}
+	\label{posteriors-main}
+\end{figure*}
+
+Consistent with predictions, for the trigger ``some'', we find a clear negative prior probabilty effect of the stronger alternative ``all'' being true, as inidcated by the probability of the negative effect of prior being $P = 0.999$ (Fig.\ref{posteriors-main}, \texttt{prior\_some} line, orange color).
+Similarly, we find a clear positive effect for the effect of speaker competence ($P =  1$, Fig.\ref{posteriors-main}, \texttt{competence\_some} line, green color).
+However, we do not find any clear effects of relevance (Fig.\ref{posteriors-main}, \texttt{relevance\_some} line, split colors).
+If anything, the data supports the hypothesis that the effect is not negative.
+We hypothesize that the absense of a clear-cut relevance effect might be due to the difficulty to manipulate this dimension contextually.
+It might be that the vignettes used in the experiment failed to make the listener relevance salient enough for any effects to be borne out.
+Further experiments should improve this manipulation, as well as additionally consider speaker relevance.
+
+For the trigger ``or'', in contrast, we only find a positive effect of competence ($P =  0.993$, Fig.\ref{posteriors-main}, \texttt{competence\_xor} line, green color).
+We find no comprehensive effects for relevance or competence.
+While the reason for the former might be the same as for ``some'', we suggest that these results might also be influenced by collinearity of the effects.
+We turn to this conjecture in light of the exploratory analyses, described below.
+Overall, these results would provide evidence against the scalar implicature hypothesis because not all factors yielded an effect on the robustness of scalar implicatures geenrated from ``some'' and ``or''.
+Similarly, the results would provide evidence against the identity hypothesis, as en effect of competence was observed for ``some'', but not for ``or''.
+
+\subsection{Exploratory analysis}
+\begin{figure*}[t]
+	\begin{center}
+		\includegraphics[width=1\linewidth]{images/posterior-effects-categorical.png}
+	\end{center}
+	\vspace{-0.3cm}
+	\caption{Exlporatory analysis with categorical predictors: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''.
+The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect).}
+	\label{posteriors-cat}
+\end{figure*}
+In addition to the main analysis, we conducted two exploratory analyses.
+First, we fit a model using categorical predictors only.
+That is, we fit a Bayesian linear mixed regression model, regressing the inference rating against our prior intuitive classifications of the vignettes on the dimensions relevance, competence and prior (high or low, respectively), the effect of trigger, and their interactions.
+The random effects structure and the priors were identical to the main model.
+The factors were sum-coded, -1 representing the low condition, 1 representing the high condition.
+This model decouples teh predictors by dichotomizing their classification and, therefore, addresses the potential collinearty issue.
+Furthermore, it provides a comparison between participants' judgements, and our intuitive prior classifications of the vignettes.
+Given this model and the data, we found very similar results for both triggers: We observe a negative effect of prior and a positive effect of competence, in line with predictions of the scalar implicature hypothesis (Fig.\ref{posteriors-}).
+However, we again observe no effects of relevance.
+If anything, the relevance effect tends in the direction opposite to predictions of the SI account for the trigger ``or''.
+\begin{figure*}[t]
+	\begin{center}
+		\includegraphics[width=1\linewidth]{images/posterior-effects-means.png}
+	\end{center}
+	\vspace{-0.3cm}
+	\caption{Exlporatory analysis with rating predictors by-vignette: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''.
+The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect).}
+	\label{posteriors-mean}
+\end{figure*}
+
+These results match the second exploratory analysis, whereby the ratings were used as predictors in the linear model, but the ratings were averaged for each vignette across participants.
+In this second  exploratory model, we again find similar effects for ``some'' and ``or'': a positive effect of comptenece, negative for prior, and an indefinite relevance effect (Fig.\ref{posteriors-mean}).
+By averaging over singel participant ratings, this model indicates that the absent competence effect in the main model might be due to by-participant variation.
+
+In order to further investigate the relevance effects and potantial collinarity concerns, we computed the exploratory pairwise correlations of all the predictors.
+Interestingly, we find a significant correlation between the prior and the relevance effects for the trigger ``or'' ($R^2 = -0.106$).
+We also found a significant correlation between competence and relevance ($R^2 = 0.127$) for ``or''.
+No correlations were found for ``some''.
+These intriguing results might indicate that these effects are closely related for the disjunction.
+Intuitively, a negative correlation between the prior and relvance of the stronger alternative ``and'' make sense: The more likely the conjunction is a priori, the more likely it is that the listener might know about it already, yielding the implicature irrelevant to the listener.
+Similarly, competence might be connected to relevance indirectly via the prior: the more likely the stronger alternative a priori, the more likely it is that a speaker knows about it, while teh irrelevance still persists for the listener.
+However, this entangled results need further research.
+
+\section{Discussion}
+In sum, the results provide moderate evidence for the scalar implicature account of ``some'' and ``or''.
+While we could confirm the effects of prior and comptence for the interpretation of ``some'' as predicted by the SI hypothesis, we could not verify them for the disjunction.
+Furthermore, our results hint at intriguing dependencies between relevance and prior effects for the interpretation of exclusive disjunctions.
+Taken together, the results might speak against the identity hypothesis, indicating that there might be subtle differences in the conceptualization of the stronger alternative for ``some'' and ``or''.
+
+\section{Appendix}
+\subsection{Vignettes}
 Example transcript of an “or”-story and the associated statements:
+
 Story: “Bill and Eric went to a bar. Eric’s wife, Lily, promised to pick them up if Eric drank any alcohol. She phones up Bill to ask about their night.”
@@
 Example transcript of a “some”-story and the associated statements:
+
 Story: “Bernard just finished his final exam.
@@
 
-The experiment can be viewed under: https://magpie-xor-some.netlify.app/
-
-\section{Results}
-
-\section{DIscussion}
-
-\bibliography{your-bibliography-file}
+\bibliography{paper.bib}
 

### 2021-11-24 writing/paper/paper_draft.tex
fix typos

@@
 Therefore, we compare experimentally the interpretation of ``some'' and ``or'', while manipulating the three factors with respect to the stronger alternatives “all”, and “and”, respectively.
-Our results indicate that competence has a robust effect of the robustness of the implicature generation for both triggers, while prior and relevance exhibit different more intricate effects differing between the triggers.
+Our results indicate that competence has a consistent effect of the robustness of the implicature generation for both triggers, while prior and relevance exhibit different, more intricate, effects varying between the triggers.
 
@@
 
-\textbf{Related experimental work: } 
+\textbf{Related experimental/modelling work: } 
 Some: \citep{degen2015investigating}, 
-
 Or: \citep{Li2021}
@@
 Lastly, high prior probability of the stronger ``all'' being true is assumed to \textit{decrease} the scalar implicature rate.
-If the currently standard scalar implicature account also holds for the interpretation of disjunctions, analogous predictions for the robustness of the implicature from ``or'' to ``A or B, but not both'' can be derived.
-In case of the disjunction, the relevant stronger alternative is ``and''.
+If the currently standard scalar implicature account also holds for the interpretation of disjunctions, analogous predictions for the robustness of the implicature from ``A or B'' to ``A or B, but not both'' can be derived.
+In case of disjunction, the relevant stronger alternative is ``and''.
 
-We hypothesize that all three factors influence the robustness of the scalar implicatures for both triggers.
-That is, we expect a higher rate of endorsements for sentences containing the upper-bounded or exclusive paraphrase of the respective triggers when the three factors are intuitively judged in the way described above.
+We hypothesize that all three factors influence the robustness of scalar implicatures for both triggers.
+That is, we expect a higher rate of endorsements for sentences containing the upper-bounded (exclusive, in case of ``or'') paraphrases of the respective triggers when the three factors are intuitively judged to be high or low, as described above.
 We, therefore, test six hypotheses, one concerning the effect of each factor for each trigger.
-Simultaneously,  we also address the Identity Hypothesis that exclusive readings of disjunctions are a scalar implicature just like the ``some → some but not all'' case.
-If the Identity Hypothesis is correct, we expect that whichever factors positively or negatively affect the implicature readings of ``some'', exactly these factors also affect the exclusive readings of ``or''.
-In other words, the prediction of the Identity Hypothesis could still be confirmed even if the hypotheses related to the effects of relevance, competence and prior are not borne out, as long as the observed results are the same for ``some'' and ``or''.
-Hence we directly compare the interpretation of the two triggers within one experiment, described in the following section.
+Simultaneously,  we also address the \textit{identity hypothesis} that exclusive readings of disjunctions are a scalar implicature just like the ``some → some but not all'' case.
+If the identity hypothesis is correct, we expect that whichever factors positively or negatively affect the implicature readings of ``some'', exactly these factors also affect the exclusive readings of ``or''.
+In other words, the prediction of the identity hypothesis could still be confirmed even if the hypotheses related to the effects of relevance, competence and prior are not borne out, as long as the observed results are the same for ``some'' and ``or''.
+To investigate this, we directly compare the interpretation of the two triggers within one experiment, described in the following section.
 
@@
 
-In this experiment we directly compare the interpretation of ``some'' and ``or'', so as to directly compare if the predictions generated by the scalar implicature account are borne out comparably for both triggers.
+In this experiment we directly compare the interpretation of ``some'' and ``or'', so as to directly explore if the predictions generated by the scalar implicature account are borne out comparably for both triggers.
 Specifically, we manipulate the \textit{contextual relevance of the stronger alternative to the listener}, the \textit{speaker's competence about the truth of ``all'' / ``and''}, and the \textit{prior probability of ``all'' / ``and'' being true}, to investigate the robustness of an exclusive reading of ``or'' (``A or B, but not both'') under these manipulations, in direct comparison to the upper-bounded reading of “some” (``some but not all'').
 
-To operationalize these factors, we design context stories which were judged to score either high or low with respect to each factor of interest (high/low prior probability of the more informative alternative $\times$ high/low speaker competence $\times$ high/low contextual relevance of the alternative).
-These dichotomozed manipulations resulted in eight unique conditions.
+To operationalize these factors, we designed context stories (vignettes) which were judged to score either high or low with respect to each factor of interest (high/low prior probability of the more informative alternative $\times$ high/low speaker competence $\times$ high/low contextual relevance of the alternative).
+Combining these dichotomozed manipulations resulted in eight unique conditions.
 
@@
 A and B are the disjuncts in case of disjunction.
-For the inference strength elicitation, participants additionally read a critical utterance of the form ``Y says to X: Y'', containing the trigger (``some'' / ``or'') below the background story, presented in a red box.
+For the inference strength elicitation, participants additionally read a critical utterance of the form ``Z says to X: Y'', containing the trigger (``some'' / ``or''), presented below the background story in a red box.
 
@@
 The continuous ratings were converted to an integer rating between 0 and 100 in the background.
-For each critical vignette, participants completed a block of 10 or 11 trials, consisting of a trial with a comprehension question statement, trials with statements eliciting relevance, comptenece and prior ratings each (the last factor was elicited with two statements for the trigger ``or''), followed by three more comprehension questions, and the critical implicature strength elicitation trial.
-Comprehension trials were visually identical to the critical trials, but the statement to be rated only referred to the content of the backgorund stories, and were designed so as to be either clearly true, clearly false or uncertain given the background story.
+For each critical vignette, participants completed a block of 10 or 11 trials, consisting of a trial with a comprehension question, trials with statements eliciting relevance, comptenece and prior ratings each (the last factor was elicited with two statements for the trigger ``or''), followed by three more comprehension questions, and the critical implicature strength elicitation trial.
+Comprehension trials were visually identical to the critical trials, but the statement to be rated only referred to the content of the background stories, and were designed so as to be either clearly true, clearly false or uncertain given the background story.
 The four comprehension statements for one vignette were sampled at random from six possible statements (two per type true / false / uncertain).
@@
 The main part of the experiment consisted of eight critical stories, randomly shuffled with eight attention check stories.
-The attention checks consist of one trial, visually matching critical trials.
-On these trials, the vignettes contain a statement which clearly writes out what participants are supposed to answer (e.g., ``Please move the slider maximally left'') in an area of text that participants have to read to complete the trial.
-Participants who fail more than two of these attention checks will not be reimbursed and their data excluded from the analysis.
+The attention checks consisted of one trial, visually matching critical trials.
+On these trials, the vignettes contained a statement which clearly wrote out what participants were supposed to answer (e.g., ``Please move the slider maximally left'') in an area of text that participants had to read to complete the trial.
+Participants who failed more than two of these attention checks were excluded from the analysis.
 %Each critical story block has the following structure: Participants read a background story, followed by a comprehension statement which is either clearly true, clearly false, or uncertain (sampled at random for each story) which they are asked to rate using the slider.
@@
 The background story remains on the screen throughout the trials.
-After the study, participants may fill out a socio-demographic questionnaire.
+After the study, participants could fill out a socio-demographic questionnaire.
 The experiment can be viewed under \texttt{https://magpie-xor-some.netlify.app/}. The preregistration for the experiment can be found under \texttt{https://osf.io/v7zjp}.
@@
 \subsection{Main analysis}
-We analysed the results using a Bayesian linear mixed effects model, regressing the target implicature strength rating against the fixed effects of predictor ratings (i.e., the relevance, comptence and prior ratings elicited by the same participant for that vignette), the effect of trigger and their two-, three-, and four-way interactions.
-Maximal computationally tractable random effects strcuture justified by the experimental design was included: we included random intercepts and random slope effects for the main effects of trigger, relevance, competence and prior by-subject, as well as random intercepts by-vignette.\footnote{Model in R syntax style: \texttt{inference-rating $\sim$ trigger * relevance * competence * prior + (1 + trigger + relevance + competence + prior | subjectID) + (1 | vignette)}}
+We analysed the results using a Bayesian linear mixed effects model, regressing the target implicature likelihood rating against the fixed effects of predictor ratings (i.e., the relevance, comptence and prior ratings elicited by the same participant for that vignette), the effect of trigger and their two-, three-, and four-way interactions.
+Maximal computationally tractable random effects structure justified by the experimental design was included: we included random intercepts and random slope effects for the main effects of trigger, relevance, competence and prior by-subject, as well as random intercepts by-vignette.\footnote{Model in R syntax style: \texttt{inference-rating $\sim$ trigger * relevance * competence * prior + (1 + trigger + relevance + competence + prior || subjectID) + (1 | vignette)}.
+For computational tractability reasons, the correlation of random effects by-subject was set to 0.}
 The categorical effect of trigger was dummy coded, using ``some'' as the reference level.
@@
 We interpret the data as providing evidence in favor of an effect (positive, negative, no effect) if the posterior probability of the effect being true is >= 0.95 (i.e., 95\% of posterior samples are in the corresponding interval).
-We speak of evidence in favor of the scalar implicature account of the exclusive disjunction interpretation if the posterior probability of the prior effect being negative, the relevance and competence effects being positive is >= 0.95.
+In particular, we speak of evidence in favor of the scalar implicature account of the exclusive disjunction interpretation if the posterior probability of the prior effect being negative, the relevance and competence effects being positive is >= 0.95.
 
@@
 If anything, the data supports the hypothesis that the effect is not negative.
-We hypothesize that the absense of a clear-cut relevance effect might be due to the difficulty to manipulate this dimension contextually.
+We hypothesize that the absense of a clear-cut relevance effect might be due to the difficulty manipulating this dimension contextually.
 It might be that the vignettes used in the experiment failed to make the listener relevance salient enough for any effects to be borne out.
@@
 We turn to this conjecture in light of the exploratory analyses, described below.
-Overall, these results would provide evidence against the scalar implicature hypothesis because not all factors yielded an effect on the robustness of scalar implicatures geenrated from ``some'' and ``or''.
+Overall, these results would provide evidence against the scalar implicature hypothesis because not all factors yielded an effect on the robustness of scalar implicatures generated from ``some'' and ``or''.
 Similarly, the results would provide evidence against the identity hypothesis, as en effect of competence was observed for ``some'', but not for ``or''.
@@
 	\vspace{-0.3cm}
-	\caption{Exlporatory analysis with categorical predictors: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''.
+	\caption{Exploratory analysis with categorical predictors: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''.
 The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect).}
@@
 First, we fit a model using categorical predictors only.
-That is, we fit a Bayesian linear mixed regression model, regressing the inference rating against our prior intuitive classifications of the vignettes on the dimensions relevance, competence and prior (high or low, respectively), the effect of trigger, and their interactions.
+That is, we fit a Bayesian linear mixed effects regression model, regressing the inference rating against our prior intuitive classifications of the vignettes on the dimensions relevance, competence and prior (high or low, respectively), the effect of trigger, and their interactions.
 The random effects structure and the priors were identical to the main model.
 The factors were sum-coded, -1 representing the low condition, 1 representing the high condition.
-This model decouples teh predictors by dichotomizing their classification and, therefore, addresses the potential collinearty issue.
-Furthermore, it provides a comparison between participants' judgements, and our intuitive prior classifications of the vignettes.
-Given this model and the data, we found very similar results for both triggers: We observe a negative effect of prior and a positive effect of competence, in line with predictions of the scalar implicature hypothesis (Fig.\ref{posteriors-}).
+This model decouples the predictors by dichotomizing their classification and, therefore, addresses the potential collinearty issue.
+Furthermore, it provides a comparison between participants' judgements and our intuitive prior classifications of the vignettes.
+
+Given this model and the data, we found very similar results for both triggers: We observe a negative effect of prior and a positive effect of competence, in line with predictions of the scalar implicature hypothesis (Fig.\ref{posteriors-cat}).
 However, we again observe no effects of relevance.
@@
 	\vspace{-0.3cm}
-	\caption{Exlporatory analysis with rating predictors by-vignette: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''.
+	\caption{Exploratory analysis with predictor ratings averaged by-vignette: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''.
 The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect).}
@@
 In this second  exploratory model, we again find similar effects for ``some'' and ``or'': a positive effect of comptenece, negative for prior, and an indefinite relevance effect (Fig.\ref{posteriors-mean}).
-By averaging over singel participant ratings, this model indicates that the absent competence effect in the main model might be due to by-participant variation.
+By averaging over single participant ratings, this model indicates that the absent competence effect in the main model might be due to by-participant variation.
 
-In order to further investigate the relevance effects and potantial collinarity concerns, we computed the exploratory pairwise correlations of all the predictors.
+In order to further investigate the relevance effects and potential collinarity concerns, we computed exploratory pairwise correlations of all the predictors.
 Interestingly, we find a significant correlation between the prior and the relevance effects for the trigger ``or'' ($R^2 = -0.106$).
@@
 These intriguing results might indicate that these effects are closely related for the disjunction.
-Intuitively, a negative correlation between the prior and relvance of the stronger alternative ``and'' make sense: The more likely the conjunction is a priori, the more likely it is that the listener might know about it already, yielding the implicature irrelevant to the listener.
-Similarly, competence might be connected to relevance indirectly via the prior: the more likely the stronger alternative a priori, the more likely it is that a speaker knows about it, while teh irrelevance still persists for the listener.
+Intuitively, a negative correlation between prior and relevance of the stronger alternative ``and'' makes sense: the more likely the conjunction is a priori, the more likely it is that the listener might know about it already, yielding the implicature irrelevant to the listener.
+Similarly, competence might be connected to relevance indirectly via the prior: the more likely the stronger alternative is a priori, the more likely it is that a speaker knows about it, while the irrelevance still persists for the listener.
 However, this entangled results need further research.
@@
 While we could confirm the effects of prior and comptence for the interpretation of ``some'' as predicted by the SI hypothesis, we could not verify them for the disjunction.
+We also did not find evidence for effects of relevance on the robustness of scalar implicature generation, given the vignettes.
 Furthermore, our results hint at intriguing dependencies between relevance and prior effects for the interpretation of exclusive disjunctions.

### 2021-11-25 writing/paper/paper_draft.tex
minor edits and comments on first write-up draft

@@
 \pdfkeywords{exclusive disjunction, scalar implicature, pragmatics}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% customization
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\usepackage[dvipsnames]{xcolor}
+\newcommand{\mf}[1]{\textcolor{BurntOrange}{[MF: #1]}}
+\newcommand{\pt}[1]{\textcolor{Cerulean}{[PT: #1]}}
+\newcommand{\bvt}[1]{\textcolor{ForestGreen}{[BvT: #1]}}
 
@@
 \subsection{Predictions}
-Prior research has shown that the robustness of implicatures from ``some'' to ``some, but not all''  predicted by the standrard scalar implicature (SI) account is influenced by three factors.
+Prior research has shown that the robustness of implicatures from ``some'' to ``some, but not all''  predicted by the standard scalar implicature (SI) account is influenced by three factors.
 First, high contextual relevance of the stronger alternative ``all'' to the listener is predicted to render more robust scalar implicatures.
@@
 On critical trials, participants were asked to rate four sentences, one per factor and one containing an upper-bounded (exclusive) paraphrase of the trigger.
-The sentences elicited judgements how likely it is that each factor was high, and how likely the critical scalar implicature was true.
+The sentences elicited judgements of how likely it is that each factor was high, and how likely the critical scalar implicature was true.
 The context stories and sentences manipulating the factors of interest were designed in an analogous fashion for both triggers and varied within-subjects, so the following descriptions apply for both triggers.
+\mf{I suggest to give one full example of a vignette (context, critical questions, possibly test questions) for each trigger here; similar to what is now in the appendix (but with nicer formatting)}
+\mf{it might be useful to have a graphical representation of the order of ``screens'' within each trial; e.g.
+similar to the example in Figure 1 here: \url{https://magpie-manual.netlify.app/00_getting_started/02_basics/}}
 
@@
 %Participants were then asked to rate statements eliciting four judgements using a slider bar: the relevance of the “all”/“and”-alternative, the competence of the speaker about the alternative, the prior probability of ”all” / “and” being true, and, crucially, the likelihood of the exclusive reading of “or” or of the upper-bounded reading of “some”.
+\mf{Last paragraph feels redundant to me here; best repeat hypotheses in results / analyze section but not necessarily here}
 
 \subsection{Procedure and Materials}
-This study is a two-by-two-by-two-by-two within-subjects rating task (relevance $\times$ competence $\times$ prior $\times$ trigger), conducted as a web-based experiment.
+This study is a $ 2 \times 2 \times 2 \times 2$ within-subjects rating task (relevance $\times$ competence $\times$ prior $\times$ trigger), conducted as a web-based experiment.
 The study proceeds as follows.
 
-On each trial, participants read a background story, followed by a sentence presented in a blue box which is meant to elicit the likelihood rating.
-These sentences were of the form ``It is important to X to know whether Y'' (relevance rating), ``Z knows whether Y'', ``If A, then B'' / ``If B , then A'' (disjunction prior rating), ``If some Y, then all Y'' (some prior rating), ``From what Z said we may conclude that W'', where X is the listener, Z is the speaker and Y is the target event in the background story, and W is the event under the upper bounded / exclusive reading.
+On each trial, participants read a background story, followed by a sentence presented in a blue box which is meant to elicit the likelihood rating.\mf{maybe include a screenshot?} These sentences were of the form ``It is important to X to know whether Y'' (relevance rating), ``Z knows whether Y'', ``If A, then B'' / ``If B , then A'' (disjunction prior rating), ``If some Y, then all Y'' (some prior rating), ``From what Z said we may conclude that W'', where X is the listener, Z is the speaker and Y is the target event in the background story, and W is the event under the upper bounded / exclusive reading.
 A and B are the disjuncts in case of disjunction.
@@
  We designed 32 different stories per trigger type (``some''~vs.~``or''), resulting in a total of 64 stories.
-Each subject sees eight randomly sampled stories (one per prior $\times$ competence $\times$ relevance condition out of four possible stories) such that they see four ``or'' and four ``some'' stories in randomized order.
+Each subject saw eight randomly sampled stories (one per prior $\times$ competence $\times$ relevance condition out of four possible stories) such that they saw four ``or'' and four ``some'' stories in randomized order.
 The assignment of conditions to the triggers is randomized between-subjects.
@@
 The background story remains on the screen throughout the trials.
-After the study, participants could fill out a socio-demographic questionnaire.
+After the study, participants could voluntarily fill out a socio-demographic questionnaire.
 The experiment can be viewed under \texttt{https://magpie-xor-some.netlify.app/}. The preregistration for the experiment can be found under \texttt{https://osf.io/v7zjp}.
@@
 Due to a coding error of the comprehension questions, 206 participants were left post applying exclusion criteria, although the preregistered target sample size post exclusions was 200 subjects.
-We conducted a power analysis in order to determine the minimal sample size required for a power of 0.8 of finding evidence for the conjunction of all factor hypotheses (relevance, competence and prior) predicted by the SI account, finding that at least 180 subjects were required.
+We conducted a power analysis in order to determine the minimal sample size required for a power of 0.8 of finding evidence for the conjunction of all factor hypotheses (relevance, competence and prior) predicted by the SI account, finding that at least 180 subjects were required.\mf{we should give some details about the power analysis or link to a document in the repository}
 In the following analyses, data from the 206 participants is analysed.\footnote{All analyses were also conducted on data from 200 subject, excluding the last six submissions.
@@
 	\caption{Main analysis results: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''.
-The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect).}
+The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect).
+\mf{better use nicer labels for $y$-axis: use ``or'' not ``xor'' and no hyphens?; same for next plots}}
 	\label{posteriors-main}
@@
 Consistent with predictions, for the trigger ``some'', we find a clear negative prior probabilty effect of the stronger alternative ``all'' being true, as inidcated by the probability of the negative effect of prior being $P = 0.999$ (Fig.\ref{posteriors-main}, \texttt{prior\_some} line, orange color).
-Similarly, we find a clear positive effect for the effect of speaker competence ($P =  1$, Fig.\ref{posteriors-main}, \texttt{competence\_some} line, green color).
+Similarly, we find a clear positive effect for speaker competence ($P =  1$, Fig.\ref{posteriors-main}, \texttt{competence\_some} line, green color).
 However, we do not find any clear effects of relevance (Fig.\ref{posteriors-main}, \texttt{relevance\_some} line, split colors).
@@
 Further experiments should improve this manipulation, as well as additionally consider speaker relevance.
+\mf{best first just state the results and use the last two sentences for ``Discussion'' section}
 
@@
 Overall, these results would provide evidence against the scalar implicature hypothesis because not all factors yielded an effect on the robustness of scalar implicatures generated from ``some'' and ``or''.
-Similarly, the results would provide evidence against the identity hypothesis, as en effect of competence was observed for ``some'', but not for ``or''.
+Similarly, the results would provide evidence against the identity hypothesis, as an effect of competence \mf{prior?} was observed for ``some'', but not for ``or''.\mf{we should explain somewhere where the threshold of 0.05 comes from; when discussing the results here we could add that (think) no effects becomes credible even if the threshold would be reduced to 0}
 
@@
 	\caption{Exploratory analysis with categorical predictors: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''.
-The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect).}
+The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect).
+\mf{I'm not sure if we should blindly reuse a threshold of 0.05 here without further justification.
+Something to think about.}}
 	\label{posteriors-cat}
@@
 	\caption{Exploratory analysis with predictor ratings averaged by-vignette: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''.
-The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect).}
+The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect).
+\mf{I'm not sure if we should blindly reuse a threshold of 0.05 here without further justification.
+Something to think about.}}
 	\label{posteriors-mean}
@@
 These results match the second exploratory analysis, whereby the ratings were used as predictors in the linear model, but the ratings were averaged for each vignette across participants.
-In this second  exploratory model, we again find similar effects for ``some'' and ``or'': a positive effect of comptenece, negative for prior, and an indefinite relevance effect (Fig.\ref{posteriors-mean}).
+In this second  exploratory model, we again find similar effects for ``some'' and ``or'': a positive effect of competence, negative for prior, and an indefinite relevance effect (Fig.\ref{posteriors-mean}).
 By averaging over single participant ratings, this model indicates that the absent competence effect in the main model might be due to by-participant variation.
@@
 No correlations were found for ``some''.
-These intriguing results might indicate that these effects are closely related for the disjunction.
+These intriguing results might indicate that these effects are closely related for disjunction.
 Intuitively, a negative correlation between prior and relevance of the stronger alternative ``and'' makes sense: the more likely the conjunction is a priori, the more likely it is that the listener might know about it already, yielding the implicature irrelevant to the listener.

### 2021-11-27 analysis/main/04_xor-some_analysis_model_comparison.Rmd
ELM draft, model comparison init commits

---
title: "Xor-Some Preregistered Study: Analyses"
author: "Polina Tsvilodub"
date: "8/31/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(brms)
library(tidyboot)
library(tidybayes)
```

Read data created in preprocessing script:
```{r, message=FALSE, warning=FALSE, results='hide'}
d_critical_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")

d_critical_zScored_wide_xor_prior_order <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors-order.csv")

d_critical_zScored_wide_xor_prior_type <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors-type.csv")

d_critical_zScored_wide_xor <- d_critical_zScored_wide %>% filter(main_type == "xor")
```

## Comparison of prior and relevance effects

Since we observed potential colinearity effects in the main preregistered analysis for the trigger "or", we further explore the effects of this colinearity on the strength of the individual effects in the regression models.

To do so, we compute individual models omitting one of the factors on the "or" data, and compute the Bayes Factor (BF) of the model omitting one of the colinear predictors, relative to the full model.

```{r, results='hide', message=FALSE, warning=FALSE}
# Full model 
# set priors
priors <- set_prior("student_t(1, 0, 2)", class = "b")

model_SI <- brm(target ~ prior*competence*relevance +
                   (1 + prior + competence + relevance || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.95),
                 cores = 4,
                 iter = 20000,
                 save_all_pars = TRUE
                )
```

```{r}
summary(model_SI)
```

```{r, results='hide', message=FALSE, warning=FALSE}
# model with prior effect only
model_prior_comp <- brm(target ~ prior*competence +
                   (1 + prior + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 20000,
                 save_all_pars = TRUE
                 )
```

```{r}
summary(model_prior_comp)
```

```{r, results='hide', message=FALSE, warning=FALSE}
# model with relevance effect only
model_rel_comp <- brm(target ~ relevance*competence +
                   (1 + relevance + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 20000,
                 save_all_pars = TRUE
                 )
```

```{r}
summary(model_rel_comp)
```

```{r, results='hide', message=FALSE, warning=FALSE}
# model WITHOUT comptenece effect, for reasons of symmetry
model_rel_pri <- brm(target ~ relevance*prior +
                   (1 + relevance + prior || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 20000,
                 save_all_pars = TRUE
                 )
```

```{r}
summary(model_rel_pri)
```

```{r, message=FALSE, warning=FALSE}
# computes the BFs 
# in favor of the prior only model over full model for XOR
bayes_factor(model_prior_comp, model_SI, silent = TRUE)
# in favor of the relevance only model over the full model for XOR
bayes_factor(model_rel_comp, model_SI, silent = TRUE)
# in favor of the relevance model over the prior model
bayes_factor(model_rel_comp, model_prior_comp, silent = TRUE)
# and in vice versa in favor of the prior model over the relevance model
bayes_factor(model_prior_comp, model_rel_comp, silent = TRUE)

# for symmetry, in favor of model without competence over full model
bayes_factor(model_rel_pri, model_SI, silent = TRUE)
bayes_factor(model_rel_pri, model_rel_comp, silent = TRUE)
bayes_factor(model_rel_pri, model_prior_comp, silent = TRUE)

```

In order to investigate the correlation between relevance and competence we observed for "or", we also compute a model without the competence effect. Since we also observed correlation between the remaining two factors prior and relevance, perhaps the best way to disentangle the effects is to fit single-predictor models, and compate them to an intercept only model.
```{r, results='hide', message=FALSE, warning=FALSE}
# intercept only model
priors_int <- set_prior("student_t(3, 0.1, 2.5)", class = "Intercept")
model_int <- brm(target ~ 1 +
                   (1 | submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors_int,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 20000,
                 save_all_pars = TRUE
                 )
# single predictor models
model_rel <- brm(target ~ relevance +
                   (1 + relevance || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 20000,
                 save_all_pars = TRUE
                 )

model_comp <- brm(target ~ competence +
                   (1 + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 20000,
                 save_all_pars = TRUE
                 )

model_pri <- brm(target ~ prior +
                   (1 + prior || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 20000,
                 save_all_pars = TRUE
                 )
```

```{r}
summary(model_int)
summary(model_rel)
summary(model_comp)
summary(model_pri)
```

Below, the BFs in favor of the single predictors over the intercept-only model are computed. If these are large (let's say >10), they indicate that the respective factor is important for explaining the data. 
```{r}
bayes_factor(model_pri, model_int, silent = TRUE)
bayes_factor(model_comp, model_int, silent = TRUE)
bayes_factor(model_rel, model_int, silent = TRUE)
```

Below, the BFs in favor of the two-predictor models over the respective single-predictor models are computed. If these are large, they would indicate that 1) the added model complexity is justified by a better fit to the data and 2) that the second factor is "necessary" to explain the data.
```{r}
bayes_factor(model_prior_comp, model_pri, silent = TRUE)
bayes_factor(model_prior_comp, model_comp, silent = TRUE)

bayes_factor(model_rel_pri, model_pri, silent = TRUE)
bayes_factor(model_rel_pri, model_rel, silent = TRUE)

bayes_factor(model_rel_comp, model_rel, silent = TRUE)
bayes_factor(model_rel_comp, model_comp, silent = TRUE)
```

### 2021-11-27 writing/ELM2/abstract.tex
ELM draft, model comparison init commits

\documentclass[11pt,letterpaper]{article}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\renewcommand{\familydefault}{\sfdefault}
\usepackage{helvet}
\pagestyle{empty}
\usepackage[kerning=true]{microtype}
\usepackage{parskip}
\usepackage{sansmath}
\usepackage{graphicx}
\usepackage{sidecap}  
\sidecaptionvpos{figure}{c}
\usepackage{float}
\usepackage{color, soul}
% Feel free to use additional packages for glosses, figures, whatnot.
\usepackage[dvipsnames]{xcolor}
\newcommand{\mf}[1]{\textcolor{BurntOrange}{[MF: #1]}}
\newcommand{\pt}[1]{\textcolor{Cerulean}{[PT: #1]}}
\newcommand{\bvt}[1]{\textcolor{ForestGreen}{[BvT: #1]}}
% The next bit is for reserving sufficient space for authors,
% affiliations, and e-mail address.  No need to change for initial
% anonymous version.  For the final version, replace the
%\toggletrue{anonymous} %with \togglefalse{anonymous} to de-anonymize.
\usepackage{etoolbox}
\newtoggle{anonymous}
\toggletrue{anonymous}

\renewcommand{\title}[1]{\textbf{#1}\\}
\newcommand{\authors}[1]{\iftoggle{anonymous}{\phantom{#1}}{#1}\\}
\newcommand{\email}[1]{\iftoggle{anonymous}{\phantom{#1}}{#1}}

\begin{document}

% First page:

% Insert title, authors, affiliations, and e-mail address in the next three lines:

\title{The role of relevance, competence and priors for scalar implicatures}
\authors{Michael Franke (University of T\"ubingen), Bob van Tiel (Radboud University), Polina Tsvilodub (Osnabr\"uck university)}
\email{pstvilodub@uos.de}

% Here goes the main text of your abstract:

If someone says "Anna ate some cookies." the hearer of the sentence might infer that Anna ate \textit{some, but not all} cookies that were in the jar. %Similarly, when hearing the sentence "Donald ate a donut or a pretzel." one might infer that Donald ate either \textit{the donut or the pretzel, but not}. 
Such an inference is an instance of \textit{scalar implicature (SI)}---an inference based on the comparison to a salient informationally stronger alternative ``all'' (Horn, 1972). That is, the listener might arrive at the upper-bounded reading of ``some'' by reasoning about a cooperative speaker: if the stronger alternative was true, a rational agent would have used the stronger utterance; but since she did not, she must have lacked the evidence for using ``all'' (Grice, 1975). Current literature also suggests an analogous account for the exclusive interpretation of ``or'', generated by comparison to the stronger alternative ``and'' (Geurts, 2010).

Prior work addresses the influence of sentence structure, or of specific linguistic markers (Li, 2021). Futhermore, literature suggests that, among others, three factors influence the robustness of scalar implicatures: (1) \textit{relevance} of the stronger alternative to the listener, (2) the \textit{competence} of the speaker about the truth of the stronger alternative, and (3) the \textit{prior probability that the stronger alternative is true} (Sperber\&Wilson, 1995; Goodman\&Stuhlmüller, 2013; Degen, Tessler\&Goodman, 2015). However, their influence was mostly investigated for individual factors. We set out to explore how these three factors influence the robustness of scalar implicatures in combination, when the SI triggers ``some'' and ``or'' are presented in rich context. Simultanously, we directly compare these two triggers with respect to their influenceability by these factors.

We conducted a web-based rating study, whereby participants read background stories which were designed to score either high or low with respect to each factor of interest, manipulated within-subjects (high/low prior probability of the more informative alternative $\times$ high/low speaker competence $\times$ high/low contextual relevance of the alternative; eight unique conditions, applied to each trigger). 
On critical trials, participants were asked to rate four sentences on sclae ranging from ``certainly true'' to ``certainly false'' (converted to 0--100), one per factor and one containing an upper-bounded (or exclusive) paraphrase of the trigger. The sentences elicited judgements of how likely it is that each factor was high, and how likely the critical scalar implicature was true (s. transcript). If the SI account is true, on average, we expect higher likelihood ratings for the scalar implicature, if the relevance and comptenece in the item were judged as high and the prior of the alternative as low. Each participant saw eight stories (four per trigger) sampled from 32 stories per trigger, randomly shuffled with visually identical attnetion checks.

We recruited 277 participants through Prolific (71  excluded following preregistered exclusion criteria). The rating were z-scored within each factor for each participant. We analysed the results using a Bayesian linear mixed effects model, regressing the target implicature likelihood rating against the fixed effects of predictor ratings (i.e., the relevance, comptence and prior ratings elicited by the same participant for that vignette), the effect of trigger and their two-, three-, and four-way interactions. Random intercepts and random slope effects for the main effects of trigger, relevance, competence and prior by-subject, as well as random intercepts by-vignette were included.

Consistent with predictions, for the trigger ``some'', we find a clear negative prior probabilty effect of the stronger alternative ``all'' being true (probability of the effect of prior being smaller than 0.05: $P = 0.999$). Similarly, we find a clear positive effect for speaker competence (effect larger than 0.05 probability: $P =  1$). We do not find any clear effects of relevance. For the trigger ``or'', we only find a positive effect of competence ($P =  0.993$). Further, we computed exploratory pairwise correlations of all the predictors. Interestingly, we find a significant correlation between the prior and the relevance effects for the trigger ``or'' ($R^2 = -0.106$). We also found a significant correlation between competence and relevance ($R^2 = 0.127$) for ``or''. No correlations were found for ``some''. 
Overall, the results provide moderate evidence for the scalar implicature account of ``some'' and ``or'', but also indicate that there might be intricate dependencies between the factors' influence on the robustness of SI, especially for disjunction interpretation.
\newpage

% Second page for additional materials, example stimuli, graphs, tables, references:

%\begin{figure}
%   \makebox[\linewidth][c]{\includegraphics[width=0.7\linewidth]{amlap2020_E1-E2.pdf}}%
% \caption{E1 (left): Means and 95\% bootstrapped confidence intervals (in all plots) of ratings (on a scale ranging 0-100) for how well a sentence with the adjective “big” (or “small”) describes a member of a subordinate category (normal-sized compared to other members of the same category in the context picture) when different nouns (color) appear in different syntactic frames (x-axis). Points represent participant means within condition.\newline 
%    E2 (right): Basic-level referent label productions (e.g., “dog” when the referent is a normal-sized Great Dane) in different syntactic frames (x-axis).}
%    \label{syntax-production}
%\end{figure}

%\begin{figure}

%    \makebox[\linewidth][c]{\includegraphics[scale=0.7]{amlap_expt3-cc-inference.pdf}}%

%    \caption{Inferred basic-level comparison classes (e.g. “...big relative to other \textit{dogs}”) when the referent (e.g., a Great Dane) appears in basic-level visual context (left panel) or in subordinate context (right panel) from a sentence where the N is “one”, the subordinate (“Great Dane”) or the basic-level (“dog”) label of the referent (colors), appearing in different syntactic positions (x-axis).}
%    \label{cc-infer}
%\end{figure}
%\begin{SCfigure}[1.7]
%    \includegraphics[ scale=0.5]{amlap_expt4_double_mod2.pdf}
 %   \caption{Pilot Results: Inferred basic-level comparison classes (e.g. ‘...big relative to other \textit{dogs}’) when the directly modified subordinate N (‘big Great Dane’) appears in different syntactic positions (x-axis).\newline
%    \newline \textbf{References:} [1] Kamp, J., In \textit{Formal semantics of natural language}, 1975  [2] Kennedy, C., \textit{Linguistics and Philosophy, 30(1)}, 2007 [3] Tessler, M. H.; Lopez-Brau, M.; Goodman N. D., In \textit{39th annual meeting of the Cognitive Science society}, 2017
%    \label{double-mod}[4] Reboul, A., In \textit{Language typology and language universals, an international handbook, vol.1}, 2001}
%\end{SCfigure}

\newpage

% Optional third page for additional information if the investigated language is not English:
\end{document}

### 2021-12-01 analysis/main/02_xor-some_prereg_plots_final.Rmd
add figures to abstract, fine tune content

@@
 left_join(d_critical_predictorsXtarget_summary, d_critical_prior_classes, by = c("title", "predictors" = "class_condition")) %>%
-  mutate(prior_class = as.factor(prior_class)) %>%
+  mutate(prior_class = factor(prior_class, levels = c(0,1), labels=c("low", "high"))) %>%
   ggplot(., aes(x = mean_predictor, y = mean_target)) +
   geom_point(aes(color = prior_class)) +
-  ggrepel::geom_text_repel(aes(label=title), size = 2.5) +
+  ggrepel::geom_text_repel(aes(label=title), size = 3.5) +
   geom_smooth(method = "lm") +
@@
   xlab("Mean predictor ratings") + 
-  facet_wrap(main_type~predictors) +
-  scale_fill_manual(name = "Prior classification of item")
+  facet_grid(cols=vars(predictors), rows=vars(main_type)) +
+  scale_fill_discrete(name = "Prior classification of item") +
+  theme(strip.text.y = element_text(size = 8))
+
+ggsave("../../writing/ELM2/byItem-ratings.pdf", width = 9, height = 6)
 ```

### 2021-12-01 analysis/main/03_xor-some_prereg_analysis_final.Rmd
add figures to abstract, fine tune content

@@
 model_SI_posteriors_long %>%
+  mutate(effect = factor(effect, 
+                         levels = c("relevance_xor", "relevance_some", "prior_xor", "prior_some", "competence_xor", "competence_some"), 
+                         labels = c("Relevance (or)", "Relevance (some)", "Prior (or)", "Prior (some)", "Competence (or)", "Competence (some)"))) %>%
   ggplot(.) + # 
@@
 
+ggsave("../../writing/ELM2/posterior-effects.pdf", width=7, height=4)
 ```

### 2021-12-01 writing/ELM2/abstract.tex
add figures to abstract, fine tune content

@@
 
-If someone says "Anna ate some cookies." the hearer of the sentence might infer that Anna ate \textit{some, but not all} cookies that were in the jar.
-%Similarly, when hearing the sentence "Donald ate a donut or a pretzel." one might infer that Donald ate either \textit{the donut or the pretzel, but not}.
-Such an inference is an instance of \textit{scalar implicature (SI)}---an inference based on the comparison to a salient informationally stronger alternative ``all'' (Horn, 1972).
-That is, the listener might arrive at the upper-bounded reading of ``some'' by reasoning about a cooperative speaker: if the stronger alternative was true, a rational agent would have used the stronger utterance; but since she did not, she must have lacked the evidence for using ``all'' (Grice, 1975).
-Current literature also suggests an analogous account for the exclusive interpretation of ``or'', generated by comparison to the stronger alternative ``and'' (Geurts, 2010).
+If someone says "Anna ate some cookies." the hearer of this sentence might infer the upper-bounded reading that Anna ate \textit{some, but not all} cookies.
+Similarly, given the sentence "Donald ate a donut or a pretzel." one might infer an exclusive interpretation, i.e., that Donald ate either \textit{the donut or the pretzel, but not both}.
+Both inferences are considered as instances of \textit{scalar implicature (SI)}---an inference based on the comparison to a salient informationally stronger alternative ``all'' or ``and'', respectively [2,3].
+That is, the listener might arrive at these interpretations by reasoning about the epistemic state of a cooperative rational speaker [4].
+%: if the stronger alternative was true, a rational agent would have used the stronger utterance; but since she did not, she must have lacked the evidence for using ``all''
+%(Grice, 1975).
+%Some prior work addresses the influence of sentence structure, or of specific linguistic markers on the robustness of SIs (Li, 2021). Moreover, 
+Literature suggests that, among others, three factors influence the robustness of scalar implicatures: (1) \textit{relevance} of the stronger alternative to the listener, (2) the \textit{competence} of the speaker about the truth of the stronger alternative, and (3) the \textit{prior probability that the stronger alternative is true} [1,3,5].
+However, their influence was mostly investigated in isolation in prior work.
+We aim to fill this gap by exploring how these three factors influence the robustness of scalar implicatures in combination, when the SI triggers ``some'' and ``or'' are presented in rich context.
+Simultanously, we directly compare whether the two triggers are both influenced by these factors as predicted by the SI account.
 
-Prior work addresses the influence of sentence structure, or of specific linguistic markers (Li, 2021).
-Futhermore, literature suggests that, among others, three factors influence the robustness of scalar implicatures: (1) \textit{relevance} of the stronger alternative to the listener, (2) the \textit{competence} of the speaker about the truth of the stronger alternative, and (3) the \textit{prior probability that the stronger alternative is true} (Sperber\&Wilson, 1995; Goodman\&Stuhlmüller, 2013; Degen, Tessler\&Goodman, 2015).
-However, their influence was mostly investigated for individual factors.
-We set out to explore how these three factors influence the robustness of scalar implicatures in combination, when the SI triggers ``some'' and ``or'' are presented in rich context.
-Simultanously, we directly compare these two triggers with respect to their influenceability by these factors.
-
-We conducted a web-based rating study, whereby participants read background stories which were designed to score either high or low with respect to each factor of interest, manipulated within-subjects (high/low prior probability of the more informative alternative $\times$ high/low speaker competence $\times$ high/low contextual relevance of the alternative; eight unique conditions, applied to each trigger).
-On critical trials, participants were asked to rate four sentences on sclae ranging from ``certainly true'' to ``certainly false'' (converted to 0--100), one per factor and one containing an upper-bounded (or exclusive) paraphrase of the trigger.
+In our web-based rating study (LINK), participants read background stories which were designed to score either high or low with respect to each factor of interest (high/low prior probability of the more informative alternative $\times$ high/low speaker competence $\times$ high/low contextual relevance of the alternative $\times$ triggers), manipulated within-subjects.
+On critical trials, participants were asked to rate four sentences on a scale ranging from ``certainly true'' to ``certainly false'' (converted to 0--100), one per factor and one containing an upper-bounded (or exclusive) paraphrase of the trigger.
 The sentences elicited judgements of how likely it is that each factor was high, and how likely the critical scalar implicature was true (s.
 transcript).
-If the SI account is true, on average, we expect higher likelihood ratings for the scalar implicature, if the relevance and comptenece in the item were judged as high and the prior of the alternative as low.
-Each participant saw eight stories (four per trigger) sampled from 32 stories per trigger, randomly shuffled with visually identical attnetion checks.
+If the SI account is true, on average, we expect higher likelihood ratings for the scalar implicature statement, if the relevance and comptenece of the item were judged as high and the prior of the stronger alternative as low.
+Each participant saw eight stories (four per trigger) sampled from 32 stories per trigger, randomly shuffled with eight visually identical attention checks and comprehension questions.
 
-We recruited 277 participants through Prolific (71  excluded following preregistered exclusion criteria).
-The rating were z-scored within each factor for each participant.
-We analysed the results using a Bayesian linear mixed effects model, regressing the target implicature likelihood rating against the fixed effects of predictor ratings (i.e., the relevance, comptence and prior ratings elicited by the same participant for that vignette), the effect of trigger and their two-, three-, and four-way interactions.
-Random intercepts and random slope effects for the main effects of trigger, relevance, competence and prior by-subject, as well as random intercepts by-vignette were included.
-
-Consistent with predictions, for the trigger ``some'', we find a clear negative prior probabilty effect of the stronger alternative ``all'' being true (probability of the effect of prior being smaller than 0.05: $P = 0.999$).
-Similarly, we find a clear positive effect for speaker competence (effect larger than 0.05 probability: $P =  1$).
-We do not find any clear effects of relevance.
-For the trigger ``or'', we only find a positive effect of competence ($P =  0.993$).
+%(71  excluded following preregistered exclusion criteria).
+We analysed data from 206 participants recruited through Prolific; their ratings were z-scored within each factor by-participant.
+We analysed the results using a Bayesian linear mixed effects model, regressing the implicature likelihood rating against the fixed effects of all factor ratings by the same participant for that vignette, the effect of trigger, all interactions and maximal tractable random effects structure.
+Participants' factor ratings agreed well with the prior classification of the items (Fig.~\ref{by-item-ratings}, colors).
+%Consistent with predictions of the SI account, for the trigger ``some'', we find a clear negative prior probabilty effect of the stronger alternative ``all'' being true (probability of the effect of prior being smaller than 0.05: $P = 0.999$).
+As predicted by the SI account, participants are more likely to infer the ``some, but not al'' reading when they judged the prior probability of ``all'' being true as low ($P=0.999$ for the effect being $<0.05$, Fig.\ref{posteriors} for all results).
+They also robustly infer the upper-bounded interpretation when judging speaker competence as high ($P=1$ for effect $>0.05$).
+%Similarly, we find a clear positive effect for speaker competence (effect larger than 0.05 probability: $P =  1$). We do not find any clear effects of relevance. 
+However, participants were only more likely to infer the ``A or B, but not both'' reading when they rated the speaker competence as high  ($P=0.993$ for effect $>0.05$).
+No relevance effects were found for both triggers.
+%For the trigger ``or'', we only find a positive effect of competence ($P =  0.993$). 
 Further, we computed exploratory pairwise correlations of all the predictors.
-Interestingly, we find a significant correlation between the prior and the relevance effects for the trigger ``or'' ($R^2 = -0.106$).
-We also found a significant correlation between competence and relevance ($R^2 = 0.127$) for ``or''.
-No correlations were found for ``some''.
-Overall, the results provide moderate evidence for the scalar implicature account of ``some'' and ``or'', but also indicate that there might be intricate dependencies between the factors' influence on the robustness of SI, especially for disjunction interpretation.
+While no correlations were found for ``some'', we find a significant correlation between the prior and the relevance effects for ``or'' ($R^2 = -0.106, p=...$), and a significant correlation between competence and relevance ($R^2 = 0.127, p=...$) for ``or''.
+This suggests that these factors might be interdependent in case of ``or''.
+Overall, our results support the SI account, but also indicate that these factors might have a more fine-grained influence on the robustness of SI generation in combination.%, especially for disjunction interpretation.
+%Overall, the results provide moderate evidence for the scalar implicature account of ``some'' and ``or'', but also indicate that there might be intricate dependencies between the factors' influence on the robustness of SI, especially for disjunction interpretation.
 \newpage
@@
 
-%\begin{figure}
-%   \makebox[\linewidth][c]{\includegraphics[width=0.7\linewidth]{amlap2020_E1-E2.pdf}}%
-% \caption{E1 (left): Means and 95\% bootstrapped confidence intervals (in all plots) of ratings (on a scale ranging 0-100) for how well a sentence with the adjective “big” (or “small”) describes a member of a subordinate category (normal-sized compared to other members of the same category in the context picture) when different nouns (color) appear in different syntactic frames (x-axis).
-Points represent participant means within condition.\newline
-%    E2 (right): Basic-level referent label productions (e.g., “dog” when the referent is a normal-sized Great Dane) in different syntactic frames (x-axis).}
-%    \label{syntax-production}
-%\end{figure}
+\begin{figure}
+   \makebox[\linewidth][c]{\includegraphics[width=0.7\linewidth]{byItem-ratings.pdf}}%
+ \caption{Relating ratings for relevance, competence and prior statements (x-axis) to
+ratings for the strength of pragmatic enrichments (y-axis). The top row shows ratings,
+averaged for each story for ``some'' (enriched to ``some, but not all''). The bottom row shows ratings, averaged for each story for ``or'' (enriched to ``A or B, but not both'').}
+    \label{by-item-ratings}
+\end{figure}
 
-%\begin{figure}
+\begin{figure}   
+    \makebox[\linewidth][c]{\includegraphics[scale=0.7]{posterior-effects.pdf}}
+    \caption{Distributions of posterior samples for the effects of each predictor for each trigger (y-axis).
+The colors indicate the fraction of the distribution density corresponding to a positive, no or a negative effect.
+\newline \textbf{References:} [1] Degen, J., Tessler, M.
+H., \& Goodman, N.
+D., In \textit{Proceedings of CogSci}, 2015 [2] Geurts, B., \textit{Quantity Implicatures}, 2010 [3] Goodman, N.
+D., \& Stuhlm\"uller, A., In \textit{Topics in Cognitive Science}, 5(1), 2013 [4] Grice, H.
+P.
+In \textit{Syntax and semantics, vol.
+3: Speech acts}, 1975 [3] Horn, L.
+R., 1972 [5] Sperber, D., \& Wilson, D., \textit{Relevance: communication and cognition}, 1995
+    }
+    \label{posteriors}
+\end{figure}
 
-%    \makebox[\linewidth][c]{\includegraphics[scale=0.7]{amlap_expt3-cc-inference.pdf}}%
-
-%    \caption{Inferred basic-level comparison classes (e.g.
-“...big relative to other \textit{dogs}”) when the referent (e.g., a Great Dane) appears in basic-level visual context (left panel) or in subordinate context (right panel) from a sentence where the N is “one”, the subordinate (“Great Dane”) or the basic-level (“dog”) label of the referent (colors), appearing in different syntactic positions (x-axis).}
-%    \label{cc-infer}
-%\end{figure}
 %\begin{SCfigure}[1.7]

### 2021-12-02 analysis/main/04_xor-some_analysis_model_comparison.Rmd
update model comparison priors

@@
 # set priors
-priors <- set_prior("student_t(1, 0, 2)", class = "b")
+priors <- set_prior("student_t(1, 0, 0.25)", class = "b")
 
@@
 # in favor of the prior only model over full model for XOR
-bayes_factor(model_prior_comp, model_SI, silent = TRUE)
+bf_pri_comp_full <- bayes_factor(model_prior_comp, model_SI, silent = TRUE)
 # in favor of the relevance only model over the full model for XOR
-bayes_factor(model_rel_comp, model_SI, silent = TRUE)
+bf_rel_comp_full <- bayes_factor(model_rel_comp, model_SI, silent = TRUE)
 # in favor of the relevance model over the prior model
-bayes_factor(model_rel_comp, model_prior_comp, silent = TRUE)
+bf_rel_comp_pri_comp <- bayes_factor(model_rel_comp, model_prior_comp, silent = TRUE)
 # and in vice versa in favor of the prior model over the relevance model
-bayes_factor(model_prior_comp, model_rel_comp, silent = TRUE)
+bf_pri_comp_rel_comp <- bayes_factor(model_prior_comp, model_rel_comp, silent = TRUE)
 
 # for symmetry, in favor of model without competence over full model
-bayes_factor(model_rel_pri, model_SI, silent = TRUE)
-bayes_factor(model_rel_pri, model_rel_comp, silent = TRUE)
-bayes_factor(model_rel_pri, model_prior_comp, silent = TRUE)
+bf_rel_pri_full <- bayes_factor(model_rel_pri, model_SI, silent = TRUE)
+bf_rel_pri_rel_comp <- bayes_factor(model_rel_pri, model_rel_comp, silent = TRUE)
+bf_rel_pri_pri_comp <- bayes_factor(model_rel_pri, model_prior_comp, silent = TRUE)
 
+```
+
+```{r}
+# try to make table by extracting numbers here
+
+bf_results <- tribble(
+  ~"Model in favor", ~"Model against", ~"BF",
+  "Prior*Comp", "Pri*Comp*Rel", round(bf_pri_comp_full[1]$bf, 1),
+  "Rel*Comp", "Pri*Comp*Rel", round(bf_rel_comp_full[1]$bf, 1),
+  "Rel*Comp", "Pri*Comp", round(bf_rel_comp_pri_comp[1]$bf, 1),
+  "Prior*Comp", "Rel*Comp", round(bf_pri_comp_rel_comp[1]$bf, 1),
+  "Rel*Pri", "Pri*Comp*Rel", round(bf_rel_pri_full[1]$bf, 1),
+  "Rel*Pri", "Rel*Comp", round(bf_rel_pri_rel_comp[1]$bf, 1),
+  "Rel*Pri", "Pri*Comp", round(bf_rel_pri_pri_comp[1]$bf, 1)
+)
+bf_results
+
+write_csv(bf_results, "../../data/main/bf_main_comparisons.csv")
 ```
@@
 # intercept only model
-priors_int <- set_prior("student_t(3, 0.1, 2.5)", class = "Intercept")
+priors_int <- set_prior("student_t(1, 0, 0.25)", class = "Intercept")
 model_int <- brm(target ~ 1 +
@@
 ```{r}
-bayes_factor(model_pri, model_int, silent = TRUE)
-bayes_factor(model_comp, model_int, silent = TRUE)
-bayes_factor(model_rel, model_int, silent = TRUE)
+bf_pri_int <- bayes_factor(model_pri, model_int, silent = TRUE)
+bf_comp_int <- bayes_factor(model_comp, model_int, silent = TRUE)
+bf_rel_int <- bayes_factor(model_rel, model_int, silent = TRUE)
 ```
@@
 ```{r}
-bayes_factor(model_prior_comp, model_pri, silent = TRUE)
-bayes_factor(model_prior_comp, model_comp, silent = TRUE)
+bf_pri_comp_pri <- bayes_factor(model_prior_comp, model_pri, silent = TRUE)
+bf_pri_comp_comp <- bayes_factor(model_prior_comp, model_comp, silent = TRUE)
 
-bayes_factor(model_rel_pri, model_pri, silent = TRUE)
-bayes_factor(model_rel_pri, model_rel, silent = TRUE)
+bf_rel_pri_pri <- bayes_factor(model_rel_pri, model_pri, silent = TRUE)
+bf_rel_pri_rel <- bayes_factor(model_rel_pri, model_rel, silent = TRUE)
 
-bayes_factor(model_rel_comp, model_rel, silent = TRUE)
-bayes_factor(model_rel_comp, model_comp, silent = TRUE)
+bf_rel_comp_rel <- bayes_factor(model_rel_comp, model_rel, silent = TRUE)
+bf_rel_comp_comp <- bayes_factor(model_rel_comp, model_comp, silent = TRUE)
 ```
+
+```{r}
+
+bf_results_explore <- tribble(
+  ~"Model in favor", ~"Model against", ~"BF",
+  "Prior", "Int", round(bf_pri_int[1]$bf, 1),
+  "Comp", "Int", round(bf_comp_int[1]$bf, 1),
+  "Rel", "Int", round(bf_rel_int[1]$bf, 1),
+  "Prior*Comp", "Prior", round(bf_pri_comp_pri[1]$bf, 1),
+  "Prior*Comp", "Comp", round(bf_pri_comp_comp[1]$bf, 1),
+  "Rel*Pri", "Prior", round(bf_rel_pri_pri[1]$bf, 1),
+  "Rel*Pri", "Rel", round(bf_rel_pri_rel[1]$bf, 1),
+  "Rel*Comp", "Rel", round(bf_rel_comp_rel[1]$bf, 1),
+  "Rel*Comp", "Comp", round(bf_rel_comp_comp[1]$bf, 1)
+)
+bf_results_explore
+
+write_csv(bf_results_explore, "../../data/main/bf_expl_comparisons.csv")
+
+bf_results %>% 
+  rbind(., bf_results_explore) %>%
+  write_csv(., "../../data/main/bf_full_comparisons.csv")
+```

### 2021-12-04 analysis/main/05_xor-some_analysis_model_compariso_grid-search-SD.Rmd
grid search over prior SD for rel vs pri, interactions BFs

---
title: "Xor-Some Preregistered Study: Analyses"
author: "Polina Tsvilodub"
date: "8/31/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(brms)
library(tidyboot)
library(tidybayes)
```

Read data created in preprocessing script:
```{r, message=FALSE, warning=FALSE, results='hide'}
d_critical_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")

d_critical_zScored_wide_xor_prior_order <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors-order.csv")

d_critical_zScored_wide_xor_prior_type <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors-type.csv")

d_critical_zScored_wide_xor <- d_critical_zScored_wide %>% filter(main_type == "xor")
```

## Comparison of prior and relevance effects

Since we observed potential colinearity effects in the main preregistered analysis for the trigger "or", we further explore the effects of this colinearity on the strength of the individual effects in the regression models.

To do so, we compute individual models omitting one of the factors on the "or" data, and compute the Bayes Factor (BF) of the model omitting one of the colinear predictors, relative to the full model.

Upon the [first comparison](https://github.com/magpie-ea/magpie-xor-experiment/blob/master/analysis/main/04_xor-some_analysis_model_comparison.md) of different models, we decide to focus on pitting the addition of the relevance effect against the addition of the prior effect to the competence effect. For more stable results, we conduct a little grid search over the SD parameter of the prior for the main effect regression coefficients, since the BF might be quite sensitive to a particular configuration. The range of the SD parameter is motivated by the theoretically informed expected effect size.

```{r}
# define a convenience function for iterating over the SDs and critical models
models_grid_search_SD = function(sd_vec, iter) {
  # create output df
  results <- tribble(
    ~"model in favor", ~"model against", ~"SD", ~"BF"
  )
  # set priors
  for (i in sd_vec) {
    dist <- paste("student_t(1, 0, ", i, ")", sep = "")
    prior <- set_prior(dist, class = "b")
    # fit rel * comp model with specific prior
    model_rel_comp <- brm(target ~ relevance*competence +
                   (1 + relevance + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = prior,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = iter,
                 save_all_pars = TRUE
                 )
    # fit prior * comp model with specific prior
    model_prior_comp <- brm(target ~ prior*competence +
                   (1 + prior + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = prior,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = iter,
                 save_all_pars = TRUE
                 )
    # compute BF
    bf <- bayes_factor(model_rel_comp, model_prior_comp, silent = TRUE)
    print(paste("SD: ", i, sep=""))
    print(paste("BF: ", bf, sep=""))
    # append to results dataframe, write out
    results <- results %>% add_row(
      "model in favor" = "Rel * Comp",
      "model against" = "Prior * Comp",
      SD = i,
      BF = round(bf[1]$bf, 4)
    )
    # dot might be a problem
    write_csv(results, 
              paste("../../data/main/grid_rel-pri_SD_0_", i*100, ".csv", sep = ""))
  }
  return (results)
}
```

Call the grid search:
```{r, message=FALSE, warning=FALSE}
grid_results <- models_grid_search_SD(c(0.05, 0.1, 0.15), 40000)

```

```{r}
grid_results

write_csv(grid_results, "../../data/main/grid_rel-pri_SD_final.csv")
```

Plot the effect of SD against the BF:
```{r}
grid_results_final <- read_csv("../../data/main/grid_rel-pri_SD_final.csv")

grid_results_final %>%
ggplot(., aes(x = SD, y = BF, color = `model in favor`)) +
  geom_point() +
  geom_line() +
  ggtitle("BF in favor of including a relevance effect over\n including a prior effect, depending on SD of prior t-distribution")

ggsave("BF_vs_SD_rel-comp_pri-comp.pdf", width = 6, height = 4)
```

For exploration, we also compute the BF in favor of a model with an interaction between the two effects against the same model without interaction estimates.
```{r, results='hide', message=FALSE, warning=FALSE}
# use rather tight prior
priors <- set_prior("student_t(1, 0, 0.1)", class = "b")
# model with prior effect only
model_prior_comp_int <- brm(target ~ prior*competence +
                   (1 + prior + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE
                 )
model_prior_comp <- brm(target ~ prior + competence +
                   (1 + prior + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE
                 )
bf_pri_comp_int <- bayes_factor(model_prior_comp_int, model_prior_comp, silent = TRUE)
```

```{r}
summary(model_prior_comp_int)
summary(model_prior_comp)
bf_pri_comp_int
```

```{r}
model_rel_comp_int <- brm(target ~ relevance*competence +
                   (1 + relevance + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE
                 )

model_rel_comp <- brm(target ~ relevance + competence +
                   (1 + relevance + competence || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.97),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE
                 )
bf_rel_comp_int <- bayes_factor(model_rel_comp_int, model_rel_comp, silent = TRUE)
```

```{r}
summary(model_rel_comp)
summary(model_rel_comp_int)
bf_rel_comp_int
```

For purpose of exploration, we also compare the full model with all interactions to a model without interactions, both with rather tight priors:
```{r, results='hide', message=FALSE, warning=FALSE}
# Full model 
# set priors
priors <- set_prior("student_t(1, 0, 0.1)", class = "b")

model_SI_ints <- brm(target ~ prior*competence*relevance +
                   (1 + prior + competence + relevance || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.95),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE
                )

model_SI <- brm(target ~ prior + competence + relevance +
                   (1 + prior + competence + relevance || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide_xor,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.95),
                 cores = 4,
                 iter = 40000,
                 save_all_pars = TRUE
                )

bf_ints <- bayes_factor(model_SI_ints, model_SI, silent = TRUE)
```

```{r}
summary(model_SI)
summary(model_SI_ints)
bf_ints
```

Export exploration results to csv:
```{r}
results_ints <- tribble(
  ~"Model in favor", ~"Model against", ~"BF", ~"SD",
  "Pri * Comp", "Pri + Comp", bf_pri_comp_int[1]$bf, "0.1",
  "Rel * Comp", "Rel + Comp", bf_rel_comp_int[1]$bf, "0.1",
  "Rel * Pri * Comp", "Rel + Pri + Comp", bf_ints[1]$bf, "0.1"
)
results_ints 
results_ints %>% write_csv(., "../../data/main/bf_interactions_comparisons.csv")
```

### 2021-12-06 05_views.js
remove identifying info from welcome page

@@ -28,8 +28,7 @@ const intro = magpieViews.view_generator("intro", {
             <br />
             <br />
             Legal information:
-            By answering the following questions, you are participating in a study
-            being performed by scientists from the University of Osnabrück.
+            By answering the following questions, you are participating in a scientific study.
             <br />
             <br />
             You must be at least 18 years old to participate.

### 2021-12-11 writing/ELM2/abstract.tex
update model comparison priors
grid search over prior SD for rel vs pri, interactions BFs
port over BvT's changes, slightly modified

@@
 Similarly, given the sentence "Donald ate a donut or a pretzel." one might infer an exclusive interpretation, i.e., that Donald ate either \textit{the donut or the pretzel, but not both}.
-Both inferences are considered as instances of \textit{scalar implicature (SI)}---an inference based on the comparison to a salient informationally stronger alternative ``all'' or ``and'', respectively [2,3].
-That is, the listener might arrive at these interpretations by reasoning about the epistemic state of a cooperative rational speaker [4].
+Both of these inferences are usually explained as a variety of \textit{scalar implicature (SI)}.
+SIs rely on lexical scales consisting of words that are ordered in terms of informativeness, like $\langle$some, all$\rangle$ and $\langle$or, and$\rangle$.
+If a speaker uses an informationally weaker term (e.g., ``some''), they may imply that the corresponding stronger alternative (e.g., ``all'') is false [2,3].
 %: if the stronger alternative was true, a rational agent would have used the stronger utterance; but since she did not, she must have lacked the evidence for using ``all''
@@
 %Some prior work addresses the influence of sentence structure, or of specific linguistic markers on the robustness of SIs (Li, 2021). Moreover, 
-Literature suggests that, among others, three factors influence the robustness of scalar implicatures: (1) \textit{relevance} of the stronger alternative to the listener, (2) the \textit{competence} of the speaker about the truth of the stronger alternative, and (3) the \textit{prior probability that the stronger alternative is true} [1,3,5].
+Crucially, prior research suggests that the robustness of SIs varies across contexts [6].
+Here, we investigate the effects of three factors that have been argued to influence the robustness of SIs: (1) \textit{relevance} of the stronger alternative to the listener, (2) the \textit{competence} of the speaker about the truth of the stronger alternative, and (3) the \textit{prior probability} that the stronger alternative is true [1,3,5].
 However, their influence was mostly investigated in isolation in prior work.
-We aim to fill this gap by exploring how these three factors influence the robustness of scalar implicatures in combination, when the SI triggers ``some'' and ``or'' are presented in rich context.
-Simultanously, we directly compare whether the two triggers are both influenced by these factors as predicted by the SI account.
+We explore how these three factors interactively influence the robustness of SIs associated with the triggers ``some'' and ``or''.
+Simultanously, we directly compare whether the two triggers are both influenced by these factors.
 
-In our web-based rating study (LINK), participants read background stories which were designed to score either high or low with respect to each factor of interest (high/low prior probability of the more informative alternative $\times$ high/low speaker competence $\times$ high/low contextual relevance of the alternative $\times$ triggers), manipulated within-subjects.
-On critical trials, participants were asked to rate four sentences on a scale ranging from ``certainly true'' to ``certainly false'' (converted to 0--100), one per factor and one containing an upper-bounded (or exclusive) paraphrase of the trigger.
-The sentences elicited judgements of how likely it is that each factor was high, and how likely the critical scalar implicature was true (s.
-transcript).
-If the SI account is true, on average, we expect higher likelihood ratings for the scalar implicature statement, if the relevance and comptenece of the item were judged as high and the prior of the stronger alternative as low.
-Each participant saw eight stories (four per trigger) sampled from 32 stories per trigger, randomly shuffled with eight visually identical attention checks and comprehension questions.
+In our web-based rating study, participants read background stories which were designed to vary in terms of the strength of the three factors of interest (prior probability $\times$ competence $\times$ relevance, for each trigger), manipulated within-subjects.
+On critical trials, participants were asked to rate three sentences on a scale ranging from ``certainly true'' to ``certainly false'' (converted to 0--100), one per factor.
+The background story ended with
+one of the characters in the story making an utterance containing ``some'' or ``or''. Participants then
+had to indicate the probability of an SI-enriched paraphrase of that utterance. We thus obtained
+judgements on the contextual factors, and on the robustness of the SI (s. \url{https://tinyurl.com/3ru9sdja} for transcripts). Based on
+the literature, we expected higher likelihood ratings for the SI-enriched paraphrase if the
+alternative was perceived as highly relevant, the speaker was judged as highly competent, and
+the stronger alternative was viewed as a priori unlikely. Each participant saw eight stories (four per trigger)
+sampled from 32 stories/trigger, randomly shuffled with eight structurally similar attention checks
+and comprehension questions.
 
@@
 %Consistent with predictions of the SI account, for the trigger ``some'', we find a clear negative prior probabilty effect of the stronger alternative ``all'' being true (probability of the effect of prior being smaller than 0.05: $P = 0.999$).
-As predicted by the SI account, participants are more likely to infer the ``some, but not al'' reading when they judged the prior probability of ``all'' being true as low ($P=0.999$ for the effect being $<0.05$, Fig.\ref{posteriors} for all results).
-They also robustly infer the upper-bounded interpretation when judging speaker competence as high ($P=1$ for effect $>0.05$).
-%Similarly, we find a clear positive effect for speaker competence (effect larger than 0.05 probability: $P =  1$). We do not find any clear effects of relevance. 
-However, participants were only more likely to infer the ``A or B, but not both'' reading when they rated the speaker competence as high  ($P=0.993$ for effect $>0.05$).
-No relevance effects were found for both triggers.
+As predicted, participants were more likely to derive the SIs of ``some'' and ``or'' when judging speaker competence as high ($P=0.999$ for ``some'',  $P=0.993$ for ``or'' for the effect being $<0.05$, Fig.\ref{posteriors} for all results).
+Participants were also more likely to derive the SI of ``some'' when judging the prior probability of ``all'' as low ($P =  1$ for effect $<0.05$), which was not the case for ``or''.
+%Interestingly, we did not observe a credible effect of prior probability in the case of ``or''.
+Finally, we did not observe credible effects of relevance for either trigger.
+
 %For the trigger ``or'', we only find a positive effect of competence ($P =  0.993$). 
@@
 This suggests that these factors might be interdependent in case of ``or''.
-Overall, our results support the SI account, but also indicate that these factors might have a more fine-grained influence on the robustness of SI generation in combination.%, especially for disjunction interpretation.
+
+These results partly confirm the predictions. First, the effects of competence show that both types
+of SIs rely on epistemic reasoning. However, only in the case of “some” did the robustness of
+the SI vary with the prior likelihood that the alternative is true. Although we did not find such an
+effect of prior probability for “or”, exploratory analyses indicate that this null effect may be due to a
+correlation with the other two factors. Perhaps most forcefully, our results speak against the idea
+that SIs are modulated by considerations of relevance. Our study thus provides insight into the
+contextual features that enter into people’s decision as to whether or not derive an SI.
+%Overall, our results support the SI account, but also indicate that these factors might have a more fine-grained influence on the robustness of SI generation in combination.%, especially for disjunction interpretation.
 %Overall, the results provide moderate evidence for the scalar implicature account of ``some'' and ``or'', but also indicate that there might be intricate dependencies between the factors' influence on the robustness of SI, especially for disjunction interpretation.
@@
 
-\begin{figure}
-   \makebox[\linewidth][c]{\includegraphics[width=0.7\linewidth]{byItem-ratings.pdf}}%
+\begin{SCfigure}[1.7]
+  % \makebox[\linewidth][c]{\includegraphics[width=0.7\linewidth]{byItem-ratings.pdf}}%
+  \includegraphics[scale=0.5]{byItem-ratings.pdf}
  \caption{Relating ratings for relevance, competence and prior statements (x-axis) to
@@
     \label{by-item-ratings}
-\end{figure}
+\end{SCfigure}
 
@@
 3: Speech acts}, 1975 [3] Horn, L.
-R., 1972 [5] Sperber, D., \& Wilson, D., \textit{Relevance: communication and cognition}, 1995
+R., \textit{On the semantic properties of logical operators in English}, 1972 [5] Sperber, D., \& Wilson, D., \textit{Relevance: communication and cognition}, 1995
     }

### 2021-12-11 analysis/main/02_xor-some_prereg_plots_final.Rmd
update narrative, update plot layout and make descriptive by predictor plots

--- 2021-11-14 analysis/main/02_xor-some_prereg_plots_final.Rmd
@@
 left_join(d_critical_predictorsXtarget_summary, d_critical_prior_classes, by = c("title", "predictors" = "class_condition")) %>%
-  mutate(prior_class = as.factor(prior_class)) %>%
+  mutate(`Prior categorization` = factor(prior_class, levels = c(0,1),
+                                         labels=c("low", "high")),
+         Trigger = factor(main_type, levels = c("some", "xor"), 
+                            labels = c("Some", "Or")),
+         predictors = factor(predictors, levels = c("relevance", "competence", "prior"),
+                             labels = c("Relevance", "Competence", "Prior"))) -> df_by_predictor_4_plotting
+
+df_by_predictor_4_plotting %>%
   ggplot(., aes(x = mean_predictor, y = mean_target)) +
-  geom_point(aes(color = prior_class)) +
-  ggrepel::geom_text_repel(aes(label=title), size = 2.5) +
+  geom_point(aes(color = `Prior categorization`)) +
+  ggrepel::geom_text_repel(aes(label=title), size = 3.5) +
   geom_smooth(method = "lm") +
@@
   xlab("Mean predictor ratings") + 
-  facet_wrap(main_type~predictors) +
-  scale_fill_manual(name = "Prior classification of item")
+  facet_grid(cols=vars(predictors), rows=vars(Trigger)) +
+  theme_bw() +
+  theme(legend.position = "top",
+        legend.text = element_text(size = 10),
+        axis.text = element_text(size = 10),
+        strip.text.x = element_text(size = 15),
+        strip.text.y = element_text(size = 15))
+
+ggsave("../../writing/ELM2/byItem-ratings.pdf", width = 9, height = 7)
+```
+
+```{r, fig.height=10, fig.width=8}
+df_by_predictor_4_plotting %>% 
+  group_by(`Prior categorization`, Trigger, predictors) %>%
+  summarize(mean = mean(mean_predictor)) -> by_item_predictors_summary
+
+df_by_predictor_4_plotting %>%
+  ggplot(., aes(x = `Prior categorization`, y = mean_predictor)) +
+  geom_point(aes(color = Trigger, fill = Trigger), size = 3.5, alpha = 0.8,  position = position_jitter(width = 0.2)) +
+  #geom_col(data = by_item_predictors_summary, aes(y = mean, x = `Prior categorization`), alpha = 0.3) +
+  ylab("Mean predictor strength rating") +
+  xlab("Prior categorization of story w.r.t. predictor") + 
+  facet_grid(cols=vars(predictors)) + # , rows=vars(main_type)
+  theme_bw() +
+  theme(legend.position = "top",
+        legend.text = element_text(size = 10),
+        axis.text = element_text(size = 10),
+        strip.text.x = element_text(size = 15),
+        strip.text.y = element_text(size = 15))
+
+ggsave("../../writing/ELM2/byItem-scatter.pdf", width = 9, height = 5)
+
+```
+
+```{r}
+df_by_predictor_4_plotting %>%
+  ggplot(., aes(x = `Prior categorization`, y = mean_predictor)) +
+  geom_point(aes(color = Trigger, fill = Trigger), size = 3.5, alpha = 0.8,  position = position_jitter(width = 0.2)) +
+  ggrepel::geom_text_repel(aes(label=title), size = 3.5) +
+  ylab("Mean predictor strength rating") +
+  xlab("Prior categorization of story w.r.t. predictor") + 
+  facet_grid(cols=vars(predictors), rows=vars(Trigger)) +  
+  theme_bw() +
+  theme(legend.position = "top",
+        legend.text = element_text(size = 10),
+        axis.text = element_text(size = 10),
+        strip.text.x = element_text(size = 15),
+        strip.text.y = element_text(size = 15))
+
+ggsave("../../writing/ELM2/byItem-scatter-wTitles.pdf", width = 9, height = 7)
+
 ```

### 2021-12-11 writing/ELM2/abstract.tex
update narrative, update plot layout and make descriptive by predictor plots

@@
 
-If someone says "Anna ate some cookies." the hearer of this sentence might infer the upper-bounded reading that Anna ate \textit{some, but not all} cookies.
-Similarly, given the sentence "Donald ate a donut or a pretzel." one might infer an exclusive interpretation, i.e., that Donald ate either \textit{the donut or the pretzel, but not both}.
-Both of these inferences are usually explained as a variety of \textit{scalar implicature (SI)}.
+If someone says ``Anna ate some cookies'', the hearer of the sentence might infer the upper-bounded reading that Anna ate \textit{some, but not all} cookies.
+Similarly, given ``Donald ate a donut or a pretzel.'', one might infer that Donald ate either \textit{the donut or the pretzel, but not both} (i.e., an exclusive interpretation).
+Both inferences are usually explained as a variety of \textit{scalar implicature} (SI).
 SIs rely on lexical scales consisting of words that are ordered in terms of informativeness, like $\langle$some, all$\rangle$ and $\langle$or, and$\rangle$.
@@
 Crucially, prior research suggests that the robustness of SIs varies across contexts [6].
-Here, we investigate the effects of three factors that have been argued to influence the robustness of SIs: (1) \textit{relevance} of the stronger alternative to the listener, (2) the \textit{competence} of the speaker about the truth of the stronger alternative, and (3) the \textit{prior probability} that the stronger alternative is true [1,3,5].
-However, their influence was mostly investigated in isolation in prior work.
-We explore how these three factors interactively influence the robustness of SIs associated with the triggers ``some'' and ``or''.
-Simultanously, we directly compare whether the two triggers are both influenced by these factors.
+We investigate the effects of three factors that have been argued to influence the robustness of SIs: (1) \textit{relevance} of the stronger alternative to the listener, (2) the \textit{competence} of the speaker about the truth of the stronger alternative, and (3) the \textit{prior probability} that the stronger alternative is true [1,3,5].
+Critically, we explore how these three factors \textit{interactively} influence the robustness of SIs associated with the triggers ``some'' and ``or''.
+%Simultanously, we directly compare whether the two triggers are both influenced by these factors.
 
-In our web-based rating study, participants read background stories which were designed to vary in terms of the strength of the three factors of interest (prior probability $\times$ competence $\times$ relevance, for each trigger), manipulated within-subjects.
+In our web-based rating study, participants read background stories which were designed to vary in terms of the strength of the three factors (\texttt{high} or \texttt{low} with respect to: prior probability$\times$competence$\times$ relevance, for each trigger), manipulated within-subjects.
 On critical trials, participants were asked to rate three sentences on a scale ranging from ``certainly true'' to ``certainly false'' (converted to 0--100), one per factor. 
-The background story ended with
-one of the characters in the story making an utterance containing ``some'' or ``or''. Participants then
-had to indicate the probability of an SI-enriched paraphrase of that utterance. We thus obtained
-judgements on the contextual factors, and on the robustness of the SI (s. \url{https://tinyurl.com/3ru9sdja} for transcripts). Based on
-the literature, we expected higher likelihood ratings for the SI-enriched paraphrase if the
-alternative was perceived as highly relevant, the speaker was judged as highly competent, and
-the stronger alternative was viewed as a priori unlikely. Each participant saw eight stories (four per trigger)
-sampled from 32 stories/trigger, randomly shuffled with eight structurally similar attention checks
-and comprehension questions.
+The story ended with one of the characters in the story making an utterance containing ``some'' or ``or''.
+Participants then had to indicate the probability of an SI-enriched paraphrase of that utterance.
+We thus obtained judgements on the contextual factors, and on the robustness of the SI (s. \url{https://tinyurl.com/3ru9sdja} for transcripts). 
+Based on the literature, we expected higher likelihood ratings for the SI-enriched paraphrase if the alternative was perceived as highly relevant, the speaker was judged as highly competent, and the stronger alternative was viewed as a priori unlikely.
+Each participant saw four stories per trigger, sampled from 32 stories/trigger, randomly shuffled with eight structurally similar attention checks and comprehension questions.
 
 %(71  excluded following preregistered exclusion criteria).
-We analysed data from 206 participants recruited through Prolific; their ratings were z-scored within each factor by-participant.
-We analysed the results using a Bayesian linear mixed effects model, regressing the implicature likelihood rating against the fixed effects of all factor ratings by the same participant for that vignette, the effect of trigger, all interactions and maximal tractable random effects structure.
-Participants' factor ratings agreed well with the prior classification of the items (Fig.~\ref{by-item-ratings}, colors).
+We analysed data from 206 participants recruited through Prolific.
+Their ratings were z-scored within each factor by-participant.
+We analysed the results using a Bayesian linear mixed effects model, regressing the implicature likelihood rating against the fixed effects of all factor ratings within-story, the effect of trigger, all interactions and maximal tractable random effects.
+Participants' factor ratings by-story agreed well with the designed classification of the stories (Fig.~\ref{by-item-ratings}, red~vs.~blue color on x-axis).
 %Consistent with predictions of the SI account, for the trigger ``some'', we find a clear negative prior probabilty effect of the stronger alternative ``all'' being true (probability of the effect of prior being smaller than 0.05: $P = 0.999$).
@@
 Finally, we did not observe credible effects of relevance for either trigger.
+%For the trigger ``or'', we only find a positive effect of competence ($P =  0.993$). 
+Given the unexpected absence of prior effects for ``or'', we computed exploratory pairwise correlations of all the predictors.
+While no correlations were found for ``some'', we found a significant correlation between the prior and relevance effects for ``or'' ($R^2 = -0.106, p<0.01$).
+%, and a significant correlation between competence and relevance ($R^2 = 0.127, p=0.001$) for ``or''.
+An exploratory model comparison of two models, one containing the relevance effect over one containing the prior effect, each combined with competence, revealed mild evidence in favor of the prior as a better explanatory factor than relevance (Bayes Factor $ = 0.27$).
+%This might be due to the difficulty of accessing relevance experimentally.
+While supporting the view that SIs for both triggers rely on epistemic reasoning affected by speaker competence, our results indicate that prior and relevance might be closely connected for the disjunction.
+Ultimately, our results suggest that the interdependence of the three factors is more complex than just the sum of the effects anticipated in the literature. 
+Our study thus outlines further directions for investigating how exactly they enter into people’s decision as to whether or not derive an SI.
 
-%For the trigger ``or'', we only find a positive effect of competence ($P =  0.993$). 
-Further, we computed exploratory pairwise correlations of all the predictors.
-While no correlations were found for ``some'', we find a significant correlation between the prior and the relevance effects for ``or'' ($R^2 = -0.106, p=...$), and a significant correlation between competence and relevance ($R^2 = 0.127, p=...$) for ``or''.
-This suggests that these factors might be interdependent in case of ``or''.
-
-These results partly confirm the predictions. First, the effects of competence show that both types
-of SIs rely on epistemic reasoning. However, only in the case of “some” did the robustness of
-the SI vary with the prior likelihood that the alternative is true. Although we did not find such an
-effect of prior probability for “or”, exploratory analyses indicate that this null effect may be due to a
-correlation with the other two factors. Perhaps most forcefully, our results speak against the idea
-that SIs are modulated by considerations of relevance. Our study thus provides insight into the
-contextual features that enter into people’s decision as to whether or not derive an SI.
+%These results partly confirm the predictions.
+First, the effects of competence show that both types of SIs rely on epistemic reasoning.
+However, only in the case of “some” did the robustness of the SI vary with the prior likelihood that the alternative is true.
+Although we did not find such an effect of prior probability for “or”, exploratory analyses indicate that this null effect may be due to a correlation with the other two factors.
+Perhaps most forcefully, our results speak against the idea that SIs are modulated by considerations of relevance.
+Our study thus provides insight into the contextual features that enter into people’s decision as to whether or not derive an SI.
 %Overall, our results support the SI account, but also indicate that these factors might have a more fine-grained influence on the robustness of SI generation in combination.%, especially for disjunction interpretation.
@@
     \makebox[\linewidth][c]{\includegraphics[scale=0.7]{posterior-effects.pdf}}
-    \caption{Distributions of posterior samples for the effects of each predictor for each trigger (y-axis).
+    \caption{Distributions of posterior samples for the effects (x-axis) of each predictor for each trigger (y-axis).
 The colors indicate the fraction of the distribution density corresponding to a positive, no or a negative effect.
-\newline \textbf{References:} [1] Degen, J., Tessler, M.
+The dashed lines indicate the Region Of Practical Equivalence that we defined as (-0.05, 0.05) in which posterior samples are considered equivalent to 0 (i.e., no effect).\newline \textbf{References:} [1] Degen, J., Tessler, M.
 H., \& Goodman, N.
@@
 3: Speech acts}, 1975 [3] Horn, L.
-R., \textit{On the semantic properties of logical operators in English}, 1972 [5] Sperber, D., \& Wilson, D., \textit{Relevance: communication and cognition}, 1995
+R., \textit{On the semantic properties of logical operators in English}, 1972 [5] Sperber, D., \& Wilson, D., \textit{Relevance: communication and cognition}, 1995 [6] Degen, J., In \textit{Semantics \& Pragmatics}, 2015
     }

### 2021-12-14 analysis/main/02_xor-some_prereg_plots_final.Rmd
shorten abstract to 2 pages, add new raw data plot, add link directly to prereg

@@
 ```
+
+## Plots for ELM2
+Plot predictor ratings against inference strength ratings by trigger:
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# add the prior predictor categorization to wide df
+df_by_predictor_4_plotting %>% select(title, predictors, prior_class) %>% 
+  mutate(
+    predictors = tolower(predictors)
+    ) -> df_predictors_prior_classes
+
+d_critical_raw_wPrior <- d_critical_zScored_wide %>% 
+  pivot_longer(
+    cols = c(relevance, prior, competence),
+    names_to = "predictors",
+    values_to = "predictor_response"
+  ) %>%
+  left_join(., df_predictors_prior_classes, by = c("title", "predictors"))
+
+  d_critical_raw_wPrior_4plot <- d_critical_raw_wPrior %>%
+    mutate(
+      `Prior categorization` = factor(prior_class, levels = c(0,1),
+                                         labels=c("low", "high")),
+         Trigger = factor(main_type, levels = c("some", "xor"), 
+                            labels = c("Some", "Or")),
+         predictors = factor(predictors, levels = c("relevance", "competence", "prior"),
+                             labels = c("Relevance", "Competence", "Prior"))
+    )
+```
+
+```{r}
+d_critical_raw_wPrior_4plot %>%  
+  ggplot(., aes(x = predictor_response, y = target)) +
+  geom_point(aes(color = `Prior categorization`), size = 2, alpha = 0.8) +
+  geom_smooth(method = "lm") +
+  ylab("") +
+  facet_grid(cols=vars(predictors), rows=vars(Trigger)) +
+  theme_bw() +
+  labs(x = "Participants' factor ratings", y = "Participants' inference likelihood ratings", color = "Prior categorization of vignettes") +
+  theme(legend.position = "top",
+        legend.title = element_text(size = 13),
+        legend.text = element_text(size = 13),
+        axis.title = element_text(size = 13),
+        strip.text.x = element_text(size = 13),
+        strip.text.y = element_text(size = 13))
+
+ggsave("../../writing/ELM2/byFactor-byPrior-raw-scatter.pdf", width = 9, height = 7)
+```

### 2021-12-14 writing/ELM2/abstract.tex
shorten abstract to 2 pages, add new raw data plot, add link directly to prereg

@@
 
-If someone says ``Anna ate some cookies'', the hearer of the sentence might infer the upper-bounded reading that Anna ate \textit{some, but not all} cookies.
+If someone says ``Anna ate some cookies'', the hearer might infer the upper-bounded reading that Anna ate \textit{some, but not all} cookies.
 Similarly, given ``Donald ate a donut or a pretzel.'', one might infer that Donald ate either \textit{the donut or the pretzel, but not both} (i.e., an exclusive interpretation).
 Both inferences are usually explained as a variety of \textit{scalar implicature} (SI).
-SIs rely on lexical scales consisting of words that are ordered in terms of informativeness, like $\langle$some, all$\rangle$ and $\langle$or, and$\rangle$.
-If a speaker uses an informationally weaker term (e.g., ``some''), they may imply that the corresponding stronger alternative (e.g., ``all'') is false [2,3].
+SIs rely on lexical scales consisting of words ordered in terms of informativeness, like $\langle$some, all$\rangle$ and $\langle$or, and$\rangle$.
+If a speaker uses an informationally weaker term (e.g., ``some''), they may imply that the corresponding stronger alternative (e.g., ``all'') is false [3,5].
 %: if the stronger alternative was true, a rational agent would have used the stronger utterance; but since she did not, she must have lacked the evidence for using ``all''
@@
 %Some prior work addresses the influence of sentence structure, or of specific linguistic markers on the robustness of SIs (Li, 2021). Moreover, 
-Crucially, prior research suggests that the robustness of SIs varies across contexts [6].
-We investigate the effects of three factors that have been argued to influence the robustness of SIs: (1) \textit{relevance} of the stronger alternative to the listener, (2) the \textit{competence} of the speaker about the truth of the stronger alternative, and (3) the \textit{prior probability} that the stronger alternative is true [1,3,5].
-Critically, we explore how these three factors \textit{interactively} influence the robustness of SIs associated with the triggers ``some'' and ``or''.
-%Simultanously, we directly compare whether the two triggers are both influenced by these factors.
+Crucially, prior research suggests that the robustness of SIs is influenced by different contextual factors [1].
+We investigate the effects of three factors: (1) the \textit{competence} of the speaker about the truth of the stronger alternative, (2) its \textit{relevance} to the listener, and (3) the \textit{prior probability} that it is true [2,4,6].
+We explore how these three factors \textit{interactively} influence the robustness of SIs associated with the triggers ``some'' and ``or''.
 
 In our web-based rating study, participants read background stories which were designed to vary in terms of the strength of the three factors (\texttt{high} or \texttt{low} with respect to: prior probability$\times$competence$\times$ relevance, for each trigger), manipulated within-subjects.
-On critical trials, participants were asked to rate three sentences on a scale ranging from ``certainly true'' to ``certainly false'' (converted to 0--100), one per factor. 
+On critical trials, participants were asked to rate three sentences on a scale ranging from ``certainly true'' to ``certainly false'' ($\widehat{=}$0--100), one per factor. 
 The story ended with one of the characters in the story making an utterance containing ``some'' or ``or''.
 Participants then had to indicate the probability of an SI-enriched paraphrase of that utterance.
-We thus obtained judgements on the contextual factors, and on the robustness of the SI (s. \url{https://tinyurl.com/3ru9sdja} for transcripts). 
-Based on the literature, we expected higher likelihood ratings for the SI-enriched paraphrase if the alternative was perceived as highly relevant, the speaker was judged as highly competent, and the stronger alternative was viewed as a priori unlikely.
-Each participant saw four stories per trigger, sampled from 32 stories/trigger, randomly shuffled with eight structurally similar attention checks and comprehension questions.
+We thus obtained judgements on the contextual factors, and on the robustness of the SI (s. \url{https://tinyurl.com/3ru9sdja} for experiment details). 
+Based on the literature, we expected higher likelihood ratings for the SI-enriched paraphrase if the alternative was perceived as highly relevant, the speaker was judged as highly competent, and the stronger alternative was viewed as a priori unlikely [2,4,6].
+Each participant saw four stories per trigger, sampled from 32 stories/triggers, randomly shuffled with eight structurally similar attention checks and comprehension questions.
 
 %(71  excluded following preregistered exclusion criteria).
-We analysed data from 206 participants recruited through Prolific.
-Their ratings were z-scored within each factor by-participant.
-We analysed the results using a Bayesian linear mixed effects model, regressing the implicature likelihood rating against the fixed effects of all factor ratings within-story, the effect of trigger, all interactions and maximal tractable random effects.
-Participants' factor ratings by-story agreed well with the designed classification of the stories (Fig.~\ref{by-item-ratings}, red~vs.~blue color on x-axis).
+We analysed data from 206 participants recruited on Prolific. Their ratings were z-scored within each factor by-participant. 
+We regressed the implicature likelihood ratings against the fixed effects of trigger, all factor ratings within-story, all interactions and maximal  random effects, using a Bayesian linear mixed effects model.
+%We analysed the results using a Bayesian linear mixed effects model, regressing the implicature likelihood rating against the fixed effects of all factor ratings within-story, the effect of trigger, all interactions and maximal tractable random effects.
+Participants' factor ratings by-story agreed well with the designed classification of the stories (Fig.\ref{by-item-ratings}, red~vs.~blue color on x-axis).
 %Consistent with predictions of the SI account, for the trigger ``some'', we find a clear negative prior probabilty effect of the stronger alternative ``all'' being true (probability of the effect of prior being smaller than 0.05: $P = 0.999$).
-As predicted, participants were more likely to derive the SIs of ``some'' and ``or'' when judging speaker competence as high ($P=0.999$ for ``some'',  $P=0.993$ for ``or'' for the effect being $<0.05$, Fig.\ref{posteriors} for all results).
-Participants were also more likely to derive the SI of ``some'' when judging the prior probability of ``all'' as low ($P =  1$ for effect $<0.05$), which was not the case for ``or''.
+As predicted, participants were more likely to derive the SIs of ``some'' and ``or'' when judging speaker competence as high ($P=0.999$ for ``some'',  $P=0.993$ for ``or'' for effect sizes being $>0.05$, Fig.\ref{posteriors} for all results).
+They were also more likely to derive the SI of ``some'' when judging the prior probability of ``all'' as low ($P= 1$ for effect $<0.05$), which was not the case for ``or''.
 %Interestingly, we did not observe a credible effect of prior probability in the case of ``or''.
@@
 %For the trigger ``or'', we only find a positive effect of competence ($P =  0.993$). 
-Given the unexpected absence of prior effects for ``or'', we computed exploratory pairwise correlations of all the predictors.
-While no correlations were found for ``some'', we found a significant correlation between the prior and relevance effects for ``or'' ($R^2 = -0.106, p<0.01$).
+Given the unexpected absence of prior effects for ``or'', we computed exploratory pairwise correlations of all predictors.
+While no correlations were found for ``some'', we found a significant correlation between the explanatory factors prior and relevance for ``or'' ($R^2=-0.106, p<0.01$).
 %, and a significant correlation between competence and relevance ($R^2 = 0.127, p=0.001$) for ``or''.
@@
 %This might be due to the difficulty of accessing relevance experimentally.
-While supporting the view that SIs for both triggers rely on epistemic reasoning affected by speaker competence, our results indicate that prior and relevance might be closely connected for the disjunction.
+While supporting the view that SIs for both triggers rely on epistemic reasoning affected by speaker competence, our results indicate that prior and relevance might be closely connected for ``or''.
 Ultimately, our results suggest that the interdependence of the three factors is more complex than just the sum of the effects anticipated in the literature. 
-Our study thus outlines further directions for investigating how exactly they enter into people’s decision as to whether or not derive an SI.
+Our study thus outlines further directions for investigating how exactly they enter into people’s decision as to whether or not to derive an SI.
 
@@
   % \makebox[\linewidth][c]{\includegraphics[width=0.7\linewidth]{byItem-ratings.pdf}}%
-  \includegraphics[scale=0.5]{byItem-ratings.pdf}
- \caption{Relating ratings for relevance, competence and prior statements (x-axis) to
-ratings for the strength of pragmatic enrichments (y-axis). The top row shows ratings,
-averaged for each story for ``some'' (enriched to ``some, but not all''). The bottom row shows ratings, averaged for each story for ``or'' (enriched to ``A or B, but not both'').}
+  \includegraphics[scale=0.5]{byFactor-byPrior-raw-scatter.pdf}
+ \caption{Relating ratings for relevance, competence and prior statements (x-axis) to ratings for the strength of pragmatic enrichments (y-axis). 
+ The top row shows ratings for ``some'' (enriched to ``some, but not all'').
+The bottom row shows ratings for ``or'' (enriched to ``A or B, but not both'').
+Ratings for stories categorized as low (red) w.r.t.~a given factor are on average lower (x-axis) than for those categorized as high (blue).
+The apparent effect of prior for ``or'' is mitigated by collinearity with relevance.}
     \label{by-item-ratings}
@@
 
+%\begin{SCfigure}[0.6]%[0.6]
+%	\includegraphics[width=0.6\textwidth]{posterior-effects.pdf}
+%\includegraphics[scale=0.5]{posterior-effects.pdf}
+%	\caption{Distributions of posterior samples for the effects (x-axis) of each predictor for each trigger (y-axis).
+The colors indicate the fraction of the distribution density corresponding to a positive, no or a negative effect.
+The dashed lines indicate the Region Of Practical Equivalence that we defined as (-0.05, 0.05) in which posterior samples are considered equivalent to 0 (i.e., no effect).}
+%	\label{posteriors}
+%\end{SCfigure}
+%\textbf{References:} [1] Kamp, J., In \textit{Formal semantics of natural language}, 1975  [2] Kennedy, C., \textit{Linguistics and Philosophy, 30(1)}, 2007 [3] Tessler, M.
+H.; Lopez-Brau, M.; Goodman N.
+D., In \textit{39th annual meeting of the Cognitive Science society}, 2017
+%    \label{double-mod}[4] Reboul, A., In \textit{Language typology and language universals, an international handbook, vol.1}, 2001
 \begin{figure}   
-    \makebox[\linewidth][c]{\includegraphics[scale=0.7]{posterior-effects.pdf}}
+    \makebox[\linewidth][c]{\includegraphics[scale=0.65]{posterior-effects.pdf}}
     \caption{Distributions of posterior samples for the effects (x-axis) of each predictor for each trigger (y-axis).
 The colors indicate the fraction of the distribution density corresponding to a positive, no or a negative effect.
-The dashed lines indicate the Region Of Practical Equivalence that we defined as (-0.05, 0.05) in which posterior samples are considered equivalent to 0 (i.e., no effect).\newline \textbf{References:} [1] Degen, J., Tessler, M.
+The dashed lines indicate the Region Of Practical Equivalence that we defined as (-0.05, 0.05) in which posterior samples are considered equivalent to 0 (i.e., no credible effect).\newline \textbf{References:} [1] Degen, J., In \textit{Semantics \& Pragmatics}, 2015 [2] Degen, J., Tessler, M.
 H., \& Goodman, N.
-D., In \textit{Proceedings of CogSci}, 2015 [2] Geurts, B., \textit{Quantity Implicatures}, 2010 [3] Goodman, N.
-D., \& Stuhlm\"uller, A., In \textit{Topics in Cognitive Science}, 5(1), 2013 [4] Grice, H.
-P.
-In \textit{Syntax and semantics, vol.
-3: Speech acts}, 1975 [3] Horn, L.
-R., \textit{On the semantic properties of logical operators in English}, 1972 [5] Sperber, D., \& Wilson, D., \textit{Relevance: communication and cognition}, 1995 [6] Degen, J., In \textit{Semantics \& Pragmatics}, 2015
-    }
+D., In \textit{Proceedings of CogSci}, 2015 [3] Geurts, B., \textit{Quantity Implicatures}, 2010 [4] Goodman, N.
+D., \& Stuhlm\"uller, A., In \textit{Topics in Cognitive Science}, 5(1), 2013 [5] Horn, L.
+R., \textit{On the semantic properties of logical operators in English}, 1972 [6] Sperber, D., \& Wilson, D., \textit{Relevance: communication and cognition}, 1995}
     \label{posteriors}

### 2021-12-15 writing/ELM2/abstract.tex
fix typos in overview, final abstract tweaks

@@
 Participants then had to indicate the probability of an SI-enriched paraphrase of that utterance.
-We thus obtained judgements on the contextual factors, and on the robustness of the SI (s. \url{https://tinyurl.com/3ru9sdja} for experiment details). 
+We thus obtained judgements on the contextual factors, and on the robustness of the SI (see \url{https://tinyurl.com/3ru9sdja} for experiment details). 
 Based on the literature, we expected higher likelihood ratings for the SI-enriched paraphrase if the alternative was perceived as highly relevant, the speaker was judged as highly competent, and the stronger alternative was viewed as a priori unlikely [2,4,6].
@@
 %We analysed the results using a Bayesian linear mixed effects model, regressing the implicature likelihood rating against the fixed effects of all factor ratings within-story, the effect of trigger, all interactions and maximal tractable random effects.
-Participants' factor ratings by-story agreed well with the designed classification of the stories (Fig.\ref{by-item-ratings}, red~vs.~blue color on x-axis).
+Participants' factor ratings by-story agreed well with the designed classification of the stories (Fig.~\ref{by-item-ratings}, red~vs.~blue color on x-axis).
 %Consistent with predictions of the SI account, for the trigger ``some'', we find a clear negative prior probabilty effect of the stronger alternative ``all'' being true (probability of the effect of prior being smaller than 0.05: $P = 0.999$).
-As predicted, participants were more likely to derive the SIs of ``some'' and ``or'' when judging speaker competence as high ($P=0.999$ for ``some'',  $P=0.993$ for ``or'' for effect sizes being $>0.05$, Fig.\ref{posteriors} for all results).
+As predicted, participants were more likely to derive the SIs of ``some'' and ``or'' when judging speaker competence as high ($P=0.999$ for ``some'',  $P=0.993$ for ``or'' for effect sizes being $>0.05$, Fig.~\ref{posteriors} for all results).
 They were also more likely to derive the SI of ``some'' when judging the prior probability of ``all'' as low ($P= 1$ for effect $<0.05$), which was not the case for ``or''.
@@
 %, and a significant correlation between competence and relevance ($R^2 = 0.127, p=0.001$) for ``or''.
-An exploratory model comparison of two models, one containing the relevance effect over one containing the prior effect, each combined with competence, revealed mild evidence in favor of the prior as a better explanatory factor than relevance (Bayes Factor $ = 0.27$).
+An exploratory model comparison of two models, one containing the relevance effect over one containing the prior effect, each combined with competence, revealed mild evidence in favor of the prior as a better explanatory factor than relevance (Bayes Factor $ = 3.70$).
 %This might be due to the difficulty of accessing relevance experimentally.
 While supporting the view that SIs for both triggers rely on epistemic reasoning affected by speaker competence, our results indicate that prior and relevance might be closely connected for ``or''.
-Ultimately, our results suggest that the interdependence of the three factors is more complex than just the sum of the effects anticipated in the literature. 
-Our study thus outlines further directions for investigating how exactly they enter into people’s decision as to whether or not to derive an SI.
+Ultimately, our results suggest that the interdependence of the three factors is more complex than just the sum of the effects anticipated in the literature, and provide further insights into how exactly they might enter into people’s decision as to whether or not to derive an SI.
 

### 2021-12-20 analysis/main/03_xor-some_prereg_analysis_final.Rmd
full SD grid search vs Int

--- 2021-11-14 analysis/main/03_xor-some_prereg_analysis_final.Rmd
@@
   scale_fill_manual(name = "Direction of effect", values = c("coral", "yellow3", "mediumseagreen")) +
-  ggtitle("Posterior effect distributions. Points indicate posterior means\nThick lines indicate 66% CrIs, thin lines 95% CrIs.")
+  labs(x = "Effect size", y = "Effect") +
+  theme(legend.position = "top",
+        legend.title = element_text(size = 12),
+        legend.text = element_text(size = 12),
+        axis.title = element_text(size = 13),
+        strip.text.x = element_text(size = 13),
+        strip.text.y = element_text(size = 13))
 
+ggsave("../../writing/ELM2/posterior-effects.pdf", width=7, height=5)
 ```

### 2021-12-20 analysis/main/05_xor-some_analysis_model_compariso_grid-search-SD.Rmd
full SD grid search vs Int

@@
 ```
+
+## Model comparisons with grid search for final paper
+Below, we conduct a similar grid search over more models.
+More specifically, we compute the BF in favor of a more complex model over the intercept-only baseline model.
+We compute these pairwise comparisons for each iteratively added explaantory factor.
+This results in 7 model comparisons.
+For each comparison, we perform a small grid search over the SD parameter of the prior t-distribution over the fixed effects regression coefficients, using the values [0.05, 0.1, 0.15].
+For each model, 40.000 samples are extracted.
+```{r}
+# define a convenience function for iterating over the SDs and critical models
+models_grid_search_SD_full = function(sd_vec, iter, trigger) {
+  # create output df
+  results <- tribble(
+    ~"model in favor", ~"model against", "~trigger", ~"SD", ~"BF"
+  )
+  # filter data by trigger
+  data <- d_critical_zScored_wide %>% filter(main_type == trigger)
+  # set priors
+  for (i in sd_vec) {
+    dist <- paste("student_t(1, 0, ", i, ")", sep = "")
+    prior <- set_prior(dist, class = "b")
+    # fit the baseline intercept only model with same prior SD
+    prior_int <- set_prior(dist, class = "Intercept")
+    model_int <- brm(target ~ 1 +
+                   (1 | submission_id) +
+                   (1 | title),
+                 data = data,
+                 prior = prior_int,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.97),
+                 cores = 4,
+                 iter = iter,
+                 save_all_pars = TRUE
+                 )
+    # fit the single predictor models
+    model_pri <- brm(target ~ prior + 
+                   (1 + prior || submission_id) +
+                   (1 | title),
+                 data = data,
+                 prior = prior,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.97),
+                 cores = 4,
+                 iter = iter,
+                 save_all_pars = TRUE
+                 )
+    model_rel <- brm(target ~ relevance + 
+                   (1 + relevance || submission_id) +
+                   (1 | title),
+                 data = data,
+                 prior = prior,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.97),
+                 cores = 4,
+                 iter = iter,
+                 save_all_pars = TRUE
+                 )
+    model_comp <- brm(target ~ competence + 
+                   (1 + competence || submission_id) +
+                   (1 | title),
+                 data = data,
+                 prior = prior,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.97),
+                 cores = 4,
+                 iter = iter,
+                 save_all_pars = TRUE
+                 )
+    # fit rel * comp model with specific prior
+    model_rel_comp <- brm(target ~ relevance*competence +
+                   (1 + relevance + competence || submission_id) +
+                   (1 | title),
+                 data = data,
+                 prior = prior,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.97),
+                 cores = 4,
+                 iter = iter,
+                 save_all_pars = TRUE
+                 )
+    # fit prior * comp model with specific prior
+    model_pri_comp <- brm(target ~ prior*competence +
+                   (1 + prior + competence || submission_id) +
+                   (1 | title),
+                 data = data,
+                 prior = prior,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.97),
+                 cores = 4,
+                 iter = iter,
+                 save_all_pars = TRUE
+                 )
+    # rel*pri model
+    model_rel_pri <- brm(target ~ prior*relevance +
+                   (1 + prior + relevance || submission_id) +
+                   (1 | title),
+                 data = data,
+                 prior = prior,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.97),
+                 cores = 4,
+                 iter = iter,
+                 save_all_pars = TRUE
+                 )
+    # fit full model
+    model_pri_rel_comp <- brm(target ~ prior*competence*relevance +
+                   (1 + prior + competence + relevance || submission_id) +
+                   (1 | title),
+                 data = data,
+                 prior = prior,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.97),
+                 cores = 4,
+                 iter = iter,
+                 save_all_pars = TRUE
+                 )
+    # compute BF
+    # iterate over list of models 
+    model_list <- c("model_rel", "model_pri", "model_comp", "model_rel_comp", "model_pri_comp", "model_rel_pri", "model_pri_rel_comp")
+    for (m in model_list) {
+      bf <- bayes_factor(m, model_int, silent = TRUE)
+      print(paste("SD: ", i, sep=""))
+      print(paste("Model in favor: ", m, sep=""))
+      print(paste("BF: ", bf, sep=""))
+      # append to results dataframe, write out
+      results <- results %>% add_row(
+        "model in favor" = m,
+        "model against" = "Int",
+        "trigger" = trigger,
+        SD = i,
+        BF = round(bf[1]$bf, 4)
+      )
+    write_csv(results, 
+              paste("../../data/main/grid_full_", trigger, "_SD_0_", i*100, ".csv", sep = ""))
+    }
+
+  }
+  return (results)
+}
+```
+
+```{r}
+# call the grid search for xor 
+grid_results_full_xor <- models_grid_search_SD_full(c(0.05, 0.1, 0.15), 40000, "xor")
+
+# call the grid search for some 
+grid_results_full_some <- models_grid_search_SD_full(c(0.05, 0.1, 0.15), 40000, "some")
+
+# write out results
+write_csv(grid_results_full_xor, "../../data/main/grid_full_xor_SD_final.csv")
+write_csv(grid_results_full_some, "../../data/main/grid_full_some_SD_final.csv")
+
+```
+
+Plot the effect of SD against the BF:
+```{r}
+grid_results_full_xor <- read_csv("../../data/main/grid_full_xor_SD_final.csv")
+grid_results_full_some <- read_csv("../../data/main/grid_full_some_SD_final.csv")
+grid_results_both <- rbind(grid_results_full_xor, grid_results_full_some)
+
+grid_results_both %>%
+ggplot(., aes(x = SD, y = BF, color = `model in favor`)) +
+  geom_point() +
+  geom_line() +
+  facet_wrap(~trigger) + 
+  ggtitle("BF in favor of including a relevance effect over\n including a prior effect, depending on SD of prior t-distribution")
+
+#ggsave("BF_vs_SD_all_vs_int.pdf", width = 8, height = 4)
+```
+
+```{r}
+
+# bf <- bayes_factor(model_rel_comp, model_prior_comp, silent = TRUE)
+#     print(paste("SD: ", i, sep=""))
+#     print(paste("BF: ", bf, sep=""))
+#     # append to results dataframe, write out
+#     results <- results %>% add_row(
+#       "model in favor" = "Rel * Comp",
+#       "model against" = "Prior * Comp",
+#       SD = i,
+#       BF = round(bf[1]$bf, 4)
+#     )
+#     # dot might be a problem
+#     write_csv(results, 
+#               paste("../../data/main/grid_rel-pri_SD_0_", i*100, ".csv", sep = ""))
+```

### 2021-12-29 analysis/main/05_xor-some_analysis_model_compariso_grid-search-SD.Rmd
full SD grid search for all xor models vs Int only model

--- 2021-12-04 analysis/main/05_xor-some_analysis_model_compariso_grid-search-SD.Rmd
@@
 ---
-title: "Xor-Some Preregistered Study: Analyses"
+title: "Xor-Some Exploratory Analyses for Preregistered Study: Model comparison SD grid search"
 author: "Polina Tsvilodub"
-date: "8/31/2021"
+date: "12/1/2021"
 output: github_document
@@
 
-Since we observed potential colinearity effects in the main preregistered analysis for the trigger "or", we further explore the effects of this colinearity on the strength of the individual effects in the regression models.
+Since we observed potential collinearity effects in the main preregistered analysis for the trigger "or", we further explore the effects of this collinearity on the strength of the individual effects in the regression models.
 
-To do so, we compute individual models omitting one of the factors on the "or" data, and compute the Bayes Factor (BF) of the model omitting one of the colinear predictors, relative to the full model.
+To do so, we compute individual models omitting one of the factors on the "or" data, and compute the Bayes Factor (BF) of the model omitting one of the collinear predictors, relative to the full model.
 
@@
 
+## Exploration of interaction effects
+
 For exploration, we also compute the BF in favor of a model with an interaction between the two effects against the same model without interaction estimates.
+This is to estimate whether the additional complexity resulting from adding an additional estimate is benefitial in terms of model fit an hence beneficial as an exploratory factor.
 ```{r, results='hide', message=FALSE, warning=FALSE}
@@
 ```
+
+## Model comparisons with grid search for final paper
+
+Below, we conduct a similar grid search over more models.
+More specifically, we compute the BF in favor of a more complex model over the intercept-only baseline model.
+We compute these pairwise comparisons for each possible combination of factors over the intercept-only model, always adding the full interaction structure to the main effects.
+This results in 7 model comparisons.
+For each comparison, we perform a small grid search over the SD parameter of the prior t-distribution for the fixed effects regression coefficients, using the values [0.05, 0.1, 0.15].
+For each model, 40.000 samples are extracted.
+These grid searches are performed separately for "or" and "some".
+```{r}
+# define a convenience function for iterating over the SDs and critical models
+models_grid_search_SD_full = function(sd_vec, iter, trigger) {
+  # create output df
+  results <- tribble(
+    ~"model in favor", ~"model against", ~"trigger", ~"SD", ~"BF", ~"log(BF)"
+  )
+  # filter data by trigger
+  data <- d_critical_zScored_wide %>% filter(main_type == trigger)
+  # models of interest
+  formula_baseline <- "target ~ 1 + (1 | submission_id) + (1 | title)"
+  formulas <- c(
+    "target ~ prior + (1 + prior || submission_id) + (1 | title)",
+    "target ~ relevance + (1 + relevance || submission_id) + (1 | title)",
+    "target ~ competence + (1 + competence || submission_id) + (1 | title)",
+    "target ~ relevance*competence + (1 + relevance + competence || submission_id) + (1 | title)",
+    "target ~ prior*competence + (1 + prior + competence || submission_id) + (1 | title)",
+    "target ~ prior*relevance + (1 + prior + relevance || submission_id) + (1 | title)",
+    "target ~ prior*competence*relevance + (1 + prior + competence + relevance || submission_id) + (1 | title)"
+  )
+  model_name_list <- c("model_pri", "model_rel", "model_comp", "model_rel_comp", "model_pri_comp", "model_rel_pri", "model_pri_rel_comp")
+  # set priors
+  for (i in sd_vec) {
+    dist <- paste("student_t(1, 0, ", i, ")", sep = "")
+    prior <- set_prior(dist, class = "b")
+    # fit the baseline intercept only model with same prior SD
+    prior_int <- set_prior(dist, class = "Intercept")
+    model_int <- brm(as.formula(formula_baseline),
+                 data = data,
+                 prior = prior_int,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.97),
+                 cores = 4,
+                 iter = iter,
+                 save_all_pars = TRUE,
+                 silent = TRUE
+                 )
+    # fit all the models
+    for (f in 1:length(formulas)) {
+      print(formulas[f])
+      model_target <- brm(as.formula(formulas[f]),
+                 data = data,
+                 prior = prior,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.97),
+                 cores = 4,
+                 iter = iter,
+                 save_all_pars = TRUE,
+                 silent = TRUE
+                 )
+      # compute the BF
+      print(paste("SD: ", i, sep=""))
+      bf <- bayes_factor(model_target, model_int, silent = TRUE)
+      print(paste("Model in favor: ", model_name_list[f], sep=""))
+      print(paste("BF: ", bf, sep=""))
+      # append to results dataframe, write out
+      results <- results %>% add_row(
+        "model in favor" = model_name_list[f],
+        "model against" = "model_Int",
+        "trigger" = trigger,
+        SD = i,
+        BF = round(bf[1]$bf, 4),
+        "log(BF)" = round(log(bf[1]$bf), 4)
+      )
+    write_csv(results, 
+              paste("../../data/main/grid_full_", trigger, "_SD_0_", i*100, ".csv", sep = ""))
+    }
+
+  }
+  return (results)
+}
+```
+
+```{r}
+# call the grid search for xor 
+grid_results_full_xor <- models_grid_search_SD_full(c(0.05, 0.1, 0.15), 40000, "xor")
+# write out results
+write_csv(grid_results_full_xor, "../../data/main/grid_full_xor_SD_final.csv")
+
+# call the grid search for some 
+grid_results_full_some <- models_grid_search_SD_full(c(0.05, 0.1, 0.15), 40000, "some")
+# write out results
+write_csv(grid_results_full_some, "../../data/main/grid_full_some_SD_final.csv")
+
+```
+
+Plot the effect of SD against the BF:
+```{r}
+grid_results_full_xor <- read_csv("../../data/main/grid_full_xor_SD_final.csv")
+grid_results_full_some <- read_csv("../../data/main/grid_full_some_SD_final.csv")
+grid_results_both <- rbind(grid_results_full_xor, grid_results_full_some)
+
+grid_results_full_xor %>%
+ggplot(., aes(x = SD, y = `log(BF)`, color = `model in favor`)) +
+  geom_point() +
+  geom_line() +
+  facet_wrap(~trigger) + 
+  ggtitle("BF in favor of including certain effects over an Intercept-only model,\ndepending on SD of the prior t-distribution over betas")
+
+ggsave("BF_vs_SD_allModels_vs_int.pdf", width = 6.5, height = 4)
+```
+
+## Additional interaction comparisons
+
+Compute the critical BF for including the effect of relevance over including the effect of prior in a two-factor model when only including the simple effects, no interactions.
+```{r}
+prior <- set_prior("student_t(1, 0, 0.05)", class = "b")
+model_rel_comp_simple <- brm(target ~ competence+relevance +
+                   (1 + competence + relevance || submission_id) +
+                   (1 | title),
+                 data = d_critical_zScored_wide %>% filter(main_type == "xor"),
+                 prior = prior,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.97),
+                 cores = 4,
+                 iter = 40000,
+                 save_all_pars = TRUE,
+                 silent=TRUE
+                 )
+
+model_pri_comp_simple <- brm(target ~ competence+prior +
+                   (1 + competence + prior || submission_id) +
+                   (1 | title),
+                 data = d_critical_zScored_wide %>% filter(main_type == "xor"),
+                 prior = prior,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.97),
+                 cores = 4,
+                 iter = 40000,
+                 save_all_pars = TRUE
+                 )
+bf_rel_pri_simple <- bayes_factor(model_rel_comp_simple, model_pri_comp_simple, silent=TRUE) 
+# read old data with interaction comparisons
+df_ints <- read_csv("./../../data/main/bf_interactions_comparisons.csv")
+glimpse(df_ints)
+df_ints %>% add_row(
+  `Model in favor` = "Rel + Comp",
+  `Model against` = "Pri + Comp",
+  BF = round(bf_rel_pri_simple[1]$bf, 4),
+  SD = 0.05
+) #%>% write_csv("./../../data/main/bf_interactions_comparisons.csv")
+```
+
+```{r}
+# TODO: use different seeds / rerun the process for a more stable BF estimate
+```

### 2021-12-30 analysis/main/05_xor-some_analysis_model_compariso_grid-search-SD.Rmd
start passing thru MFs comments, add screenshots, power analysis prose

@@
 # call the grid search for some 
-grid_results_full_some <- models_grid_search_SD_full(c(0.05, 0.1, 0.15), 40000, "some")
+#grid_results_full_some <- models_grid_search_SD_full(c(0.05, 0.1, 0.15), 40000, "some")
 # write out results
-write_csv(grid_results_full_some, "../../data/main/grid_full_some_SD_final.csv")
+#write_csv(grid_results_full_some, "../../data/main/grid_full_some_SD_final.csv")
 
@@
 grid_results_full_xor <- read_csv("../../data/main/grid_full_xor_SD_final.csv")
-grid_results_full_some <- read_csv("../../data/main/grid_full_some_SD_final.csv")
-grid_results_both <- rbind(grid_results_full_xor, grid_results_full_some)
+#grid_results_full_some <- read_csv("../../data/main/grid_full_some_SD_final.csv")
+#grid_results_both <- rbind(grid_results_full_xor, grid_results_full_some)
 

### 2021-12-30 writing/paper/paper_draft.tex
start passing thru MFs comments, add screenshots, power analysis prose

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\usepackage{gb4e}
+\noautomath
 \usepackage[dvipsnames]{xcolor}
@@
 The context stories and sentences manipulating the factors of interest were designed in an analogous fashion for both triggers and varied within-subjects, so the following descriptions apply for both triggers.
-\mf{I suggest to give one full example of a vignette (context, critical questions, possibly test questions) for each trigger here; similar to what is now in the appendix (but with nicer formatting)}
+%\mf{I suggest to give one full example of a vignette (context, critical questions, possibly test questions) for each trigger here; similar to what is now in the appendix (but with nicer formatting)}
 \mf{it might be useful to have a graphical representation of the order of ``screens'' within each trial; e.g.
-similar to the example in Figure 1 here: \url{https://magpie-manual.netlify.app/00_getting_started/02_basics/}}
+similar to the example in Figure 1 here: \url{https://magpie-manual.netlify.app/00_getting_started/02_basics/}}.
+Example stories for each trigger can be found below.
 
-If our factor hypotheses are true, on average, we expect higher likelihood ratings for the statements eliciting relevance, competence and prior judgements given items in the respective high conditions.
+\begin{exe}
+\ex	\textbf{Example transcript of an ``or''-story and the associated statements}\newline
+\textit{Title and classification:} ``Lily's husband'' (Prior -- / Competence + / Relevance --)\newline
+\textit{Story:} ``Bill and Eric went to a bar. Eric’s wife, Lily, promised to pick them up if Eric drank any alcohol. She phones up Bill to ask about their night.''\newline
+\textit{Example comprehension statements:} ``Eric and Lily went to a bar together.'' (false), ``Eric and Bill went to see a football game.'' (false), ``Eric and Bill go out to a bar every weekend.'' (uncertain), ``Eric's wife has a driver’s licence.'' (true)\newline
+\textit{Competence statement:} ``Bill knows whether Eric drank both wine and vodka.''\newline
+\textit{Relevance statement:} ``It is important for Lily to know whether her husband drank both wine and vodka.''\newline
+\textit{Prior statements:} ``If Eric drank wine, it is likely that he drank vodka as well.'', ``If Eric drank vodka, it is likely that he drank wine as well.''\newline
+\textit{Critical utterance:} ``Bill tells Lily: 'Your husband drank wine or vodka.'''\newline
+\textit{Implicature statement:} ``From what Bill said we may conclude that Eric did not drink both wine and vodka.''
+\end{exe}
+
+\begin{exe}
+\ex \textbf{Example transcript of a “some”-story and the associated statements}\newline
+\textit{Title and classification:} ``Final exam'' (Prior -- / Competence -- / Relevance +)\newline
+\textit{Story}: ``Bernard just finished his final exam.
+He isn't the brightest student, but if he passes this one, he will be through to the next grade.
+During the break, he goes to the teacher's lounge.
+He cannot wait to ask his teacher whether he passed.
+Bernard's teacher only just started correcting his exam.''\newline
+\textit{Example comprehension statements:} ``Bernard wants to get to the next grade.'' (true), ``Bernard wants to know his grade for his college application.'' (false), ``Bernard took a final exam in biology.'' (uncertain), ``Bernard is not the best student of his class.'' (true)\newline
+\textit{Competence statement:} ``Bernard’s teacher knows whether all of the answers Bernard gave on the exam are correct.''\newline
+\textit{Prior statement:} ``If at least some of the answers Bernard gave on the exam are correct, then all of them are correct.''\newline
+\textit{Relevance statement:} ``It is important for Bernard to know whether all of the answers he gave on the exam are correct.''\newline
+\textit{Critical utterance:} ``Bernard’s teacher says: 'Some of the answers you gave are correct.'''\newline
+\textit{Implicature statement:} ``From what Bernard’s teacher said we may conclude that not all of Bernard’s answers are correct.''\newline
+\end{exe}
+
+%If our factor hypotheses are true, on average, we expect higher likelihood ratings for the statements eliciting relevance, competence and prior judgements given items in the respective high conditions.
 Critically, we expect higher likelihood ratings for the scalar implicature, if the relevance and comptenece in the item were judged as high and the prior of the alternative as low.
 %Participants were then asked to rate statements eliciting four judgements using a slider bar: the relevance of the “all”/“and”-alternative, the competence of the speaker about the alternative, the prior probability of ”all” / “and” being true, and, crucially, the likelihood of the exclusive reading of “or” or of the upper-bounded reading of “some”.
-\mf{Last paragraph feels redundant to me here; best repeat hypotheses in results / analyze section but not necessarily here}
+%\mf{Last paragraph feels redundant to me here; best repeat hypotheses in results / analyze section but not necessarily here}
 
@@
 
-On each trial, participants read a background story, followed by a sentence presented in a blue box which is meant to elicit the likelihood rating.\mf{maybe include a screenshot?} These sentences were of the form ``It is important to X to know whether Y'' (relevance rating), ``Z knows whether Y'', ``If A, then B'' / ``If B , then A'' (disjunction prior rating), ``If some Y, then all Y'' (some prior rating), ``From what Z said we may conclude that W'', where X is the listener, Z is the speaker and Y is the target event in the background story, and W is the event under the upper bounded / exclusive reading.
+On each trial, participants read a background story, followed by a sentence presented in a blue box which is meant to elicit the likelihood rating (Fig.~\ref{screenshot-xor-comp}).
+These sentences were of the form ``It is important to X to know whether Y'' (relevance rating), ``Z knows whether Y'', ``If A, then B'' / ``If B , then A'' (disjunction prior rating), ``If some Y, then all Y'' (some prior rating), ``From what Z said we may conclude that W'', where X is the listener, Z is the speaker and Y is the target event in the background story, and W is the event under the upper bounded / exclusive reading.
 A and B are the disjuncts in case of disjunction.
-For the inference strength elicitation, participants additionally read a critical utterance of the form ``Z says to X: Y'', containing the trigger (``some'' / ``or''), presented below the background story in a red box.
+For the inference strength elicitation, participants additionally read a critical utterance of the form ``Z says to X: Y'', containing the trigger (``some'' / ``or''), presented below the background story in a red box (Fig.~\ref{screenshot-xor-target}).
+
+\begin{figure}[h]
+ 	\begin{center}
+ 		\includegraphics[width=0.8\linewidth]{images/screenshot_xor_comp.png}
+ 	\end{center}
+ 	\vspace{-0.3cm}
+ 	\caption{Example view of a trial eliciting the competence rating for an ``or''-vignette.}
+ 	\label{screenshot-xor-comp}
+\end{figure}
+\begin{figure}[h]
+	\begin{center}
+		\includegraphics[width=0.8\linewidth]{images/screenshot_xor_target.png}
+	\end{center}
+	\vspace{-0.3cm}
+	\caption{Example view of a trial eliciting the critical inference strength rating for an ``or''-vignette.}
+	\label{screenshot-xor-target}
+\end{figure}
 
@@
 Following our preregistered exclusion criteria, we excluded 15 participants for not indicating their native language, 3 participants for completing the study in under 8 minutes, 26 participants for failing more than 20\% of the attention checks, and 27 for failing more than 20\% of the comprehension questions.
-Due to a coding error of the comprehension questions, 206 participants were left post applying exclusion criteria, although the preregistered target sample size post exclusions was 200 subjects.
-We conducted a power analysis in order to determine the minimal sample size required for a power of 0.8 of finding evidence for the conjunction of all factor hypotheses (relevance, competence and prior) predicted by the SI account, finding that at least 180 subjects were required.\mf{we should give some details about the power analysis or link to a document in the repository}
+Due to a coding error of the comprehension questions, 206 participants were left post applying exclusion criteria, although the preregistered target sample size post exclusions was 200 subjects (see the power analysis section below for justification details).
+%\mf{we should give some details about the power analysis or link to a document in the repository}
 In the following analyses, data from the 206 participants is analysed.\footnote{All analyses were also conducted on data from 200 subject, excluding the last six submissions.
 No noteworthy quantitative or qualitative differences of the results were observed.}
+
+\subsection{Sample size justification}
+We conducted a power analysis in order to determine the minimal sample size required for a power of 0.8 of finding evidence for the conjunction of all factor hypotheses (relevance, competence and prior) predicted by the SI account, finding that at least 180 subjects were required.
+
+More specifically, we conducted a simulation-based Bayesian power analysis which aimed at determining how many participants are required in order to detect a theoretically motivated conjunctive effect of the predictors prior, relevance and competence with a confidence of at least 0.8 \citep[e.g.,][]{degen2015investigating}.
+The simulations were based on an assumed effect size $\beta = 0.15$ for all predictors, and a Region Of Practical Equivalence (ROPE) $\delta = 0.05$ for judging evidence for the directionality of a coefficient given the data \citep{kruschke2014doing}.
+The simulations were further based on the maximal desired model justified theoretically and by the design.\footnote{The maximal model in R syntax is: \texttt{rating = trigger * prior * competence * relevance}.
+The random effects were omitted for computational tractability reasons and due to the lack of a well-founded estimates of group-level effect sizes.}
+
+For this analysis, hypothetical experimental data representing the predictor effect rating was simulated by drawing four samples from a standard Normal distribution $\mathcal{N}(0,1)$ per simulated subject.
+Next, the target inference strength rating was modelled as a sample from a normal distribution with the mean computed as the sum of the simulated predictor values weighted by the assumed effect size (0.15).
+Then, the desired maximal Bayesian regression model was computed on the simulated data.
+The described steps were performed for a number of subjects ranging from 100 up to 340, increased iteratively in steps of 20.
+For each number of subjects, the model was re-computed on the simulated samples for $k=100$ iterations.
+For each iteration $i_k$, the conjunctive hypothesis of interest was tested.
+The power for the given number of participants was calculated as the proportion of iterations for which the conjunctive hypothesis was credible (i.e., if the posterior probability ($P(\beta_X > \delta \mid D)$) was at least 0.95, for $\delta = 0.05$ was the parameter that defined our ROPE for the conjunction of all three predictor coefficients).
+
+We found that at least 180 subjects were required. To account for potential data collection artifacts, we aimed to recruit 200 participants passing our exclusion criteria.
 
@@
 This is done to counterbalance possible by-participant responsiveness to particular factors, while keeping an anlysis advantageous to the scalar implicature account, since the standardization is applied across triggers.
+
+If our factor hypotheses are true, on average, we expect higher likelihood ratings for the statements eliciting relevance, competence and prior judgements given items in the respective high conditions.
+Critically, we expect higher likelihood ratings for the scalar implicature, if the relevance and comptenece in the item were judged as high and the prior of the alternative as low.
 
@@
 For all regression coefficients we used a wide and uninformative prior given by a $t$-distribution with mean 0, standard deviation of 2 and 1 degree of freedom.
-The model was fitted using the R \texttt{brms} package \citep{burkner2017brms}.\footnote{The data and the analyses can be found under \texttt{https://github.com/magpie-ea/magpie-xor-experiment}}
+The model was fitted using the R \texttt{brms} package \citep{burkner2017brms}.\footnote{The data and the analyses can be found under\newline \texttt{https://github.com/magpie-ea/magpie-xor-experiment}}
 

### 2022-01-16 analysis/main/03_xor-some_prereg_analysis_final.Rmd
address MFs comments, update plots, grid search

@@
 
+For the paper: Investigate whether we get the same results (qualitatively) if we set the ROPE to 0:
+```{r}
+posterior_hypotheses_nullROPE <- model_SI_posteriors %>% 
+  select(prior_xor, prior_some, 
+         competence_xor, competence_some,
+         relevance_xor, relevance_some) %>%
+  gather(key, val) %>%
+  group_by(key) %>% mutate(positive = mean(val > 0),
+                           negative = mean(val < 0)) %>% # no = mean(val = 0)
+  summarise(positive_eff = mean(positive),
+            negative_eff = mean(negative) # no_eff = mean(no)
+            )
+posterior_hypotheses_nullROPE
+```
+
 #### Perform BF analysis
@@
 
-ggsave("../../writing/ELM2/posterior-effects.pdf", width=7, height=5)
+# ggsave("../../writing/ELM2/posterior-effects.pdf", width=7, height=5)
 ```
@@
 model_xor_cat_posteriors_long %>%
+  mutate(effect = factor(effect, 
+                         levels = c("relevance_xor", "relevance_some", "prior_xor", "prior_some", "competence_xor", "competence_some"), 
+                         labels = c("Relevance (or)", "Relevance (some)", "Prior (or)", "Prior (some)", "Competence (or)", "Competence (some)"))) %>%
   ggplot(.) + # 
@@
   scale_fill_manual(name = "Direction of effect", values = c("coral", "yellow3", "mediumseagreen")) +
-  ggtitle("Posterior effect distributions. Points indicate posterior means\nThick lines indicate 66% CrIs, thin lines 95% CrIs.")
+  labs(x = "Effect size", y = "Effect") +
+  theme(legend.position = "top",
+        legend.title = element_text(size = 12),
+        legend.text = element_text(size = 12),
+        axis.title = element_text(size = 13),
+        strip.text.x = element_text(size = 13),
+        strip.text.y = element_text(size = 13))
 
+#ggsave("../../writing/paper/images/posterior-effects-categorical.pdf", width=7, height=5)
 ```
@@
 model_mean_posteriors_long %>%
+  mutate(effect = factor(effect, 
+                         levels = c("relevance_xor", "relevance_some", "prior_xor", "prior_some", "competence_xor", "competence_some"), 
+                         labels = c("Relevance (or)", "Relevance (some)", "Prior (or)", "Prior (some)", "Competence (or)", "Competence (some)"))) %>%
   ggplot(.) + # 
@@
   scale_fill_manual(name = "Direction of effect", values = c("coral", "yellow3", "mediumseagreen")) +
-  ggtitle("Posterior effect distributions. Points indicate posterior means\nThick lines indicate 66% CrIs, thin lines 95% CrIs.")
+  labs(x = "Effect size", y = "Effect") +
+  theme(legend.position = "top",
+        legend.title = element_text(size = 12),
+        legend.text = element_text(size = 12),
+        axis.title = element_text(size = 13),
+        strip.text.x = element_text(size = 13),
+        strip.text.y = element_text(size = 13))
 

### 2022-01-16 writing/paper/paper_draft.tex
address MFs comments, update plots, grid search

@@
 The simulations were further based on the maximal desired model justified theoretically and by the design.\footnote{The maximal model in R syntax is: \texttt{rating = trigger * prior * competence * relevance}.
-The random effects were omitted for computational tractability reasons and due to the lack of a well-founded estimates of group-level effect sizes.}
+The random effects were omitted for computational tractability reasons and due to the lack of well-founded estimates of group-level effect sizes.}
 
@@
 We check whether the posterior estimate of each effect is in either one of three intervals: (1) negative effect (<= -0.05), (2) no effect (between -0.05 and 0.05), or (3) positive effect (>= 0.05).
+We set the threshold for considering an effect as positive or negative to ±0.05 because we consider 0.05 to be the Region of Practical Equivalence (ROPE) for the effect sizes we expect based on literature \citep{kruschke2014doing}.
 We interpret the data as providing evidence in favor of an effect (positive, negative, no effect) if the posterior probability of the effect being true is >= 0.95 (i.e., 95\% of posterior samples are in the corresponding interval).
+The probabilities of the respective coefficients lying in a particular interval are reported below (i.e., for instance, if $P=0.95$ is reported, it means that 95\% of the posterior samples of the given coefficient are in the respective interval).
 In particular, we speak of evidence in favor of the scalar implicature account of the exclusive disjunction interpretation if the posterior probability of the prior effect being negative, the relevance and competence effects being positive is >= 0.95.
@@
 	\begin{center}
-		\includegraphics[width=1\linewidth]{images/posterior-effects-main.png}
+		\includegraphics[width=1\linewidth]{images/posterior-effects-main.pdf}
 	\end{center}
@@
 The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect).
-\mf{better use nicer labels for $y$-axis: use ``or'' not ``xor'' and no hyphens?; same for next plots}}
+Points indicate posterior means, thick lines indicate 66\% credible intervals, thin lines indicate 95\% credible intervals.}
 	\label{posteriors-main}
@@
 
-Consistent with predictions, for the trigger ``some'', we find a clear negative prior probabilty effect of the stronger alternative ``all'' being true, as inidcated by the probability of the negative effect of prior being $P = 0.999$ (Fig.\ref{posteriors-main}, \texttt{prior\_some} line, orange color).
-Similarly, we find a clear positive effect for speaker competence ($P =  1$, Fig.\ref{posteriors-main}, \texttt{competence\_some} line, green color).
-However, we do not find any clear effects of relevance (Fig.\ref{posteriors-main}, \texttt{relevance\_some} line, split colors).
+Consistent with predictions, for the trigger ``some'', we find a clear negative prior probabilty effect of the stronger alternative ``all'' being true, as inidcated by the probability of the negative effect of prior being $P = 0.999$ (Fig.\ref{posteriors-main}, \texttt{Prior~(some)} line, orange color).
+Similarly, we find a clear positive effect for speaker competence ($P =  1$, Fig.\ref{posteriors-main}, \texttt{Competence~(some)} line, green color).
+However, we do not find any clear effects of relevance (Fig.\ref{posteriors-main}, \texttt{Relevance~(some)} line, split colors).
 If anything, the data supports the hypothesis that the effect is not negative.
-We hypothesize that the absense of a clear-cut relevance effect might be due to the difficulty manipulating this dimension contextually.
-It might be that the vignettes used in the experiment failed to make the listener relevance salient enough for any effects to be borne out.
-Further experiments should improve this manipulation, as well as additionally consider speaker relevance.
-\mf{best first just state the results and use the last two sentences for ``Discussion'' section}
 
-For the trigger ``or'', in contrast, we only find a positive effect of competence ($P =  0.993$, Fig.\ref{posteriors-main}, \texttt{competence\_xor} line, green color).
+For the trigger ``or'', in contrast, we only find a positive effect of competence ($P =  0.993$, Fig.\ref{posteriors-main}, \texttt{Competence~(or)} line, green color).
 We find no comprehensive effects for relevance or competence.
@@
 Overall, these results would provide evidence against the scalar implicature hypothesis because not all factors yielded an effect on the robustness of scalar implicatures generated from ``some'' and ``or''.
-Similarly, the results would provide evidence against the identity hypothesis, as an effect of competence \mf{prior?} was observed for ``some'', but not for ``or''.\mf{we should explain somewhere where the threshold of 0.05 comes from; when discussing the results here we could add that (think) no effects becomes credible even if the threshold would be reduced to 0}
+Similarly, the results would provide evidence against the identity hypothesis, as an effect of prior was observed for ``some'', but not for ``or''.
 
@@
 	\begin{center}
-		\includegraphics[width=1\linewidth]{images/posterior-effects-categorical.png}
+		\includegraphics[width=1\linewidth]{images/posterior-effects-categorical.pdf}
 	\end{center}
@@
 The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect).
+Points indicate posterior means, thick lines indicate 66\% credible intervals, thin lines indicate 95\% credible intervals.
 \mf{I'm not sure if we should blindly reuse a threshold of 0.05 here without further justification.
@@
 	\begin{center}
-		\includegraphics[width=1\linewidth]{images/posterior-effects-means.png}
+		\includegraphics[width=1\linewidth]{images/posterior-effects-means.pdf}
 	\end{center}
@@
 
+Finally, we conduct an exploratory analysis when setting the ROPE to 0.
+That is, we investigate whether the posterior estimates for the regression coefficients of interest are positive ($>0$) or negative ($<0$) and interpret the data as providing evidence in favor of a particular direction if at least 95\% of posterior samples for the respective coefficent are positive or negative.
+As in the main analysis, we find credible positive effects of competence for both ``some'' ($P = 1$) and ``or'' ($P = 1$), and a credible negative effect of prior for ``some'' ($P = 1$), but not for ``or'' ($P=0.91$).
+However, in contrast to the main analysis, we do find a credible positive effect of relevance for ``or'' ($P=0.978$), but not for ``some'' ($P=0.829$).
+\pt{We should include justification for using a ROPE in the first place here}
+
 \section{Discussion}
@@
 We also did not find evidence for effects of relevance on the robustness of scalar implicature generation, given the vignettes.
+We hypothesize that the absense of a clear-cut relevance effect might be due to the difficulty manipulating this dimension contextually.
+It might be that the vignettes used in the experiment failed to make the listener relevance salient enough for any effects to be borne out.
+Further experiments should improve this manipulation, as well as additionally consider speaker relevance.
+
 Furthermore, our results hint at intriguing dependencies between relevance and prior effects for the interpretation of exclusive disjunctions.
@@
 \subsection{Vignettes}
-Example transcript of an “or”-story and the associated statements:
-
-Story: “Bill and Eric went to a bar. Eric’s wife, Lily, promised to pick them up if Eric drank any alcohol. She phones up Bill to ask about their night.”
-Example comprehension statements: “Eric and Lily went to a bar together.” (false), “Eric and Bill went to see a football game.” (false), “Eric and Bill go out to a bar every weekend.” (uncertain), “Eric's wife has a driver’s licence.” (true)
-Competence statement: “Bill knows whether Eric drank both wine and vodka.”
-Relevance statement: “It is important for Lily to know whether her husband drank both wine and vodka.”
-Prior statements: “If Eric drank wine, it is likely that he drank vodka as well.”, “If Eric drank vodka, it is likely that he drank wine as well.”
-Critical utterance: “Bill tells Lily: 'Your husband drank wine or vodka.'”
-Implicature statement: “From what Bill said we may conclude that Eric did not drink both wine and vodka.”
-
-Example transcript of a “some”-story and the associated statements:
-
-Story: “Bernard just finished his final exam.
-He isn't the brightest student, but if he passes this one, he will be through to the next grade.
-During the break, he goes to the teacher's lounge.
-He cannot wait to ask his teacher whether he passed.
-Bernard's teacher only just started correcting his exam.”
-Example comprehension statements: “Bernard wants to get to the next grade.” (true), “Bernard wants to know his grade for his college application.” (false), “Bernard took a final exam in biology.” (uncertain), “Bernard is not the best student of his class.” (true)
-Competence statement: “Bernard’s teacher knows whether all of the answers Bernard gave on the exam are correct.”
-Prior statement: “If at least some of the answers Bernard gave on the exam are correct, then all of them are correct.”
-Relevance statement: “It is important for Bernard to know whether all of the answers he gave on the exam are correct.”
-Critical utterance: “Bernard’s teacher says: ‘Some of the answers you gave are correct.’”
-Implicature statement: “From what Bernard’s teacher said we may conclude that not all of Bernard’s answers are correct.”
+\pt{Include all the vignettes?}
 

### 2022-01-30 analysis/main/03_xor-some_prereg_analysis_final.Rmd
finish model comps, port over old intro

@@
         strip.text.y = element_text(size = 13))
-
+#ggsave("../../writing/paper/images/posterior-effects-means.pdf", width=7, height=5)
 ```

### 2022-01-30 writing/paper/paper_draft.tex
finish model comps, port over old intro

@@
 General intuition, introduction \citep[e.g.,][]{horn1972semantic}.
+ \pt{ported over from old paper below}
+Introductions to logic usually distinguish between two readings of “or”: an inclusive
+and an exclusive reading (e.g., McCawley, 1981; Copi \& Cohen, 2005). The inclusive
+reading corresponds to the meaning of logical disjunction. According to this reading,
+“A or B” is true if at least one and possibly both of A and B are true. The exclusive
+reading is more strict in that it excludes the possibility that both A and B are true. So
+on its exclusive reading, “A or B” is true if exactly one of A and B is true. The presence of these two readings might suggest that “or” is associated with two
+different lexical entries (e.g., Basson \& O’Connor, 1960; Baum, 1996; Rescher, 1964).
+There are, however, good reasons to reject such a lexicalist approach. Perhaps the most
+compelling reason is that, in some contexts, “or” can only receive one interpretation.
+To illustrate, consider the following sentence, in which “or” occurs in the scope of
+negation: (2) Joe does not support Donald or Hillary.
+This sentence has only one interpretation, namely that Joe supports neither Donald nor
+Hillary (cf. Crain, 2008). This interpretation corresponds to the inclusive reading of
+“or”. It does not have an interpretation corresponding to the exclusive reading; in other
+words, it does not have a reading according to which Joe either supports both Donald
+and Hillary, or neither of them. Since “or” is systematically monosemous in certain
+contexts—negation being one of them—it follows that the two readings of “or” cannot
+simply be due to a lexical ambiguity.
+An attractive alternative to the lexicalist approach stems from Grice’s (1975) theory
+of conversational implicature. According to Grice, natural conversation is governed by
+the assumption that speakers are cooperative; that is, they attempt to further the purpose
+of the discourse by means of their utterances. Speakers are cooperative by adhering
+to four maxims that enjoin their utterances to be truthful, informative, relevant,
+and clear. Sometimes hearers have to make ancillary assumptions in order to align a
+speaker’s utterance with the assumption of cooperativity. Such ancillary assumptions
+are called conversational implicatures. Horn (1972) was the first to argue that the exclusive reading of “or” can be explained
+as a conversational implicature, too, based on the assumption that the primary
+meaning of “or” is inclusive. To illustrate, consider (1) again. Assuming that the primary
+meaning of “or” is inclusive, the speaker of (1) could have been more informative,
+and hence cooperative, by uttering the “and”-alternative “Joe supports Donald
+and Hillary.” Why didn’t she? Presumably because she does not believe that the “and”-
+alternative is true. This is a weak inference that is compatible with a situation in which
+the speaker is unsure about whether or not Joe supports both Donald and Hillary. This
+weak inference can be strengthened if there is reason to believe that the speaker knows
+whether or not Joe supports both Donald and Hillary. This assumption is often called
+the competence assumption (e.g., Geurts, 2010; Russell, 2006; Schulz \& van Rooij,
+2006; Zimmermann, 2000). If the competence assumption is sufficiently plausible, it
+follows that, according to the speaker, Joe does not support both Donald and Hillary.
+This specific kind of conversational implicature is often called a scalar implicature
+because it is assumed that “or” forms a lexical scale with “and”, and that alternatives
+are generated by substituting the scalemates “or” and “and”. Other lexical scales are
+hsome, alli, hwarm, hoti, and hintelligent, brillianti.
+
+The implicature account straightforwardly explains the absence of exclusive readings
+under negation. Consider (2) again. Here, unlike in the case of (1), the “and”-
+alternative is less informative than the utterance itself. After all, (2) is only compatible
+with a situation in which Joe supports neither Donald nor Hillary, whereas the
+“and”-alternative is also consistent with situations in which Joe supports exactly one
+of Donald and Hillary. So the speaker was already maximally informative and hence
+no implicature is derived.
+Although the implicature account has since become the standard in the literature
+(e.g., Chevallier, Noveck, Nazir, Bott, Lanzetti, \& Sperber, 2008; Chierchia, Fox, \&
+Spector, 2012; Geurts, 2010; Sauerland, 2004), it is not without its problems, as we
+will see in the next section. Therefore, we set out to experimentally test the adequacy
+of the implicature account. To that end, we investigated the effect of three factors on the
+robustness of exclusive readings: (i) the relevance of the “and”-alternative to the hearer,
+(ii) the competence of the speaker about the truth of the “and”-alternative, and (iii) the
+prior probability that the “and”-alternative is true. These factors are usually taken to
+influence the robustness of scalar implicatures. Hence, if the implicature account is
+correct, we expect these factors to influence the robustness of the exclusive reading of
+“or”, too, For comparison, we also investigated how relevance, competence, and prior probability
+influence the robustness of a bona fide scalar implicature, namely the inference
+from “some” to “not all”. To illustrate, consider:
+(4) Some of my friends support Donald.
+An utterance of this sentence may convey that not all of the speaker’s friends support
+Donald. The derivation of this upper-bounded construal runs analogous to that of the
+exclusive reading of “or”: the speaker could have been more informative by saying
+“All of my friends support Donald”. Why didn’t she? Presumably because she does. not believe that this alternative is true. If, moreover, the speaker knows whether or
+not all of her friends support Donald—i.e., if the speaker is competent in the relevant
+sense—it follows that she believes that not all of her friends support Donald.
+As we will see, it turns out that there are marked differences in how relevance,
+competence, and prior probability influence the robustness of the exclusive reading of
+“or” compared to how they influence the robustness of the inference from “some” to
+“not all”. These findings will encourage us to reconsider the implicature account of the
+exclusive reading of “or”. In the next section, we explain our three factors of interest in more detail. Afterwards, we outline the details of our experiments and discuss the results.
 
@@
 
+\subsection{Predictor ratings}
+
+\begin{figure}[h]
+	\begin{center}
+		\includegraphics[width=1\linewidth]{images/byFactor-byPrior-raw-scatter.pdf}
+	\end{center}
+	\vspace{-0.3cm}
+	\caption{Relating ratings for relevance, competence and prior statements (x-axis) to ratings for the strength of pragmatic enrichments (y-axis).
+The top row shows ratings for “some” (enriched to “some, but not all”).
+The bottom row shows ratings for “or” (enriched to “A or
+B, but not both”). Ratings for stories categorized as low (red) w.r.t. a given factor are on average lower (x-axis) than for those categorized as high (blue). The apparent effect of prior for
+“or” is mitigated by collinearity with relevance.}
+	\label{main-raw-ratings}
+\end{figure}
+
+Participants’ factor ratings by-story agreed well with the designed classification of the stories (Fig. \ref{main-raw-ratings}, red vs. blue color on x-axis).
+
 \subsection{Main analysis}
+
 We analysed the results using a Bayesian linear mixed effects model, regressing the target implicature likelihood rating against the fixed effects of predictor ratings (i.e., the relevance, comptence and prior ratings elicited by the same participant for that vignette), the effect of trigger and their two-, three-, and four-way interactions.
@@
 
-\begin{figure*}[t]
+\begin{figure}[h]
 	\begin{center}
@@
 	\label{posteriors-main}
-\end{figure*}
+\end{figure}
 
@@
 \subsection{Exploratory analysis}
-\begin{figure*}[t]
+\subsubsection{Models with alternative predictors}
+\begin{figure}[h]
 	\begin{center}
@@
 	\label{posteriors-cat}
-\end{figure*}
-In addition to the main analysis, we conducted two exploratory analyses.
+\end{figure}
+
+In addition to the main analysis, we conducted three exploratory regression analyses.
 First, we fit a model using categorical predictors only.
@@
 If anything, the relevance effect tends in the direction opposite to predictions of the SI account for the trigger ``or''.
-\begin{figure*}[t]
+
+\begin{figure}[h]
 	\begin{center}
@@
 	\label{posteriors-mean}
-\end{figure*}
+\end{figure}
 
@@
 \pt{We should include justification for using a ROPE in the first place here}
+
+\subsubsection{Model comparisons}
+Because the correlations hint at possible collinearity effects between relevance and prior for the trigger ``or'', we conduct exploratory model comparisons.
+That is, in order to check whether certain factors among relevance, competence and prior have a stronger influence on the inference stength ratings, we compare regression models of different complexity.
+We take a Bayesian approach to model comparison here in terms of their Bayes factors \citep{rouder2012default}.
+We compare all regression models that can be built with the three explanatory factors as main factors against a baseline intercept-only model.\footnote{If the model contains more than one main effect, all interactions between main effects are always included in the model, as well.
+All justified and computationally tractable random effects are added to each model, respectively.}
+Following literature, we interpret the evidence in favor of a particular model as strong if the odds (i.e., the Bayes Factor) are between are above 10, as mild, if they are between 3 and 10, and as absent if the odds are between 3 and 1/3 \citep[e.g.,][]{lodewyckx2011tutorial}.
+
+\begin{figure}[h]
+	\begin{center}
+		\includegraphics[width=0.8\linewidth]{images/BF_vs_SD_allModels_vs_int.pdf}
+	\end{center}
+	\vspace{-0.3cm}
+	\caption{Bayes factor comparison of different main factor combinations.}
+	\label{bf-grid-search}
+\end{figure}
+
+To investigate the stability of the Bayes factors of a certain model over the baseline, we conduct a grid search over the standard deviation parameter of the prior student-t distribution over the main effect regression coefficients.
+More precisely, given that, based on prior experimental work, the expected effect size for the main factors was around $\beta=0.15$, the standard deviation parameters considered in the grid search included {0.05, 0.1, 0.15}.
+Figure \ref{bf-grid-search} shows the Bayes factors against the intercept-only model of all the regression models plotted against the different standard deviation parameters of the priors.
+It indicates that, in general, the Bayes factor does not change strongly dependent on the tightness of the prior distribution.
+It further suggests that the intercept-only model is a better fit to the data than all of the explored regression models, but more so for models containing both relevance and prior as explanatory factors.
+This corroborates the finding based on the correlation, indicating that relevance and prior might be co-dependent and explain each other away.
+We also observe that models containing the effect of relevance are slightly less likely given the data than models containing the effect of prior, but the difference is negligible.
+However, these results might also be due to the fact that additional model complexity of the regression models overpowers potential improvement in model fit.
+
+\begin{figure}[h]
+	\begin{center}
+		\includegraphics[width=0.8\linewidth]{images/BF_vs_SD_rel-comp_pri-comp.pdf}
+	\end{center}
+	\vspace{-0.3cm}
+	\caption{Bayes factor of the relevance*competence model over the prior*competence model.}
+	\label{bf-rel-pri}
+\end{figure}
+
+Therefore, to zoom in on the comparison between the effects of relevance and prior on inference strength ratings we conduct further exploratory model comparisons.
+Specifically, we compare two-factor models, comparing relevance over prior, each combined with competence, respectively.
+As can be seen in Figure \ref{bf-rel-pri}, the Bayes factor varies more strongly depending on the standard deviation parameter of the prior distribution over the main regression coefficients.
+For $SD=0.05$, we find $BF=0.27$, indicating that ``prior'' model is around three times more likely than the ``relevance'' model, which can we consider mild evidence in favor of prior being the better predictor for the inference strength of exclusive disjunctions.
+Yet further investigation of the interaction between relavance and competence effects with respect to their influence on exclusive disjunctions is needed.
 

### 2022-02-08 writing/paper/paper_draft.tex
pass over intro section and exploratory analyses

@@
 
-General intuition, introduction \citep[e.g.,][]{horn1972semantic}.
+\pt{Framing suggestion: start out with a general introduction on scalar implicature as applicable to both some and or,  give some details on why xor is also interesting.
+Ported over from abstract.
+Review prior work focusing on separate factors influencing SIs.
+Motivate our study.}
+
+If someone says “Anna ate some cookies”, the hearer might infer the upper-bounded reading that Anna ate some, but not all cookies.
+Modern theories usually treat this inference as one of the prime examples of \textit{scalar implicature} (SI) \citep{horn1972semantic}.
+More specifically, the SI relies on a lexical scale consisting of words ordered in terms of informativeness (or, strength), like $\langle$some, all$\rangle$.
+When interlocutors encounter the weaker expression ``some'', the stronger ``all'' functions as a salient alternative \citep{matsumoto1995conversational}.
+
+Theories of pragmatics suggest that interlocutors reason about these scales, guided by conversational pronciples introduced by \citet{grice1975logic}.
+According to \citet{grice1975logic}, natural conversation is governed by
+the assumption that speakers are cooperative; that is, they attempt to further the purpose
+of the discourse by means of their utterances. Speakers are cooperative by adhering
+to four maxims that enjoin their utterances to be truthful, informative, relevant,
+and clear. Sometimes hearers have to make ancillary assumptions in order to align a
+speaker’s utterance with the assumption of cooperativity. Such ancillary assumptions
+are called conversational implicatures.
+%Recent theories of pragmatics suggest that cooperative interlocutors reason about each other's utterances, guided by certain \textit{conversational maximes} \citep{grice1975logic}. 
+That is, in case of ``some'', assuming a rational speaker who wishes to be maximally cooperative, the listener of the utterance would infer that if the stronger alternative (i.e., ``all'') was true, the speaker would have used the stronger utterance; but since she did not, she must have lacked the evidence for using ``all'' \citep{grice1975logic, goodman2016pragmatic}.
+
+Similarly to the first example, given the utterance “Donald ate a donut or a pretzel.”, the listener might infer that Donald ate either the donut or the pretzel, but not both (i.e., an \textit{exclusive} interpretation).
+%This inference can also be explained as a variety of scalar implicature, relying on the scale $\langle$or, and$\rangle$.
+If a speaker uses an informationally weaker term (e.g., “some”), they may imply that the corresponding stronger alternative (e.g., “and”) is false [3,5].
 \pt{ported over from old paper below}
+
 Introductions to logic usually distinguish between two readings of “or”: an inclusive
@@
 reading is more strict in that it excludes the possibility that both A and B are true. So
-on its exclusive reading, “A or B” is true if exactly one of A and B is true. The presence of these two readings might suggest that “or” is associated with two
+on its exclusive reading, “A or B” is true if exactly one of A and B is true.
+
+The standard account of exclusive readings of disjunction in the literature also stems from Grice’s (1975) theory of conversational implicature (e.g., Chevallier, Noveck, Nazir, Bott, Lanzetti, \& Sperber, 2008; Chierchia, Fox, \& Spector, 2012; Geurts, 2010; Sauerland, 2004).
+\citet{horn1972semantic} was the first to argue that the exclusive reading of “or” can be explained as a conversational implicature, too, based on the assumption that the primary meaning of “or” is inclusive.
+To illustrate, consider (1) again.
+Assuming that the primary
+meaning of “or” is inclusive, the speaker of (1) could have been more informative,
+and hence cooperative, by uttering the “and”-alternative “Joe supports Donald
+and Hillary.” Why didn’t she? Presumably because she does not believe that the “and”-
+alternative is true. This is a weak inference that is compatible with a situation in which
+the speaker is unsure about whether or not Joe supports both Donald and Hillary. This
+weak inference can be strengthened if there is reason to believe that the speaker knows
+whether or not Joe supports both Donald and Hillary. This assumption is often called
+the competence assumption (e.g., Geurts, 2010; Russell, 2006; Schulz \& van Rooij,
+2006; Zimmermann, 2000). If the competence assumption is sufficiently plausible, it
+follows that, according to the speaker, Joe does not support both Donald and Hillary.
+For this conversational implicature it is assumed that “or” forms a lexical scale with “and” %, and that alternatives are generated by substituting the scalemates “or” and “and”.
+Other lexical scales are <some, all>, <warm, hot>, and <intelligent, brilliant>.
+
+Alternative theories propose that the presence of these two readings might suggest that “or” is associated with two
 different lexical entries (e.g., Basson \& O’Connor, 1960; Baum, 1996; Rescher, 1964).
@@
 simply be due to a lexical ambiguity.
-An attractive alternative to the lexicalist approach stems from Grice’s (1975) theory
-of conversational implicature. According to Grice, natural conversation is governed by
-the assumption that speakers are cooperative; that is, they attempt to further the purpose
-of the discourse by means of their utterances. Speakers are cooperative by adhering
-to four maxims that enjoin their utterances to be truthful, informative, relevant,
-and clear. Sometimes hearers have to make ancillary assumptions in order to align a
-speaker’s utterance with the assumption of cooperativity. Such ancillary assumptions
-are called conversational implicatures. Horn (1972) was the first to argue that the exclusive reading of “or” can be explained
-as a conversational implicature, too, based on the assumption that the primary
-meaning of “or” is inclusive. To illustrate, consider (1) again. Assuming that the primary
-meaning of “or” is inclusive, the speaker of (1) could have been more informative,
-and hence cooperative, by uttering the “and”-alternative “Joe supports Donald
-and Hillary.” Why didn’t she? Presumably because she does not believe that the “and”-
-alternative is true. This is a weak inference that is compatible with a situation in which
-the speaker is unsure about whether or not Joe supports both Donald and Hillary. This
-weak inference can be strengthened if there is reason to believe that the speaker knows
-whether or not Joe supports both Donald and Hillary. This assumption is often called
-the competence assumption (e.g., Geurts, 2010; Russell, 2006; Schulz \& van Rooij,
-2006; Zimmermann, 2000). If the competence assumption is sufficiently plausible, it
-follows that, according to the speaker, Joe does not support both Donald and Hillary.
-This specific kind of conversational implicature is often called a scalar implicature
-because it is assumed that “or” forms a lexical scale with “and”, and that alternatives
-are generated by substituting the scalemates “or” and “and”. Other lexical scales are
-hsome, alli, hwarm, hoti, and hintelligent, brillianti.
 
@@
 Spector, 2012; Geurts, 2010; Sauerland, 2004), it is not without its problems, as we
-will see in the next section. Therefore, we set out to experimentally test the adequacy
-of the implicature account. To that end, we investigated the effect of three factors on the
-robustness of exclusive readings: (i) the relevance of the “and”-alternative to the hearer,
-(ii) the competence of the speaker about the truth of the “and”-alternative, and (iii) the
-prior probability that the “and”-alternative is true. These factors are usually taken to
-influence the robustness of scalar implicatures. Hence, if the implicature account is
-correct, we expect these factors to influence the robustness of the exclusive reading of
-“or”, too, For comparison, we also investigated how relevance, competence, and prior probability
-influence the robustness of a bona fide scalar implicature, namely the inference
-from “some” to “not all”. To illustrate, consider:
-(4) Some of my friends support Donald.
-An utterance of this sentence may convey that not all of the speaker’s friends support
-Donald. The derivation of this upper-bounded construal runs analogous to that of the
-exclusive reading of “or”: the speaker could have been more informative by saying
-“All of my friends support Donald”. Why didn’t she? Presumably because she does. not believe that this alternative is true. If, moreover, the speaker knows whether or
-not all of her friends support Donald—i.e., if the speaker is competent in the relevant
-sense—it follows that she believes that not all of her friends support Donald.
-As we will see, it turns out that there are marked differences in how relevance,
-competence, and prior probability influence the robustness of the exclusive reading of
-“or” compared to how they influence the robustness of the inference from “some” to
-“not all”. These findings will encourage us to reconsider the implicature account of the
-exclusive reading of “or”. In the next section, we explain our three factors of interest in more detail. Afterwards, we outline the details of our experiments and discuss the results.
+will see in the next section. 
+\pt{end of section from old paper}
+
+Crucially, prior research suggests that the robustness of SIs is influenced by different contextual factors.
 
@@
 
-SI accounts and treatments of ``some'' and ``or''.
-
 \textbf{Related experimental/modelling work: } 
-Some: \citep{degen2015investigating}, 
+Some: \citep{degen2015investigating, goodman2016pragmatic}, 
 Or: \citep{Li2021}
@@
 Lastly, high prior probability of the stronger ``all'' being true is assumed to \textit{decrease} the scalar implicature rate.
+\pt{If the framing is okay, add some motivation and predictions regarding investigating the conjunction of these factors.}
 If the currently standard scalar implicature account also holds for the interpretation of disjunctions, analogous predictions for the robustness of the implicature from ``A or B'' to ``A or B, but not both'' can be derived.
@@
 
-More specifically, we conducted a simulation-based Bayesian power analysis which aimed at determining how many participants are required in order to detect a theoretically motivated conjunctive effect of the predictors prior, relevance and competence with a confidence of at least 0.8 \citep[e.g.,][]{degen2015investigating}.
-The simulations were based on an assumed effect size $\beta = 0.15$ for all predictors, and a Region Of Practical Equivalence (ROPE) $\delta = 0.05$ for judging evidence for the directionality of a coefficient given the data \citep{kruschke2014doing}.
+More specifically, we conducted a simulation-based Bayesian power analysis which aimed at determining how many participants are required in order to detect a theoretically motivated conjunctive effect of the predictors prior, relevance and competence with a confidence of at least 0.8.
+The simulations were based on an assumed effect size $\beta = 0.15$ for all predictors \citep[e.g.,][]{degen2015investigating}, and a Region Of Practical Equivalence (ROPE) $\delta = 0.05$ for judging evidence for the directionality of a coefficient given the data \citep{kruschke2014doing}.
+That is, we determined how many participants would be necessary such that the regression coefficients |$\beta_X$| for the three predictors would be larger than $\delta$.
 The simulations were further based on the maximal desired model justified theoretically and by the design.\footnote{The maximal model in R syntax is: \texttt{rating = trigger * prior * competence * relevance}.
@@
 
-For this analysis, hypothetical experimental data representing the predictor effect rating was simulated by drawing four samples from a standard Normal distribution $\mathcal{N}(0,1)$ per simulated subject.
+For this analysis, hypothetical experimental data representing the predictor effect rating was simulated by drawing four samples from a standard Normal distribution $\mathcal{N}(0,1)$ per simulated subject per trigger.
 Next, the target inference strength rating was modelled as a sample from a normal distribution with the mean computed as the sum of the simulated predictor values weighted by the assumed effect size (0.15).
@@
 For each iteration $i_k$, the conjunctive hypothesis of interest was tested.
-The power for the given number of participants was calculated as the proportion of iterations for which the conjunctive hypothesis was credible (i.e., if the posterior probability ($P(\beta_X > \delta \mid D)$) was at least 0.95, for $\delta = 0.05$ was the parameter that defined our ROPE for the conjunction of all three predictor coefficients).
+The power for the given number of participants was calculated as the proportion of iterations for which the conjunctive hypothesis was credible (i.e., if the posterior probability $P(\beta_X > \delta \mid D)$ was at least 0.95, for $\delta = 0.05$ was the parameter that defined our ROPE for the conjunction of all three predictor coefficients).
+The target proportion was 0.8.
 
@@
 The top row shows ratings for “some” (enriched to “some, but not all”).
-The bottom row shows ratings for “or” (enriched to “A or
-B, but not both”). Ratings for stories categorized as low (red) w.r.t. a given factor are on average lower (x-axis) than for those categorized as high (blue). The apparent effect of prior for
-“or” is mitigated by collinearity with relevance.}
+The bottom row shows ratings for “or” (enriched to “A or B, but not both”).
+Ratings for stories categorized as low (red) w.r.t.
+a given factor are on average lower (x-axis) than for those categorized as high (blue).
+The apparent effect of prior for “or” is mitigated by collinearity with relevance.}
 	\label{main-raw-ratings}
@@
 
-Participants’ factor ratings by-story agreed well with the designed classification of the stories (Fig. \ref{main-raw-ratings}, red vs. blue color on x-axis).
+Participants’ factor ratings by-story agreed well with the designed classification of the stories (Fig.
+\ref{main-raw-ratings}, red vs.
+blue color on x-axis).
+That is, the ratings for the relevance, comptence and prior statements were not distributed uniformly across the stories, but align with the prior categorization of the story for the respective factor (Fig.
+\ref{main-raw-ratings}, columns).
+This provides validation for our experimental manipulation of the explanatory factors.
 
@@
 
-Given this model and the data, we found very similar results for both triggers: We observe a negative effect of prior and a positive effect of competence, in line with predictions of the scalar implicature hypothesis (Fig.\ref{posteriors-cat}).
+Given this model and the data, we found very similar results for both triggers: We observe a negative effect of prior and a positive effect of competence, in line with predictions of the scalar implicature hypothesis (Fig.~\ref{posteriors-cat}).
 However, we again observe no effects of relevance.
@@
 These results match the second exploratory analysis, whereby the ratings were used as predictors in the linear model, but the ratings were averaged for each vignette across participants.
-In this second  exploratory model, we again find similar effects for ``some'' and ``or'': a positive effect of competence, negative for prior, and an indefinite relevance effect (Fig.\ref{posteriors-mean}).
-By averaging over single participant ratings, this model indicates that the absent competence effect in the main model might be due to by-participant variation.
+In this second  exploratory model, we again find similar effects for ``some'' and ``or'': a positive effect of competence, negative for prior, and an indefinite relevance effect (Fig.~\ref{posteriors-mean}).
+By averaging over single participant ratings, this model indicates that the absent competence effect for the disjunction in the main model might be due to by-participant variation.
 
@@
 We also found a significant correlation between competence and relevance ($R^2 = 0.127$) for ``or''.
-No correlations were found for ``some''.
+\pt{We should talk about the latter correlation in more detail at some point.} No correlations were found for ``some''.
 These intriguing results might indicate that these effects are closely related for disjunction.
@@
 Similarly, competence might be connected to relevance indirectly via the prior: the more likely the stronger alternative is a priori, the more likely it is that a speaker knows about it, while the irrelevance still persists for the listener.
-However, this entangled results need further research.
+However, these speculations call for further research.
 
@@
 However, in contrast to the main analysis, we do find a credible positive effect of relevance for ``or'' ($P=0.978$), but not for ``some'' ($P=0.829$).
-\pt{We should include justification for using a ROPE in the first place here}
+\pt{We should include justification for using a ROPE in the first place here.
+Some more discussion needs to be added too.}
 
@@
 Because the correlations hint at possible collinearity effects between relevance and prior for the trigger ``or'', we conduct exploratory model comparisons.
-That is, in order to check whether certain factors among relevance, competence and prior have a stronger influence on the inference stength ratings, we compare regression models of different complexity.
+That is, in order to check whether certain factors among relevance, competence and prior have a stronger influence on the inference stength ratings than other, we compare regression models of different complexity.
 We take a Bayesian approach to model comparison here in terms of their Bayes factors \citep{rouder2012default}.
@@
 All justified and computationally tractable random effects are added to each model, respectively.}
-Following literature, we interpret the evidence in favor of a particular model as strong if the odds (i.e., the Bayes Factor) are between are above 10, as mild, if they are between 3 and 10, and as absent if the odds are between 3 and 1/3 \citep[e.g.,][]{lodewyckx2011tutorial}.
+Following literature, we interpret the evidence in favor of a particular model as strong if the odds (i.e., the Bayes Factor) are above 10 or below 1/10; as mild, if they are between 3 and 10 or 1/3 and 1/10; and as absent if the odds are between 3 and 1/3 \citep[e.g.,][]{lodewyckx2011tutorial}.
 
@@
 	\begin{center}
-		\includegraphics[width=0.8\linewidth]{images/BF_vs_SD_allModels_vs_int.pdf}
+		\includegraphics[width=\linewidth]{images/BF_vs_SD_allModels_vs_int.pdf}
 	\end{center}
 	\vspace{-0.3cm}
-	\caption{Bayes factor comparison of different main factor combinations.}
+	\caption{Bayes factor comparison of different main factor combinations (colors) for different standard deviation parameter values of the prior $t-$distribution over the main regression coefficients (x-axis).
+The log Bayes Factor (y-axis) is computed in favor the more complex model over an intercept-only model.
+\pt{Model names will be made human-readable, the title will be removed.}}
 	\label{bf-grid-search}
@@
 
-To investigate the stability of the Bayes factors of a certain model over the baseline, we conduct a grid search over the standard deviation parameter of the prior student-t distribution over the main effect regression coefficients.
+To investigate the stability of the Bayes factors of a certain model over the baseline, we conduct a grid search over the standard deviation parameter of the prior student-$t$ distribution over the main effect regression coefficients.
 More precisely, given that, based on prior experimental work, the expected effect size for the main factors was around $\beta=0.15$, the standard deviation parameters considered in the grid search included {0.05, 0.1, 0.15}.
@@
 	\begin{center}
-		\includegraphics[width=0.8\linewidth]{images/BF_vs_SD_rel-comp_pri-comp.pdf}
+		\includegraphics[width=\linewidth]{images/BF_vs_SD_rel-comp_pri-comp.pdf}
 	\end{center}
 	\vspace{-0.3cm}
-	\caption{Bayes factor of the relevance*competence model over the prior*competence model.}
+	\caption{Bayes factor of the \texttt{relevance*competence} model over the \texttt{prior*competence model} (y-axis), plotted against different standard deviation parameter values of the prior $t-$distribution over the main regression coefficients (x-axis).
+\pt{The title will be removed.}}
 	\label{bf-rel-pri}
@@
 As can be seen in Figure \ref{bf-rel-pri}, the Bayes factor varies more strongly depending on the standard deviation parameter of the prior distribution over the main regression coefficients.
-For $SD=0.05$, we find $BF=0.27$, indicating that ``prior'' model is around three times more likely than the ``relevance'' model, which can we consider mild evidence in favor of prior being the better predictor for the inference strength of exclusive disjunctions.
+For $SD=0.05$, we find $BF=0.27$, indicating that the model containing  the predictors \texttt{prior*competence } is around three times more likely than the \texttt{relevance*competence} model, which can we consider mild evidence in favor of prior being the better predictor for the inference strength of exclusive disjunctions.
 Yet further investigation of the interaction between relavance and competence effects with respect to their influence on exclusive disjunctions is needed.
@@
 Taken together, the results might speak against the identity hypothesis, indicating that there might be subtle differences in the conceptualization of the stronger alternative for ``some'' and ``or''.
+\pt{Focus on the aspect of interactions between the predictors more strongly.}
 

### 2022-03-31 writing/paper/paper_draft.tex
minor edits and comments

@@
 \usepackage[dvipsnames]{xcolor}
+\usepackage{multirow,booktabs}
 \newcommand{\mf}[1]{\textcolor{BurntOrange}{[MF: #1]}}
@@
 \author[]{% As many authors as you like, each separated by \AND.
-  \spauthor{Michael Franke \\ \institute{Institute1}} \AND
-  \spauthor{Bob van Tiel \\ \institute{Institute2}} \AND
-  \spauthor{Polina Tsvilodub \\ \institute{Institute3}}}
+  \spauthor{Polina Tsvilodub \\ \institute{Institute}} \AND
+  \spauthor{Bob van Tiel \\ \institute{Institute}} \AND
+  \spauthor{Michael Franke \\ \institute{Department of Linguistics \\ University of Tübigen}}}
 
@@
 Current standard accounts of disjunction interpretation posit that the exclusive interpretation of disjunctions is an instance of \textit{scalar implicature}.
-Prior work suggests that, among others, three factors influence the robustness of scalar implicatures: (1) \textit{relevance} of the stronger alternative to the listener, (2) the \textit{competence} of the speaker about the truth of the stronger alternative, and (3) the \textit{prior probability} that the stronger alternative is true.
+Prior work suggests that, among others, three factors influence the robustness of scalar implicatures: (i) \textit{relevance} of the stronger alternative to the listener, (ii) the \textit{competence} of the speaker about the truth of the stronger alternative, and (iii) the \textit{prior probability} that the stronger alternative is true.
 While their influence was investigated experimentally for the interpretation of ``some'', the evidence is less conclusive in the case of disjunction interpretation.
@@
 
-Theories of pragmatics suggest that interlocutors reason about these scales, guided by conversational pronciples introduced by \citet{grice1975logic}.
+Theories of pragmatics suggest that interlocutors reason about these scales, guided by conversational principles introduced by \citet{grice1975logic}.
 According to \citet{grice1975logic}, natural conversation is governed by
@@
 %Recent theories of pragmatics suggest that cooperative interlocutors reason about each other's utterances, guided by certain \textit{conversational maximes} \citep{grice1975logic}. 
-That is, in case of ``some'', assuming a rational speaker who wishes to be maximally cooperative, the listener of the utterance would infer that if the stronger alternative (i.e., ``all'') was true, the speaker would have used the stronger utterance; but since she did not, she must have lacked the evidence for using ``all'' \citep{grice1975logic, goodman2016pragmatic}.
+That is, in case of ``some'', assuming a rational speaker who wishes to be maximally cooperative, the listener of the utterance would infer that if the stronger alternative (i.e., ``all'') was true, the speaker would have used the stronger utterance; but since she did not, she must have lacked the evidence for using ``all''P.
 
@@
 Lastly, high prior probability of the stronger ``all'' being true is assumed to \textit{decrease} the scalar implicature rate.
-\pt{If the framing is okay, add some motivation and predictions regarding investigating the conjunction of these factors.}
+\pt{If the framing is okay, add some motivation and predictions regarding investigating the conjunction of these factors.} \mf{yes, good idea}
 If the currently standard scalar implicature account also holds for the interpretation of disjunctions, analogous predictions for the robustness of the implicature from ``A or B'' to ``A or B, but not both'' can be derived.
@@
 
-In this experiment we directly compare the interpretation of ``some'' and ``or'', so as to directly explore if the predictions generated by the scalar implicature account are borne out comparably for both triggers.
+In this experiment we directly compare the interpretation of ``some'' and ``or'', so as to explore if the predictions generated by the scalar implicature account are borne out comparably for both triggers.
 Specifically, we manipulate the \textit{contextual relevance of the stronger alternative to the listener}, the \textit{speaker's competence about the truth of ``all'' / ``and''}, and the \textit{prior probability of ``all'' / ``and'' being true}, to investigate the robustness of an exclusive reading of ``or'' (``A or B, but not both'') under these manipulations, in direct comparison to the upper-bounded reading of “some” (``some but not all'').
+
+\mf{The exposition of the experiment meanders quite a bit between high-level description, materials and procedure. I would advocate a tidier organization: high-level, materials, then procedure.}
 
@@
 \subsection{Sample size justification}
+\mf{I suggest to move this paragraph to the appendix}
 We conducted a power analysis in order to determine the minimal sample size required for a power of 0.8 of finding evidence for the conjunction of all factor hypotheses (relevance, competence and prior) predicted by the SI account, finding that at least 180 subjects were required.
@@
 More specifically, we conducted a simulation-based Bayesian power analysis which aimed at determining how many participants are required in order to detect a theoretically motivated conjunctive effect of the predictors prior, relevance and competence with a confidence of at least 0.8.
-The simulations were based on an assumed effect size $\beta = 0.15$ for all predictors \citep[e.g.,][]{degen2015investigating}, and a Region Of Practical Equivalence (ROPE) $\delta = 0.05$ for judging evidence for the directionality of a coefficient given the data \citep{kruschke2014doing}.
-That is, we determined how many participants would be necessary such that the regression coefficients |$\beta_X$| for the three predictors would be larger than $\delta$.
+The simulations were based on an assumed effect size $\beta = 0.15$ for all predictors assuming $z$-scored data \citep[e.g.,][]{degen2015investigating}, and a Region Of Practical Equivalence (ROPE) $\delta = 0.05$ for judging evidence for the directionality of a coefficient given the data \citep{kruschke2014doing}.
+That is, we determined how many participants would be necessary such that for each factor $X$ the absolute value of the regression coefficients |$\beta_X$| for main factor $X$ would be larger than $\delta$.
 The simulations were further based on the maximal desired model justified theoretically and by the design.\footnote{The maximal model in R syntax is: \texttt{rating = trigger * prior * competence * relevance}.
@@
 
+\mf{Give section overview: 4.1 plots and descriptive stats, 4.2 Confirmatory analyses following preregistration; 4.3 additional exploratory analyses (some of which preregistered?)}
+
 \subsection{Predictor ratings}
+
+\mf{I think that it would be great to have a plot, accompanying Figure~\ref{main-raw-ratings}, which shows ``idealized data according to assumed hypotheses''.
+So, how would the plot in the Figure look like if all of our predictions were borne out precisely?
+We could sample ``fictitious'' data and make a plot exactly like in Figure~\ref{posteriors-main}.}
 
@@
 \ref{main-raw-ratings}, red vs.
-blue color on x-axis).
+blue color on $x$-axis).
 That is, the ratings for the relevance, comptence and prior statements were not distributed uniformly across the stories, but align with the prior categorization of the story for the respective factor (Fig.
-\ref{main-raw-ratings}, columns).
+\ref{main-raw-ratings}).
 This provides validation for our experimental manipulation of the explanatory factors.
 
-\subsection{Main analysis}
+\subsection{Confirmatory analyses following preregistration}
 
-We analysed the results using a Bayesian linear mixed effects model, regressing the target implicature likelihood rating against the fixed effects of predictor ratings (i.e., the relevance, comptence and prior ratings elicited by the same participant for that vignette), the effect of trigger and their two-, three-, and four-way interactions.
+We analysed the results using a Bayesian linear mixed effects model, regressing the target implicature rating against the fixed effects of predictor ratings (i.e., the relevance, comptence and prior ratings elicited by the same participant for that vignette), the effect of trigger and their two-, three-, and four-way interactions.
 Maximal computationally tractable random effects structure justified by the experimental design was included: we included random intercepts and random slope effects for the main effects of trigger, relevance, competence and prior by-subject, as well as random intercepts by-vignette.\footnote{Model in R syntax style: \texttt{inference-rating $\sim$ trigger * relevance * competence * prior + (1 + trigger + relevance + competence + prior || subjectID) + (1 | vignette)}.
@@
 We focus on six slope coefficients of the regression model, for the effects of relevance, competence and prior, respectively, once for ``some'' and once for ``or''.
-We check whether the posterior estimate of each effect is in either one of three intervals: (1) negative effect (<= -0.05), (2) no effect (between -0.05 and 0.05), or (3) positive effect (>= 0.05).
-We set the threshold for considering an effect as positive or negative to ±0.05 because we consider 0.05 to be the Region of Practical Equivalence (ROPE) for the effect sizes we expect based on literature \citep{kruschke2014doing}.
+We check whether the posterior estimate of each effect is in either one of three intervals: (1) negative effect ($\le$ -0.05), (2) no effect (between -0.05 and 0.05), or (3) positive effect ($\ge$ 0.05).
+We set the threshold for considering an effect as positive or negative to ±0.05 because we consider 0.05 to be the Region of Practical Equivalence (ROPE) for the effect sizes we expect \citep{kruschke2014doing}.
 We interpret the data as providing evidence in favor of an effect (positive, negative, no effect) if the posterior probability of the effect being true is >= 0.95 (i.e., 95\% of posterior samples are in the corresponding interval).
 The probabilities of the respective coefficients lying in a particular interval are reported below (i.e., for instance, if $P=0.95$ is reported, it means that 95\% of the posterior samples of the given coefficient are in the respective interval).
-In particular, we speak of evidence in favor of the scalar implicature account of the exclusive disjunction interpretation if the posterior probability of the prior effect being negative, the relevance and competence effects being positive is >= 0.95.
+In particular, we speak of evidence in favor of the scalar implicature account of the exclusive disjunction interpretation if the posterior probability of the prior effect being negative, the relevance and competence effects being positive is $\ge$ 0.95.
 
@@
 
+\begin{table}
+  \centering
+   \begin{tabular}{llccc}
+           &            & negative & no effect & positive \\ \midrule
+    {some} & relevance  & ...      & ...       & ...      \\
+           & competence & 0        & 0         & 1        \\
+           & prior      & 0        & 0.001     & 0.999    \\
+    {or}   & relevance  & ...      & ...       & ...      \\
+           & competence & ...      & ...       & ...      \\
+           & prior      & ...      & ...       & ...      \\
+   \end{tabular}
+  \caption{Results of confirmatory analysis. \mf{Fill me, please.}}
+  \label{tab:results-confirmatory}
+\end{table}
+
+Results from this analysis are summarized in Table
 Consistent with predictions, for the trigger ``some'', we find a clear negative prior probabilty effect of the stronger alternative ``all'' being true, as inidcated by the probability of the negative effect of prior being $P = 0.999$ (Fig.\ref{posteriors-main}, \texttt{Prior~(some)} line, orange color).
@@
 
-\subsection{Exploratory analysis}
+\subsection{Exploratory analyses}
+
 \subsubsection{Models with alternative predictors}
+
 \begin{figure}[h]
@@
 
-In addition to the main analysis, we conducted three exploratory regression analyses.
+In order to further investigate the relevance effects and potential collinarity concerns, we computed exploratory pairwise correlations of all the predictors.
+Interestingly, we find a significant correlation between the prior and the relevance effects for the trigger ``or'' ($R^2 = -0.106$).
+We also found a significant correlation between competence and relevance ($R^2 = 0.127$) for ``or''.
+\pt{We should talk about the latter correlation in more detail at some point.} No correlations were found for ``some''.
+These intriguing results might indicate that these effects are closely related for disjunction.
+Intuitively, a negative correlation between prior and relevance of the stronger alternative ``and'' makes sense: the more likely the conjunction is a priori, the more likely it is that the listener might know about it already, yielding the implicature irrelevant to the listener.
+Similarly, competence might be connected to relevance indirectly via the prior: the more likely the stronger alternative is a priori, the more likely it is that a speaker knows about it, while the irrelevance still persists for the listener.
+However, these speculations call for further research.
+\mf{I suggest removing the last sentence here and to pick this up in the general discussion, together with Bob's worry that we lack an explanation for why there is no such correlation in the case of ``some''.}
+
+In addition to the main analysis, we conducted three exploratory regression analyses, which shed more light on the concerns of collinearity.
 First, we fit a model using categorical predictors only.
@@
 
-In order to further investigate the relevance effects and potential collinarity concerns, we computed exploratory pairwise correlations of all the predictors.
-Interestingly, we find a significant correlation between the prior and the relevance effects for the trigger ``or'' ($R^2 = -0.106$).
-We also found a significant correlation between competence and relevance ($R^2 = 0.127$) for ``or''.
-\pt{We should talk about the latter correlation in more detail at some point.} No correlations were found for ``some''.
-These intriguing results might indicate that these effects are closely related for disjunction.
-Intuitively, a negative correlation between prior and relevance of the stronger alternative ``and'' makes sense: the more likely the conjunction is a priori, the more likely it is that the listener might know about it already, yielding the implicature irrelevant to the listener.
-Similarly, competence might be connected to relevance indirectly via the prior: the more likely the stronger alternative is a priori, the more likely it is that a speaker knows about it, while the irrelevance still persists for the listener.
-However, these speculations call for further research.
-
 Finally, we conduct an exploratory analysis when setting the ROPE to 0.
@@
 Because the correlations hint at possible collinearity effects between relevance and prior for the trigger ``or'', we conduct exploratory model comparisons.
-That is, in order to check whether certain factors among relevance, competence and prior have a stronger influence on the inference stength ratings than other, we compare regression models of different complexity.
+That is, in order to check whether certain factors among relevance, competence and prior have a stronger influence on the inference stength ratings than others, we compare regression models of different complexity.
 We take a Bayesian approach to model comparison here in terms of their Bayes factors \citep{rouder2012default}.
@@
 
-To investigate the stability of the Bayes factors of a certain model over the baseline, we conduct a grid search over the standard deviation parameter of the prior student-$t$ distribution over the main effect regression coefficients.
-More precisely, given that, based on prior experimental work, the expected effect size for the main factors was around $\beta=0.15$, the standard deviation parameters considered in the grid search included {0.05, 0.1, 0.15}.
+To investigate the stability of the Bayes factors of a certain model over the baseline, we conduct a grid search over the standard deviation parameter of the prior $t$-distribution over the main effect regression coefficients.
+More precisely, given that the expected effect size for the main factors was around $\beta=0.15$, the standard deviation parameters considered in the grid search included {0.05, 0.1, 0.15}.
 Figure \ref{bf-grid-search} shows the Bayes factors against the intercept-only model of all the regression models plotted against the different standard deviation parameters of the priors.
@@
 As can be seen in Figure \ref{bf-rel-pri}, the Bayes factor varies more strongly depending on the standard deviation parameter of the prior distribution over the main regression coefficients.
-For $SD=0.05$, we find $BF=0.27$, indicating that the model containing  the predictors \texttt{prior*competence } is around three times more likely than the \texttt{relevance*competence} model, which can we consider mild evidence in favor of prior being the better predictor for the inference strength of exclusive disjunctions.
+For $SD=0.05$, we find $BF=0.27$, indicating that the model containing  the predictors \texttt{prior*competence } is made around three times more likely than the \texttt{relevance*competence} model, which can we consider mild evidence in favor of prior being the better predictor for the inference strength of exclusive disjunctions.
 Yet further investigation of the interaction between relavance and competence effects with respect to their influence on exclusive disjunctions is needed.

### 2022-03-31 analysis/main/01_xor-some_prereg_preprocessing_final.Rmd
added comments on preprocessing

@@
 ```{r, echo=FALSE}
-d_critical_clean <- d_full_clean %>% filter(condition == "critical")
+d_critical_clean <- d_full_clean %>% filter(condition == "critical") %>% 
+  mutate(
+    block = case_when(
+      block == "comp" ~ "competence", 
+      block == "rel"  ~ "relevance", 
+      block == "pri"  ~ "prior",
+      T ~ block) ,
+    block_extended = case_when(
+      block == "some" ~ "target", 
+      block == "xor"  ~ "target", 
+      T ~ block) )
+
+# write_csv(d_critical_clean, "../../data/main/results_prereg_final_tidy-MF.csv")
+
+# MF: Careful! This step introduces tripling each individual `response` measure
+
 d_critical_long <- d_critical_clean %>% 
@@
 
-d_critical_long <- d_critical_long %>% 
-  mutate(block = ifelse(block == "comp", "competence", 
-                        ifelse(block == "rel", "relevance", ifelse(block == "pri", "prior", block) )))
-```
-
-Create more extensive condition labels.
-
-``` {r}
-# extending 'conditions' labels to include whether the utterance was present or not
-d_critical_long <- d_critical_long %>% 
-  mutate(
-    block_extended =  ifelse(block %in% c("some", "xor"), "target", block)
-  )
 ```
@@
 Z-scoring:
+
+MF: Here is a problem. The `d_critical_long` contains every `response` measure three times.
+    This is unproblematic for the mean, but not of the SD.
+
 ```{r}
+# MF: ooriginal code
+
 d_critical_zScored <- d_critical_long %>% group_by(submission_id, block_extended) %>%
@@
          response_centered = ifelse(is.na(response_centered), 0, response_centered))
+```
 
+MF: To see this problem, compare the following (the first is wrong and what we used so far, the second correct):
+
+```{r}
+
+d_critical_long %>% 
+  select(submission_id, block_extended, response) %>% 
+  group_by(submission_id, block_extended) %>%
+  summarize(mean = mean(response),
+            sd = sd(response)) %>% 
+  mutate(data_used = 'long') %>% 
+  arrange(submission_id, block_extended)
+```
+
+```{r}
+d_critical_clean %>% 
+  select(submission_id, block_extended, response) %>% 
+  group_by(submission_id, block_extended) %>%
+  summarize(mean = mean(response),
+            sd = sd(response)) %>% 
+  mutate(data_used = 'short') %>% 
+  arrange(submission_id, block_extended)
+```
+
+We should therefore use the following (but now the table doesn't contain the high/low classifications as individual rows):
+
+```{r, eval = F}
+# MF: corrected code; not executed
+
+d_critical_zScored <- d_critical_clean %>% group_by(submission_id, block_extended) %>%
+  mutate(block_mean = mean(response),
+         block_sd = sd(response),
+         response_centered = (response - block_mean)/block_sd,
+         # catch the cases where sd is 0 
+         response_centered = ifelse(is.na(response_centered), 0, response_centered))
+```
+
+Wrangling into wide format.
+
+```{r}
 d_critical_zScored_wide <- d_critical_zScored %>% 

### 2022-04-07 analysis/main/01_xor-some_prereg_preprocessing_final.Rmd
exclude subjects resilient to context, remove duplicates

--- 2021-11-14 analysis/main/01_xor-some_prereg_preprocessing_final.Rmd
@@
          ) %>% filter(passed_main == F)
+# now check for critical trials only
+d_main_critical_fail <- d_main %>% group_by(submission_id) %>% 
+  filter(condition == "critical") %>%
+  mutate(passed_main = case_when(max(response) - min(response) <= 10 ~ FALSE,
+                                 TRUE ~ TRUE)
+         ) %>% filter(passed_main == F)
+# now check implicature trials only 
+d_main_implicature_fail <- d_main %>% group_by(submission_id) %>% 
+  filter((block == "xor") | (block == "some")) %>%
+  mutate(passed_main = case_when(max(response) - min(response) <= 10 ~ FALSE,
+                                 TRUE ~ TRUE)
+         ) %>% filter(passed_main == F)
+# check if this failure is the case for certain vignettes more often then for others
+d_main_implicature_fail %>% ungroup() %>% group_by(ID, block) %>% count() %>% View()
+# --> it seems that the vignettes 28 (xor) and 45 (some) have this issue the most
+# check how often the single participants do this
+d_main_implicature_fail %>% count() %>% View()
+# --> all of them do this on all of the trials - these are 15 people 
+# write out a csv with their IDs for an exploratory exclusion
+#d_main_implicature_fail %>% count() %>% write_csv("./../../data/main/submission-ids_implicature_fails.csv")
+
 cat(" Subjects providing the same ratings throughout the trials: ", d_main_fail %>% distinct(submission_id) %>% pull() %>% length())
@@
 
+# write out tidy data without the participants providing only 100 ratings on the critical trials
+d_clean_exclude_implicatureFails <- anti_join(d_full_clean, d_main_implicature_fail, by = "submission_id")
+
 cat(" Nr. of participants left after cleaning: ", d_full_clean %>% distinct(submission_id) %>% pull() %>% length())
+
+cat(" Nr. of participants left after excluding participants failing implicature manipulations: ", d_clean_exclude_implicatureFails %>% distinct(submission_id) %>% pull() %>% length())
+
 ```
@@
 d_full_clean %>% write_csv("./../../data/main/results_prereg_final_tidy.csv")
+
+d_clean_exclude_implicatureFails %>% write_csv("./../../data/main/results_prereg_final_tidy_exclude_implcatureFailures.csv")
 ```
@@
          response_centered = ifelse(is.na(response_centered), 0, response_centered))
-
+        # response_centered = ifelse(is.na(response_centered), (response-block_mean), response_centered)
 d_critical_zScored_wide <- d_critical_zScored %>% 
@@
   )
+
+# assume that the mistake in the z-scoring is the absence of dropping the duplicates
+# or is it the point where the 0 SDs are caught?
+d_critical_zScored_noDuplicates <- d_critical_long %>% 
+  filter((block_extended == class_condition) | (class_condition =="relevance" & block_extended == "target") )
+
+d_critical_noDuplicates_zScored <- d_critical_zScored_noDuplicates %>% group_by(submission_id, block_extended) %>%
+  mutate(block_mean = mean(response),
+         block_sd = sd(response),
+         response_centered = (response - block_mean)/block_sd,
+         # catch the cases where sd is 0 
+         response_centered = ifelse(is.na(response_centered), 0, response_centered))
 
@@
 d_critical_zScored_wide %>% write_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
+d_critical_noDuplicates_zScored %>% write_csv("./../../data/main/results_prereg_tidy_noDuplicates_zScored_long.csv")
 

### 2022-04-17 analysis/main/01_xor-some_prereg_preprocessing_final.Rmd
fix analyses and plots with correctly z-scored data

@@
 # check if this failure is the case for certain vignettes more often then for others
-d_main_implicature_fail %>% ungroup() %>% group_by(ID, block) %>% count() %>% View()
+d_main_implicature_fail %>% ungroup() %>% group_by(ID, block) %>% count() %>% arrange(., desc(n)) # View()
 # --> it seems that the vignettes 28 (xor) and 45 (some) have this issue the most
 # check how often the single participants do this
-d_main_implicature_fail %>% count() %>% View()
+d_main_implicature_fail %>% count() # %>% View()
 # --> all of them do this on all of the trials - these are 15 people 
@@
 
+d_critical_noDuplicates_zScored_wide <- d_critical_noDuplicates_zScored %>% 
+  select(submission_id, title, main_type, block_extended, response_centered) %>% 
+  unique() %>% 
+  pivot_wider(
+    names_from = block_extended, 
+    values_from = response_centered, 
+    values_fn = mean # getting means for double prior measurement in "xor"
+  ) 
+```
+
+Z-score data when excluding participants who failed the implicature ratings: 
+```{r}
+d_critical_clean_noImplFails <- d_clean_exclude_implicatureFails %>% filter(condition == "critical")
+d_critical_long_noImplFails <- d_critical_clean_noImplFails %>% 
+  pivot_longer(c(competence, relevance, prior), 
+               names_to = "class_condition", 
+               values_to = "prior_class")
+
+d_critical_long_noImplFails <- d_critical_long_noImplFails %>% 
+  mutate(block = ifelse(block == "comp", "competence", 
+                        ifelse(block == "rel", "relevance", ifelse(block == "pri", "prior", block) )))
+
+d_critical_long_noImplFails <- d_critical_long_noImplFails %>% 
+  mutate(
+    block_extended =  ifelse(block %in% c("some", "xor"), "target", block)
+  )
+
+# z-scoring, excluding duplicates
+d_critical_zScored_noImplFails_noDuplicates <- d_critical_long_noImplFails %>% 
+  filter((block_extended == class_condition) | (class_condition =="relevance" & block_extended == "target") )
+
+d_critical_noImplFails_noDuplicates_zScored <- d_critical_zScored_noImplFails_noDuplicates %>% group_by(submission_id, block_extended) %>%
+  mutate(block_mean = mean(response),
+         block_sd = sd(response),
+         response_centered = (response - block_mean)/block_sd,
+         # catch the cases where sd is 0 
+         response_centered = ifelse(is.na(response_centered), 0, response_centered))
+
+d_critical_noImplFails_noDuplicates_zScored_wide <- d_critical_noImplFails_noDuplicates_zScored %>% 
+  select(submission_id, title, main_type, block_extended, response_centered) %>% 
+  unique() %>% 
+  pivot_wider(
+    names_from = block_extended, 
+    values_from = response_centered, 
+    values_fn = mean # getting means for double prior measurement in "xor"
+  ) 
 ```
@@
 d_critical_noDuplicates_zScored %>% write_csv("./../../data/main/results_prereg_tidy_noDuplicates_zScored_long.csv")
+d_critical_noDuplicates_zScored_wide %>% write_csv("./../../data/main/results_prereg_tidy_noDuplicates_zScored_wide.csv")
+d_critical_noImplFails_noDuplicates_zScored %>%
+  write_csv("./../../data/main/results_prereg_tidy_noDuplicates_noImplFails_zScored_long.csv")
+d_critical_noImplFails_noDuplicates_zScored_wide %>%
+  write_csv("./../../data/main/results_prereg_tidy_noDuplicates_noImplFails_zScored_wide.csv")
 

### 2022-04-17 analysis/main/02_xor-some_prereg_plots_final.Rmd
fix analyses and plots with correctly z-scored data

@@
 ```{r, warning=FALSE, message=FALSE, results='hide'}
-d_critical_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
-d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")
+d_critical_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_noDuplicates_zScored_wide.csv")
+#d_critical_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
+d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_noDuplicates_zScored_long.csv")
+#d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")
 d_test <- read_csv("./../../data/main/results_prereg_raw_final_test.csv")
@@
 
-ggsave("../../writing/ELM2/byItem-scatter-wTitles.pdf", width = 9, height = 7)
+#ggsave("../../writing/ELM2/byItem-scatter-wTitles.pdf", width = 9, height = 7)
 
@@
 
-ggsave("../../writing/ELM2/byFactor-byPrior-raw-scatter.pdf", width = 9, height = 7)
+#ggsave("../../writing/ELM2/byFactor-byPrior-raw-scatter.pdf", width = 9, height = 7)
 ```
+
+### Simulated plots for the paper
+
+Below, some simulations are conducted to create plots for the paper.

### 2022-04-17 analysis/main/03_xor-some_prereg_analysis_final.Rmd
fix analyses and plots with correctly z-scored data

--- 2022-01-16 analysis/main/03_xor-some_prereg_analysis_final.Rmd
@@
 d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")
+
+# add correctly filtered dataframes 
+d_critical_noDuplicates_zScored <- read_csv("./../../data/main/results_prereg_tidy_noDuplicates_zScored_long.csv")
+d_critical_noDuplicates_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_noDuplicates_zScored_wide.csv")
+d_critical_noImplFails_noDuplicates_zScored <-
+  read_csv("./../../data/main/results_prereg_tidy_noDuplicates_noImplFails_zScored_long.csv")
+d_critical_noImplFails_noDuplicates_zScored_wide <-
+  read_csv("./../../data/main/results_prereg_tidy_noDuplicates_noImplFails_zScored_wide.csv")
 
@@
 
+Now perform the main confirmatory analysis on correctly z-scored data (no duplicates), to double-check that the results don't change qualitatively.
+```{r, results='hide', message=FALSE, warning=FALSE}
+# Full model 
+d_critical_noDuplicates_zScored_wide <- d_critical_noDuplicates_zScored_wide %>% mutate(
+  main_type = as.factor(main_type)
+)
+# some = 0, xor = 1
+contrasts(d_critical_noDuplicates_zScored_wide$main_type)
+
+# set priors
+priors <- set_prior("student_t(1, 0, 2)", class = "b")
+
+model_SI_no_dupl <- brm(target ~ prior*competence*relevance* main_type +
+                   (1 + prior + competence + relevance + main_type || submission_id) +
+                   (1 | title),
+                 data = d_critical_noDuplicates_zScored_wide,
+                 prior = priors,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.95),
+                 cores = 4,
+                 iter = 3000)
+```
+
+```{r}
+summary(model_SI_no_dupl)
+```
+
+Extract contrasts:
+```{r}
+model_SI_no_dupl %>% spread_draws(b_Intercept, b_prior, b_competence, b_relevance,
+                          b_main_typexor, `b_prior:competence`,
+                          `b_prior:relevance`, `b_competence:relevance`,
+                          `b_prior:main_typexor`, `b_competence:main_typexor`,
+                          `b_relevance:main_typexor`, `b_prior:competence:relevance`, 
+                          `b_prior:competence:main_typexor`, `b_prior:relevance:main_typexor`,
+                          `b_competence:relevance:main_typexor`, 
+                          `b_prior:competence:relevance:main_typexor`) %>% 
+  mutate(
+    prior_xor = b_prior + `b_prior:main_typexor`,
+    prior_some = b_prior,
+    competence_xor = b_competence + `b_competence:main_typexor`,
+    competence_some = b_competence,
+    relevance_xor =  b_relevance + `b_relevance:main_typexor`,
+    relevance_some = b_relevance
+  ) -> model_SI_no_dupl_posteriors
+
+# check P that the effects are positive / negative / no effect present
+posterior_hypotheses_no_dupl <- model_SI_no_dupl_posteriors %>% 
+  select(prior_xor, prior_some, 
+         competence_xor, competence_some,
+         relevance_xor, relevance_some) %>%
+  gather(key, val) %>%
+  group_by(key) %>% mutate(positive = mean(val > 0.05),
+                           negative = mean(val < -0.05),
+                           no = mean(val %>% between(-0.05, 0.05))) %>%
+  summarise(negative_eff = mean(negative),
+            no_eff = mean(no),
+            positive_eff = mean(positive)
+            )
+posterior_hypotheses_no_dupl
+
+# write out for reading values in paper
+#posterior_hypotheses_no_dupl %>% write_csv("./../../writing/paper/data/posterior_h_confirmatory.csv")
+```
+Also check if the 0 ROPE results change on the correctly z-scored data:
+```{r}
+posterior_hypotheses_nullROPE_no_dupl <- model_SI_no_dupl_posteriors %>% 
+  select(prior_xor, prior_some, 
+         competence_xor, competence_some,
+         relevance_xor, relevance_some) %>%
+  gather(key, val) %>%
+  group_by(key) %>% mutate(positive = mean(val > 0),
+                           negative = mean(val < 0)) %>% # no = mean(val = 0)
+  summarise(positive_eff = mean(positive),
+            negative_eff = mean(negative) # no_eff = mean(no)
+            )
+posterior_hypotheses_nullROPE_no_dupl
+#posterior_hypotheses_nullROPE_no_dupl %>% write_csv("./../../writing/paper/data/posterior_h_confirmatory_0-ROPE.csv")
+```
+
+Now investigate if the confirmatory results change if excluding participants who were resilient to contextual manipulations:
+```{r, results='hide', message=FALSE, warning=FALSE}
+# Full model 
+d_critical_noImplFails_noDuplicates_zScored_wide <- d_critical_noImplFails_noDuplicates_zScored_wide %>% mutate(
+  main_type = as.factor(main_type)
+)
+# some = 0, xor = 1
+contrasts(d_critical_noImplFails_noDuplicates_zScored_wide$main_type)
+
+# set priors
+priors <- set_prior("student_t(1, 0, 2)", class = "b")
+
+model_SI_no_dupl_no_impl <- brm(target ~ prior*competence*relevance* main_type +
+                   (1 + prior + competence + relevance + main_type || submission_id) +
+                   (1 | title),
+                 data = d_critical_noImplFails_noDuplicates_zScored_wide,
+                 prior = priors,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.95),
+                 cores = 4,
+                 iter = 3000)
+```
+
+```{r}
+summary(model_SI_no_dupl_no_impl)
+```
+
+Contrasts:
+```{r}
+model_SI_no_dupl_no_impl %>% spread_draws(b_Intercept, b_prior, b_competence, b_relevance,
+                          b_main_typexor, `b_prior:competence`,
+                          `b_prior:relevance`, `b_competence:relevance`,
+                          `b_prior:main_typexor`, `b_competence:main_typexor`,
+                          `b_relevance:main_typexor`, `b_prior:competence:relevance`, 
+                          `b_prior:competence:main_typexor`, `b_prior:relevance:main_typexor`,
+                          `b_competence:relevance:main_typexor`, 
+                          `b_prior:competence:relevance:main_typexor`) %>% 
+  mutate(
+    prior_xor = b_prior + `b_prior:main_typexor`,
+    prior_some = b_prior,
+    competence_xor = b_competence + `b_competence:main_typexor`,
+    competence_some = b_competence,
+    relevance_xor =  b_relevance + `b_relevance:main_typexor`,
+    relevance_some = b_relevance
+  ) -> model_SI_no_dupl_no_impl_posteriors
+
+# check P that the effects are positive / negative / no effect present
+posterior_hypotheses_no_dupl_no_impl <- model_SI_no_dupl_no_impl_posteriors %>% 
+  select(prior_xor, prior_some, 
+         competence_xor, competence_some,
+         relevance_xor, relevance_some) %>%
+  gather(key, val) %>%
+  group_by(key) %>% mutate(positive = mean(val > 0.05),
+                           negative = mean(val < -0.05),
+                           no = mean(val %>% between(-0.05, 0.05))) %>%
+  summarise(positive_eff = mean(positive),
+            negative_eff = mean(negative),
+            no_eff = mean(no))
+posterior_hypotheses_no_dupl_no_impl
+```
+
+0 ROPE:
+```{r}
+posterior_hypotheses_nullROPE_no_dupl_no_impl <- model_SI_no_dupl_no_impl_posteriors %>% 
+  select(prior_xor, prior_some, 
+         competence_xor, competence_some,
+         relevance_xor, relevance_some) %>%
+  gather(key, val) %>%
+  group_by(key) %>% mutate(positive = mean(val > 0),
+                           negative = mean(val < 0)) %>% # no = mean(val = 0)
+  summarise(positive_eff = mean(positive),
+            negative_eff = mean(negative) # no_eff = mean(no)
+            )
+posterior_hypotheses_nullROPE_no_dupl_no_impl
+```
+
 #### Perform BF analysis
@@
 test_conjunction_of_all_hypotheses(posterior_hypotheses)
+
+# applied to correctly z-scored data
+test_conjunction_of_all_hypotheses(posterior_hypotheses_no_dupl)
+test_conjunction_of_all_hypotheses(posterior_hypotheses_no_dupl_no_impl)
+
 ```
@@
 ```{r, fig.height = 6, fig.width=9}
-model_SI_posteriors %>% select(prior_xor, prior_some,
+# model_SI_posteriors
+model_SI_no_dupl_posteriors %>% select(prior_xor, prior_some,
                                competence_xor, competence_some,
@@
 
-# ggsave("../../writing/ELM2/posterior-effects.pdf", width=7, height=5)
+#ggsave("../../writing/ELM2/posterior-effects.pdf", width=7, height=5)
+#ggsave("../../writing/ELM2/posterior-effects-noDupl.pdf", width=7, height=5)
 ```
@@
 ```{r, results='hide', message=FALSE, warning=FALSE}
-d_critical_zScored %>% 
-  select(submission_id, title, main_type, block_extended, response_centered, prior_class, class_condition) %>% 
+#d_critical_zScored %>% 
+d_critical_noDuplicates_zScored %>%
+  select(submission_id, title, main_type, response_centered, prior_class, class_condition, block_extended) %>% # ,
   unique() %>% 
@@
     values_from = prior_class
-  )  %>%
+  )  -> d_critical_wide_cat_zScored
+
+# wrangle prior categorization of the items for the correctly z-scored data 
+d_critical_wide_cat_zScored %>%
+  replace_na(list(relevance = 0, competence = 0, prior = 0)) %>%
+  group_by(submission_id, title) %>%
+  filter(block_extended != "target") %>%
+  summarise_at(c( "relevance", "competence", "prior"), funs(sum)) %>%
+  mutate(prior = ifelse(prior == 2, 1, prior)) -> d_critical_wide_cat_zScored_priors
+
+d_critical_wide_cat_zScored_targets <- d_critical_wide_cat_zScored %>%
+  select(submission_id, title, main_type, block_extended, response_centered) %>%
+  filter(block_extended == "target") %>% select(-block_extended)
+
+d_critical_wide_cat_zScored_full <- left_join(d_critical_wide_cat_zScored_priors, d_critical_wide_cat_zScored_targets, by = c("submission_id", "title") )
+
+d_critical_wide_cat_zScored_full %>%
   mutate(relevance = factor(relevance, levels = c(1, 0)),
@@
          prior = factor(prior, levels = c(1, 0)),
-         main_type = as.factor(main_type)) %>% 
-  filter(block_extended == "target") -> d_critical_wide_cat_zScored
+         main_type = as.factor(main_type)) -> d_critical_wide_cat_zScored #%>% 
+#  filter(block_extended == "target") -> d_critical_wide_cat_zScored
 # sum code predictors
@@
 posterior_hypotheses_cat
+#posterior_hypotheses_cat %>% write_csv("./../../writing/paper/data/posterior_h_exploratory_categorical.csv")
 ```
@@
 #ggsave("../../writing/paper/images/posterior-effects-categorical.pdf", width=7, height=5)
+#ggsave("../../writing/paper/images/posterior-effects-categorical-noDupl.pdf", width=7, height=5)
 ```
@@
 ```{r, results='hide', message=FALSE, warning=FALSE}
-d_critical_zScored_wide_byItem <- d_critical_zScored_wide %>% ungroup() %>%
+d_critical_zScored_wide_byItem <- d_critical_noDuplicates_zScored_wide %>% ungroup() %>%
   group_by(title, main_type) %>%
@@
 posterior_hypotheses_means
+#posterior_hypotheses_means %>% write_csv("./../../writing/paper/data/posterior_h_exploratory_means.csv")
 ```
@@
         strip.text.y = element_text(size = 13))
-
+#ggsave("../../writing/paper/images/posterior-effects-means.pdf", width=7, height=5)
+#ggsave("../../writing/paper/images/posterior-effects-means-noDupl.pdf", width=7, height=5)
 ```
 
+[NOTE: Analyses below are not replicated on data z-scored correctly.]
 A non-preregistered, but nevertheless possibly interesting exploratory analysis including the two individual prior ratings of the conditionals in the xor condition as predictors (as opposed to the mean of the two used in the preceding models):

### 2022-04-17 writing/ELM2/abstract.tex
fix analyses and plots with correctly z-scored data

@@
 \newtoggle{anonymous}
-\toggletrue{anonymous}
+\togglefalse{anonymous}
 
@@
 \title{The role of relevance, competence and priors for scalar implicatures}
-\authors{Michael Franke (University of T\"ubingen), Bob van Tiel (Radboud University), Polina Tsvilodub (Osnabr\"uck university)}
-\email{pstvilodub@uos.de}
-
+\authors{Polina Tsvilodub (Osnabr\"uck University, ptsvilodub@uos.de), Bob van Tiel (Radboud University), Michael Franke (University of T\"ubingen)}
+%\email{}
 % Here goes the main text of your abstract:
-
 If someone says ``Anna ate some cookies'', the hearer might infer the upper-bounded reading that Anna ate \textit{some, but not all} cookies.
@@
 While supporting the view that SIs for both triggers rely on epistemic reasoning affected by speaker competence, our results indicate that prior and relevance might be closely connected for ``or''.
-Ultimately, our results suggest that the interdependence of the three factors is more complex than just the sum of the effects anticipated in the literature, and provide further insights into how exactly they might enter into people’s decision as to whether or not to derive an SI.
+Ultimately, our results suggest that the interdependence of the three factors is more complex than just the sum of the effects anticipated in the literature, and provide further insights into how they might enter into people’s decision as to whether or not to derive an SI.
 

### 2022-04-17 analysis/main/02_xor-some_prereg_plots_final.Rmd
create dist and posterior plots on simulated data

@@
 Below, some simulations are conducted to create plots for the paper.
+Simulate data similarly to the power analysis, assuming an idealized effect size of 0.2 (somewhat larger than in power analysis, for better visibility).
+Modify the sampling a little to create low and high condition data.
+```{r}
+# simulate data from 200 participants with the function from the power analysis
+create_fake_data <- function(N = 100, seed = 1) {
+  set.seed(seed)
+  # optionally changed to 0.25
+  beta_rel  <- 0.25
+  beta_comp <- 0.25
+  beta_pri  <- -0.25
+  map_df(
+    1:N, 
+    function(i) {
+      tibble(
+        subj    = rep(str_c("subj_",i), times = 8),
+        trigger = rep(c('or', 'some'), times= 4), 
+        rel     = rnorm(8),
+        comp    = rnorm(8),
+        pri     = rnorm(8)
+      ) %>% 
+        group_by(subj) %>% 
+        mutate(
+          rel     = (rel-mean(rel)) / sd(rel),
+          comp    = (comp-mean(comp)) / sd(comp),
+          pri     = (pri-mean(pri)) / sd(pri)
+        ) %>% 
+        ungroup() %>% 
+        mutate(
+          target  = rnorm(8, beta_rel * rel + beta_comp * comp + beta_pri * pri)
+        )
+    }
+  )
+}
+
+fake_data_scatter <- create_fake_data(200, 1234)
+```
+
+```{r}
+fake_data_long <- fake_data_scatter %>% pivot_longer(
+  c(comp, rel, pri), 
+               names_to = "Predictors", 
+               values_to = "Predictor_response"
+) %>%
+  mutate(
+    prior_categorization = ifelse(Predictor_response <= 0, 0, 1),
+    prior_categorization = as.factor(prior_categorization)
+  ) %>%
+    mutate(
+      `Prior categorization` = factor(prior_categorization, levels = c(0,1),
+                                         labels=c("low", "high")),
+         Trigger = factor(trigger, levels = c("some", "or"), 
+                            labels = c("Some", "Or")),
+         Predictors = factor(Predictors, levels = c("rel", "comp", "pri"),
+                             labels = c("Relevance", "Competence", "Prior"))
+    )
+
+# write out for fitting the idealized model in the analysis script
+#fake_data_long %>% write_csv("../../data/main/idealized_simulation_long_N200_B025.csv")
+
+fake_data_long %>%
+  ggplot(., aes(x = Predictor_response, y = target)) +
+  geom_point(aes(color = `Prior categorization`), size = 2, alpha = 0.8) +
+  geom_smooth(method = "lm") +
+  ylab("") +
+  facet_grid(cols=vars(Predictors), rows=vars(Trigger)) +
+  theme_bw() +
+  labs(x = "Participants' factor ratings", y = "Participants' inference likelihood ratings", color = "Prior categorization of vignettes") +
+  theme(legend.position = "top",
+        legend.title = element_text(size = 13),
+        legend.text = element_text(size = 13),
+        axis.title = element_text(size = 13),
+        strip.text.x = element_text(size = 13),
+        strip.text.y = element_text(size = 13)) 
+#ggsave("../../writing/paper/images/idealized_data_N200_b025.pdf",  width = 9, height = 7)
+```

### 2022-04-17 analysis/main/03_xor-some_prereg_analysis_final.Rmd
create dist and posterior plots on simulated data

@@
 ```
+
+### Analysis on fake simulated data for visualization purposes
+No random effects here since none were simulated.
+```{r}
+fake_data <- read_csv("../../data/main/idealized_simulation_long_N200_B025.csv")
+
+# Full model 
+fake_data_wide <- fake_data %>% select(-prior_categorization, -`Prior categorization`, -Trigger ) %>% pivot_wider(
+  names_from = Predictors, values_from = Predictor_response
+) %>% 
+  mutate(
+    pri = Prior,
+    comp = Competence,
+    rel = Relevance,
+  main_type = factor(trigger, levels=c("some", "or"))
+)
+
+# some = 0, xor = 1
+contrasts(fake_data_wide$main_type)
+
+# set priors
+priors <- set_prior("student_t(1, 0, 2)", class = "b")
+
+model_fake <- brm(target ~ pri*comp*rel* main_type,
+                 data = fake_data_wide,
+                 prior = priors,
+                 sample_prior = T,
+                 control = list(adapt_delta = 0.9),
+                 cores = 4,
+                 iter = 3000)
+
+# posteriors
+model_fake %>% spread_draws(b_Intercept, b_pri, b_comp, b_rel,
+                          b_main_typeor, `b_pri:comp`,
+                          `b_pri:rel`, `b_comp:rel`,
+                          `b_pri:main_typeor`, `b_comp:main_typeor`,
+                          `b_rel:main_typeor`, `b_pri:comp:rel`, 
+                          `b_pri:comp:main_typeor`, `b_pri:rel:main_typeor`,
+                          `b_comp:rel:main_typeor`, 
+                          `b_pri:comp:rel:main_typeor`) %>% 
+  mutate(
+    prior_xor = b_pri+ `b_pri:main_typeor`,
+    prior_some = b_pri,
+    competence_xor = b_comp + `b_comp:main_typeor`,
+    competence_some = b_comp,
+    relevance_xor =  b_rel + `b_rel:main_typeor`,
+    relevance_some = b_rel
+  ) -> model_fake_posteriors
+
+```
+
+```{r}
+model_fake_posteriors %>% select(prior_xor, prior_some,
+                               competence_xor, competence_some,
+                               relevance_xor, relevance_some) %>%
+  pivot_longer(cols = everything(), names_to = "effect", values_to = "value") -> model_fake_posteriors_long
+
+model_fake_posteriors_long %>%
+  mutate(effect = factor(effect, 
+                         levels = c("relevance_xor", "relevance_some", "prior_xor", "prior_some", "competence_xor", "competence_some"), 
+                         labels = c("Relevance (or)", "Relevance (some)", "Prior (or)", "Prior (some)", "Competence (or)", "Competence (some)"))) %>%
+  ggplot(.) + # 
+  stat_halfeye(aes(y = effect, x = value, 
+                   fill = stat(case_when(x %>% between(-0.05, 0.05) ~ "no effect", x < -0.05 ~ "negative effect", x > 0.05 ~ "positive effect"))
+                   ), 
+               alpha = .7) +
+  geom_vline(xintercept = c(-.05, .05), linetype = "dashed") +
+  scale_fill_manual(name = "Direction of effect", values = c("coral", "mediumseagreen")) + # , "yellow3"
+  labs(x = "Effect size", y = "Effect") +
+  theme(legend.position = "top",
+        legend.title = element_text(size = 12),
+        legend.text = element_text(size = 12),
+        axis.title = element_text(size = 13),
+        strip.text.x = element_text(size = 13),
+        strip.text.y = element_text(size = 13))
+
+#ggsave("../../writing/paper/images/posterior-effects-idealized-sim.pdf", width=7, height=5)
+```

### 2022-05-10 analysis/main/01_xor-some_prereg_preprocessing_final.Rmd
plotting and explorations for paper and presentation

@@
 ```
+```{r}
+# look at raw ratings
+# df to be used before applying z-scoring in cell above 
+d_critical_zScored_noDuplicates %>%
+  filter(block_extended != "target") %>%
+  ungroup() %>%
+  group_by(block, prior_class, main_type) %>%
+  summarize(mean_rating = mean(response)) -> by_type_factor_means
+
+d_critical_zScored_noDuplicates %>%
+  filter(block_extended!="target") %>%
+  ggplot(., aes(x=main_type, y = response)) +
+  geom_jitter(aes(color=prior_class)) +
+  geom_point(data = by_type_factor_means, aes(x=main_type, y = mean_rating), color="red") +
+  facet_grid(prior_class~block)
+```
 

### 2022-05-10 analysis/main/02_xor-some_prereg_plots_final.Rmd
additional exploratory plotting
plotting and explorations for paper and presentation

--- 2022-04-17 analysis/main/02_xor-some_prereg_plots_final.Rmd
@@
 Below, some simulations are conducted to create plots for the paper.
+Simulate data similarly to the power analysis, assuming an idealized effect size of 0.2 (somewhat larger than in power analysis, for better visibility).
+Modify the sampling a little to create low and high condition data.
+```{r}
+# simulate data from 200 participants with the function from the power analysis
+create_fake_data <- function(N = 100, seed = 1) {
+  set.seed(seed)
+  # optionally changed to 0.25
+  beta_rel  <- 0.25
+  beta_comp <- 0.25
+  beta_pri  <- -0.25
+  map_df(
+    1:N, 
+    function(i) {
+      tibble(
+        subj    = rep(str_c("subj_",i), times = 8),
+        trigger = rep(c('or', 'some'), times= 4), 
+        prior = rep(c("low", "high"), each=4),
+        rel     = c(rnorm(4, mean=-1), rnorm(4, mean =1)),
+       # rel_high = rnorm(4, mean = 0.15),
+        comp    = c(rnorm(4, mean=-1), rnorm(4, mean=1)),
+       # comp_high=rnorm(4, mean=0.15),
+        pri     = c(rnorm(4, mean=1), rnorm(4, mean=-1))
+    #    pri_high= rnorm(4, mean=0.15)
+      ) %>% 
+        group_by(subj) %>% 
+        mutate(
+          rel     = (rel-mean(rel)) / sd(rel),
+          comp    = (comp-mean(comp)) / sd(comp),
+          pri     = (pri-mean(pri)) / sd(pri)
+        ) %>% 
+        ungroup() %>% 
+        mutate(
+          target  = rnorm(8, beta_rel * rel + beta_comp * comp + beta_pri * pri)
+        )
+    }
+  )
+}
+
+fake_data_scatter <- create_fake_data(200, 1234)
+```
+
+```{r}
+fake_data_long <- fake_data_scatter %>% pivot_longer(
+  c(comp, rel, pri), 
+               names_to = "Predictors", 
+               values_to = "Predictor_response"
+) %>%
+  mutate(
+    prior_categorization = ifelse(Predictor_response <= 0, 0, 1),
+    prior_categorization = as.factor(prior_categorization)
+  ) %>%
+    mutate(
+      `Prior categorization` = factor(prior_categorization, levels = c(0,1),
+                                         labels=c("low", "high")),
+         Trigger = factor(trigger, levels = c("some", "or"), 
+                            labels = c("Some", "Or")),
+         Predictors = factor(Predictors, levels = c("rel", "comp", "pri"),
+                             labels = c("Relevance", "Competence", "Prior"))
+    )
+
+# write out for fitting the idealized model in the analysis script
+# fake_data_long %>% write_csv("../../data/main/idealized_simulation_long_N200_B025.csv")
+
+fake_data_long %>%
+  ggplot(., aes(x = Predictor_response, y = target)) +
+  geom_point(aes(color = prior), size = 2, alpha = 0.8) +
+  geom_smooth(method = "lm") +
+  ylab("") +
+  facet_grid(cols=vars(Predictors), rows=vars(Trigger)) +
+  theme_bw() +
+  labs(x = "Participants' factor ratings", y = "Participants' inference likelihood ratings", color = "Prior categorization of vignettes") +
+  theme(legend.position = "top",
+        legend.title = element_text(size = 13),
+        legend.text = element_text(size = 13),
+        axis.title = element_text(size = 13),
+        strip.text.x = element_text(size = 13),
+        strip.text.y = element_text(size = 13)) 
+#ggsave("../../writing/paper/images/idealized_data_N200_b025.pdf",  width = 9, height = 7)
+```

### 2022-05-10 analysis/main/03_xor-some_prereg_analysis_final.Rmd
plotting and explorations for paper and presentation

@@
   scale_fill_manual(name = "Direction of effect", values = c("coral", "yellow3", "mediumseagreen")) +
+  scale_x_continuous(breaks=c(-0.25, 0, 0.25)) +
   labs(x = "Effect size", y = "Effect") +
@@
 #ggsave("../../writing/ELM2/posterior-effects-noDupl.pdf", width=7, height=5)
+#ggsave("../../writing/paper/images/posterior-effects-noDupl-breaks.png", width=7, height=5)
+
 ```
@@
   geom_vline(xintercept = c(-.05, .05), linetype = "dashed") +
-  scale_fill_manual(name = "Direction of effect", values = c("coral", "mediumseagreen")) + # , "yellow3"
+  scale_fill_manual(name = "Direction of effect", values = c("coral", "yellow3", "mediumseagreen")) + 
+  scale_x_continuous( breaks=c(-0.25, 0, 0.25)) +
   labs(x = "Effect size", y = "Effect") +
@@
 
-#ggsave("../../writing/paper/images/posterior-effects-idealized-sim.pdf", width=7, height=5)
+model_fake_posteriors %>% 
+  select(prior_xor, prior_some, 
+         competence_xor, competence_some,
+         relevance_xor, relevance_some) %>%
+  gather(key, val) %>%
+  group_by(key) %>% mutate(positive = mean(val > 0.05),
+                           negative = mean(val < -0.05),
+                           no = mean(val %>% between(-0.05, 0.05))) %>%
+  summarise(positive_eff = mean(positive),
+            negative_eff = mean(negative),
+            no_eff = mean(no))
+
+#ggsave("../../writing/paper/images/posterior-effects-idealized-sim-015.pdf", width=7, height=5)
 ```

### 2022-05-10 analysis/main/02_xor-some_prereg_plots_final.Rmd

@@
 ```{r, warning=FALSE, message=FALSE, results='hide'}
-d_critical_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_noDuplicates_zScored_wide.csv")
-#d_critical_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
-d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_noDuplicates_zScored_long.csv")
-#d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")
+d_critical_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
+d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")
 d_test <- read_csv("./../../data/main/results_prereg_raw_final_test.csv")
@@
 
-#ggsave("../../writing/ELM2/byItem-scatter-wTitles.pdf", width = 9, height = 7)
+ggsave("../../writing/ELM2/byItem-scatter-wTitles.pdf", width = 9, height = 7)
 
@@
   filter(d_critical_zScored_wide, main_type == "some") %>% ungroup() %>%  
+    select(prior, competence, relevance, target)
+)
+```
+
+Plot the pairwise correlations of the predictors and the inference rating for "or": 
+```{r, message=FALSE, warning=FALSE}
+GGally::ggpairs(
+  filter(d_critical_zScored_wide, main_type == "xor") %>% ungroup() %>%  
     select(prior, competence, relevance, target)

### 2022-05-10 analysis/main/03_xor-some_prereg_analysis_final.Rmd

@@
 d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")
-
-# add correctly filtered dataframes 
-d_critical_noDuplicates_zScored <- read_csv("./../../data/main/results_prereg_tidy_noDuplicates_zScored_long.csv")
-d_critical_noDuplicates_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_noDuplicates_zScored_wide.csv")
-d_critical_noImplFails_noDuplicates_zScored <-
-  read_csv("./../../data/main/results_prereg_tidy_noDuplicates_noImplFails_zScored_long.csv")
-d_critical_noImplFails_noDuplicates_zScored_wide <-
-  read_csv("./../../data/main/results_prereg_tidy_noDuplicates_noImplFails_zScored_wide.csv")
 
@@
 
-Now perform the main confirmatory analysis on correctly z-scored data (no duplicates), to double-check that the results don't change qualitatively.
-```{r, results='hide', message=FALSE, warning=FALSE}
-# Full model 
-d_critical_noDuplicates_zScored_wide <- d_critical_noDuplicates_zScored_wide %>% mutate(
-  main_type = as.factor(main_type)
-)
-# some = 0, xor = 1
-contrasts(d_critical_noDuplicates_zScored_wide$main_type)
-
-# set priors
-priors <- set_prior("student_t(1, 0, 2)", class = "b")
-
-model_SI_no_dupl <- brm(target ~ prior*competence*relevance* main_type +
-                   (1 + prior + competence + relevance + main_type || submission_id) +
-                   (1 | title),
-                 data = d_critical_noDuplicates_zScored_wide,
-                 prior = priors,
-                 sample_prior = T,
-                 control = list(adapt_delta = 0.95),
-                 cores = 4,
-                 iter = 3000)
-```
-
-```{r}
-summary(model_SI_no_dupl)
-```
-
-Extract contrasts:
-```{r}
-model_SI_no_dupl %>% spread_draws(b_Intercept, b_prior, b_competence, b_relevance,
-                          b_main_typexor, `b_prior:competence`,
-                          `b_prior:relevance`, `b_competence:relevance`,
-                          `b_prior:main_typexor`, `b_competence:main_typexor`,
-                          `b_relevance:main_typexor`, `b_prior:competence:relevance`, 
-                          `b_prior:competence:main_typexor`, `b_prior:relevance:main_typexor`,
-                          `b_competence:relevance:main_typexor`, 
-                          `b_prior:competence:relevance:main_typexor`) %>% 
-  mutate(
-    prior_xor = b_prior + `b_prior:main_typexor`,
-    prior_some = b_prior,
-    competence_xor = b_competence + `b_competence:main_typexor`,
-    competence_some = b_competence,
-    relevance_xor =  b_relevance + `b_relevance:main_typexor`,
-    relevance_some = b_relevance
-  ) -> model_SI_no_dupl_posteriors
-
-# check P that the effects are positive / negative / no effect present
-posterior_hypotheses_no_dupl <- model_SI_no_dupl_posteriors %>% 
-  select(prior_xor, prior_some, 
-         competence_xor, competence_some,
-         relevance_xor, relevance_some) %>%
-  gather(key, val) %>%
-  group_by(key) %>% mutate(positive = mean(val > 0.05),
-                           negative = mean(val < -0.05),
-                           no = mean(val %>% between(-0.05, 0.05))) %>%
-  summarise(negative_eff = mean(negative),
-            no_eff = mean(no),
-            positive_eff = mean(positive)
-            )
-posterior_hypotheses_no_dupl
-
-# write out for reading values in paper
-#posterior_hypotheses_no_dupl %>% write_csv("./../../writing/paper/data/posterior_h_confirmatory.csv")
-```
-Also check if the 0 ROPE results change on the correctly z-scored data:
-```{r}
-posterior_hypotheses_nullROPE_no_dupl <- model_SI_no_dupl_posteriors %>% 
-  select(prior_xor, prior_some, 
-         competence_xor, competence_some,
-         relevance_xor, relevance_some) %>%
-  gather(key, val) %>%
-  group_by(key) %>% mutate(positive = mean(val > 0),
-                           negative = mean(val < 0)) %>% # no = mean(val = 0)
-  summarise(positive_eff = mean(positive),
-            negative_eff = mean(negative) # no_eff = mean(no)
-            )
-posterior_hypotheses_nullROPE_no_dupl
-#posterior_hypotheses_nullROPE_no_dupl %>% write_csv("./../../writing/paper/data/posterior_h_confirmatory_0-ROPE.csv")
-```
-
-Now investigate if the confirmatory results change if excluding participants who were resilient to contextual manipulations:
-```{r, results='hide', message=FALSE, warning=FALSE}
-# Full model 
-d_critical_noImplFails_noDuplicates_zScored_wide <- d_critical_noImplFails_noDuplicates_zScored_wide %>% mutate(
-  main_type = as.factor(main_type)
-)
-# some = 0, xor = 1
-contrasts(d_critical_noImplFails_noDuplicates_zScored_wide$main_type)
-
-# set priors
-priors <- set_prior("student_t(1, 0, 2)", class = "b")
-
-model_SI_no_dupl_no_impl <- brm(target ~ prior*competence*relevance* main_type +
-                   (1 + prior + competence + relevance + main_type || submission_id) +
-                   (1 | title),
-                 data = d_critical_noImplFails_noDuplicates_zScored_wide,
-                 prior = priors,
-                 sample_prior = T,
-                 control = list(adapt_delta = 0.95),
-                 cores = 4,
-                 iter = 3000)
-```
-
-```{r}
-summary(model_SI_no_dupl_no_impl)
-```
-
-Contrasts:
-```{r}
-model_SI_no_dupl_no_impl %>% spread_draws(b_Intercept, b_prior, b_competence, b_relevance,
-                          b_main_typexor, `b_prior:competence`,
-                          `b_prior:relevance`, `b_competence:relevance`,
-                          `b_prior:main_typexor`, `b_competence:main_typexor`,
-                          `b_relevance:main_typexor`, `b_prior:competence:relevance`, 
-                          `b_prior:competence:main_typexor`, `b_prior:relevance:main_typexor`,
-                          `b_competence:relevance:main_typexor`, 
-                          `b_prior:competence:relevance:main_typexor`) %>% 
-  mutate(
-    prior_xor = b_prior + `b_prior:main_typexor`,
-    prior_some = b_prior,
-    competence_xor = b_competence + `b_competence:main_typexor`,
-    competence_some = b_competence,
-    relevance_xor =  b_relevance + `b_relevance:main_typexor`,
-    relevance_some = b_relevance
-  ) -> model_SI_no_dupl_no_impl_posteriors
-
-# check P that the effects are positive / negative / no effect present
-posterior_hypotheses_no_dupl_no_impl <- model_SI_no_dupl_no_impl_posteriors %>% 
-  select(prior_xor, prior_some, 
-         competence_xor, competence_some,
-         relevance_xor, relevance_some) %>%
-  gather(key, val) %>%
-  group_by(key) %>% mutate(positive = mean(val > 0.05),
-                           negative = mean(val < -0.05),
-                           no = mean(val %>% between(-0.05, 0.05))) %>%
-  summarise(positive_eff = mean(positive),
-            negative_eff = mean(negative),
-            no_eff = mean(no))
-posterior_hypotheses_no_dupl_no_impl
-```
-
-0 ROPE:
-```{r}
-posterior_hypotheses_nullROPE_no_dupl_no_impl <- model_SI_no_dupl_no_impl_posteriors %>% 
-  select(prior_xor, prior_some, 
-         competence_xor, competence_some,
-         relevance_xor, relevance_some) %>%
-  gather(key, val) %>%
-  group_by(key) %>% mutate(positive = mean(val > 0),
-                           negative = mean(val < 0)) %>% # no = mean(val = 0)
-  summarise(positive_eff = mean(positive),
-            negative_eff = mean(negative) # no_eff = mean(no)
-            )
-posterior_hypotheses_nullROPE_no_dupl_no_impl
-```
-
 #### Perform BF analysis
@@
 test_conjunction_of_all_hypotheses(posterior_hypotheses)
-
-# applied to correctly z-scored data
-test_conjunction_of_all_hypotheses(posterior_hypotheses_no_dupl)
-test_conjunction_of_all_hypotheses(posterior_hypotheses_no_dupl_no_impl)
-
 ```
@@
 ```{r, fig.height = 6, fig.width=9}
-# model_SI_posteriors
-model_SI_no_dupl_posteriors %>% select(prior_xor, prior_some,
+model_SI_posteriors %>% select(prior_xor, prior_some,
                                competence_xor, competence_some,
@@
 ```{r, results='hide', message=FALSE, warning=FALSE}
-#d_critical_zScored %>% 
-d_critical_noDuplicates_zScored %>%
-  select(submission_id, title, main_type, response_centered, prior_class, class_condition, block_extended) %>% # ,
+d_critical_zScored %>% 
+  select(submission_id, title, main_type, block_extended, response_centered, prior_class, class_condition) %>% 
   unique() %>% 
@@
     values_from = prior_class
-  )  -> d_critical_wide_cat_zScored
-
-# wrangle prior categorization of the items for the correctly z-scored data 
-d_critical_wide_cat_zScored %>%
-  replace_na(list(relevance = 0, competence = 0, prior = 0)) %>%
-  group_by(submission_id, title) %>%
-  filter(block_extended != "target") %>%
-  summarise_at(c( "relevance", "competence", "prior"), funs(sum)) %>%
-  mutate(prior = ifelse(prior == 2, 1, prior)) -> d_critical_wide_cat_zScored_priors
-
-d_critical_wide_cat_zScored_targets <- d_critical_wide_cat_zScored %>%
-  select(submission_id, title, main_type, block_extended, response_centered) %>%
-  filter(block_extended == "target") %>% select(-block_extended)
-
-d_critical_wide_cat_zScored_full <- left_join(d_critical_wide_cat_zScored_priors, d_critical_wide_cat_zScored_targets, by = c("submission_id", "title") )
-
-d_critical_wide_cat_zScored_full %>%
+  )  %>%
   mutate(relevance = factor(relevance, levels = c(1, 0)),
@@
          prior = factor(prior, levels = c(1, 0)),
-         main_type = as.factor(main_type)) -> d_critical_wide_cat_zScored #%>% 
-#  filter(block_extended == "target") -> d_critical_wide_cat_zScored
+         main_type = as.factor(main_type)) %>% 
+  filter(block_extended == "target") -> d_critical_wide_cat_zScored
 # sum code predictors
@@
 ```
-
 ```{r}
@@
 posterior_hypotheses_cat
-#posterior_hypotheses_cat %>% write_csv("./../../writing/paper/data/posterior_h_exploratory_categorical.csv")
 ```
@@
 #ggsave("../../writing/paper/images/posterior-effects-categorical.pdf", width=7, height=5)
-#ggsave("../../writing/paper/images/posterior-effects-categorical-noDupl.pdf", width=7, height=5)
 ```
@@
 ```{r, results='hide', message=FALSE, warning=FALSE}
-d_critical_zScored_wide_byItem <- d_critical_noDuplicates_zScored_wide %>% ungroup() %>%
+d_critical_zScored_wide_byItem <- d_critical_zScored_wide %>% ungroup() %>%
   group_by(title, main_type) %>%
@@
 posterior_hypotheses_means
-#posterior_hypotheses_means %>% write_csv("./../../writing/paper/data/posterior_h_exploratory_means.csv")
 ```
@@
 #ggsave("../../writing/paper/images/posterior-effects-means.pdf", width=7, height=5)
-#ggsave("../../writing/paper/images/posterior-effects-means-noDupl.pdf", width=7, height=5)
 ```
 
-[NOTE: Analyses below are not replicated on data z-scored correctly.]
 A non-preregistered, but nevertheless possibly interesting exploratory analysis including the two individual prior ratings of the conditionals in the xor condition as predictors (as opposed to the mean of the two used in the preceding models):

### 2022-05-10 analysis/main/01_xor-some_prereg_preprocessing_final.Rmd
replicate MFs prior-means-first analysis

@@
 
+# perform the conceptually cleaner procedure: average over prior ratings first, then z-score
+d_critical_zScored_noDuplicates %>% unique() %>% 
+  select(submission_id, title, main_type, block_extended, response) %>%
+  pivot_wider(
+    names_from = block_extended, 
+    values_from = response, 
+    values_fn = mean # getting means for double prior measurement in "xor"
+  ) %>%
+  pivot_longer(
+    cols=c(relevance, competence, prior, target),
+    names_to = "block_extended",
+    values_to = "response"
+  ) -> d_critical_zScored_meanFirst_noDupl
+
+d_critical_zScored_meanFirst_noDupl_zScored <- d_critical_zScored_meanFirst_noDupl %>%  
+  group_by(submission_id, block_extended) %>%
+  mutate(block_mean = mean(response),
+         block_sd = sd(response),
+         response_centered = (response - block_mean)/block_sd,
+         # catch the cases where sd is 0 
+         response_centered = ifelse(is.na(response_centered), 0, response_centered))
+
+d_critical_zScored_meanFirst_noDupl_zScored_wide <- d_critical_zScored_meanFirst_noDupl_zScored %>% 
+  select(submission_id, title, main_type, block_extended, response_centered) %>% 
+  unique() %>% 
+  pivot_wider(
+    names_from = block_extended, 
+    values_from = response_centered, 
+    values_fn = mean # getting means for double prior measurement in "xor"
+  )
+
 d_critical_noDuplicates_zScored_wide <- d_critical_noDuplicates_zScored %>% 
@@
 
+d_critical_zScored_meanFirst_noDupl_zScored %>%
+  write_csv("./../../data/main/results_prereg_tidy_noDuplicates_zScored_meansFirst_long.csv")
+d_critical_zScored_meanFirst_noDupl_zScored_wide %>%
+  write_csv("./../../data/main/results_prereg_tidy_noDuplicates_zScored_meansFirst_wide.csv")
 ```

### 2022-05-11 analysis/main/02_xor-some_prereg_plots_final.Rmd
plot w means first data

@@
 d_critical_zScored_wide_xor_priors_type <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors-type.csv")
+
+d_critical_zScored_meanFirst_noDupl_zScored_wide <- 
+  read_csv("./../../data/main/results_prereg_tidy_noDuplicates_zScored_meansFirst_wide.csv")
+
 ```
@@
 d_critical_zScored_wide %>%
+  ggplot(., aes(x = prior, y = target)) +
+  geom_point(size = 2, alpha = 0.7) +
+  geom_smooth(method = "lm") +
+  ylab("") +
+  facet_wrap(~main_type, ncol = 1) -> p.pri.z
+
+gridExtra::grid.arrange(p.rel.z, p.comp.z, p.pri.z, ncol = 3) 
+```
+Replicate plot above on data where means were computed over the prior ratings for or before z-scoring:
+```{r}
+d_critical_zScored_meanFirst_noDupl_zScored_wide %>%
+  ggplot(., aes(x = relevance, y = target)) +
+  geom_point(size = 2, alpha = 0.7) +
+  ylab("Inference strength ratings") +
+  geom_smooth(method = "lm") +
+  facet_wrap(~main_type, ncol = 1) -> p.rel.z
+
+d_critical_zScored_meanFirst_noDupl_zScored_wide %>%
+  ggplot(., aes(x = competence, y = target)) +
+  geom_point(size = 2, alpha = 0.7) +
+  geom_smooth(method = "lm") +
+  ylab("") +
+  facet_wrap(~main_type, ncol = 1) -> p.comp.z
+
+d_critical_zScored_meanFirst_noDupl_zScored_wide %>%
   ggplot(., aes(x = prior, y = target)) +

### 2022-05-11 writing/paper/paper_draft.tex
plot w means first data

@@
   \spauthor{Bob van Tiel \\ \institute{Institute}} \AND
-  \spauthor{Michael Franke \\ \institute{Department of Linguistics \\ University of Tübigen}}}
+  \spauthor{Michael Franke \\ \institute{Department of Linguistics \\ University of Tübingen}}}
 
@@
 
+\begin{figure}[h]
+	\begin{center}
+		\includegraphics[width=1\linewidth]{images/idealized_data_N200_b015.pdf}
+	\end{center}
+	\vspace{-0.3cm}
+	\caption{Simulated ratings for relevance, competence and prior statements (x-axis) in relation to ratings for the strength of pragmatic enrichments (y-axis).
+Ratings for stories categorized as low (red) w.r.t.
+a given factor are on average lower (x-axis) than for those categorized as high (blue).}
+	\label{main-sim-ratings}
+\end{figure}
+
 Participants’ factor ratings by-story agreed well with the designed classification of the stories (Fig.
@@
 \ref{main-raw-ratings}).
-This provides validation for our experimental manipulation of the explanatory factors.
+This provides validation for our experimental manipulation of the explanatory factors, as can be seen when comparing Fig.
+\ref{main-raw-ratings} with Fig.
+\ref{main-sim-ratings} which depicts idealized data (N=200) simulating perfectly borne out hypotheses.
+\pt{Like this?}
 
@@
 In particular, we speak of evidence in favor of the scalar implicature account of the exclusive disjunction interpretation if the posterior probability of the prior effect being negative, the relevance and competence effects being positive is $\ge$ 0.95.
+Example simulated posterior distributions over effect size samples which would confirm all our hypotheses can be seen in Fig.
+\ref{posteriors-sim}.
 
@@
 	\begin{center}
-		\includegraphics[width=1\linewidth]{images/posterior-effects-main.pdf}
+		\includegraphics[width=1\linewidth]{images/posterior-effects-idealized-sim.pdf}
+	\end{center}
+	\vspace{-0.3cm}
+	\caption{ \pt{How's this?} Simulated main analysis results which would confirm all hypotheses: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''.
+The effect size was set to 0.25 and matches Fig.
+\ref{main-sim-ratings} for visualization purposes.
+The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect).
+Points indicate posterior means, thick lines indicate 66\% credible intervals, thin lines indicate 95\% credible intervals.}
+	\label{posteriors-sim}
+\end{figure}
+
+\begin{figure}[h]
+	\begin{center}
+		\includegraphics[width=1\linewidth]{images/posterior-effects-noDupl.pdf}
 	\end{center}
@@
            &            & negative & no effect & positive \\ \midrule
-    {some} & relevance  & ...      & ...       & ...      \\
+    {some} & relevance  & 0.019    & 0.592       & 0.389      \\
            & competence & 0        & 0         & 1        \\
-           & prior      & 0        & 0.001     & 0.999    \\
-    {or}   & relevance  & ...      & ...       & ...      \\
-           & competence & ...      & ...       & ...      \\
-           & prior      & ...      & ...       & ...      \\
+           & prior      & 0.999        & 0.001     & 0    \\
+    {or}   & relevance  & 0     & 0.228       & 0.772      \\
+           & competence & 0      & 0.007       & 0.993      \\
+           & prior      & 0.619    & 0.373       & 0.008      \\
    \end{tabular}
-  \caption{Results of confirmatory analysis. \mf{Fill me, please.}}
+  \caption{Results of confirmatory analysis.}
   \label{tab:results-confirmatory}
@@
 
-Results from this analysis are summarized in Table
+Results from this analysis are summarized in Table \ref{tab:results-confirmatory}.
 Consistent with predictions, for the trigger ``some'', we find a clear negative prior probabilty effect of the stronger alternative ``all'' being true, as inidcated by the probability of the negative effect of prior being $P = 0.999$ (Fig.\ref{posteriors-main}, \texttt{Prior~(some)} line, orange color).
@@
 	\begin{center}
-		\includegraphics[width=1\linewidth]{images/posterior-effects-categorical.pdf}
+		\includegraphics[width=1\linewidth]{images/posterior-effects-categorical-noDupl.pdf}
 	\end{center}
@@
 	\begin{center}
-		\includegraphics[width=1\linewidth]{images/posterior-effects-means.pdf}
+		\includegraphics[width=1\linewidth]{images/posterior-effects-means-noDupl.pdf}
 	\end{center}
@@
 Some more discussion needs to be added too.}
+
+\pt{We have a generally higher rate of implicature generations for ``some'' than for ``or'' (across conditions): the average inference likelihood rating for the former is 0.179, and -0.179 for the latter.
+Maybe something to mention in the discussion}
 

### 2022-05-12 analysis/main/mf-exploration.Rmd
added exploration files (role of by-item REs)

---
title: "Exploring influence of REs on models fits"
author: "Michael Franke"
date: "12/05/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = F, message = FALSE, warning = FALSE, error = FALSE, fig.width = 5, fig.align = "center")
```

```{r libraries, message = FALSE, warning = FALSE, include = FALSE}
library(tidyverse)
library(brms)
library(tidyboot)
library(tidybayes)
```

# Data

We use the data with the "new" z-scoring for the prior data for 'or'.

```{r}
d_critical_clean <- read_csv("../../data/main/results_prereg_tidy_noDuplicates_zScored_meansFirst_wide.csv") %>% 
  mutate(main_type = factor(main_type))
head(d_critical_clean)
```

We also want additional information about the vignettes, i.e., their intuitive pre-classifications.

```{r }
vignettes <- read_csv("./../../data/main/results_73_xor-some-Prolific-main_N275_anonym.csv") %>% 
  select(title, main_type, relevance, competence, prior) %>% 
  unique() %>% 
  filter(! is.na(main_type)) %>% 
  rename(relCat = relevance,
         compCat = competence,
         priCat = prior)
head(vignettes)
```

# Functions for fitting models

We are interested in comparing the following models, specified here in terms of `brms` formulas:

```{r }
## full model with REs for subjects and items
f_full <- formula(
  target ~ prior * competence * relevance * main_type +
    (1 + prior + competence + relevance + main_type || submission_id) +
    (1 | title)
)

## model with RE for 'title' only (by-item random intercepts)
f_RETitle <- formula(
  target ~ prior * competence * relevance * main_type +
    (1 | title)
)

## RE for 'submject_id' only (by-subj random intercepts)
f_RESubj <- formula(
  target ~ prior * competence * relevance * main_type +
    (1 + prior + competence + relevance + main_type || submission_id)
)

## model without REs; fixed effects only 
f_noREs <- formula(
  target ~ prior * competence * relevance * main_type
)
```

Here's a function for running the models:

```{r }

run_model <- function(form, data){

  ## some = 0, xor = 1
  contrasts(data$main_type)

  ## set priors
  priors <- set_prior("student_t(1, 0, 2)", class = "b")

  ## run and return model
  brm(
    formula      = form,
    data         = data,
    prior        = priors,
    sample_prior = T,
    control      = list(adapt_delta = 0.95),
    cores        = 4,
    iter         = 3000
  )

} 
```

And here is a function for returning the relevant result summary from a fitted model:

```{r }
## function to get prediction for a model fit
get_predictions <- function(model_SI) {
  model_SI %>% spread_draws(b_Intercept, b_prior, b_competence, b_relevance,
                            b_main_typexor, `b_prior:competence`,
                            `b_prior:relevance`, `b_competence:relevance`,
                            `b_prior:main_typexor`, `b_competence:main_typexor`,
                            `b_relevance:main_typexor`, `b_prior:competence:relevance`, 
                            `b_prior:competence:main_typexor`, `b_prior:relevance:main_typexor`,
                            `b_competence:relevance:main_typexor`, 
                            `b_prior:competence:relevance:main_typexor`) %>% 
    mutate(
      prior_xor = b_prior + `b_prior:main_typexor`,
      prior_some = b_prior,
      competence_xor = b_competence + `b_competence:main_typexor`,
      competence_some = b_competence,
      relevance_xor =  b_relevance + `b_relevance:main_typexor`,
      relevance_some = b_relevance
    ) -> model_SI_posteriors

  # check P that the effects are positive / negative / no effect present
  posterior_hypotheses <- model_SI_posteriors %>% 
    select(prior_xor, prior_some, 
           competence_xor, competence_some,
           relevance_xor, relevance_some) %>%
    gather(key, val) %>%
    group_by(key) %>% mutate(positive = mean(val > 0.05),
                             negative = mean(val < -0.05),
                             no = mean(val %>% between(-0.05, 0.05))) %>%
    summarise(positive_eff = mean(positive),
              no_eff = mean(no),
              negative_eff = mean(negative))
  posterior_hypotheses
}
```

# Fitting and inspecting the relevant models

Model fits:

```{r results = 'hide'}
model_SI         <- run_model(f_full,d_critical_clean) 
model_SI_RETitle <- run_model(f_RETitle,d_critical_clean)
model_SI_RESubj  <- run_model(f_RESubj,d_critical_clean)
model_SI_noREs   <- run_model(f_noREs,d_critical_clean)
```

We can now inspect the results for each model, and find that **including by-item random intercepts determines whether we see a main effect of prior for the 'or' items, irrespective of whether we include or omit the by-subject random effects**.
We may conclude from this that it is really the by-item random effects that "explain away" the main effect of prior on 'or' items.

```{r }
## We compare the results for three model which differ only in the random effects
## they take into account. We care about a potential main effect of 'prior' in the
## 'or' condition.

## (i)  full REs (subj. & item)  -> no effect 
get_predictions(model_SI)

## (ii) no subj., just item REs  -> no effect 
get_predictions(model_SI_RETitle)

## (iii) just subj., no item REs -> strong effect 
get_predictions(model_SI_RESubj)

## (iv) no REs                   -> strong effect 
get_predictions(model_SI_noREs)
```

# Probing which vignettes are most influential in producing/hiding the effect

Which vignettes does the model estimate large random by-item effect for?
Here's a function that prints the top 6 most extreme estimates for the random effects in question.

```{r}

extract_REs <- function(model_fit) {
  randEff <- brms::ranef(model_fit)$title %>% as.data.frame() %>% 
    as_tibble(rownames = "title")
  vignettes  %>% 
    full_join(randEff, by= 'title')
}

```

Applying this function to both model with by-item random effects, we see that these top six do not differ much between whether we include the by-subject random effects or not.

```{r}
## full model
extract_REs(model_SI) %>% arrange(Estimate.Intercept)
extract_REs(model_SI) %>% arrange(-Estimate.Intercept)

## model without by-subject REs
extract_REs(model_SI_RETitle) %>% arrange(Estimate.Intercept)
extract_REs(model_SI_RETitle) %>% arrange(-Estimate.Intercept)

```

To further find out how the addition of by-item REs affects the model fit, we can try fittin a model *with* by-item REs, but omitting vignettes which receive rather high/low estimates for their RE.
Here we gather the titles of the top-$n$ highest or lowest ranked vignettes for each trigger word.

```{r}
get_extreme_vignettes <- function(model = model_SI, n=5, trigger_type = 'some', high = T) {
  extract_REs(model) %>% arrange(Estimate.Intercept * ifelse(high, 1, -1)) %>% 
    filter(main_type == trigger_type) %>% 
    pull(title) %>% .[1:n]
}

hi_some <- get_extreme_vignettes(trigger_type = 'some', high = T)
lo_some <- get_extreme_vignettes(trigger_type = 'some', high = F)
hi_or   <- get_extreme_vignettes(trigger_type = 'xor', high = T)
lo_or   <- get_extreme_vignettes(trigger_type = 'xor', high = F)
```

So, let's fit some model on reduced data sets, excluding "extreme vignettes".

```{r, results = 'hide'}

predictions_for_reduced_data <- function(excluded_vignettes) {
  run_model(
    f_full, 
    d_critical_clean %>% filter(! title %in% excluded_vignettes)
  ) %>% get_predictions()  
}

predictions_for_reduced_data(hi_some)
predictions_for_reduced_data(lo_some)
predictions_for_reduced_data(hi_or)
predictions_for_reduced_data(lo_or)

```

It seems that the vignettes for 'some' which are estimated to have particularly high random intercepts are most conducive for whether the main effect shows or not.

### 2022-05-13 analysis/main/mf-exploration.Rmd
add joint posterior plots

@@
   target ~ prior * competence * relevance * main_type
+)
+
+## full model with RE ineractions by-subject
+f_full_inters <- formula(
+  target ~ prior * competence * relevance * main_type +
+    (1 + prior * competence * relevance * main_type || submission_id) +
+    (1 | title)
 )
@@
 model_SI_noREs   <- run_model(f_noREs,d_critical_clean)
+model_SI_fullREs <- run_model(f_full_inters, d_critical_clean)
 ```
@@
 It seems that the vignettes for 'some' which are estimated to have particularly high random intercepts are most conducive for whether the main effect shows or not.
+
+## Polina's shenanigans
+
+Check the joint posterior of fixed and random effects, to investigate which of them might cause the pattern of results that we observe. Analysis suggested by Roger.
+The following plots shows the joint posterior of fixed and random effects in the xor and some conditions, respectively.
+
+``` {r}
+# extract posterior contrast of interest of fixed effects
+posteriors <- model_SI_fullREs %>%
+  spread_draws(b_prior, `b_prior:main_typexor`, b_relevance, `b_relevance:main_typexor`, b_competence, `b_competence:main_typexor`,
+               sd_submission_id__prior, `sd_submission_id__prior:main_typexor`,
+               sd_submission_id__Intercept, sd_title__Intercept, 
+               sd_submission_id__relevance, `sd_submission_id__relevance:main_typexor`) %>%
+    # extract contrasts of interest, especially effect of syntax by-trial
+    mutate(prior_xor = b_prior + `b_prior:main_typexor`,
+           prior_some = b_prior,
+           relevance_xor = b_relevance + `b_relevance:main_typexor`,
+           relevance_some = b_relevance,
+           competence_xor = b_competence + `b_competence:main_typexor`,
+           competence_some = b_competence,
+           random_prior_xor = sd_submission_id__prior + `sd_submission_id__prior:main_typexor`,
+           random_prior_some = sd_submission_id__prior,
+           random_relevance_xor = sd_submission_id__relevance + `sd_submission_id__relevance:main_typexor`,
+           random_relevance_some = sd_submission_id__relevance
+           ) %>% # subject vs predicate syntax
+    select(prior_xor, prior_some, relevance_xor, relevance_some, competence_xor, competence_some,
+           random_prior_xor, random_prior_some, sd_submission_id__Intercept, sd_title__Intercept, random_relevance_xor, random_relevance_some) %>%
+    gather(key, val)
+
+```
+
+Helper function for extracting and plotting a heat map of two effects, given a df containing all relevant posterior samples / effects, the effects and a title for the plot. 
+```{r}
+get_and_plot_2d_posterior <- function(posteriors, fe, re, title) {
+  fe_selected <- posteriors %>% filter(key == fe) %>% rename(fe = val)
+  re_selected <- posteriors %>% filter(key == re) %>% rename(re = val) %>% select(-key)
+
+  df <- tibble(fe_selected, re_selected)
+
+  p <- ggplot(df, aes(x = fe, y = re)) +
+  stat_bin2d(aes(fill = after_stat(density))) +
+  ggtitle(title)
+
+  return (p)
+}
+```
+
+First, we explore how the fixed effects of prior behave against by-title random intercepts and by-subject random slope effects of prior.
+```{r}
+get_and_plot_2d_posterior(posteriors, "prior_xor", "sd_title__Intercept", "FE of prior for XOR against by-story random intercepts")
+#ggsave("FE_prior_RE_story_xor.pdf", width=7, height=5)
+
+get_and_plot_2d_posterior(posteriors, "prior_some", "sd_title__Intercept", "FE of prior for SOME against by-story random intercepts")
+
+get_and_plot_2d_posterior(posteriors, "prior_xor", "random_prior_xor", "FE of prior for XOR against by-subj prior RE")
+
+get_and_plot_2d_posterior(posteriors, "prior_some", "random_prior_some", "FE of prior for SOME against by-subj prior RE")
+```
+
+Now, double check if there is a "credit assignment" problem for other factors.
+```{r}
+get_and_plot_2d_posterior(posteriors, "competence_xor", "sd_title__Intercept", "FE of competence for XOR against by-story random intercepts")
+
+get_and_plot_2d_posterior(posteriors, "competence_some", "sd_title__Intercept", "FE of competence for SOME against by-story random intercepts")
+
+get_and_plot_2d_posterior(posteriors, "relevance_xor", "sd_title__Intercept", "FE of relevance for XOR against by-story random intercepts")
+
+get_and_plot_2d_posterior(posteriors, "relevance_some", "sd_title__Intercept", "FE of relevance for SOME against by-story random intercepts")
+```
+
+Explore an xor-only model.
+```{r}
+f_xor <- formula(
+  target ~ prior * competence * relevance +
+    (1 + prior + competence + relevance || submission_id) +
+    (1 | title)
+)
+model_xor<- run_model(f_xor, d_critical_clean %>% filter(main_type == "xor")) 
+summary(model_xor)
+```
+
+```{r}
+# get the probability of a negative effect for or
+model_xor %>% spread_draws(b_prior) %>%
+  select(b_prior) %>%
+  gather(key, val) %>%
+  group_by(key) %>% mutate(positive = mean(val > 0.05),
+                             negative = mean(val < -0.05),
+                             no = mean(val %>% between(-0.05, 0.05))) %>%
+    summarise(positive_eff = mean(positive),
+              no_eff = mean(no),
+              negative_eff = mean(negative))
+
+```
